---
title: "Still Face Raw Import"
date: "2026-08-20"
editor: 
  markdown: 
    wrap: 72
execute:
  freeze: true
---

# Overview

This page documents the technical import and scoring pipeline for the
**6-month Still Face behavioral-coding data**. The source files are
per-subject Excel event logs containing timestamped starts/stops for
coded infant and caregiver behaviors across Baseline, Still Face, and
Reunion.

::: {.callout-tip appearance="simple" icon="false"}
**Already read the Variable Dictionary?** [Jump directly to the
technical workflow](#sec-technical-workflow).
:::

# Still Face Paradigm

The Still Face Paradigm [@tronick] is a brief, structured interaction
task used to study early social-emotional regulation and the
caregiver-infant relationship. It has three phases, or episodes, which
are described below:

- **Baseline (bl)** - the mother engages in normal face-to-face play
  with her infant, establishing each dyad's typical interactive style
- *a 15 second transition, where the mother looks away before returning
  her gaze towards her infant*
- **Still Face (sf)** - holding a neutral, unmoving expression for a set
  interval, while remaining present but unresponsive
- **Reunion (re)** - the mother resumes normal interaction, and the
  dyad's recovery is observed

Because infants expect consistent, responsive interaction, the temporary
withdrawal of maternal responsiveness acts as a mild stressor that
reveals how the infant regulates distress and how the mother-infant dyad
repairs afterward.

Trained coders capture these interactions by continuously marking the
onset and offset of discrete behavioral states (e.g., gaze direction,
facial affect, vocalization, and motor behavior) throughout each
episode. The resulting raw event logs contain timestamped start/stop
events for each behavior and subject. The import pipeline validates and
transforms those event-level records into the scored variables
summarized in the dictionary below.

::: {.callout-tip appearance="simple" icon="false"}
**Scientific relevance of Individual and dyadic differences**: the
*magnitude* of the still-face effect and the *speed/completeness* of
reunion recovery are used as indicators of infant self-regulation,
mother-infant synchrony, and attachment-relevant processes. Blunted
still-face reactivity, minimal recovery, or atypical patterns (e.g.,
disorganized behavior, dysregulated affect) have been linked in the
literature to risk factors such as maternal depression, prenatal
substance exposure, and early adversity.
:::

::: {.callout-note appearance="simple" icon="false"}
### Portfolio Navigation

**1. Variable Dictionary & Project Overview** → **2. Raw Import &
Scoring Pipeline**

[Variable Dictionary](variable-dictionary.qmd) · [Raw Import & Scoring
Pipeline](6M_Still_Face_Raw_Import.qmd)
:::

# Technical Workflow {#sec-technical-workflow}

The companion [Variable Dictionary](variable-dictionary.qmd) introduces
the paradigm, shows a raw-data example, and explains how coded behaviors
map onto analysis variables.

**The pipeline:**

- imports every subject's raw event log;
- validates file IDs and expected Subject/Behavior codes;
- identifies Baseline, Still Face, and Reunion episode boundaries;
- assigns each coded event to the appropriate episode;
- computes duration, count, and joint-attention measures;
- reshapes one subject at a time from event-level long data to one-row
  scored output;
- prefers consensus-coded files when present and prevents duplicate
  consensus sources;
- validates the merged dataset and compares calculated episode
  boundaries against the hand-checked coding log; and
- produces one wide scored dataset for downstream analysis.

## 1. Reference Tables

### Episode markers

Certain codes are only used to mark the start and end of an episode, and
we can use these to mark the edges between Baseline, Still Face, and
Reunion. This allows us to assign each row of behavior to one of the
three episodes. They are not themselves scored behavior.

::: {.callout-tip appearance="simple" icon="false"}
Returning to the raw data example, the behavior marker codes are in bold
and the baseline and 15 second transition can therefore be assigned to
the codes between them.
:::

```{r}
#| label: file-paths
#| include: false

# EDIT THIS LINE: point PROJECT_ROOT at your own "ff (cast.casbox )" Box folder.
PROJECT_ROOT <- "C:/Users/rcm5550/Box/ff (cast.casbox )"

# Raw (non-consensus) event log files -- one ID-named subfolder per subject.
PATH_EVENT_LOGS <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding/Still Face - SFF/Coded Still Face Videos - XLX Files/6 Month XLX Files/General")

# Reliability Files
PATH_EVENT_LOGS_RELIABILITY <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding/Still Face - SFF/Coded Still Face Videos - XLX Files/6 Month XLX Files/Reliability")

# Reliability Training Files
PATH_EVENT_LOGS_TRAINING <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding/Still Face - SFF/Coded Still Face Videos - XLX Files/6 Month XLX Files/Training")

# Coding Log Path
PATH_CODING_LOG  <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding/Still Face - SFF/Coding Log SFF Still Face.xlsx")


```

**Load Packages**
```{r}
library(readxl)     # Import raw Excel event logs and supporting workbooks
library(flextable)  # Format variable dictionaries and summary tables for rendered output
library(officer)    # Support flextable rendering and document-compatible table formatting
library(dplyr)      # Clean, transform, join, summarize, and validate behavioral data
library(tidyr)      # Reshape event-level data between long and wide formats
library(stringr)    # Clean and standardize text fields, file names, and variable labels
library(purrr)      # Iterate over subject files and apply scoring functions across datasets
```


```{r}
#| label: still-face-data-episodes
#| echo: false

library(readxl)
library(flextable)
library(officer)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)

raw <- read_excel("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Still Face/6M Event Logs/111_SFF_SF6M_64 - Event Logs.xlsx")

# Two illustrative windows from the raw BORIS export:
# 1) The opening seconds of Baseline, showing several behaviors starting at once
# 2) The Baseline -> Still Face transition, showing the epoch flags the syntax reads
demo <- raw %>%
  filter(Time_Relative_sf <= 2.12 | (Time_Relative_sf >= 119.995 & Time_Relative_sf <= 120.02)) %>%
  transmute(
    Segment  = if_else(Time_Relative_sf <= 2.12, "Opening of Baseline", "Baseline \u2192 Still Face transition"),
    `Time (s)`     = round(Time_Relative_sf, 3),
    `Duration (s)` = round(Duration_sf, 3),
    Subject,
    Behavior,
    `Event Type` = Event_Type
  )

ft_episodes <- flextable(demo) %>%
  merge_v(j = "Segment") %>%
  theme_box() %>%
  set_table_properties(layout = "fixed") %>%
  width(j = "Segment", width = 1.1) %>%
  width(j = "Time (s)", width = 0.85) %>%
  width(j = "Duration (s)", width = 0.9) %>%
  width(j = "Subject", width = 0.65) %>%
  width(j = "Behavior", width = 2.2) %>%
  width(j = "Event Type", width = 0.8) %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%
  align(j = c("Time (s)", "Duration (s)"), align = "right", part = "all") %>%
  font(j = "Behavior", fontname = "Consolas") %>%
  bold(i = ~ grepl("^15 S|^BASELINE|^REUNION", Behavior), j = "Behavior") %>%
  hline(i = 8, border = fp_border(color = "#999999", width = 1.5))

ft_episodes
```

```{r}
#| label: episode-markers-list
# Full list of episode marker codes
EPISODE_MARKERS <- c(
  "BASELINE 15 S REUNION NO COPE",
  "15 S NO INFANT VOCAL",
  "15 S NO FACIAL EXPRESSION",
  "15 S STILL FACE NO AFFECT",
  "15 S STILL FACE NO GAZE",
  "15 S STILL FACE NO TALKING/TOUCHING",
  "REUNION TRANSITION NO EXPRESSION",
  "REUNION TRANSITION NO VOCAL")
```

### Behavior reference table

`BEHAVIOR_REFERENCE` lists every Subject/Behavior combination we expect
to see in a raw file. Built into the pipeline is a data validation
check - the import process stops if a file contains a Subject/Behavior
code that isn't in this table -- most likely cause would be a typo in
the coding software, or an addition of a code after the scoring pipeline
was created.

`category` groups related behaviors together for readability and
organization. `var_stem` is the short suffix that ends up in the final
prefixed variable names (see the @score-spec section below).

```{r}
#| label: behavior-reference
BEHAVIOR_REFERENCE <- tribble(
  ~Subject,  ~Behavior,                          ~category,               ~var_stem,

  # --- Infant gaze: who the infant is looking at ---
  # Infant looking at mom
  "Infant",  "Looking at Mom",                   "infant_gaze",        "yeslook",
  # Infant looking away from mom
  "Infant",  "Looking Away",                     "infant_gaze",        "notlook",
  # Infant's gaze could not be coded (face/eyes not visible on video)
  "Infant",  "Infant Gaze Not Visible",           "infant_gaze",        "gaze_nv",

  # --- Infant facial expression ---
  # Neutral facial expression
  "Infant",  "Neutral Expression",               "infant_expression",  "neuexp",
  # Smiling / positive facial expression
  "Infant",  "Smile/Positive Expression",        "infant_expression",  "posexp",
  # Frowning / negative facial expression
  "Infant",  "Frown/Negative Expression",        "infant_expression",  "negexp",
  # Infant's face could not be coded (not visible on video)
  "Infant",  "Infant Face Not Visible",           "infant_expression",  "face_nv",

  # --- Infant vocalization ---
  # Infant not vocalizing
  "Infant",  "Infant Silent",                    "infant_vocal",       "silvoc",
  # Neutral vocalization
  "Infant",  "Neutral Vocal",                    "infant_vocal",       "neuvoc",
  # Laughing / positive vocalization
  "Infant",  "Laughing/Positive Vocal",          "infant_vocal",       "posvoc",
  # Crying / negative vocalization
  "Infant",  "Crying/Negative Vocal",            "infant_vocal",       "negvoc",

  # --- Infant orientation (Still Face episode only) ---
  # Infant oriented toward an object rather than the parent
  "Infant",  "Object Orientation",               "infant_orientation", "obj",
  # Infant oriented toward the parent
  "Infant",  "Parent Orientation",               "infant_orientation", "par",

  # --- Infant discrete self-soothing/escape actions (Still Face only) ---
  # Infant engaging in a self-comforting behavior
  "Infant",  "Infant Comforting",                "infant_action",      "comf",
  # Infant engaging in an escape behavior (e.g., arching away)
  "Infant",  "Infant Escape",                    "infant_action",      "esc",

  # --- Mother/parent affect ---
  # High-intensity negative affect
  "Mother",  "High Negative Affect",             "mother_affect",      "negaffH",
  # Low-intensity negative affect
  "Mother",  "Negative Affect",                  "mother_affect",      "negaffL",
  # Neutral affect
  "Mother",  "Neutral Affect",                   "mother_affect",      "neuaff",
  # Low-intensity positive affect
  "Mother",  "Positive Affect",                  "mother_affect",      "posaffL",
  # High-intensity positive affect
  "Mother",  "High Positive Affect",             "mother_affect",      "posaffH",
  # Affect not visible
  "Mother", "Mom Affect Not Visible",            "mother_affect",      "noaffect",  # ID107

  # --- Mother talking/touching the infant ---
  # Mom is talking to and/or touching the infant
  "Mother",  "Mom Talking/Touching Infant",      "mother_talktouch",   "tlktch",
  # Mom is NOT talking to or touching the infant
  "Mother",  "Mom NOT Talking/Touching Infant",  "mother_talktouch",   "nottlktch"
)
```

## 2. Import one subject's raw event log

This section builds the tools to import in a single subject's raw Excel
file and do an initial data quality check before the raw data is scored.

**Get the subject ID from a file name**

```{r}
#| label: extract-id-function
# Every event log file name starts with the subject's numeric ID
# (E.g. "567_SFF_SF6M - Raw Data.xlsx" -> "567")

# This function pulls  ID out of the file name
extract_id_from_filename <- function(filename) {

  # Grab the 3 digits at the very start of the (base, not full-path) file name
  id <- str_extract(basename(filename), "^\\d+")

  # If no leading digits were found, stops rather than continuing with a missing ID
  if (is.na(id)) {
    stop("Could not identify a subject ID from file: ", basename(filename))
  }

  # Returns the extracted ID
  id
}
```

**Import one subject's file and run basic checks**

```{r}
#| label: import-function
# ── Import and Validate One Raw Event Log ───────────────────────────────────────
import_event_log <- function(filename_full) {

    # ── Identify the Source File ──────────────────────────────────────────────────
  # Keep the short file name around for readable error messages
  filename_short <- basename(filename_full)

  # Get the subject ID using function built above
  filename_id <- extract_id_from_filename(filename_short)

    # ── Import the Raw Excel Export ───────────────────────────────────────────────
  # Read the raw Excel export
  data <- read_excel(filename_full) %>%
    as.data.frame()  # Convert from a tibble to a plain data frame
  
    # ── Detect and Remove Blank Template Rows ─────────────────────────────────────
  # Identify blank template rows: no Behavior coded at all, and both the
  # Relative time and duration are exactly 0 -- these are empty leftover
  # Rows in the export
  blank_rows <- data %>%
    filter(is.na(Behavior), Time_Relative_sf == 0, Duration_sf == 0)
  
  if (nrow(blank_rows) > 0) {
    # Record this file's blank rows in the running log (<<- writes to the
    # BLANK_ROW_LOG for lead coder to review)
    BLANK_ROW_LOG <<- bind_rows(
      BLANK_ROW_LOG,
      tibble(ID = filename_id, file = filename_short, n_blank_rows = nrow(blank_rows))
    )
    # Drop  blank rows
    data <- data %>%
      filter(!(is.na(Behavior) & Time_Relative_sf == 0 & Duration_sf == 0))
  }

    # ── Standardize Time and Record Order ─────────────────────────────────────────
  # Time_Relative_hms comes in as a full datetime with a fake Excel-created date
  # Reformat to "HH:MM:SS" text
  data <- data %>%
    mutate(Time_Relative_hms = format(Time_Relative_hms, "%H:%M:%S"))

  # Number every row in a new column called "Code"
  # Creates an event-order count the script uses to find episode boundaries
  data <- data %>%
    tibble::rowid_to_column("Code") %>%
    # Duplicate the "Subject ID" column into a plain, easy-to-type "ID" column
    mutate(ID = as.character(`Subject ID`))  

    # ── Validate Subject ID ───────────────────────────────────────────────────────
  # Collect every distinct ID value found inside the file itself
  internal_ids <- unique(data$ID)

  # Data quality check: stop if the file contains more than one Subject ID
  if (length(internal_ids) != 1) {
    stop(
      "File does not contain exactly one Subject ID: ", filename_short,
      ". Found: ", paste(internal_ids, collapse = ", ")
    )
  }

  # Data quality check: Stop if the ID inside the file doesn't match the ID in the file name
  if (internal_ids != filename_id) {
    stop(
      "File name ID (", filename_id, ") does not match internal Subject ID (",
      internal_ids, ") in file: ", filename_short
    )
  }

  # Pull out every Subject/Behavior combination in this file that ISN'T a episode marker
  observed_behaviors <- data %>%
    filter(!(Behavior %in% EPISODE_MARKERS)) %>%
    distinct(Subject, Behavior)

  # Compare those observed combinations against BEHAVIOR_REFERENCE, keeping only the ones that do NOT have a match there
  unmapped <- observed_behaviors %>%
    anti_join(BEHAVIOR_REFERENCE, by = c("Subject", "Behavior"))

  # Data quality check: Stop if anything unmapped was found
  if (nrow(unmapped) > 0) {
    stop(
      "Unrecognized Subject/Behavior combination(s) in file ", filename_short,
      ": ", paste(unmapped$Subject, unmapped$Behavior, sep = "/", collapse = "; "),
      ". Add to BEHAVIOR_REFERENCE or fix the coding before re-running."
    )
  }

  # Return the cleaned, validated data for this one subject
  data
}
```

## 3. Episode anchors

"Anchors" are the `Code` numbers marking where each episode (Baseline,
Still Face, Reunion) starts and ends for one subject.

::: callout-important
Everything after this depends on these episode boundaries being found
correctly, so this function stops immediately if an episode marker is
missing or the anchors come out of order.
:::

```{r}
#| label: episode-anchor-function
# ── Identify Episode Anchors ────────────────────────────────────────────────────
get_episode_anchors <- function(data) {

  # Subject ID, kept for clearer error messages below
  filename_note <- unique(data$ID)

  # Baseline starts at the very first coded row in the file
  bl_start <- min(data$Code)

  # Reunion ends at the very last coded row in the file
  reunion_end <- max(data$Code)

  # "15 S NO INFANT VOCAL" brackets the baseline-to-still-face transition
  # Buffer: baseline ends where this marker STARTS, and still face doesn't
  # Begin until this marker STOPS. The gap in between (a ~15-30s "do not code"
  # Transition) is correctly left unclassified (episode = NA)

  # Every "State start" row for this marker
  no_infant_vocal_start <- data %>%
    filter(Behavior == "15 S NO INFANT VOCAL", Event_Type == "State start")

  # Every "State stop" row for this marker
  no_infant_vocal_stop <- data %>%
    filter(Behavior == "15 S NO INFANT VOCAL", Event_Type == "State stop")

  # Data quality check: Stop if either the start or the stop event is missing
  if (nrow(no_infant_vocal_start) == 0 | nrow(no_infant_vocal_stop) == 0) {
    stop("Could not find both a start and a stop for '15 S NO INFANT VOCAL' marker for ID: ", filename_note)
  }

  # Baseline ends at the first "start" row of the marker
  bl_end <- first(no_infant_vocal_start$Code)

  # Still Face begins at the first "stop" row of the marker
  sf_start <- first(no_infant_vocal_stop$Code)

  # Same start/stop logic for the still-face-to-reunion transition buffer,
  # This time using the "REUNION TRANSITION NO EXPRESSION" marker

  # Every "State start" row for this marker
  reunion_transition_start <- data %>%
    filter(Behavior == "REUNION TRANSITION NO EXPRESSION", Event_Type == "State start")

  # Every "State stop" row for this marker
  reunion_transition_stop <- data %>%
    filter(Behavior == "REUNION TRANSITION NO EXPRESSION", Event_Type == "State stop")

  # Data quality check: Stop if either the start or the stop event is missing
  if (nrow(reunion_transition_start) == 0 | nrow(reunion_transition_stop) == 0) {
    stop("Could not find both a start and a stop for 'REUNION TRANSITION NO EXPRESSION' marker for ID: ", filename_note)
  }

  # Still Face ends at the first "start" row of this marker
  sf_end <- first(reunion_transition_start$Code)

  # Reunion begins at the first "stop" row of this marker
  reunion_start <- first(reunion_transition_stop$Code)

  # Bundle all six anchor points into a one-row tibble, labelled with the subject ID
  anchors <- tibble(
    ID = filename_note,
    bl_start = bl_start, bl_end = bl_end,
    sf_start = sf_start, sf_end = sf_end,
    reunion_start = reunion_start, reunion_end = reunion_end
  )

  # Put the six anchor points in one vector, in the order they SHOULD occur
  anchor_sequence <- c(bl_start, bl_end, sf_start, sf_end, reunion_start, reunion_end)

  # Data quality check: Stop if that sequence is incorrect
  # (E.g. a marker miscoded earlier/later than it should be)
  if (is.unsorted(anchor_sequence, strictly = FALSE)) {
    stop("Episode anchors are not in increasing order for ID: ", filename_note)
  }

  # Return the anchors tibble
  anchors
}
```

## 4. Label every row with its episode

`episode` records which episode each row belongs to: `1` = Baseline,
`2`= Still Face, `3` = Reunion. Rows that fall in the transition buffers
between episodes (including the boundary marker rows themselves) are
expected to come out as `NA` to be correctly excluded from scoring.

```{r}
#| label: episode-function
# ── Assign Events to Episodes ───────────────────────────────────────────────────
assign_episode <- function(data, anchors) {
  data %>%
    mutate(episode = case_when(
      # Row falls within the Baseline anchor range
      Code >= anchors$bl_start & Code <= anchors$bl_end ~ 1,
      # Row falls within the Still Face anchor range
      Code >= anchors$sf_start & Code <= anchors$sf_end ~ 2,
      # Row falls within the Reunion anchor range
      Code >= anchors$reunion_start & Code <= anchors$reunion_end ~ 3,
      # Anything else (transition buffers, marker rows) is left as missing
      TRUE ~ NA_real_
    ))
}
```

## 5. Put it together: import + anchor + label one subject

```{r}
#| label: label-raw-function
# ── Run the Single-Subject Import Workflow ──────────────────────────────────────
import_and_label_one_subject <- function(filename_full) {

  # Import and validate this subject's raw file
  data <- import_event_log(filename_full)

  # Find this subject's episode boundary "anchors"
  anchors <- get_episode_anchors(data)

  # Add the episode (episode) label to every row
  data <- assign_episode(data, anchors)

  # Keep only the columns scoring actually needs
  data %>%
    select(ID, Code, Subject, Behavior, Event_Type, Time_Relative_sf, Time_Relative_hms, Duration_sf, episode)
}
```

## 6. Scoring building blocks

Duration of behavior is only calculated from "State start" rows, as
these rows have the full duration of that state, so "State stop" rows
are filtered out here to avoid double-counting the same time span.

We then create four small functions that every score in the final
dataset is built from: - total duration of a behavior - count of a
behavior - overlap between two behaviors (for joint attention) - check
for whether an episode had any real coding at allTotal seconds one
subject spent in one behavior during one episode.

```{r}
#| label: duration-function
# ── Score Behavior Duration ─────────────────────────────────────────────────────
compute_duration <- function(data, subj, behavior, episode_num) {
  data %>%
    # Keep only rows matching this subject, this behavior, "State start" events, and this episode
    filter(Subject == subj, Behavior == behavior,
           Event_Type == "State start", episode == episode_num) %>%
    # Add up the duration of every matching row
    summarise(total = sum(Duration_sf)) %>%
    # Pull that single total out as a plain number
    pull(total)
}
```

```{r}
#| label: count-function
# ── Score Behavior Counts ───────────────────────────────────────────────────────
compute_count <- function(data, subj, behavior, episode_num) {
  data %>%
    # Keep only rows matching this subject, this behavior, "State start"
    # Events, and this episode -- each such row is one "entry" into the
    # Behavior
    filter(Subject == subj, Behavior == behavior,
           Event_Type == "State start", episode == episode_num) %>%
    # Count the matching rows
    nrow()
}
```

```{r}
#| label: overlap-function
# ── Score Behavioral Overlap ──────────────────────────────────────────────────── (i.e. mom talking/touching WHILE infant looks at mom)
compute_overlap_duration <- function(data, subject_a, behavior_a, subject_b, behavior_b, episode) {

  # Build every interval (start, end) for behavior A in this episode
  intervals_a <- data %>%
    filter(Subject == subject_a, Behavior == behavior_a,
           Event_Type == "State start", episode == episode) %>%
    # Start_a = when it began; end_a = start plus how long it lasted
    transmute(start_a = Time_Relative_sf, end_a = Time_Relative_sf + Duration_sf)

  # Build every interval (start, end) for behavior B in this episode
  intervals_b <- data %>%
    filter(Subject == subject_b, Behavior == behavior_b,
           Event_Type == "State start", episode == episode) %>%
    # Start_b = when it began; end_b = start plus how long it lasted
    transmute(start_b = Time_Relative_sf, end_b = Time_Relative_sf + Duration_sf)

  # If either behavior never occurred in this episode, there can be no
  # Overlap -- returns 0 rather than trying to cross an empty table
  if (nrow(intervals_a) == 0 | nrow(intervals_b) == 0) {
    return(0)
  }

  # Pair every interval of A with every interval of B (a full cross join),
  crossing(intervals_a, intervals_b) %>%
    # Then compute how much each pair overlaps: the later of the two
    # Starts to the earlier of the two ends, floored at 0 (pmax/pmin) so
    # Non-overlapping pairs contribute nothing negative
    mutate(overlap = pmax(0, pmin(end_a, end_b) - pmax(start_a, start_b))) %>%
    # Add up the overlap across every pair,
    summarise(total = sum(overlap)) %>%
    # And pull that single total out as a plain number
    pull(total)
}
```

```{r}
#| label: episode-has-codes-function
# ── Check Whether an Episode Contains Valid Codes ───────────────────────────────: Distinguishes "episode was coded and at least one behavior happened" resulting in behaviors not coded as 0 from "episode was never coded at all" and then behaviors are coded as NA
episode_has_codes <- function(data, episode_num) {
  data %>%
    # Rows in this episode that are NOT boundary markers
    filter(episode == episode_num, !(Behavior %in% EPISODE_MARKERS)) %>%
    # TRUE if there's at least one such row
    nrow() > 0
}
```

```{r}
#| label: blank-uncoded-episodes-function
# Lookup: episode number -> the variable-name prefix used for that episode
# 6M Still Face - Baseline: v2sfbl; 6M Still Face - Still Face: v2sfsf; 6M Still Face - Reunion: v2sfre;
EPISODE_PREFIX <- c(`1` = "v2sfbl", `2` = "v2sfsf", `3` = "v2sfre")

# ── Blank Uncoded Episode Scores ────────────────────────────────────────────────
# V2sf<episode>... column for any episode that had zero real codes (per
# Episode_has_codes() above), so an uncoded episode reads as missing rather
# Than as all zeros
blank_uncoded_episodes <- function(scores, data) {
  # Check each of the three episodes in turn
  for (episode in c(1, 2, 3)) {
    # Only act on episodes that were never actually coded
    if (!episode_has_codes(data, episode)) {
      # Get this episode's variable-name prefix (e.g. "v2sfsf")
      prefix <- EPISODE_PREFIX[[as.character(episode)]]
      # Find every column in `scores` that starts with that prefix
      cols_to_blank <- names(scores)[str_starts(names(scores), prefix)]
      # Set every one of those columns to NA
      scores[cols_to_blank] <- NA_real_
    }
  }
  # Return the (possibly blanked) scores tibble
  scores
}
```

## 7. Score spec: one row per duration/count variable

`SCORE_SPEC` is the master list that drives the long-to-wide reshape in
the next section. Each row has: "final variable name = subject's
behavior, in this episode, as this metric (duration or count)."

```{r}
#| label: score-spec
SCORE_SPEC <- tribble(
  ~var_name,          ~subject,  ~behavior,                          ~episode, ~metric,

  # --- Infant vocal: baseline / still face / reunion ---
   # Negative vocal duration, one row per episode
  "v2sfblnegvoc      ", "Infant", "Crying/Negative Vocal", 1, "duration",
  "v2sfsfnegvoc      ", "Infant", "Crying/Negative Vocal", 2, "duration",
  "v2sfrenegvoc      ", "Infant", "Crying/Negative Vocal", 3, "duration",
  
   # Neutral vocal duration, one row per episode
  "v2sfblneuvoc      ", "Infant", "Neutral Vocal", 1, "duration",
  "v2sfsfneuvoc      ", "Infant", "Neutral Vocal", 2, "duration",
  "v2sfreneuvoc      ", "Infant", "Neutral Vocal", 3, "duration",
 
   # Positive vocal duration, one row per episode
  "v2sfblposvoc      ", "Infant", "Laughing/Positive Vocal", 1, "duration",
  "v2sfsfposvoc      ", "Infant", "Laughing/Positive Vocal", 2, "duration",
  "v2sfreposvoc      ", "Infant", "Laughing/Positive Vocal", 3, "duration",

   # Silent duration, one row per episode
  "v2sfblsilvoc      ", "Infant", "Infant Silent", 1, "duration",
  "v2sfsfsilvoc      ", "Infant", "Infant Silent", 2, "duration",
  "v2sfresilvoc      ", "Infant", "Infant Silent", 3, "duration",

  # ------------------------------------------------------------------------------
 
  # --- Infant expression: baseline / still face / reunion ---
   # Negative expression duration, one row per episode
  "v2sfblnegexp      ", "Infant", "Frown/Negative Expression", 1, "duration",
  "v2sfsfnegexp      ", "Infant", "Frown/Negative Expression", 2, "duration",
  "v2sfrenegexp      ", "Infant", "Frown/Negative Expression", 3, "duration",
 
   # Neutral expression duration, one row per episode
  "v2sfblneuexp      ", "Infant", "Neutral Expression", 1, "duration",
  "v2sfsfneuexp      ", "Infant", "Neutral Expression", 2, "duration",
  "v2sfreneuexp      ", "Infant", "Neutral Expression", 3, "duration",
 
   # Positive expression duration, one row per episode
  "v2sfblposexp      ", "Infant", "Smile/Positive Expression", 1, "duration",
  "v2sfsfposexp      ", "Infant", "Smile/Positive Expression", 2, "duration",
  "v2sfreposexp      ", "Infant", "Smile/Positive Expression", 3, "duration",

  # ------------------------------------------------------------------------------
 
  # --- Infant gaze: baseline / reunion only (blank at still face) ---
   # Looking-at-mom duration, baseline and reunion
  "v2sfblyeslook     ", "Infant", "Looking at Mom", 1, "duration",
  "v2sfreyeslook     ", "Infant", "Looking at Mom", 3, "duration",
  
   # Looking-away duration, baseline and reunion
  "v2sfblnotlook     ", "Infant", "Looking Away", 1, "duration",
  "v2sfrenotlook     ", "Infant", "Looking Away", 3, "duration",

  # --- Infant orientation: still face only ---
   # Object-oriented duration, still face only
  "v2sfsfobj         ", "Infant", "Object Orientation", 2, "duration",
  
   # Parent-oriented duration, still face only
  "v2sfsfpar         ", "Infant", "Parent Orientation", 2, "duration",
  
  # ------------------------------------------------------------------------------
  
  # --- Infant discrete actions: still face only ---
   # Comforting duration, still face only
  "v2sfsfcomf        ", "Infant", "Infant Comforting", 2, "duration",
 
   # Escape duration, still face only
  "v2sfsfesc         ", "Infant", "Infant Escape", 2, "duration",
 
   # Comforting COUNT (number of separate instances), still face only
  "v2sfsfcomfcount   ", "Infant", "Infant Comforting", 2, "count",
 
   # Escape COUNT (number of separate instances), still face only
  "v2sfsfesccount    ", "Infant", "Infant Escape", 2, "count",

  # ------------------------------------------------------------------------------
  
  # --- Mom affect: baseline / reunion only (blank at still face) ---
   # High negative affect duration, baseline and reunion
  "v2sfblnegaffH     ", "Mother", "High Negative Affect", 1, "duration",
  "v2sfrenegaffH     ", "Mother", "High Negative Affect", 3, "duration",
  
  # Low negative affect duration, baseline and reunion
  "v2sfblnegaffL     ", "Mother", "Negative Affect", 1, "duration",
  "v2sfrenegaffL     ", "Mother", "Negative Affect", 3, "duration",
 
   # Neutral affect duration, baseline and reunion
  "v2sfblneuaff      ", "Mother", "Neutral Affect", 1, "duration",
  "v2sfreneuaff      ", "Mother", "Neutral Affect", 3, "duration",
 
   # Low positive affect duration, baseline and reunion
  "v2sfblposaffL     ", "Mother", "Positive Affect", 1, "duration",
  "v2sfreposaffL     ", "Mother", "Positive Affect", 3, "duration",
 
   # High positive affect duration, baseline and reunion
  "v2sfblposaffH     ", "Mother", "High Positive Affect", 1, "duration",
  "v2sfreposaffH     ", "Mother", "High Positive Affect", 3, "duration",

  # ------------------------------------------------------------------------------
  
  # --- Mom talk/touch: baseline / reunion only ---
  # (Still Face is excluded by task design -- mom isn't allowed to
  # Talk/touch during Still Face, so there's nothing to score there.)
  # Talking/touching duration, baseline and reunion
  "v2sfbltlktch      ", "Mother", "Mom Talking/Touching Infant", 1, "duration",
  "v2sfretlktch      ", "Mother", "Mom Talking/Touching Infant", 3, "duration",
 
   # NOT talking/touching duration, baseline and reunion
  "v2sfblnotlktch    ", "Mother", "Mom NOT Talking/Touching Infant", 1, "duration",
  "v2sfrenotlktch    ", "Mother", "Mom NOT Talking/Touching Infant", 3, "duration"
  
) %>%
  # The extra spaces padded onto every var_name above are only there so the
  # Table lines up visually in the source code. Without trimming them off
  # Here, those spaces become part of the actual variable name, and
  # Pivot_wider() (below) bakes trailing-space column names into the
  # Output -- which then breaks every `scores$v2sf...` reference downstream
  mutate(var_name = str_trim(var_name))
```

## 8. Reshape one subject to wide, and add joint attention

This is the long-to-wide transformation step: for one subject, look up
every value listed in `SCORE_SPEC`, then pivot those values into one
wide row.

```{r}
#| label: compute-spec-values
# ── Transform Event-Level Data to Scored Variables ──────────────────────────────
compute_spec_values_one_subject <- function(data) {
  SCORE_SPEC %>%
    mutate(
      # Pmap_dbl() walks the four spec columns (subject, behavior, episode, metric) row by row and returns one number per row
      value = pmap_dbl(
        list(subject, behavior, episode, metric),
        function(subj, behavior, episode, metric) {
          if (metric == "duration") {
            # This row wants a total duration
            compute_duration(data, subj, behavior, episode)
          } else {
            # This row wants a count
            compute_count(data, subj, behavior, episode)
          }
        }
      )
    ) %>%
    # Keep only the variable name and its computed value
    select(var_name, value) %>%
    # Pivot from one-row-per-variable (long) to one-row-total (wide), using var_name as the new column names
    pivot_wider(names_from = var_name, values_from = value)
}
```

```{r}
#| label: raw-scores-function
# ── Calculate One Subject's Full Score Set ──────────────────────────────────────
compute_raw_scores_one_subject <- function(data) {

  # This subject's ID (data has already been filtered to one subject)
  subject_id <- unique(data$ID)

  # Every SCORE_SPEC-driven duration/count value, reshaped wide
  spec_values <- compute_spec_values_one_subject(data)

  # Start a one-row tibble with just the ID,
  scores <- tibble(ID = subject_id) %>%
    # Then attach all of the SCORE_SPEC columns next to it
    bind_cols(spec_values)

  # Joint attention: Baseline and Reunion only
  scores$v2sfbljntatn <- compute_overlap_duration(
    data, "Mother", "Mom Talking/Touching Infant", "Infant", "Looking at Mom", episode = 1
  )
  scores$v2sfrejntatn <- compute_overlap_duration(
    data, "Mother", "Mom Talking/Touching Infant", "Infant", "Looking at Mom", episode = 3
  )

  # If an episode has zero behavior codes, make all behaviors missing (NA)
  scores <- blank_uncoded_episodes(scores, data)

  # Pull subject's episode anchors (start/end code numbers)
  anchors <- get_episode_anchors(data)

  # Add episode boundaries as new data rows
  scores <- add_episode_boundary_times(scores, data, anchors)

  # Return this subject's completed one-row wide dataset
  scores
}
```

**Episode start/end times**

```{r}
#| label: episode-boundary-times-function
# ── Attach Episode Boundary Times ───────────────────────────────────────────────
add_episode_boundary_times <- function(scores, data, anchors) {

  # One row per episode: which number it is, its prefix, and which anchor
  # Columns mark its start/end Code numbers
  episode_info <- tribble(
    ~episode, ~prefix,   ~start_col,       ~end_col,
    1,        "v2sfbl",  "bl_start",       "bl_end",
    2,        "v2sfsf",  "sf_start",       "sf_end",
    3,        "v2sfre",  "reunion_start",  "reunion_end"
  )

  # Walk through the three episodes one at a time
  for (i in seq_len(nrow(episode_info))) {

    # This episode's number (1/2/3)
    episode <- episode_info$episode[i]
    # This episode's variable-name prefix (e.g. "v2sfbl")
    prefix <- episode_info$prefix[i]
    # The Code number where this episode starts, from `anchors`
    start_code <- anchors[[episode_info$start_col[i]]]
    # The Code number where this episode ends, from `anchors`
    end_code <- anchors[[episode_info$end_col[i]]]

    # Look up the literal boundary values directly -- anchors always
    # Point to a real row (get_episode_anchors() already stops if a
    # Marker is missing), so these are always available regardless of
    # Whether any behavior was actually coded inside the episode
    start_hms <- data %>% filter(Code == start_code) %>% pull(Time_Relative_hms)
    end_hms   <- data %>% filter(Code == end_code)   %>% pull(Time_Relative_hms)
    start_sec <- data %>% filter(Code == start_code) %>% pull(Time_Relative_sf)
    end_sec   <- data %>% filter(Code == end_code)   %>% pull(Time_Relative_sf)

    # Blank the boundary times only when the episode is truly empty:
    # Start and end land on the exact same moment AND no real (non-marker)
    # Behavior was coded anywhere in the episode. An episode with a real,
    # Nonzero span but no coded behavior inside it keeps its boundary
    # Times -- only its duration/count variables get blanked elsewhere,
    # By blank_uncoded_episodes()
    if (start_sec == end_sec && !episode_has_codes(data, episode)) {
      start_hms <- NA_character_
      end_hms   <- NA_character_
      start_sec <- NA_real_
      end_sec   <- NA_real_
    }

    # Write all four boundary columns into `scores`, using this episode's
    # Prefix to build the right column names (e.g. "v2sfblepistarthms")
    scores[[paste0(prefix, "epistarthms")]] <- start_hms
    scores[[paste0(prefix, "epiendhms")]]   <- end_hms
    scores[[paste0(prefix, "epistart")]]    <- start_sec
    scores[[paste0(prefix, "epiend")]]      <- end_sec
    # Episode length in seconds (end minus start; NA automatically if
    # Either side is NA)
    scores[[paste0(prefix, "epi")]]         <- end_sec - start_sec
  }

  # Return `scores` with all the boundary-time columns added
  scores
}
```

## 9. Validate the merged raw dataset

```{r}
#| label: validate-function
# ── Validate the Merged Scored Dataset ──────────────────────────────────────────
validate_raw_dataset <- function(dataset) {

  # Count how many rows each ID appears in, then keep only IDs that show
  # Up more than once
  duplicate_ids <- dataset %>%
    count(ID) %>%
    filter(n > 1)

  # Stop if any duplicate IDs were found
  if (nrow(duplicate_ids) > 0) {
    stop("Raw dataset contains duplicate IDs: ", paste(duplicate_ids$ID, collapse = ", "))
  }

  # Otherwise, report that the check passed
  message("Still Face raw: duplicate ID check passed.")
}
```

## 10. Run it: import, score, and merge every subject

```{r}
#| label: find-files
# ── Discover Raw and Consensus Files ────────────────────────────────────────────
files_full <- list.files(
  path = PATH_EVENT_LOGS,
  pattern = "Event.?Logs\\.xlsx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

if (length(files_full) == 0) {
  stop("No event log files found in: ", PATH_EVENT_LOGS)
}

# Find consensus files in each of the two possible consensus folders
files_consensus_reliability <- list.files(
  path = PATH_EVENT_LOGS_RELIABILITY,
  pattern = "Event.?Logs\\.xlsx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)
files_consensus_training <- list.files(
  path = PATH_EVENT_LOGS_TRAINING,
  pattern = "Event.?Logs\\.xlsx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

# One row per original file: its subject ID and its path
original_lookup <- tibble(
  ID     = map_chr(files_full, extract_id_from_filename),
  path   = files_full,
  source = "original"
)

# One row per consensus file, tagging which of the two folders it came from
consensus_lookup <- bind_rows(
  tibble(
    ID     = map_chr(files_consensus_reliability, extract_id_from_filename),
    path   = files_consensus_reliability,
    source = "consensus (Reliability)"
  ),
  tibble(
    ID     = map_chr(files_consensus_training, extract_id_from_filename),
    path   = files_consensus_training,
    source = "consensus (Training)"
  )
)

# Stop if any ID has more than one consensus file
duplicate_consensus_ids <- consensus_lookup %>%
  count(ID) %>%
  filter(n > 1)

if (nrow(duplicate_consensus_ids) > 0) {
  stop(
    "More than one consensus file found for ID(s): ",
    paste(duplicate_consensus_ids$ID, collapse = ", ")
  )
}

# ── Resolve Raw Versus Consensus Sources ────────────────────────────────────────
file_lookup <- bind_rows(consensus_lookup, original_lookup) %>%
  distinct(ID, .keep_all = TRUE) %>%
  arrange(ID)

consensus_ids_used <- file_lookup %>% filter(source != "original") %>% pull(ID, source)

if (length(consensus_ids_used) > 0) {
  message(
    "Using consensus file in place of the original for ID(s): ",
    paste(names(consensus_ids_used) %||% consensus_ids_used, collapse = ", ")
  )
}

# Final list of file paths to import and score
files_final <- file_lookup$path
```

```{r}
#| label: blank-row-log
# Running log of blank/empty template rows found and dropped, across
# Every subject -- rebuilt fresh each time the pipeline runs
BLANK_ROW_LOG <- tibble(ID = character(), file = character(), n_blank_rows = integer())
```

```{r}
#| label: process-all-subjects
# ── Process and Merge All Subjects ──────────────────────────────────────────────
# Set, then stack every subject's row into a single dataset
SF_RAW <- map_dfr(files_final, function(filename_full) {
  # Import, validate, and label this one subject's raw data
  data <- import_and_label_one_subject(filename_full)
  # Compute this subject's one-row wide score set
  compute_raw_scores_one_subject(data)
}) %>%
  # Sort the merged dataset by subject ID
  arrange(ID)

# Confirm there are no duplicate subject IDs in the merged dataset
validate_raw_dataset(SF_RAW)
```

```{r}
#| label: show-blank-rows
# Show the blank-row log in the knitted output
if (nrow(BLANK_ROW_LOG) > 0) {
  knitr::kable(BLANK_ROW_LOG, caption = "Subjects with Blank Rows Dropped")
} else {
  message("No blank rows found in this run.")
}
```

## 11. Episode boundary check

```{r}
#| label: episode-timing-check
# ── Compare Calculated Episode Boundaries With Coding Log ───────────────────────
# Wide (one column per episode/boundary) to long (one row per
# ID x episode), and compute each episode's duration
EPISODE_TIMING_CHECK <- SF_RAW %>%
  select(ID,
         v2sfblepistarthms, v2sfblepiendhms, v2sfblepistart, v2sfblepiend,
         v2sfsfepistarthms, v2sfsfepiendhms, v2sfsfepistart, v2sfsfepiend,
         v2sfreepistarthms, v2sfreepiendhms, v2sfreepistart, v2sfreepiend) %>%
  pivot_longer(
    cols = -ID,
    names_to = c("episode_code", ".value"),
    names_pattern = "v2sf(bl|sf|re)epi(starthms|endhms|start|end)"
  ) %>%
  rename(start_hms = starthms, end_hms = endhms, start_sec = start, end_sec = end) %>%
  mutate(
    episode = case_when(
      episode_code == "bl" ~ "Baseline",
      episode_code == "sf" ~ "Still Face",
      episode_code == "re" ~ "Reunion"
    ),
    duration_sec = end_sec - start_sec
  ) %>%
  select(ID, episode, start_hms, end_hms, start_sec, end_sec, duration_sec) %>%
  arrange(ID, episode)

# Read the "6 Months" tab, skipping the first two rows so row 3 becomes the header
CODING_LOG_RAW <- read_excel(
  PATH_CODING_LOG,
  sheet = "6 Months",
  skip = 2
)

# Keep only the five columns needed from coding log, by position (A=1, K=11, L=12, M=13, P=16), and give them clear names
ACTUAL_START_STOP <- CODING_LOG_RAW %>%
  select(1, 11, 12, 13, 16) %>%
  setNames(c("ID", "Baseline", "Still Face", "Reunion", "COMMENTS")) %>%
  mutate(ID = as.character(ID))

# Reshape the three episode range columns to long format for the
# Mmss comparison below. COMMENTS stays in the wide ACTUAL_START_STOP
# Table and gets joined back in at the end
actual_long <- ACTUAL_START_STOP %>%
  pivot_longer(
    cols = c(Baseline, `Still Face`, Reunion),
    names_to = "episode",
    values_to = "range_string"
  ) %>%
  mutate(
    range_split        = str_split_fixed(range_string, " - ", 2),
    actual_start_mmss  = str_trim(range_split[, 1]),
    actual_end_mmss    = str_trim(range_split[, 2])
  ) %>%
  select(ID, episode, actual_start_mmss, actual_end_mmss) %>%
  pivot_longer(
    cols = c(actual_start_mmss, actual_end_mmss),
    names_to = "boundary",
    values_to = "actual_mmss"
  ) %>%
  mutate(boundary = if_else(boundary == "actual_start_mmss", "Start", "End"))

# Reshape R's own calculated times the same way, so the two sides can be
# Joined boundary-by-boundary
r_long <- EPISODE_TIMING_CHECK %>%
  select(ID, episode, start_hms, end_hms, start_sec, end_sec) %>%
  pivot_longer(
    cols = c(start_hms, end_hms, start_sec, end_sec),
    names_to = c("boundary", ".value"),
    names_pattern = "(start|end)_(hms|sec)"
  ) %>%
  rename(r_mmss = hms, r_sec = sec) %>%
  mutate(boundary = if_else(boundary == "start", "Start", "End"))

# Convert time variables to the same format
to_seconds <- function(x) {
  parts <- str_split(str_trim(x), ":")
  map_dbl(parts, function(p) {
    p <- as.numeric(p)
    if (length(p) == 2) {
      p[1] * 60 + p[2]  # M:SS
    } else if (length(p) == 3) {
      p[1] * 3600 + p[2] * 60 + p[3]  # H:MM:SS
    } else {
      NA_real_
    }
  })
}

# Join R's boundaries to the reference boundaries and classify each one
EPISODE_BOUNDARY_COMPARISON <- actual_long %>%
  full_join(r_long, by = c("ID", "episode", "boundary")) %>%
  filter(ID %in% SF_RAW$ID) %>%
  mutate(
    episode      = factor(episode, levels = c("Baseline", "Still Face", "Reunion")),
    boundary     = factor(boundary, levels = c("Start", "End")),
    actual_sec   = to_seconds(actual_mmss),
    match = case_when(
      is.na(r_mmss)                     ~ "NO CODES IN RAW",
      is.na(actual_mmss)                ~ "NO REFERENCE TIME",
      actual_sec == r_sec               ~ "MATCH",
      TRUE                              ~ "NOT-MATCH"
    ),
    sec_diff = abs(actual_sec - r_sec)
  ) %>%
  select(ID, episode, boundary, actual_mmss, r_mmss, actual_sec, r_sec, sec_diff, match) %>%
  arrange(ID, episode, boundary)


# Filter based on episodes that do not match manual coding log
# Only the non-matches worth chasing down — off by more than 4 seconds
EPISODE_BOUNDARY_FLAGGED <- EPISODE_BOUNDARY_COMPARISON %>%
  filter(match == "NOT-MATCH", sec_diff > 4)

# Preview in the knitted output
knitr::kable(EPISODE_BOUNDARY_FLAGGED, caption = "Episode Boundary Comparison: Raw Data vs. Coding Log (Flagged Subjects)")
```

## 12. Final review {#sec-preview-raw}

This portfolio page does not syntax for final variable scoring (e.g.
total duration across episodes, proportions scores) That syntax was
written SPSS syntax file, per request of lead behavioral coder.

```{r}
#| label: final-review
# ── Summarize Final Scored Dataset ──────────────────────────────────────────────
# Variables (columns) it has (minus 1 to exclude the ID column itself)
knitr::kable(tibble(
  n_subjects = n_distinct(SF_RAW$ID),
  n_variables = ncol(SF_RAW) - 1
), caption = "Final Merged Dataset Summary")
```
