---
title: "Still Face Variable Dictionary"
toc-depth: 2
bibliography: references.bib
---

# Still Face Paradigm

The Still Face Paradigm [@tronick] is a brief, structured interaction task used to study early social-emotional regulation and the caregiver-infant relationship. It has three phases, or episodes, which are described below:

-   **Baseline (bl)** - the mother engages in normal face-to-face play with her infant, establishing each dyad's typical interactive style
-   *a 15 second transition, where the mother looks away before returning her gaze towards her infant*
-   **Still Face (sf)** - holding a neutral, unmoving expression for a set interval, while remaining present but unresponsive
-   **Reunion (re)** - the mother resumes normal interaction, and the dyad's recovery is observed

Beacuse infants expect consistent, responsive interaction; the mother withdrawing it is a mild stressor that reveals how the infant regulates distress, and how the mother-infant dyad repairs afterward.

Observational coding, the data the below table is drawn from, captures this by having trained coders mark the onset and offset of discrete behavioral states (e.g. gaze direction, facial affect, vocalization, motor behavior) continuously throughout each episode. This results in raw data with timestamped start/stop events per behavior per subject, as seen in the sample raw data in @still-face-data-example. This data must be imported, cleaned, and transformed into variables in the sample data dictionary table.

The code used to create the sample data is @still-face-data-example-code and the code to create the replicated data dictionary table is @variable-dictionary-code.

::: {.callout-tip appearance="simple" icon="false"}
**Scientific relevance of Individual and dyadic differences**: the *magnitude* of the still-face effect and the *speed/completeness* of reunion recovery are used as indicators of infant self-regulation, mother-infant synchrony, and attachment-relevant processes. Blunted still-face reactivity, minimal recovery, or atypical patterns (e.g., disorganized behavior, dysregulated affect) have been linked in the literature to risk factors such as maternal depression, prenatal substance exposure, and early adversity.
:::

## Raw Data Example

```{r}
#| label: load-packages
#| warning: false
library(readxl)
library(flextable)
library(officer)
library(dplyr)
```

```{r}
#| label: still-face-data-example
#| echo: false
#| warning: false

raw <- read_excel("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Still Face/6M Event Logs/111_SFF_SF6M_64 - Event Logs.xlsx")

demo <- raw %>%
  filter(Time_Relative_sf <= 2.12 | (Time_Relative_sf >= 119.995 & Time_Relative_sf <= 120.02)) %>%
  transmute(
    `Time (s)`     = round(Time_Relative_sf, 3),
    `Duration (s)` = round(Duration_sf, 3),
    Subject,
    Behavior,
    `Event Type` = Event_Type
  )

ft_example <- flextable(demo) %>%
  theme_box() %>%
  set_table_properties(layout = "fixed") %>%
  width(j = "Time (s)", width = 0.85) %>%
  width(j = "Duration (s)", width = 0.9) %>%
  width(j = "Subject", width = 0.65) %>%
  width(j = "Behavior", width = 2.2) %>%
  width(j = "Event Type", width = 0.8) %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%
  align(j = c("Time (s)", "Duration (s)"), align = "right", part = "all") %>%
  font(j = "Behavior", fontname = "Consolas") %>%
  hline(i = 8, border = fp_border(color = "#999999", width = 1.5))

ft_example
```

## Data Dictionary Sample

```{r}
#| label: variable-dictionary
#| echo: false
#| warning: false

var_dict <- tibble::tribble(
  ~Domain,  ~Variable,                                      ~Baseline,        ~`Still Face`,     ~Reunion,          ~Total,
  "Infant", "Duration crying/negative vocal",                "blnegvoc",       "sfnegvoc",        "renegvoc",        "totnegvoc",
  "Infant", "Duration neutral vocal",                        "blneuvoc",       "sfneuvoc",        "reneuvoc",        "totneuvoc",
  "Infant", "Duration silent vocal",                         "blsilvoc",       "sfsilvoc",        "resilvoc",        "totsilvoc",
  "Infant", "Duration laughing/positive vocal",              "blposvoc",       "sfposvoc",        "reposvoc",        "totposvoc",
  "Infant", "Duration frown/negative expression",            "blnegexp",       "sfnegexp",        "renegexp",        "totnegexp",
  "Infant", "Duration neutral expression",                   "blneuexp",       "sfneuexp",        "reneuexp",        "totneuexp",
  "Infant", "Duration smile/positive expression",            "blposexp",       "sfposexp",        "reposexp",        "totposexp",
  "Infant", "Proportion of episode crying/negative vocal",   "blnegvocepi",    "sfnegvocepi",     "renegvocepi",     "totnegvocepi",
  "Infant", "Proportion of vocals that are negative",        "blnegvocprop",   "sfnegvocprop",    "renegvocprop",    "totnegvocprop"
)

prefix_colors <- c(bl = "#332288", sf = "#88CCEE", re = "#44AA99", tot = "#117733")
root1_words   <- c("neg","neu","sil","pos","yes","not","att","tlk","jnt","obj","par","comf","esc")
root1_color   <- "#882255"
root2_colors  <- c(lktch = "#88CCEE", look = "#CC6677", voc = "#CC6677", exp = "#CC6677",
                    tch = "#332288", atn = "#44AA99", aff = "#AA4499", end = "#CC6677")

match_first <- function(rest, candidates) {
  candidates <- candidates[order(-nchar(candidates))]
  for (cand in candidates) if (startsWith(rest, cand)) return(cand)
  NA_character_
}

segment_var <- function(x) {
  if (is.na(x) || x %in% c("n/a", "")) return(data.frame(text = x, color = "black", bold = TRUE))
  prefix <- match_first(x, names(prefix_colors))
  rest   <- sub(paste0("^", prefix), "", x)
  root1  <- match_first(rest, root1_words)
  rest2  <- if (!is.na(root1)) sub(paste0("^", root1), "", rest) else rest
  root2  <- match_first(rest2, names(root2_colors))
  suffix <- if (!is.na(root2)) sub(paste0("^", root2), "", rest2) else rest2
  segs <- data.frame(text = prefix, color = prefix_colors[[prefix]], bold = TRUE)
  if (!is.na(root1)) segs <- rbind(segs, data.frame(text = root1, color = root1_color, bold = TRUE))
  if (!is.na(root2)) segs <- rbind(segs, data.frame(text = root2, color = root2_colors[[root2]], bold = TRUE))
  if (nzchar(suffix)) segs <- rbind(segs, data.frame(text = suffix, color = "black", bold = TRUE))
  segs
}

desc_word_patterns <- list(
  list(pattern = "negative",   color = root1_color,           bold = FALSE),
  list(pattern = "neutral",    color = root1_color,           bold = FALSE),
  list(pattern = "silent",     color = root1_color,           bold = FALSE),
  list(pattern = "positive",   color = root1_color,           bold = FALSE),
  list(pattern = "vocal",      color = root2_colors[["voc"]], bold = FALSE),
  list(pattern = "expression", color = root2_colors[["exp"]], bold = FALSE),
  list(pattern = "episode",    color = "black",               bold = TRUE)
)

segment_desc <- function(text) {
  matches <- list()
  for (p in desc_word_patterns) {
    locs <- gregexpr(p$pattern, text, fixed = TRUE)[[1]]
    if (locs[1] == -1) next
    lens <- attr(locs, "match.length")
    for (k in seq_along(locs)) {
      matches[[length(matches) + 1]] <- list(start = locs[k], end = locs[k] + lens[k] - 1,
                                               color = p$color, bold = p$bold)
    }
  }
  if (length(matches) == 0) return(data.frame(text = text, color = "black", bold = FALSE))
  matches <- matches[order(sapply(matches, `[[`, "start"))]
  segs <- data.frame(text = character(0), color = character(0), bold = logical(0))
  cursor <- 1
  for (m in matches) {
    if (m$start > cursor) segs <- rbind(segs, data.frame(text = substr(text, cursor, m$start - 1), color = "black", bold = FALSE))
    segs <- rbind(segs, data.frame(text = substr(text, m$start, m$end), color = m$color, bold = m$bold))
    cursor <- m$end + 1
  }
  if (cursor <= nchar(text)) segs <- rbind(segs, data.frame(text = substr(text, cursor, nchar(text)), color = "black", bold = FALSE))
  segs
}

apply_segments <- function(ft, i, j, segs) {
  chunks <- lapply(seq_len(nrow(segs)), function(k) {
    as_chunk(segs$text[k], props = fp_text(color = segs$color[k], bold = segs$bold[k]))
  })
  mk_par(ft, i = i, j = j, value = do.call(as_paragraph, chunks))   # mk_par, not compose — see below
}


ft_dictionary <- flextable(var_dict) %>%
  merge_v(j = "Domain") %>%
  theme_box() %>%
  autofit() %>%
  set_table_properties(layout = "fixed") %>%
  width(j = "Domain", width = 0.6) %>%
  width(j = "Variable", width = 2.3) %>%
  width(j = c("Baseline","Still Face","Reunion","Total"), width = 1.1) %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%
  color(j = "Baseline",    color = prefix_colors[["bl"]],  part = "header") %>%
  color(j = "Still Face",  color = prefix_colors[["sf"]],  part = "header") %>%
  color(j = "Reunion",     color = prefix_colors[["re"]],  part = "header") %>%
  color(j = "Total",       color = prefix_colors[["tot"]], part = "header")

var_cols <- c("Baseline", "Still Face", "Reunion", "Total")
for (i in seq_len(nrow(var_dict))) {
  ft_dictionary <- apply_segments(ft_dictionary, i, "Variable", segment_desc(var_dict$Variable[i]))
  for (col in var_cols) {
    ft_dictionary <- apply_segments(ft_dictionary, i, col, segment_var(var_dict[[col]][i]))
  }
}

ft_dictionary
```

```{r}
#| label: still-face-data-example-code
#| code-summary: "data example code"

demo <- raw %>%
  filter(Time_Relative_sf <= 2.12 | (Time_Relative_sf >= 119.995 & Time_Relative_sf <= 120.02)) %>%
  transmute(
    Segment  = if_else(Time_Relative_sf <= 2.12, "Opening of Baseline", "Baseline \u2192 Still Face transition"),
    `Time (s)`     = round(Time_Relative_sf, 3),
    `Duration (s)` = round(Duration_sf, 3),
    Subject,
    Behavior,
    `Event Type` = Event_Type
  )

ft_example <- flextable(demo) %>%
  merge_v(j = "Segment") %>%
  theme_box() %>%
  set_table_properties(layout = "fixed") %>%
  width(j = "Segment", width = 1.1) %>%
  width(j = "Time (s)", width = 0.85) %>%
  width(j = "Duration (s)", width = 0.9) %>%
  width(j = "Subject", width = 0.65) %>%
  width(j = "Behavior", width = 2.2) %>%
  width(j = "Event Type", width = 0.8) %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%
  align(j = c("Time (s)", "Duration (s)"), align = "right", part = "all") %>%
  font(j = "Behavior", fontname = "Consolas") %>%
  color(i = ~ `Event Type` == "State start", j = "Event Type", color = "#117733") %>%
  color(i = ~ `Event Type` == "State stop",  j = "Event Type", color = "#882255") %>%
  bold(i = ~ grepl("^15 S|^BASELINE|^REUNION", Behavior), j = "Behavior") %>%
  hline(i = 8, border = fp_border(color = "#999999", width = 1.5))
```

```{r}
#| label: variable-dictionary-code
#| code-summary: "data dictionary code"

var_dict <- tibble::tribble(
  ~Domain,  ~Variable,                                      ~Baseline,        ~`Still Face`,     ~Reunion,          ~Total,
  "Infant", "Duration crying/negative vocal",                "blnegvoc",       "sfnegvoc",        "renegvoc",        "totnegvoc",
  "Infant", "Duration neutral vocal",                        "blneuvoc",       "sfneuvoc",        "reneuvoc",        "totneuvoc",
  "Infant", "Duration silent vocal",                         "blsilvoc",       "sfsilvoc",        "resilvoc",        "totsilvoc",
  "Infant", "Duration laughing/positive vocal",              "blposvoc",       "sfposvoc",        "reposvoc",        "totposvoc",
  "Infant", "Duration frown/negative expression",            "blnegexp",       "sfnegexp",        "renegexp",        "totnegexp",
  "Infant", "Duration neutral expression",                   "blneuexp",       "sfneuexp",        "reneuexp",        "totneuexp",
  "Infant", "Duration smile/positive expression",            "blposexp",       "sfposexp",        "reposexp",        "totposexp",
  "Infant", "Proportion of episode crying/negative vocal",   "blnegvocepi",    "sfnegvocepi",     "renegvocepi",     "totnegvocepi",
  "Infant", "Proportion of vocals that are negative",        "blnegvocprop",   "sfnegvocprop",    "renegvocprop",    "totnegvocprop"
)

prefix_colors <- c(bl = "#332288", sf = "#88CCEE", re = "#44AA99", tot = "#117733")
root1_words   <- c("neg","neu","sil","pos","yes","not","att","tlk","jnt","obj","par","comf","esc")
root1_color   <- "#882255"
root2_colors  <- c(lktch = "#88CCEE", look = "#CC6677", voc = "#CC6677", exp = "#CC6677",
                    tch = "#332288", atn = "#44AA99", aff = "#AA4499", end = "#CC6677")

match_first <- function(rest, candidates) {
  candidates <- candidates[order(-nchar(candidates))]
  for (cand in candidates) if (startsWith(rest, cand)) return(cand)
  NA_character_
}

segment_var <- function(x) {
  if (is.na(x) || x %in% c("n/a", "")) return(data.frame(text = x, color = "black", bold = TRUE))
  prefix <- match_first(x, names(prefix_colors))
  rest   <- sub(paste0("^", prefix), "", x)
  root1  <- match_first(rest, root1_words)
  rest2  <- if (!is.na(root1)) sub(paste0("^", root1), "", rest) else rest
  root2  <- match_first(rest2, names(root2_colors))
  suffix <- if (!is.na(root2)) sub(paste0("^", root2), "", rest2) else rest2
  segs <- data.frame(text = prefix, color = prefix_colors[[prefix]], bold = TRUE)
  if (!is.na(root1)) segs <- rbind(segs, data.frame(text = root1, color = root1_color, bold = TRUE))
  if (!is.na(root2)) segs <- rbind(segs, data.frame(text = root2, color = root2_colors[[root2]], bold = TRUE))
  if (nzchar(suffix)) segs <- rbind(segs, data.frame(text = suffix, color = "black", bold = TRUE))
  segs
}

desc_word_patterns <- list(
  list(pattern = "negative",   color = root1_color,           bold = FALSE),
  list(pattern = "neutral",    color = root1_color,           bold = FALSE),
  list(pattern = "silent",     color = root1_color,           bold = FALSE),
  list(pattern = "positive",   color = root1_color,           bold = FALSE),
  list(pattern = "vocal",      color = root2_colors[["voc"]], bold = FALSE),
  list(pattern = "expression", color = root2_colors[["exp"]], bold = FALSE),
  list(pattern = "episode",    color = "black",               bold = TRUE)
)

segment_desc <- function(text) {
  matches <- list()
  for (p in desc_word_patterns) {
    locs <- gregexpr(p$pattern, text, fixed = TRUE)[[1]]
    if (locs[1] == -1) next
    lens <- attr(locs, "match.length")
    for (k in seq_along(locs)) {
      matches[[length(matches) + 1]] <- list(start = locs[k], end = locs[k] + lens[k] - 1,
                                               color = p$color, bold = p$bold)
    }
  }
  if (length(matches) == 0) return(data.frame(text = text, color = "black", bold = FALSE))
  matches <- matches[order(sapply(matches, `[[`, "start"))]
  segs <- data.frame(text = character(0), color = character(0), bold = logical(0))
  cursor <- 1
  for (m in matches) {
    if (m$start > cursor) segs <- rbind(segs, data.frame(text = substr(text, cursor, m$start - 1), color = "black", bold = FALSE))
    segs <- rbind(segs, data.frame(text = substr(text, m$start, m$end), color = m$color, bold = m$bold))
    cursor <- m$end + 1
  }
  if (cursor <= nchar(text)) segs <- rbind(segs, data.frame(text = substr(text, cursor, nchar(text)), color = "black", bold = FALSE))
  segs
}

apply_segments <- function(ft, i, j, segs) {
  chunks <- lapply(seq_len(nrow(segs)), function(k) {
    as_chunk(segs$text[k], props = fp_text(color = segs$color[k], bold = segs$bold[k]))
  })
  mk_par(ft, i = i, j = j, value = do.call(as_paragraph, chunks))   # mk_par, not compose — see below
}

ft_dictionary <- flextable(var_dict) %>%
  merge_v(j = "Domain") %>%
  theme_box() %>%
  autofit() %>%
  set_table_properties(layout = "fixed") %>%
  width(j = "Domain", width = 0.6) %>%
  width(j = "Variable", width = 2.3) %>%
  width(j = c("Baseline","Still Face","Reunion","Total"), width = 1.1) %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%
  color(j = "Baseline",    color = prefix_colors[["bl"]],  part = "header") %>%
  color(j = "Still Face",  color = prefix_colors[["sf"]],  part = "header") %>%
  color(j = "Reunion",     color = prefix_colors[["re"]],  part = "header") %>%
  color(j = "Total",       color = prefix_colors[["tot"]], part = "header")

var_cols <- c("Baseline", "Still Face", "Reunion", "Total")
for (i in seq_len(nrow(var_dict))) {
  ft_dictionary <- apply_segments(ft_dictionary, i, "Variable", segment_desc(var_dict$Variable[i]))
  for (col in var_cols) {
    ft_dictionary <- apply_segments(ft_dictionary, i, col, segment_var(var_dict[[col]][i]))
  }
}

```
