---
title: Baseline Synchrony Calculation
toc-depth: 3
echo: true
warning: false
---

```{r}

#| label: packages

library(here)
library(nlme)
library(tidyverse)
library(ggplot2) 
library(lavaan)
library(semTools)
library(jtools)
library(haven)
library(modsem)
library(forecast)
library(gt)
```

# Load Data

Long-format positive- and negative-valence baseline warmth data, produced by the [Baseline Synchrony Processing](Baseline-Synchrony_Processing.qmd) pipeline, are read in from where that processed data already lives on disk (this page does not regenerate them). Import code is included but not rendered below.

```{r}
#| label: load data
#| echo: false

BL_WH_Long <- read_csv("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/SFF Synchrony/Processed Data/BL_WH_Long.csv")
BL_WH_Pos_Long <- read_csv("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/SFF Synchrony/Processed Data/BL_WH_Pos_Long.csv")
BL_WH_Neg_Long <- read_csv("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/SFF Synchrony/Processed Data/BL_WH_Neg_Long.csv")

```

# Positive Interaction

## Synchrony Calculation

```{r}
# Function to compute time-lagged cross-correlation synchrony between dyad members
compute_blpwh_sync <- function(df, lag = 20, n_boot = 1000) {
  
  # Remove rows with missing data for both mom and partner variables
  df <- df %>% filter(!is.na(blmpwh), !is.na(blfpwh))
  
  # Check if there's enough data to compute lagged correlations
  if (nrow(df) < lag + 1) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # --- STEP 1: Pre-whitening via ARIMA to remove autocorrelation ---
  # Fit ARIMA models to each person's time series to account for their individual temporal patterns
  fit_mom <- tryCatch(auto.arima(df$blmpwh), error = function(e) NULL)
  fit_partner <- tryCatch(auto.arima(df$blfpwh), error = function(e) NULL)
  
  # If either model fails, return NA results
  if (is.null(fit_mom) || is.null(fit_partner)) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # Extract residuals (pre-whitened series) - these represent deviations from each person's own temporal pattern
  res_mom <- residuals(fit_mom)
  res_partner <- residuals(fit_partner)
  
  # Combine residuals and remove any remaining missing values
  residuals_df <- tibble(res_mom, res_partner) %>% drop_na()
  if (nrow(residuals_df) < lag + 1) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # --- STEP 2: Compute observed cross-correlation function (CCF) ---
  # CCF examines how mom's behavior at time t correlates with partner's behavior at time t+lag
  # Negative lags: mom leads partner; Positive lags: partner leads mom
  ccf_result <- ccf(residuals_df$res_mom, residuals_df$res_partner, lag.max = lag, plot = FALSE)
  observed_acf <- ccf_result$acf  # Cross-correlation values at each lag
  lags <- ccf_result$lag           # Lag values (e.g., -20 to +20)
  
  # --- STEP 3: Bootstrap to establish confidence intervals ---
  # Create matrix to store bootstrap results
  boot_corrs <- matrix(NA, nrow = n_boot, ncol = length(lags))
  n <- nrow(residuals_df)
  
  # Generate null distribution by randomly shuffling residuals (breaks temporal coupling)
  for (i in 1:n_boot) {
    x_boot <- sample(residuals_df$res_mom, n, replace = TRUE)      # Resample mom's residuals
    y_boot <- sample(residuals_df$res_partner, n, replace = TRUE)  # Resample partner's residuals
    boot_ccf <- ccf(x_boot, y_boot, lag.max = lag, plot = FALSE)   # Compute CCF on shuffled data
    boot_corrs[i, ] <- boot_ccf$acf  # Store correlation values
  }
  
  # Calculate 95% confidence intervals from bootstrap distribution
  lower_ci <- apply(boot_corrs, 2, quantile, probs = 0.025, na.rm = TRUE)
  upper_ci <- apply(boot_corrs, 2, quantile, probs = 0.975, na.rm = TRUE)
  
  # Determine significance: observed correlation falls outside 95% CI
  significant <- (observed_acf < lower_ci) | (observed_acf > upper_ci)
  
  # Return tibble with lag values, correlations, CIs, and significance
  tibble(
    lag = as.numeric(lags),
    acf = as.numeric(observed_acf),
    lower_ci = lower_ci,
    upper_ci = upper_ci,
    significant = significant
  )
}

library(purrr)

# Nest data by couple ID (creates list-column with each dyad's time series)
blpwh_sync_id <- BL_WH_Long %>%
  group_by(ID) %>%
  nest()

# Apply the synchrony function to each couple's data
blpwh_sync_id <- blpwh_sync_id %>%
  mutate(blpwh_sync_ccf = map(data, compute_blpwh_sync))

# Unnest results to get one row per couple per lag value
blpwh_sync_results <- unnest(blpwh_sync_id, blpwh_sync_ccf)

blpwh_sync_results$lag <- as.numeric(blpwh_sync_results$lag)
```

## Max Synchrony

```{r}


#blpwh_sync_results <- unnest(blpwh_sync_id, blpwh_sync_ccf)

# Extract max synchrony for each couple
blpwh_sync_max <- blpwh_sync_results %>%
  group_by(ID) %>%
  # Find the row with maximum absolute correlation for each couple
  slice_max(abs(acf), n = 1, with_ties = FALSE) %>%
  select(ID, 
         max_lag = lag,           # Lag at which max occurs
         max_sync = acf,          # Correlation value (with sign)
         max_abs_sync = acf,      # Will convert to absolute
         lower_ci, 
         upper_ci, 
         significant) %>%
  mutate(max_abs_sync = abs(max_sync)) %>%  # Add absolute value column
  ungroup()

# View results
head(blpwh_sync_max)
```


### Max Synchrony Interpreation

```{r}

# Add interpretation columns
blpwh_sync_max_interpret <- blpwh_sync_max %>%
  mutate(
    # Who leads?
    leader = case_when(
      max_lag > 0 ~ "Mom leads",
      max_lag < 0 ~ "Partner leads",
      max_lag == 0 ~ "Concurrent",
      TRUE ~ NA_character_
    ),
    # How much lead time?
    lead_time_seconds = abs(max_lag),
    # Direction of relationship
    relationship = case_when(
      max_sync > 0 ~ "Positive (same direction)",
      max_sync < 0 ~ "Negative (opposite direction)",
      TRUE ~ NA_character_
    ),
    # Strength categorization (example thresholds)
    strength = case_when(
      max_abs_sync >= 0.5 ~ "Strong",
      max_abs_sync >= 0.3 ~ "Moderate",
      max_abs_sync >= 0.1 ~ "Weak",
      TRUE ~ "Very weak"
    )
  )

# View interpreted results
blpwh_sync_max_interpret %>%
  select(ID, max_sync, max_lag, leader, lead_time_seconds, 
         relationship, strength, significant)
```

## Descriptives

```{r}
# Descriptive statistics across all couples
blpwh_sync_max_interpret_summary <- blpwh_sync_max_interpret %>%
  summarise(
    n_couples = n(),
    n_significant = sum(significant, na.rm = TRUE),
    pct_significant = mean(significant, na.rm = TRUE) * 100,
    
    mean_max_sync = mean(max_abs_sync, na.rm = TRUE),
    sd_max_sync = sd(max_abs_sync, na.rm = TRUE),
    
    mean_lag = mean(max_lag, na.rm = TRUE),
    sd_lag = sd(max_lag, na.rm = TRUE),
    
    n_mom_leads = sum(max_lag > 0, na.rm = TRUE),
    n_partner_leads = sum(max_lag < 0, na.rm = TRUE),
    n_concurrent = sum(max_lag == 0, na.rm = TRUE),
    
    n_positive = sum(max_sync > 0, na.rm = TRUE),
    n_negative = sum(max_sync < 0, na.rm = TRUE)
  )

print(blpwh_sync_max_interpret_summary)
```


## Plot Max Synchrony

```{r}
library(ggplot2)

# Distribution of maximum synchrony values
blpwh_hist_maxsync <- ggplot(blpwh_sync_max_interpret, aes(x = max_abs_sync, fill = significant)) +
  geom_histogram(bins = 30, alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
  labs(title = "Distribution of Maximum Synchrony Across Couples",
       x = "Maximum Absolute Correlation",
       y = "Number of Couples",
       fill = "Significant",
       caption = "Histogram of each couple's maximum absolute time-lagged cross-correlation (positive interaction), colored by statistical significance.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blpwh_hist_maxsync
ggsave("../../figures/baseline-warmth-synchrony/calculation/positive/blpwh_hist_maxsync.png", plot = blpwh_hist_maxsync, width = 8, height = 6, dpi = 300)

# Distribution of lags where max occurs
blpwh_hist_lag <- ggplot(blpwh_sync_max_interpret, aes(x = max_lag, fill = leader)) +
  geom_histogram(binwidth = 1, alpha = 0.7) +
  scale_fill_manual(values = c("Mom leads" = "#CC6677", "Partner leads" = "#88CCEE", "Concurrent" = "#009988"), name = "Leader") +
  labs(title = "Lag at Maximum Synchrony",
       x = "Lag (seconds)",
       y = "Number of Couples",
       fill = "Leader",
       caption = "Histogram of the time lag (in seconds) at which each couple's maximum synchrony occurs, colored by which partner leads.") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blpwh_hist_lag
ggsave("../../figures/baseline-warmth-synchrony/calculation/positive/blpwh_hist_lag.png", plot = blpwh_hist_lag, width = 8, height = 6, dpi = 300)

# Scatterplot: max correlation vs lag
blpwh_scatter_sync <- ggplot(blpwh_sync_max_interpret,
       aes(x = max_lag, y = max_sync, color = significant)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Maximum Synchrony: Correlation vs Lag",
       x = "Lag at Maximum (seconds)",
       y = "Correlation Coefficient",
       color = "Significant",
       caption = "Each point is one couple's maximum correlation plotted against the lag at which it occurs, colored by statistical significance.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blpwh_scatter_sync
ggsave("../../figures/baseline-warmth-synchrony/calculation/positive/blpwh_scatter_sync.png", plot = blpwh_scatter_sync, width = 8, height = 6, dpi = 300)
```

### Grid Panel

```{r}
# Show multiple couples in a grid
blpwh_sync_ccfplot <- function(blpwh_sync_results, n_couples = 12) {
  # Select first n couples
  selected_ids <- unique(blpwh_sync_results$ID)[1:min(n_couples, length(unique(blpwh_sync_results$ID)))]

  blpwh_sync_results %>%
    filter(ID %in% selected_ids) %>%
    ggplot(aes(x = lag, y = acf)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.3) +
    geom_line(color = "#CC6677", size = 0.5) +
    geom_point(aes(color = significant), size = 1) +
    geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
                alpha = 0.15, fill = "#CC6677") +
    scale_color_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
    facet_wrap(~ ID, scales = "free_y", ncol = 3) +
    labs(title = "Cross-Correlation Functions Across Couples",
         x = "Lag (seconds)",
         y = "Cross-Correlation",
         color = "Significant",
         caption = "Full cross-correlation function by lag for a sample of couples, with 95% confidence bands and significant lags highlighted.") +
    theme_minimal(base_size = 16) +
    theme(legend.position = "bottom",
          strip.text = element_text(face = "bold", size = 14),
          axis.title = element_text(size = 14), axis.text = element_text(size = 12),
          panel.grid = element_blank())
}

# Display grid
blpwh_sync_ccfplot_p <- blpwh_sync_ccfplot(blpwh_sync_results, n_couples = 3)
blpwh_sync_ccfplot_p
ggsave("../../figures/baseline-warmth-synchrony/calculation/positive/blpwh_sync_ccfplot.png", plot = blpwh_sync_ccfplot_p, width = 13, height = 5, dpi = 300)
```


## Calculate Average

```{r}
# Compute mean CCF profile across all couples
blpwh_sync_avgccf <- blpwh_sync_results %>%
  group_by(lag) %>%
  summarise(
    mean_acf = mean(acf, na.rm = TRUE),
    se_acf = sd(acf, na.rm = TRUE) / sqrt(n()),
    lower_se = mean_acf - 1.96 * se_acf,
    upper_se = mean_acf + 1.96 * se_acf,
    pct_significant = mean(significant, na.rm = TRUE) * 100,
    .groups = "drop"
  )


```

### Plot Average

```{r}
# Plot average CCF
blpwh_sync_avgccf_plot <- ggplot(blpwh_sync_avgccf, aes(x = lag, y = mean_acf)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_ribbon(aes(ymin = lower_se, ymax = upper_se),
              alpha = 0.3, fill = "#CC6677") +
  geom_line(color = "#CC6677", size = 1.2) +
  geom_point(aes(size = pct_significant), color = "#CC6677") +
  scale_size_continuous(range = c(1, 4), name = "% Couples\nSignificant") +
  labs(title = "Average Cross-Correlation Function Across All Couples",
       subtitle = "Shaded area represents 95% confidence interval",
       x = "Lag (seconds)\n← Partner leads | Mom leads →",
       y = "Mean Cross-Correlation",
       caption = "Mean cross-correlation function across all couples, with point size showing the percentage of couples significant at that lag.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blpwh_sync_avgccf_plot
ggsave("../../figures/baseline-warmth-synchrony/calculation/positive/blpwh_sync_avgccf_plot.png", plot = blpwh_sync_avgccf_plot, width = 8, height = 6, dpi = 300)
```

## Concurrent Synchrony

```{r}
# Extract concurrent (lag = 0)
blpwh_sync_concurrent <- blpwh_sync_results %>%
  filter(lag == 0) %>%
  select(ID, concurrent_acf = acf, concurrent_sig = significant,
         concurrent_lower = lower_ci, concurrent_upper = upper_ci)

# Extract time-lagged (lag != 0)
blpwh_sync_lagged <- blpwh_sync_results %>% filter(lag != 0)

#max synchrony across non-zero lags
blpwh_sync_lagmax <- blpwh_sync_lagged %>%
  group_by(ID) %>%
  # Find the row with maximum absolute correlation for each couple
  slice_max(abs(acf), n = 1, with_ties = FALSE) %>%
  select(ID, 
         max_lag = lag,           # Lag at which max occurs
         max_sync = acf,          # Correlation value (with sign)
         max_abs_sync = acf,      # Will convert to absolute
         lower_ci, 
         upper_ci, 
         significant) %>%
  mutate(max_abs_sync = abs(max_sync)) %>%  # Add absolute value column
  ungroup()

# View results
knitr::kable(head(blpwh_sync_lagmax, 10), caption = "Maximum Time-Lagged Synchrony, Positive Interaction (first 10 couples)")
```

## Time-Lagged

```{r}
#| label: pos time lag to synchrony


##uses first sig lag
blpwh_sync_timetolag <- blpwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    has_significant = any(significant, na.rm = TRUE),
    time_lag_to_sync = if (any(significant, na.rm = TRUE)) {
      abs(lag[which(significant)[1]])
    } else {
      NA_integer_
    },
    .groups = "drop"
  )
```


### Strongest Time-Lag

```{r}

#uses strongest sig lag
blpwh_sync_summary <- blpwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    # Basic counts
    n_significant_lags = sum(significant),
    has_any_significant = any(significant),
    
    # If significant, what's the strongest?
    strongest_lag = if(any(significant)) {
      lag[which.max(abs(acf[significant]))]
    } else NA_real_,
    
    strongest_acf = if(any(significant)) {
      acf[significant][which.max(abs(acf[significant]))]
    } else NA_real_,
    
    # Specific patterns
    has_concurrent = any(significant & lag == 0),
    has_mom_leads = any(significant & lag > 0),
    has_partner_leads = any(significant & lag < 0),
    
    .groups = "drop"
  )


```

## Summary Table - IDs

```{r}
library(gt)
library(dplyr)

# Create the formatted table
blpwh_sync_summarytable <- blpwh_sync_summary %>%
  # Select and order columns for display
  select(ID, n_significant_lags, strongest_lag, strongest_acf,
         has_concurrent, has_mom_leads, has_partner_leads) %>%
  # Take first 20 couples for display (or use head(10) for fewer)
  head(20) %>%
  
  # Create gt table
  gt() %>%
  
  # Add title and subtitle
  tab_header(
    title = "Dyadic Synchrony Analysis",
    subtitle = "Cross-Correlation Results by Couple"
  ) %>%
  
  # Format numeric columns
  fmt_number(
    columns = c(strongest_acf),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(strongest_lag),
    decimals = 0
  ) %>%
  
  # Rename columns with readable labels
  cols_label(
    ID = "Couple ID",
    n_significant_lags = "# Significant Lags",
    strongest_lag = "Strongest Lag (sec)",
    strongest_acf = "Correlation (r)",
    has_concurrent = "Concurrent Sync",
    has_mom_leads = "Mom Leads",
    has_partner_leads = "Partner Leads"
  ) %>%
  
  # Color code the boolean columns
  data_color(
    columns = c(has_concurrent, has_mom_leads, has_partner_leads),
    colors = scales::col_factor(
      palette = c("lightgray", "lightgreen"),
      domain = c(FALSE, TRUE)
    )
  ) %>%
  
  # Color code correlation strength
  data_color(
    columns = strongest_acf,
    colors = scales::col_numeric(
      palette = c("white", "#CC6677", "#332288"),
      domain = c(0, max(blpwh_sync_summary$strongest_acf, na.rm = TRUE))
    )
  ) %>%
  
  # Color code lag direction
  data_color(
    columns = strongest_lag,
    colors = scales::col_numeric(
      palette = c("#88CCEE", "white", "#CC6677"),
      domain = c(
        min(blpwh_sync_summary$strongest_lag, na.rm = TRUE),
        max(blpwh_sync_summary$strongest_lag, na.rm = TRUE)
      )
    )
  ) %>%
  
  # Bold the column headers
  tab_style(
    style = list(
      cell_text(weight = "bold", color = "white"),
      cell_fill(color = "#CC6677")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  
  # Style the title
  tab_style(
    style = cell_text(weight = "bold", size = "large"),
    locations = cells_title(groups = "title")
  ) %>%
  
  # Add alternating row colors
  tab_style(
    style = cell_fill(color = "#F5F5F5"),
    locations = cells_body(rows = seq(2, 20, 2))
  ) %>%
  
  # Add footnotes
  tab_footnote(
    footnote = "Significant lags determined by bootstrap confidence intervals (α = .05)",
    locations = cells_column_labels(columns = n_significant_lags)
  ) %>%
  tab_footnote(
    footnote = "Positive lag = Mom leads; Negative lag = Partner leads; Zero = Concurrent",
    locations = cells_column_labels(columns = strongest_lag)
  ) %>%
  
  # Add source note
  tab_source_note(
    source_note = "Analysis based on pre-whitened ARIMA residuals with 1000 bootstrap iterations"
  ) %>%
  
  # Table size
  tab_options(
    table.font.size = px(14),
    column_labels.font.size = px(14),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(13)
  ) %>%

  # Adjust column widths
  cols_width(
    ID ~ px(90),
    n_significant_lags ~ px(130),
    strongest_lag ~ px(140),
    strongest_acf ~ px(120),
    has_concurrent ~ px(120),
    has_mom_leads ~ px(120),
    has_partner_leads ~ px(140)
  )

# Display the table
blpwh_sync_summarytable
```

## Summary Table - All

```{r}
# Create a summary statistics table instead of individual couples
blpwh_sync_summarystats <- blpwh_sync_summary %>%
  summarise(
    n_couples = n(),
    
    # Synchrony prevalence
    pct_any_significant = mean(has_any_significant, na.rm = TRUE) * 100,
    pct_concurrent = mean(has_concurrent, na.rm = TRUE) * 100,
    pct_mom_leads = mean(has_mom_leads, na.rm = TRUE) * 100,
    pct_partner_leads = mean(has_partner_leads, na.rm = TRUE) * 100,
    
    # Lag characteristics
    mean_n_sig_lags = mean(n_significant_lags, na.rm = TRUE),
    median_strongest_lag = median(strongest_lag, na.rm = TRUE),
    
    # Correlation strength
    mean_strongest_acf = mean(abs(strongest_acf), na.rm = TRUE),
    median_strongest_acf = median(abs(strongest_acf), na.rm = TRUE)
  ) %>%
  
  # Reshape for better display
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  
  # Create gt table
  gt() %>%
  tab_header(
    title = "Dyadic Synchrony: Summary Statistics",
    subtitle = paste0("N = ", nrow(blpwh_sync_summary), " couples")
  ) %>%
  
  # Format values
  fmt_number(
    columns = Value,
    decimals = 2
  ) %>%
  
  # Improve metric labels
  text_transform(
    locations = cells_body(columns = Metric),
    fn = function(x) {
      case_when(
        x == "n_couples" ~ "Total Couples",
        x == "pct_any_significant" ~ "% with Significant Synchrony",
        x == "pct_concurrent" ~ "% with Concurrent Synchrony",
        x == "pct_mom_leads" ~ "% with Mom Leading",
        x == "pct_partner_leads" ~ "% with Partner Leading",
        x == "mean_n_sig_lags" ~ "Mean # Significant Lags",
        x == "median_strongest_lag" ~ "Median Strongest Lag (sec)",
        x == "mean_strongest_acf" ~ "Mean Correlation Strength",
        x == "median_strongest_acf" ~ "Median Correlation Strength",
        TRUE ~ x
      )
    }
  ) %>%
  
  # Style
  tab_style(
    style = list(
      cell_text(weight = "bold", color = "white"),
      cell_fill(color = "#CC6677")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  
  #tab_style(
 #   style = cell_fill(color = "#F5F5F5"),
 #   locations = cells_body(rows = seq(2, 10, 2))
 # ) %>%
  
  cols_label(
    Metric = "Statistic",
    Value = "Value"
  ) %>%
  
  tab_source_note(
    source_note = "Pre-whitened cross-correlation with bootstrap significance testing"
  )

# Display
blpwh_sync_summarystats
```

# Compare Approaches

```{r}
# Compare the three approaches side-by-side for a few couples
blpwh_sync_compare <- blpwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    # Approach 1: First significant lag
    first_sig_lag = if(any(significant)) abs(lag[which(significant)[1]]) else NA_real_,
    
    # Approach 2: Categorical
    parent_led = any(significant & lag > 0),
    infant_led = any(significant & lag < 0),
    
    # Approach 3: Strongest
    strongest_lag = if(any(significant)) lag[which.max(abs(acf[significant]))] else NA_real_,
    strongest_r = if(any(significant)) acf[significant][which.max(abs(acf[significant]))] else NA_real_,
    
    .groups = "drop"
  ) %>%
  head(10) %>%
  
  gt() %>%
  tab_header(
    title = "Comparison of Three Synchrony Approaches",
    subtitle = "First 10 couples"
  ) %>%
  fmt_number(columns = c(first_sig_lag, strongest_lag), decimals = 0) %>%
  fmt_number(columns = strongest_r, decimals = 3) %>%
  cols_label(
    ID = "Couple",
    first_sig_lag = "First Sig Lag",
    parent_led = "Parent Led",
    infant_led = "Infant Led",
    strongest_lag = "Strongest Lag",
    strongest_r = "Strongest r"
  ) %>%
  tab_spanner(
    label = "Approach 1: First",
    columns = first_sig_lag
  ) %>%
  tab_spanner(
    label = "Approach 2: Categorical",
    columns = c(parent_led, infant_led)
  ) %>%
  tab_spanner(
    label = "Approach 3: Strongest",
    columns = c(strongest_lag, strongest_r)
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold", color = "white"),
                 cell_fill(color = "#CC6677")),
    locations = cells_column_labels(everything())
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold", color = "white"),
                 cell_fill(color = "#332288")),
    locations = cells_column_spanners(spanners = everything())
  )

blpwh_sync_compare
```

# Negative Interaction

## Synchrony Calculation

```{r}
# Function to compute time-lagged cross-correlation synchrony between dyad members
compute_blnwh_sync <- function(df, lag = 20, n_boot = 1000) {
  
  # Remove rows with missing data for both mom and partner variables
  df <- df %>% filter(!is.na(blmnwh), !is.na(blfnwh))
  
  # Check if there's enough data to compute lagged correlations
  if (nrow(df) < lag + 1) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # --- STEP 1: Pre-whitening via ARIMA to remove autocorrelation ---
  # Fit ARIMA models to each person's time series to account for their individual temporal patterns
  fit_mom <- tryCatch(auto.arima(df$blmnwh), error = function(e) NULL)
  fit_partner <- tryCatch(auto.arima(df$blfnwh), error = function(e) NULL)
  
  # If either model fails, return NA results
  if (is.null(fit_mom) || is.null(fit_partner)) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # Extract residuals (pre-whitened series) - these represent deviations from each person's own temporal pattern
  res_mom <- residuals(fit_mom)
  res_partner <- residuals(fit_partner)
  
  # Combine residuals and remove any remaining missing values
  residuals_df <- tibble(res_mom, res_partner) %>% drop_na()
  if (nrow(residuals_df) < lag + 1) return(tibble(lag = NA, acf = NA, lower_ci = NA, upper_ci = NA, significant = NA))
  
  # --- STEP 2: Compute observed cross-correlation function (CCF) ---
  # CCF examines how mom's behavior at time t correlates with partner's behavior at time t+lag
  # Negative lags: partner leads mom; Positive lags: mom leads partner
  ccf_result <- ccf(residuals_df$res_mom, residuals_df$res_partner, lag.max = lag, plot = FALSE)
  observed_acf <- ccf_result$acf  # Cross-correlation values at each lag
  lags <- ccf_result$lag           # Lag values (e.g., -20 to +20)
  
  # --- STEP 3: Bootstrap to establish confidence intervals ---
  # Create matrix to store bootstrap results
  boot_corrs <- matrix(NA, nrow = n_boot, ncol = length(lags))
  n <- nrow(residuals_df)
  
  # Generate null distribution by randomly shuffling residuals (breaks temporal coupling)
  for (i in 1:n_boot) {
    x_boot <- sample(residuals_df$res_mom, n, replace = TRUE)      # Resample mom's residuals
    y_boot <- sample(residuals_df$res_partner, n, replace = TRUE)  # Resample partner's residuals
    boot_ccf <- ccf(x_boot, y_boot, lag.max = lag, plot = FALSE)   # Compute CCF on shuffled data
    boot_corrs[i, ] <- boot_ccf$acf  # Store correlation values
  }
  
  # Calculate 95% confidence intervals from bootstrap distribution
  lower_ci <- apply(boot_corrs, 2, quantile, probs = 0.025, na.rm = TRUE)
  upper_ci <- apply(boot_corrs, 2, quantile, probs = 0.975, na.rm = TRUE)
  
  # Determine significance: observed correlation falls outside 95% CI
  significant <- (observed_acf < lower_ci) | (observed_acf > upper_ci)
  
  # Return tibble with lag values, correlations, CIs, and significance
  tibble(
    lag = as.numeric(lags),
    acf = as.numeric(observed_acf),
    lower_ci = lower_ci,
    upper_ci = upper_ci,
    significant = significant
  )
}

library(purrr)

# Nest data by couple ID (creates list-column with each dyad's time series)
blnwh_sync_id <- BL_WH_Long %>%
  group_by(ID) %>%
  nest()

# Apply the synchrony function to each couple's data
blnwh_sync_id <- blnwh_sync_id %>%
  mutate(blnwh_sync_ccf = map(data, compute_blnwh_sync))

# Unnest results to get one row per couple per lag value
blnwh_sync_results <- unnest(blnwh_sync_id, blnwh_sync_ccf)

blnwh_sync_results$lag <- as.numeric(blnwh_sync_results$lag)
```

## Max Synchrony

```{r}


#blnwh_sync_results <- unnest(blnwh_sync_id, blnwh_sync_ccf)

# Extract max synchrony for each couple
blnwh_sync_max <- blnwh_sync_results %>%
  group_by(ID) %>%
  # Find the row with maximum absolute correlation for each couple
  slice_max(abs(acf), n = 1, with_ties = FALSE) %>%
  select(ID, 
         max_lag = lag,           # Lag at which max occurs
         max_sync = acf,          # Correlation value (with sign)
         max_abs_sync = acf,      # Will convert to absolute
         lower_ci, 
         upper_ci, 
         significant) %>%
  mutate(max_abs_sync = abs(max_sync)) %>%  # Add absolute value column
  ungroup()

# View results
head(blnwh_sync_max)
```

### Max Synchrony Interpretation

```{r}

# Add interpretation columns
blnwh_sync_max_interpret <- blnwh_sync_max %>%
  mutate(
    # Who leads?
    leader = case_when(
      max_lag > 0 ~ "Mom leads",
      max_lag < 0 ~ "Partner leads",
      max_lag == 0 ~ "Concurrent",
      TRUE ~ NA_character_
    ),
    #leader numerical
    leaderN = case_when(
      max_lag > 0 ~ "1" , #mom
      max_lag < 0 ~ "2" , #partner
      max_lag == 0 ~ "3", #concurrent
      TRUE ~ NA_character_
    ),
    # How much lead time?
    lead_time_seconds = abs(max_lag),
    # Direction of relationship
    relationship = case_when(
      max_sync > 0 ~ "Positive (same direction)",
      max_sync < 0 ~ "Negative (opposite direction)",
      TRUE ~ NA_character_
    ),
    # Strength categorization (example thresholds)
    strength = case_when(
      max_abs_sync >= 0.5 ~ "Strong",
      max_abs_sync >= 0.3 ~ "Moderate",
      max_abs_sync >= 0.1 ~ "Weak",
      TRUE ~ "Very weak"
    )
  )

# View interpreted results
blnwh_sync_max_interpret_preview <- blnwh_sync_max_interpret %>%
  select(ID, max_sync, max_lag, leader, lead_time_seconds,
         relationship, strength, significant)

knitr::kable(head(blnwh_sync_max_interpret_preview, 10), caption = "Interpreted Maximum Synchrony, Negative Interaction (first 10 couples)")
```

## Descriptives

```{r}
# Descriptive statistics across all couples
blnwh_sync_max_interpret_summary <- blnwh_sync_max_interpret %>%
  summarise(
    n_couples = n(),
    n_significant = sum(significant, na.rm = TRUE),
    pct_significant = mean(significant, na.rm = TRUE) * 100,
    
    mean_max_sync = mean(max_abs_sync, na.rm = TRUE),
    sd_max_sync = sd(max_abs_sync, na.rm = TRUE),
    
    mean_lag = mean(max_lag, na.rm = TRUE),
    sd_lag = sd(max_lag, na.rm = TRUE),
    
    n_mom_leads = sum(max_lag > 0, na.rm = TRUE),
    n_partner_leads = sum(max_lag < 0, na.rm = TRUE),
    n_concurrent = sum(max_lag == 0, na.rm = TRUE),
    
    n_positive = sum(max_sync > 0, na.rm = TRUE),
    n_negative = sum(max_sync < 0, na.rm = TRUE)
  )

print(blnwh_sync_max_interpret_summary)
```

## Plot Max Synchrony

```{r}
library(ggplot2)

# Distribution of maximum synchrony values
blnwh_hist_maxsync <- ggplot(blnwh_sync_max_interpret, aes(x = max_abs_sync, fill = significant)) +
  geom_histogram(bins = 30, alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
  labs(title = "Distribution of Maximum Synchrony Across Couples",
       x = "Maximum Absolute Correlation",
       y = "Number of Couples",
       fill = "Significant",
       caption = "Histogram of each couple's maximum absolute time-lagged cross-correlation (negative interaction), colored by statistical significance.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blnwh_hist_maxsync
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_hist_maxsync.png", plot = blnwh_hist_maxsync, width = 8, height = 6, dpi = 300)

# Distribution of lags where max occurs
blnwh_hist_lag <- ggplot(blnwh_sync_max_interpret, aes(x = max_lag, fill = leader)) +
  geom_histogram(binwidth = 1, alpha = 0.7) +
  scale_fill_manual(values = c("Mom leads" = "#CC6677", "Partner leads" = "#88CCEE", "Concurrent" = "#009988"), name = "Leader") +
  labs(title = "Lag at Maximum Synchrony",
       x = "Lag (seconds)",
       y = "Number of Couples",
       fill = "Leader",
       caption = "Histogram of the time lag (in seconds) at which each couple's maximum synchrony occurs, colored by which partner leads.") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blnwh_hist_lag
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_hist_lag.png", plot = blnwh_hist_lag, width = 8, height = 6, dpi = 300)

# Scatterplot: max correlation vs lag
blnwh_scatter_sync <- ggplot(blnwh_sync_max_interpret,
       aes(x = max_lag, y = max_sync, color = significant)) +
  geom_point(size = 3, alpha = 0.6) +
  scale_color_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Maximum Synchrony: Correlation vs Lag",
       x = "Lag at Maximum (seconds)",
       y = "Correlation Coefficient",
       color = "Significant",
       caption = "Each point is one couple's maximum correlation plotted against the lag at which it occurs, colored by statistical significance.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blnwh_scatter_sync
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_scatter_sync.png", plot = blnwh_scatter_sync, width = 8, height = 6, dpi = 300)
```


### Grid Panel

```{r}
# Show multiple couples in a grid
blnwh_sync_ccfplot <- function(blnwh_sync_results, n_couples = 12) {
  # Select first n couples
  selected_ids <- unique(blnwh_sync_results$ID)[1:min(n_couples, length(unique(blnwh_sync_results$ID)))]

  blnwh_sync_results %>%
    filter(ID %in% selected_ids) %>%
    ggplot(aes(x = lag, y = acf)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.3) +
    geom_line(color = "#CC6677", size = 0.5) +
    geom_point(aes(color = significant), size = 1) +
    geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
                alpha = 0.15, fill = "#CC6677") +
    scale_color_manual(values = c("FALSE" = "#BBBBBB", "TRUE" = "#CC3311"), name = "Significant") +
    facet_wrap(~ ID, scales = "free_y", ncol = 3) +
    labs(title = "Cross-Correlation Functions Across Couples",
         x = "Lag (seconds)",
         y = "Cross-Correlation",
         color = "Significant",
         caption = "Full cross-correlation function by lag for a sample of couples, with 95% confidence bands and significant lags highlighted.") +
    theme_minimal(base_size = 16) +
    theme(legend.position = "bottom",
          strip.text = element_text(face = "bold", size = 14),
          axis.title = element_text(size = 14), axis.text = element_text(size = 12),
          panel.grid = element_blank())
}

# Display grid
blnwh_sync_ccfplot_p <- blnwh_sync_ccfplot(blnwh_sync_results, n_couples = 3)
blnwh_sync_ccfplot_p
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_sync_ccfplot.png", plot = blnwh_sync_ccfplot_p, width = 13, height = 5, dpi = 300)
```

## Calculate Average

```{r}
# Compute mean CCF profile across all couples
blnwh_sync_avgccf <- blnwh_sync_results %>%
  group_by(lag) %>%
  summarise(
    mean_acf = mean(acf, na.rm = TRUE),
    se_acf = sd(acf, na.rm = TRUE) / sqrt(n()),
    lower_se = mean_acf - 1.96 * se_acf,
    upper_se = mean_acf + 1.96 * se_acf,
    pct_significant = mean(significant, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# Plot average CCF
blnwh_sync_avgccf_plot <- ggplot(blnwh_sync_avgccf, aes(x = lag, y = mean_acf)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_ribbon(aes(ymin = lower_se, ymax = upper_se),
              alpha = 0.3, fill = "#CC6677") +
  geom_line(color = "#CC6677", size = 1.2) +
  geom_point(aes(size = pct_significant), color = "#CC6677") +
  scale_size_continuous(range = c(1, 4), name = "% Couples\nSignificant") +
  labs(title = "Average Cross-Correlation Function Across All Couples",
       subtitle = "Shaded area represents 95% confidence interval",
       x = "Lag (seconds)\n← Partner leads | Mom leads →",
       y = "Mean Cross-Correlation",
       caption = "Mean cross-correlation function across all couples, with point size showing the percentage of couples significant at that lag.") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
blnwh_sync_avgccf_plot
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_sync_avgccf_plot.png", plot = blnwh_sync_avgccf_plot, width = 8, height = 6, dpi = 300)
```

### Plot Average

```{r}
# Heatmap showing all couples' CCF profiles. Couples are identified only by
# their row order (not their real ID) so no participant identifiers appear
# on the public site, and the y-axis is left unlabeled for the same reason.
blnwh_sync_heatmap <- function(blnwh_sync_results, n_couples = 50) {
  selected_ids <- unique(blnwh_sync_results$ID)[1:min(n_couples, length(unique(blnwh_sync_results$ID)))]

  blnwh_sync_results %>%
    filter(ID %in% selected_ids) %>%
    mutate(couple_row = factor(match(ID, selected_ids))) %>%
    ggplot(aes(x = lag, y = couple_row, fill = acf)) +
    geom_tile() +
    scale_fill_gradient2(low = "#88CCEE", mid = "white", high = "#CC6677",
                         midpoint = 0, limits = c(-1, 1),
                         name = "Correlation") +
    labs(title = "Cross-Correlation Heatmap: Random Assortment of Couples in the Study",
         x = "Lag (seconds)",
         y = NULL,
         caption = "Correlation at every lag for a random assortment of couples in the study; couples are not individually identified. Red indicates positive correlation, blue negative, centered at zero.") +
    theme_minimal(base_size = 16) +
    theme(legend.position = "bottom", axis.title = element_text(size = 14),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          axis.text.x = element_text(size = 12),
          panel.grid = element_blank())
}

blnwh_sync_heatmap_p <- blnwh_sync_heatmap(blnwh_sync_results, n_couples = 30)
blnwh_sync_heatmap_p
ggsave("../../figures/baseline-warmth-synchrony/calculation/negative/blnwh_sync_heatmap.png", plot = blnwh_sync_heatmap_p, width = 10, height = 10, dpi = 300)
```

## Concurrent Synchrony

```{r}
# Extract concurrent (lag = 0)
blnwh_sync_concurrent <- blnwh_sync_results %>%
  filter(lag == 0) %>%
  select(ID, concurrent_acf = acf, concurrent_sig = significant,
         concurrent_lower = lower_ci, concurrent_upper = upper_ci)

knitr::kable(head(blnwh_sync_concurrent, 10), caption = "Concurrent Synchrony, Negative Interaction (first 10 couples)")
```

## Time Lagged Synchrony

```{r}

# Extract time-lagged (lag != 0)
blnwh_sync_lagged <- blnwh_sync_results %>% filter(lag != 0)

#max synchrony across non-zero lags
blnwh_sync_lagmax <- blnwh_sync_lagged %>%
  group_by(ID) %>%
  # Find the row with maximum absolute correlation for each couple
  slice_max(abs(acf), n = 1, with_ties = FALSE) %>%
  select(ID, 
         max_lag = lag,           # Lag at which max occurs
         max_sync = acf,          # Correlation value (with sign)
         max_abs_sync = acf,      # Will convert to absolute
         lower_ci, 
         upper_ci, 
         significant) %>%
  mutate(max_abs_sync = abs(max_sync)) %>%  # Add absolute value column
  ungroup()

# View results
knitr::kable(head(blnwh_sync_lagmax, 10), caption = "Maximum Time-Lagged Synchrony, Negative Interaction (first 10 couples)")
```


```{r}
#| label: neg time lag to synchrony


##uses first sig lag
blnwh_sync_timetolag <- blnwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    has_significant = any(significant, na.rm = TRUE),
    time_lag_to_sync = if (any(significant, na.rm = TRUE)) {
      abs(lag[which(significant)[1]])
    } else {
      NA_integer_
    },
    .groups = "drop"
  )
```

## Summary Table - ID

```{r}

#uses strongest sig lag
blnwh_sync_summary <- blnwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    # Basic counts
    n_significant_lags = sum(significant),
    has_any_significant = any(significant),
    
    # If significant, what's the strongest?
    strongest_lag = if(any(significant)) {
      lag[which.max(abs(acf[significant]))]
    } else NA_real_,
    
    strongest_acf = if(any(significant)) {
      acf[significant][which.max(abs(acf[significant]))]
    } else NA_real_,
    
    # Specific patterns
    has_concurrent = any(significant & lag == 0),
    has_mom_leads = any(significant & lag > 0),
    has_partner_leads = any(significant & lag < 0),
    
    .groups = "drop"
  )


```

## Summary Table - All

```{r}
library(gt)
library(dplyr)

# Create the formatted table
blnwh_sync_summarytable <- blnwh_sync_summary %>%
  # Select and order columns for display
  select(ID, n_significant_lags, strongest_lag, strongest_acf,
         has_concurrent, has_mom_leads, has_partner_leads) %>%
  # Take first 20 couples for display (or use head(10) for fewer)
  head(20) %>%
  
  # Create gt table
  gt() %>%
  
  # Add title and subtitle
  tab_header(
    title = "Dyadic Synchrony Analysis",
    subtitle = "Cross-Correlation Results by Couple"
  ) %>%
  
  # Format numeric columns
  fmt_number(
    columns = c(strongest_acf),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(strongest_lag),
    decimals = 0
  ) %>%
  
  # Rename columns with readable labels
  cols_label(
    ID = "Couple ID",
    n_significant_lags = "# Significant Lags",
    strongest_lag = "Strongest Lag (sec)",
    strongest_acf = "Correlation (r)",
    has_concurrent = "Concurrent Sync",
    has_mom_leads = "Mom Leads",
    has_partner_leads = "Partner Leads"
  ) %>%
  
  # Color code the boolean columns
  data_color(
    columns = c(has_concurrent, has_mom_leads, has_partner_leads),
    colors = scales::col_factor(
      palette = c("lightgray", "lightgreen"),
      domain = c(FALSE, TRUE)
    )
  ) %>%
  
  # Color code correlation strength
  data_color(
    columns = strongest_acf,
    colors = scales::col_numeric(
      palette = c("white", "#CC6677", "#332288"),
      domain = c(0, max(blnwh_sync_summary$strongest_acf, na.rm = TRUE))
    )
  ) %>%
  
  # Color code lag direction
  data_color(
    columns = strongest_lag,
    colors = scales::col_numeric(
      palette = c("#88CCEE", "white", "#CC6677"),
      domain = c(
        min(blnwh_sync_summary$strongest_lag, na.rm = TRUE),
        max(blnwh_sync_summary$strongest_lag, na.rm = TRUE)
      )
    )
  ) %>%
  
  # Bold the column headers
  tab_style(
    style = list(
      cell_text(weight = "bold", color = "white"),
      cell_fill(color = "#CC6677")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  
  # Style the title
  tab_style(
    style = cell_text(weight = "bold", size = "large"),
    locations = cells_title(groups = "title")
  ) %>%
  
  # Add alternating row colors
  tab_style(
    style = cell_fill(color = "#F5F5F5"),
    locations = cells_body(rows = seq(2, 20, 2))
  ) %>%
  
  # Add footnotes
  tab_footnote(
    footnote = "Significant lags determined by bootstrap confidence intervals (α = .05)",
    locations = cells_column_labels(columns = n_significant_lags)
  ) %>%
  tab_footnote(
    footnote = "Positive lag = Mom leads; Negative lag = Partner leads; Zero = Concurrent",
    locations = cells_column_labels(columns = strongest_lag)
  ) %>%
  
  # Add source note
  tab_source_note(
    source_note = "Analysis based on pre-whitened ARIMA residuals with 1000 bootstrap iterations"
  ) %>%
  
  # Table size
  tab_options(
    table.font.size = px(14),
    column_labels.font.size = px(14),
    heading.title.font.size = px(16),
    heading.subtitle.font.size = px(13)
  ) %>%

  # Adjust column widths
  cols_width(
    ID ~ px(90),
    n_significant_lags ~ px(130),
    strongest_lag ~ px(140),
    strongest_acf ~ px(120),
    has_concurrent ~ px(120),
    has_mom_leads ~ px(120),
    has_partner_leads ~ px(140)
  )

# Display the table
blnwh_sync_summarytable
```

```{r}
# Create a summary statistics table instead of individual couples
blnwh_sync_summarystats <- blnwh_sync_summary %>%
  summarise(
    n_couples = n(),
    
    # Synchrony prevalence
    pct_any_significant = mean(has_any_significant, na.rm = TRUE) * 100,
    pct_concurrent = mean(has_concurrent, na.rm = TRUE) * 100,
    pct_mom_leads = mean(has_mom_leads, na.rm = TRUE) * 100,
    pct_partner_leads = mean(has_partner_leads, na.rm = TRUE) * 100,
    
    # Lag characteristics
    mean_n_sig_lags = mean(n_significant_lags, na.rm = TRUE),
    median_strongest_lag = median(strongest_lag, na.rm = TRUE),
    
    # Correlation strength
    mean_strongest_acf = mean(abs(strongest_acf), na.rm = TRUE),
    median_strongest_acf = median(abs(strongest_acf), na.rm = TRUE)
  ) %>%
  
  # Reshape for better display
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  
  # Create gt table
  gt() %>%
  tab_header(
    title = "Dyadic Synchrony: Summary Statistics",
    subtitle = paste0("N = ", nrow(blnwh_sync_summary), " couples")
  ) %>%
  
  # Format values
  fmt_number(
    columns = Value,
    decimals = 2
  ) %>%
  
  # Improve metric labels
  text_transform(
    locations = cells_body(columns = Metric),
    fn = function(x) {
      case_when(
        x == "n_couples" ~ "Total Couples",
        x == "pct_any_significant" ~ "% with Significant Synchrony",
        x == "pct_concurrent" ~ "% with Concurrent Synchrony",
        x == "pct_mom_leads" ~ "% with Mom Leading",
        x == "pct_partner_leads" ~ "% with Partner Leading",
        x == "mean_n_sig_lags" ~ "Mean # Significant Lags",
        x == "median_strongest_lag" ~ "Median Strongest Lag (sec)",
        x == "mean_strongest_acf" ~ "Mean Correlation Strength",
        x == "median_strongest_acf" ~ "Median Correlation Strength",
        TRUE ~ x
      )
    }
  ) %>%
  
  # Style
  tab_style(
    style = list(
      cell_text(weight = "bold", color = "white"),
      cell_fill(color = "#CC6677")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  
  #tab_style(
 #   style = cell_fill(color = "#F5F5F5"),
 #   locations = cells_body(rows = seq(2, 10, 2))
 # ) %>%
  
  cols_label(
    Metric = "Statistic",
    Value = "Value"
  ) %>%
  
  tab_source_note(
    source_note = "Pre-whitened cross-correlation with bootstrap significance testing"
  )

# Display
blnwh_sync_summarystats
```

```{r}
# Compare the three approaches side-by-side for a few couples
blnwh_sync_compare <- blnwh_sync_results %>%
  mutate(significant = as.vector(significant)) %>%
  filter(!is.na(significant)) %>%
  group_by(ID) %>%
  summarise(
    # Approach 1: First significant lag
    first_sig_lag = if(any(significant)) abs(lag[which(significant)[1]]) else NA_real_,
    
    # Approach 2: Categorical
    parent_led = any(significant & lag > 0),
    infant_led = any(significant & lag < 0),
    
    # Approach 3: Strongest
    strongest_lag = if(any(significant)) lag[which.max(abs(acf[significant]))] else NA_real_,
    strongest_r = if(any(significant)) acf[significant][which.max(abs(acf[significant]))] else NA_real_,
    
    .groups = "drop"
  ) %>%
  head(10) %>%
  
  gt() %>%
  tab_header(
    title = "Comparison of Three Synchrony Approaches",
    subtitle = "First 10 couples"
  ) %>%
  fmt_number(columns = c(first_sig_lag, strongest_lag), decimals = 0) %>%
  fmt_number(columns = strongest_r, decimals = 3) %>%
  cols_label(
    ID = "Couple",
    first_sig_lag = "First Sig Lag",
    parent_led = "Parent Led",
    infant_led = "Infant Led",
    strongest_lag = "Strongest Lag",
    strongest_r = "Strongest r"
  ) %>%
  tab_spanner(
    label = "Approach 1: First",
    columns = first_sig_lag
  ) %>%
  tab_spanner(
    label = "Approach 2: Categorical",
    columns = c(parent_led, infant_led)
  ) %>%
  tab_spanner(
    label = "Approach 3: Strongest",
    columns = c(strongest_lag, strongest_r)
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold", color = "white"),
                 cell_fill(color = "#CC6677")),
    locations = cells_column_labels(everything())
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold", color = "white"),
                 cell_fill(color = "#332288")),
    locations = cells_column_spanners(spanners = everything())
  )

blnwh_sync_compare
```
