---
title: Synchrony Processing
toc: true
toc-depth: 3
bibliography: references.bib
warning: false
editor: 
  markdown: 
    wrap: 72
---

# Overview

This is the second page of a three-page workflow examining
**moment-to-moment dyadic dynamics in baseline couple interaction warmth**.
It begins with the validated wide-format baseline warmth dataset created on the
[Import & Reliability](Import_BL_WH_Syntax_annotated_v4.qmd) page and prepares the second-by-second
trajectories used in the synchrony analysis.

The processing workflow:

1. reshapes the wide SPSS file from one row per couple with one column per
   second into one row per couple-second;
2. recodes the missing-value placeholder to `NA`;
3. creates separate positive- and negative-discussion person-level long
   datasets for mom and partner;
4. saves the three analysis ready long files used by the synchrony calculation;
   and
5. plots warmth trajectories for the full sample and a representative subset
   as a visual data-quality and interpretive check.

::: {.callout-note appearance="simple" icon="false"}
### Workflow

**1. Import & Reliability** → **2. Processing & Visualization** → **3. Synchrony Calculation**

[Import & Reliability](Import_BL_WH_Syntax_annotated_v4.qmd) · [Processing & Visualization](Baseline-Synchrony_Processing.qmd) · [Synchrony Calculation](Baseline-Synchrony_Calculation.qmd)
:::


The preceding page documents the original CARMA import, validation, and
reliability workflow: [Import & Reliability](Import_BL_WH_Syntax_annotated_v4.qmd). The next page
uses the processed long-format files to calculate concurrent and time-lagged
synchrony: [Synchrony Calculation](Baseline-Synchrony_Calculation.qmd).

# Data Cleaning

**Load Packages**

```{r}
#| label: packages
#| warning: false
#| output: false

# Load packages used for reshaping, visualization, time-series analysis, modeling, and formatted output
library(here)
library(nlme)
library(tidyverse)
library(ggplot2) 
library(lavaan)
library(semTools)
library(jtools)
library(haven)
library(modsem)
library(forecast)
library(gt)
```

```{r}
#| label: load data
#| include: false

BL_WH <- read_sav("C:/Users/rcm5550/Box/ff (cast.casbox )/DATA & SYNTAX/COMPUTED DATA/Baseline/Observed couple relationship/BL_WH.sav")

# Demo dataset is used only for rendered row-level previews

demo_path <- "C:/Users/rcm5550/Desktop/portfolio-starter/docs/projects/baseline-warmth-synchrony/BL_WH_demo.csv"
BL_WH_Demo <- read.csv(
  demo_path,
  check.names = FALSE)

```

```{r}
#| label: preview-dataset
#| include: false
#| eval: false

# Row-level preview suppressed for the public site since it would print
# Real values even with the code hidden. Flip eval to true to preview locally
knitr::kable(head(BL_WH, n = 5)[1:6], caption = "Raw Baseline Warmth Data")
```

## Transformation

The raw file has one row per couple and separate columns for each second
of the interaction (e.g. `blmnwh.6`, `blmnwh.7`, ...). This section
reshapes that into a single long-format table (one row per
couple-second), recodes the dataset's missing-value placeholder to `NA`,
and then splits the negative- and positive-discussion variables into
their own person-level long datasets for plotting and later synchrony
calculation.

**Reshape Wide to Long**

```{r}
#| label: wide to long

# Transform data wide to long
BL_WH_WidetoLong <- BL_WH %>%
  pivot_longer(
    # Select columns like blmnwh.6, blmnwh.7, etc
    cols = matches("^[^.]+\\.\\d+$"), 
    
    # Split variable name based on period
    names_to = c(".value", "second"),  
    names_sep = "\\." ) %>%
  
  # Convert second to numeric
  mutate(second = as.numeric(second))
```


```{r}
#| include: false 
# Transform demo data wide to long for displayed previews only
BL_WH_Demo_WidetoLong <- BL_WH_Demo %>%
  pivot_longer(
    cols = matches("^[^.]+\\.\\d+$"),
    names_to = c(".value", "second"),
    names_sep = "\\.") %>%
  mutate(second = as.numeric(second))

# Create demo long dataset for displayed previews only
BL_WH_Demo_Long <- BL_WH_Demo_WidetoLong %>%
  select(ID, second, blmnwh, blmpwh, blfnwh, blfpwh)

BL_WH_Demo_Long$blmnwh[BL_WH_Demo_Long$blmnwh == 0] <- NA
BL_WH_Demo_Long$blmpwh[BL_WH_Demo_Long$blmpwh == 0] <- NA
BL_WH_Demo_Long$blfnwh[BL_WH_Demo_Long$blfnwh == 0] <- NA
BL_WH_Demo_Long$blfpwh[BL_WH_Demo_Long$blfpwh == 0] <- NA
```

**Recode Missing Values**

```{r}
#| label: missing data

# Select the four warmth variables, second, ID
BL_WH_Long <- BL_WH_WidetoLong %>%
  select(ID, second, blmnwh, blmpwh, blfnwh, blfpwh)

# Set 0 as missing
BL_WH_Long$blmnwh[BL_WH_Long$blmnwh == 0] <- NA
BL_WH_Long$blmpwh[BL_WH_Long$blmpwh == 0] <- NA
BL_WH_Long$blfnwh[BL_WH_Long$blfnwh == 0] <- NA
BL_WH_Long$blfpwh[BL_WH_Long$blfpwh == 0] <- NA
```

```{r}
# echo: false

knitr::kable(
  head(BL_WH_Demo_Long, 5),
  caption = "Long Baseline Warmth Data (Demo)")
```


**Create Negative Interaction Dataset**

```{r}
#| label: create negative interaction dataset

# Create long format for negative interaction
# (Blmnwh = mom, blfnwh = partner)

BL_WH_Neg_Long <- BL_WH_Long %>%
  select(ID, second, blmnwh, blfnwh) %>%
  tidyr::pivot_longer(
    cols = c(blmnwh, blfnwh),
    names_to = "source",
    values_to = "blnwh") %>%
  mutate(
    personF = case_when(
      source == "blmnwh" ~ "mom",  # Mom
      source == "blfnwh" ~ "partner" ),  # Partner
    personN = case_when(
      source == "blmnwh" ~ 1,  # Mom
      source == "blfnwh" ~ 2 ))  # Partner

# Set personF as factor
BL_WH_Neg_Long$personF <- as.factor(BL_WH_Neg_Long$personF)

# Set personN as factor
BL_WH_Neg_Long$personN <- as.numeric(BL_WH_Neg_Long$personN)
```

```{r}
#| echo: false
# Create demo negative-interaction dataset for displayed preview only
BL_WH_Demo_Neg_Long <- BL_WH_Demo_Long %>%
  select(ID, second, blmnwh, blfnwh) %>%
  pivot_longer(
    cols = c(blmnwh, blfnwh),
    names_to = "source",
    values_to = "blnwh") %>%
  mutate(
    personF = case_when(
      source == "blmnwh" ~ "mom",
      source == "blfnwh" ~ "partner"),
    personN = case_when(
      source == "blmnwh" ~ 1,
      source == "blfnwh" ~ 2))

BL_WH_Demo_Neg_Long$personF <- as.factor(BL_WH_Demo_Neg_Long$personF)
BL_WH_Demo_Neg_Long$personN <- as.numeric(BL_WH_Demo_Neg_Long$personN)

knitr::kable(
  head(BL_WH_Demo_Neg_Long, 20),
  caption = "Long Baseline Warmth Negative Interaction Data (Demo)")
```

**Create Positive Interaction Dataset**

```{r}
#| label: create positive interaction dataset

# Create long format for positive interaction
# (Blmpwh = mom, blfpwh = partner)

BL_WH_Pos_Long <- BL_WH_Long %>%
  select(ID, second, blmpwh, blfpwh) %>%
  tidyr::pivot_longer(
    cols = c(blmpwh, blfpwh),
    names_to = "source",
    values_to = "blpwh") %>%
  mutate(
    personF = case_when(
      source == "blmpwh" ~ "mom",  # Mom
      source == "blfpwh" ~ "partner" ),  # Partner
    personN = case_when(
      source == "blmpwh" ~ 1,  # Mom
      source == "blfpwh" ~ 2 ))  # Partner

# Set personF as factor
BL_WH_Pos_Long$personF <- as.factor(BL_WH_Pos_Long$personF)

# Set personN as factor
BL_WH_Pos_Long$personN <- as.numeric(BL_WH_Pos_Long$personN)
```

```{r}
#| echo: false

# Create demo positive-interaction dataset for displayed preview only
BL_WH_Demo_Pos_Long <- BL_WH_Demo_Long %>%
  select(ID, second, blmpwh, blfpwh) %>%
  pivot_longer(
    cols = c(blmpwh, blfpwh),
    names_to = "source",
    values_to = "blpwh") %>%
  mutate(
    personF = case_when(
      source == "blmpwh" ~ "mom",
      source == "blfpwh" ~ "partner"),
    personN = case_when(
      source == "blmpwh" ~ 1,
      source == "blfpwh" ~ 2))

BL_WH_Demo_Pos_Long$personF <- as.factor(BL_WH_Demo_Pos_Long$personF)
BL_WH_Demo_Pos_Long$personN <- as.numeric(BL_WH_Demo_Pos_Long$personN)

knitr::kable(
  head(BL_WH_Demo_Pos_Long, 20),
  caption = "Long Baseline Warmth Positive Interaction Data (Demo)")
```

The three long-format datasets are saved in a hidden export chunk so the
visible page stays focused on transformation and quality checks while the
[Synchrony Calculation](Baseline-Synchrony_Calculation.qmd) page can
read the exact processed files produced here.

```{r}
#| label: save-processed-data
#| include: false

# ── Save Analysis-Ready Long Files ────────────────────────────────────────────
# Define the directory consumed by the synchrony calculation page
SYNC_OUTPUT_DIR <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/SFF Synchrony/Processed Data"

# Create the output directory when it does not already exist
dir.create(SYNC_OUTPUT_DIR, recursive = TRUE, showWarnings = FALSE)

# Save the complete couple-second dataset used as the shared processed source
write_csv(BL_WH_Long, file.path(SYNC_OUTPUT_DIR, "BL_WH_Long.csv"))

# Save the positive-discussion person-level trajectories
write_csv(BL_WH_Pos_Long, file.path(SYNC_OUTPUT_DIR, "BL_WH_Pos_Long.csv"))

# Save the negative-discussion person-level trajectories
write_csv(BL_WH_Neg_Long, file.path(SYNC_OUTPUT_DIR, "BL_WH_Neg_Long.csv"))
```

## Panel Plots {#secPanel}

The two figures below show second-by-second warmth trends for the full
sample, split by person (mom vs. partner) and by discussion type
(negative vs. positive). The two figures after that repeat the same
comparison for a representative subset of couples, faceted individually,
so that couple-level variation around the sample trend is visible rather
than smoothed away.

**Warmth During Negative Discussion**

```{r}
#| label: bl-negative-warmth
#| code-fold: true
#| code-summary: "figure code"

# Plot all negative-discussion warmth trajectories to inspect dyad-level patterns and potential data anomalies
fig_blnegative_warmth <- ggplot(data = BL_WH_Neg_Long, aes(second, blnwh, color = personF)) +
  geom_smooth(linewidth = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Negative Discussion") +
  xlab("Time") +
  ylab("Average Warmth") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blnegative-warmth
#| echo: false
#| fig-width: 14
#| fig-height: 8

fig_blnegative_warmth
```

```{r}
#| label: blnegative-warmth-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blnegative-warmth.png", plot = fig_blnegative_warmth, width = 8, height = 6, dpi = 300)
```

```{r}
#| label: bl-positive-warmth
#| code-fold: true
#| code-summary: "figure code"

# Plot all positive-discussion warmth trajectories to inspect dyad-level patterns and potential data anomalies
fig_blpositive_warmth <- ggplot(data = BL_WH_Pos_Long, aes(second, blpwh, color = personF)) +
  geom_smooth(linewidth = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Positive Discussion") +
  xlab("Seconds") +
  ylab("Average Warmth") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blpositive-warmth
#| echo: false
#| #| fig-width: 14
#| fig-height: 8

fig_blpositive_warmth
```

```{r}
#| label: fig-blpositive-warmth-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blpositive-warmth.png", plot = fig_blpositive_warmth, width = 8, height = 6, dpi = 300)
```

```{r}
#| label: create-subset-for-plots
#| include: false
# Select Sample Couples for Facet Plots
# Create filtered dataset for plot

# Select three specific couples for illustrative facet plots
sample_ids <- c(112, 254, 137)

# Assign fixed pseudonymous IDs above 250 for display
sample_id_map <- c(
  "112" = "364",
  "254" = "527",
  "137" = "418"
)

BL_WH_Neg_Long_Filtered <- BL_WH_Neg_Long %>%
  filter(ID %in% sample_ids)

BL_WH_Pos_Long_Filtered <- BL_WH_Pos_Long %>%
  filter(ID %in% sample_ids)
```

```{r}
#| label: blnegative-warmth-subset
#| code-fold: true
#| code-summary: "figure code"

# Plot a smaller negative-discussion subset so within-couple temporal patterns remain visually interpretable
fig_blnegative_warmth_subset <- ggplot(data = BL_WH_Neg_Long_Filtered, aes(second, blnwh, color = personF)) +
  geom_line(lwd = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Negative Discussion (Sample Couples)") +
  xlab("Time - 0 to 600 seconds") +
  ylab("No (1) to High Warmth (7)") +
  facet_wrap(
    ~ ID,
    ncol = 3,
    labeller = labeller(ID = sample_id_map)) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blnegative-warmth-subset
#| echo: false
#| fig-width: 14
#| fig-height: 8


fig_blnegative_warmth_subset
```

```{r}
#| label: fig-blnegative-warmth-subset-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blnegative-warmth-subset.png", plot = fig_blnegative_warmth_subset, width = 13, height = 5, dpi = 300)
```

```{r}
#| label: blpositive-warmth-subset
#| code-fold: true

# Plot a smaller positive-discussion subset so within-couple temporal patterns remain visually interpretable
fig_blpositive_warmth_subset <- ggplot(data = BL_WH_Pos_Long_Filtered, aes(second, blpwh, color = personF)) +  geom_line(lwd = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Positive Discussion (Sample Couples)") +
  xlab("Time - 0 to 600 seconds") +
  ylab("No (1) to High Warmth (7)") +
  facet_wrap(
    ~ ID,
    ncol = 3,
    labeller = labeller(ID = sample_id_map)) +  
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blpositive-warmth-subset
#| echo: false
#| fig-width: 14
#| fig-height: 8

fig_blpositive_warmth_subset
```

```{r}
#| label: fig-blpositive-warmth-subset-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blpositive-warmth-subset.png", plot = fig_blpositive_warmth_subset, width = 13, height = 5, dpi = 300)
```
