---
title: Baseline Synchrony Processing
toc: true
toc-depth: 3
warning: false
editor: 
  markdown: 
    wrap: 72
---

# Overview

This page performs the initial cleaning and reshaping of the baseline
couple interaction warmth data, ahead of the second-by-second synchrony
calculation carried out on [Baseline Synchrony
Calculation](Baseline-Synchrony_Calculation.qmd). Four steps happen
here:

1.  the raw wide-format SPSS export (one row per couple, one column per
    second) is reshaped into long format (one row per couple-second);
2.  missing-value codes are recoded to `NA`;
3.  the data is split into separate negative- and positive-discussion
    long datasets, one row per person per second; and
4.  warmth trajectories over time are plotted, both for the full sample
    and a representative subset in [Panel Plots].

# Data Cleaning

**Load Packages**

```{r}
#| label: packages
#| warning: false
#| output: false
library(here)
library(nlme)
library(tidyverse)
library(ggplot2) 
library(lavaan)
library(semTools)
library(jtools)
library(haven)
library(modsem)
library(forecast)
library(gt)
```

```{r}
#| label: load data
#| include: false

BL_WH <- read_sav("C:/Users/rcm5550/Box/ff (cast.casbox )/DATA & SYNTAX/COMPUTED DATA/Baseline/Observed couple relationship/BL_WH.sav")
```

```{r}
#| label: preview-dataset
#| include: false
#| eval: false

# Row-level preview suppressed for the public site since it would print
# real values even with the code hidden. Flip eval to true to preview locally.
knitr::kable(head(BL_WH, n = 5)[1:6], caption = "Raw Baseline Warmth Data")
```

## Transformation

The raw file has one row per couple and separate columns for each second
of the interaction (e.g. `blmnwh.6`, `blmnwh.7`, ...). This section
reshapes that into a single long-format table (one row per
couple-second), recodes the dataset's missing-value placeholder to `NA`,
and then splits the negative- and positive-discussion variables into
their own person-level long datasets for plotting and later synchrony
calculation.

**Reshape Wide to Long**

```{r}
#| label: wide to long

# Transform data wide to long
BL_WH_WidetoLong <- BL_WH %>%
  pivot_longer(
    # Select columns like blmnwh.6, blmnwh.7, etc.
    cols = matches("^[^.]+\\.\\d+$"), 
    
    # Split variable name based on period
    names_to = c(".value", "second"),  
    names_sep = "\\." ) %>%
  
  # Convert second to numeric
  mutate(second = as.numeric(second))
```

**Recode Missing Values**

```{r}
#| label: missing data

# Select the four warmth variables, second, ID
BL_WH_Long <- BL_WH_WidetoLong %>%
  select(ID, second, blmnwh, blmpwh, blfnwh, blfpwh)

# Set 0 as missing
BL_WH_Long$blmnwh[BL_WH_Long$blmnwh == 0] <- NA
BL_WH_Long$blmpwh[BL_WH_Long$blmpwh == 0] <- NA
BL_WH_Long$blfnwh[BL_WH_Long$blfnwh == 0] <- NA
BL_WH_Long$blfpwh[BL_WH_Long$blfpwh == 0] <- NA

# Preview dataset
knitr::kable(head(BL_WH_Long, 5), caption = "Long Baseline Warmth Data")
```

**Create Negative Interaction Dataset**

```{r}
#| label: create negative interaction dataset

# Create long format for negative interaction 
##(blmnwh = mom, blfnwh = partner)

BL_WH_Neg_Long <- BL_WH_Long %>%
  select(ID, second, blmnwh, blfnwh) %>%
  tidyr::pivot_longer(
    cols = c(blmnwh, blfnwh),
    names_to = "source",
    values_to = "blnwh") %>%
  mutate(
    personF = case_when(
      source == "blmnwh" ~ "mom",        #mom
      source == "blfnwh" ~ "partner" ),  #partner
    personN = case_when(
      source == "blmnwh" ~ 1,    #mom
      source == "blfnwh" ~ 2 ))  #partner 

# Set personF as factor
BL_WH_Neg_Long$personF <- as.factor(BL_WH_Neg_Long$personF)

# Set personN as factor
BL_WH_Neg_Long$personN <- as.numeric(BL_WH_Neg_Long$personN)

# Preview dataset
knitr::kable(head(BL_WH_Neg_Long, 20), caption = "Long Baseline Warmth Negative Interaction Data")
```

**Create Positive Interaction Dataset**

```{r}
#| label: create positive interaction dataset

# Create long format for positive interaction 
##(blmpwh = mom, blfpwh = partner)

BL_WH_Pos_Long <- BL_WH_Long %>%
  select(ID, second, blmpwh, blfpwh) %>%
  tidyr::pivot_longer(
    cols = c(blmpwh, blfpwh),
    names_to = "source",
    values_to = "blpwh") %>%
  mutate(
    personF = case_when(
      source == "blmpwh" ~ "mom",        #mom
      source == "blfpwh" ~ "partner" ),  #partner
    personN = case_when(
      source == "blmpwh" ~ 1,    #mom
      source == "blfpwh" ~ 2 ))  #partner 

# Set personF as factor
BL_WH_Pos_Long$personF <- as.factor(BL_WH_Pos_Long$personF)

# Set personN as factor
BL_WH_Pos_Long$personN <- as.numeric(BL_WH_Pos_Long$personN)

# Preview dataset
knitr::kable(head(BL_WH_Pos_Long, 20), caption = "Long Baseline Warmth Positive Interaction Data")
```

These three long-format datasets (`BL_WH_Long`, `BL_WH_Pos_Long`,
`BL_WH_Neg_Long`) are not re-exported here; the [Baseline Synchrony
Calculation](Baseline-Synchrony_Calculation.qmd) page reads the same
processed data from where it's already saved.

## Panel Plots {#secPanel}

The two figures below show second-by-second warmth trends for the full
sample, split by person (mom vs. partner) and by discussion type
(negative vs. positive). The two figures after that repeat the same
comparison for a representative subset of couples, faceted individually,
so that couple-level variation around the sample trend is visible rather
than smoothed away.

**Warmth During Negative Discussion**

```{r}
#| label: bl-negative-warmth
#| code-fold: true
#| code-summary: "figure code"

fig_blnegative_warmth <- ggplot(data = BL_WH_Neg_Long, aes(second, blnwh, color = personF)) +
  geom_smooth(linewidth = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Negative Discussion") +
  xlab("Time") +
  ylab("Average Warmth") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blnegative-warmth
#| echo: false
#| fig-width: 14
#| fig-height: 8

fig_blnegative_warmth
```

```{r}
#| label: blnegative-warmth-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blnegative-warmth.png", plot = fig_blnegative_warmth, width = 8, height = 6, dpi = 300)
```

```{r}
#| label: bl-positive-warmth
#| code-fold: true
#| code-summary: "figure code"

fig_blpositive_warmth <- ggplot(data = BL_WH_Pos_Long, aes(second, blpwh, color = personF)) +
  geom_smooth(linewidth = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Positive Discussion") +
  xlab("Seconds") +
  ylab("Average Warmth") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blpositive-warmth
#| echo: false
#| #| fig-width: 14
#| fig-height: 8

fig_blpositive_warmth
```

```{r}
#| label: fig-blpositive-warmth-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blpositive-warmth.png", plot = fig_blpositive_warmth, width = 8, height = 6, dpi = 300)
```

```{r}
#| label: create-subset-for-plots
#| include: false
# Select Sample Couples for Facet Plots
#create filtered dataset for plot

BL_WH_Neg_Long_Filtered <- BL_WH_Neg_Long %>%
  filter(ID %in% c(110, 112, 149, 162, 172, 178, 175, 182, 183, 209, 246, 254, 279, 303))

BL_WH_Pos_Long_Filtered <- BL_WH_Pos_Long %>%
  filter(ID %in% c(110, 112, 149, 162, 172, 178, 175, 182, 183, 209, 246, 254, 279, 303))
```

```{r}
#| label: blnegative-warmth-subset
#| code-fold: true
#| code-summary: "figure code"

fig_blnegative_warmth_subset <- ggplot(data = BL_WH_Neg_Long_Filtered %>% filter(ID %in% head(unique(ID), 3)), aes(second, blnwh, color = personF)) +
  geom_line(lwd = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Negative Discussion (Sample Couples)") +
  xlab("Time - 0 to 600 seconds") +
  ylab("No (1) to High Warmth (7)") +
  facet_wrap(~ ID, ncol = 3) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blnegative-warmth-subset
#| echo: false
#| fig-width: 14
#| fig-height: 8


fig_blnegative_warmth_subset
```

```{r}
#| label: fig-blnegative-warmth-subset-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blnegative-warmth-subset.png", plot = fig_blnegative_warmth_subset, width = 13, height = 5, dpi = 300)
```

```{r}
#| label: blpositive-warmth-subset
#| code-fold: true

fig_blpositive_warmth_subset <- ggplot(data = BL_WH_Pos_Long_Filtered %>% filter(ID %in% head(unique(ID), 3)), aes(second, blpwh, color = personF)) +
  geom_line(lwd = 1.5) +
  scale_color_manual(values = c("mom" = "#CC6677", "partner" = "#88CCEE"), name = "Person") +
  labs(title = "Couple Warmth During Positive Discussion (Sample Couples)") +
  xlab("Time - 0 to 600 seconds") +
  ylab("No (1) to High Warmth (7)") +
  facet_wrap(~ ID, ncol = 3) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.title = element_text(size = 16), axis.text = element_text(size = 14),
        panel.grid = element_blank())
```

```{r}
#| label: fig-blpositive-warmth-subset
#| echo: false
#| fig-width: 14
#| fig-height: 8

fig_blpositive_warmth_subset
```

```{r}
#| label: fig-blpositive-warmth-subset-save
#| include: false

ggsave("../../figures/baseline-warmth-synchrony/processing/fig-blpositive-warmth-subset.png", plot = fig_blpositive_warmth_subset, width = 13, height = 5, dpi = 300)
```
