---
title: "Baseline Warmth Data Import, Processing, and Reliability"
date: "2026-08-10"
bibliography: references.bib
---

# Purpose

This file performs two related tasks for **baseline observational warmth coding**.

First, it imports the original baseline CARMA files, processes each video beginning at **second 6**, creates the four Mom/Partner × Negative/Positive warmth datasets in R, merges them by couple ID, and saves one final SPSS dataset (`BL_WH.sav`). The workflow includes data-quality stop points so that invalid warmth values, duplicate IDs, duplicate seconds, or unexpected CARMA timing cannot silently enter the final dataset.

Second, sections 5–9 below import the separate reliability-coded CARMA files, process them using the same second-by-second cleaning rules, match them to the corresponding original-coded observations, and calculate inter-rater reliability. Reliability data stay in R and are not added to `BL_WH.sav`.

Warmth values of **0** are converted to `NA` during processing so that SPSS reads them as system-missing values and they are excluded from average warmth scores.

# 1. Setup

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#| label: file-paths
#| include: false
# EDIT THIS LINE: replace the user-specific portion with your own Box location.
PROJECT_ROOT <- "C:/Users/rcm5550/Box/ff (cast.casbox )"

# Original baseline warmth CARMA files are stored within one subfolder per couple.
PATH_CARMA_BASELINE_WARMTH <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding/Couple Interaction Coding (warmth, withdrawal, hostility)",
  "Carma Files/Baseline/Warmth"
)

# Final merged SPSS file lives in the baseline computed-data folder.
PATH_OUTPUT_BL_WH <- file.path(
  PROJECT_ROOT,
  "DATA & SYNTAX/COMPUTED DATA/Baseline/Observed couple relationship/BL_WH.sav"
)

# Positive warmth reliability files.
PATH_WARMTH_POS_RELIABILITY <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding",
  "Couple Interaction Coding (warmth, withdrawal, hostility)",
  "Reliability/Baseline/Warmth/Positive"
)

# Negative warmth reliability files.
# EDIT ONLY IF the Negative reliability folder uses a different folder structure.
PATH_WARMTH_NEG_RELIABILITY <- file.path(
  PROJECT_ROOT,
  "FF Staff Folders Outcomes/Observational Coding",
  "Couple Interaction Coding (warmth, withdrawal, hostility)",
  "Reliability/Baseline/Warmth/Negative"
)
```

## Load required R packages

These packages are used to import files, reshape the second-by-second warmth data, combine datasets, and (in the source project) export the final SPSS `.sav` file. `tidyverse` is used for file processing, reshaping, and matching. `haven` creates the SPSS dataset. `irr` is used for the reliability calculation.

```{r}
#| label: libraries
#| include: false
library(tidyverse)
library(haven)
library(irr)
```

# 2. Define Reusable Import and Processing Functions

## Extract the couple ID from a file name {#sec-id-function}

Original CARMA files are stored in couple folders such as `Couple 564/101BL_M_NEG_WH_FL.csv`.

The function below extracts the **three-digit couple ID from the beginning of the CSV file name**.

The script stops if a file does not begin with a three-digit ID rather than assigning an incorrect ID.

```{r}
#| label: id-function
extract_couple_id <- function(filename) {

  # Extract the first three digits from the base file name.
  id <- str_extract(basename(filename), "^\\d{3}")

  # Stop if a three-digit ID cannot be identified.
  if (is.na(id)) {
    stop("Could not identify a three-digit couple ID from file: ", basename(filename))
  }

  id
}
```

## Process one CARMA warmth file {#sec-single-file-function}

This function contains the core cleaning rules that should be identical for **every CARMA file**. It imports one CSV and returns a standardized long dataset containing one row per coded second.

For each file, the function:

1.  Imports the first two CARMA columns as `second` and `warmth`.
2.  Converts the second column to numeric and explicitly locates **second 6**.
3.  Stops if second 6 is missing or appears more than once.
4.  Keeps the coding window beginning exactly at second 6, regardless of how many metadata rows precede it.
5.  Stops if duplicate seconds occur within the retained coding window.
6.  Extracts the couple ID from the beginning of the file name.
7.  Converts warmth to numeric and stops if an observed value is outside the valid **0–7** range.
8.  Truncates valid warmth values to whole-number codes to match the existing processing convention.
9.  Converts warmth `0` to `NA`.

Keeping these rules in one function ensures that CARMA files are processed in exactly the same way.

```{r}
#| label: single-file-function
process_single_warmth_file <- function(filename_full) {

  # Keep the base file name for readable error messages and documentation.
  filename_short <- basename(filename_full)

  # Import the full CARMA export without assuming a fixed number of metadata rows.
  data <- read.csv(filename_full) %>%
    as.data.frame()

  # Stop if the CARMA export does not contain at least two columns.
  if (ncol(data) < 2) {
    stop("Expected at least two columns in file: ", filename_short)
  }

  # Retain and standardize the two CARMA columns used in processing.
  data <- data[, 1:2]
  names(data) <- c("second", "warmth")

  # Convert the second column to numeric so the start of the coding window can be located.
  data <- data %>%
    mutate(second = suppressWarnings(as.numeric(as.character(second))))

  # Locate the single row corresponding to second 6.
  second_6_rows <- which(!is.na(data$second) & data$second == 6)

  # Stop if the start of the required coding window cannot be identified uniquely.
  if (length(second_6_rows) == 0) {
    stop("Could not find second 6 in file: ", filename_short)
  }

  if (length(second_6_rows) > 1) {
    stop("Second 6 appears more than once in file: ", filename_short)
  }

  # Keep only the intended coding window from second 6 through second 600.
  # If a CARMA file extended beyond 600 seconds, later seconds are excluded.
  data <- data[second_6_rows:nrow(data), , drop = FALSE] %>%
    filter(second <= 600)

  # Remove rows after the coded time series that do not contain a numeric second.
  data <- data %>%
    filter(!is.na(second))

  # Stop if seconds are duplicated within the retained coding window.
  duplicate_seconds <- unique(data$second[duplicated(data$second)])

  if (length(duplicate_seconds) > 0) {
    stop(
      "Duplicate second values found after second 6 in file ",
      filename_short, ": ", paste(duplicate_seconds, collapse = ", ")
    )
  }

  # Confirm that the retained coding window begins at second 6.
  if (nrow(data) == 0 || data$second[1] != 6) {
    stop("Retained coding window does not begin at second 6 in file: ", filename_short)
  }

  # Extract the three-digit couple ID from the beginning of the file name.
  couple_id <- extract_couple_id(filename_short)

  # Convert warmth values to numeric using the same order as the original working syntax.
  data <- data %>%
    mutate(warmth = suppressWarnings(as.numeric(as.character(warmth))))

  # Remove decimals before validation, matching manual of whole number codes.
  data <- data %>%
    mutate(warmth = trunc(warmth))

  # Stop if any processed nonmissing warmth value falls outside the valid 0-7 coding range.
  invalid_warmth <- data %>%
    filter(!is.na(warmth), warmth < 0 | warmth > 7)

  if (nrow(invalid_warmth) > 0) {
    stop(
      "Processed warmth value outside the valid 0-7 range for ID ",
      couple_id, " in file: ", filename_short,
      ". Offending second(s): ",
      paste(invalid_warmth$second, collapse = ", "),
      ". Value(s): ",
      paste(invalid_warmth$warmth, collapse = ", ")
    )
  }

  # Convert code 0 to R missing so SPSS imports it as system missing.
  data <- data %>%
    mutate(warmth = na_if(warmth, 0))

  # Return one standardized row per coded second.
  data %>%
    transmute(
      ID = couple_id,
      second = second,
      warmth = warmth,
      source_file = filename_short
    )
}
```

## Process one original baseline warmth condition

This wrapper finds all original CARMA files matching one condition (`BL_M_NEG`, `BL_P_NEG`, `BL_M_POS`, or `BL_P_POS`), sends every file through the [single-file processing function](#sec-single-file-function), and then converts the standardized long data into the wide format used in the final SPSS dataset.

Before processing, the function checks that only one source file exists for each couple within that condition. Duplicate IDs in the finished wide dataset are checked later by the [dataset validation function](#sec-validation-function).

```{r}
#| label: production-wrapper-function
process_warmth_files <- function(path, pattern, variable_prefix, average_name) {

  # Find all CSV files matching the requested original warmth condition.
  files_full <- list.files(
    path = path,
    pattern = paste0(pattern, ".*\\.csv$"),
    full.names = TRUE,
    recursive = TRUE,
    ignore.case = TRUE
  )

  # Stop if no matching files are found.
  if (length(files_full) == 0) {
    stop("No files found for pattern: ", pattern)
  }

  # Extract the couple ID from each matching base file name.
  file_ids <- map_chr(files_full, extract_couple_id)

  # Stop before processing if more than one source file maps to the same couple ID.
  duplicate_source_ids <- unique(file_ids[duplicated(file_ids)])

  if (length(duplicate_source_ids) > 0) {
    stop(
      "Duplicate source files found for ", pattern, " IDs: ",
      paste(duplicate_source_ids, collapse = ", ")
    )
  }

  # Process every source file with the common second-by-second cleaning function.
  dataset_long <- map_dfr(files_full, process_single_warmth_file)

  # Reshape the condition from one row per second (long) to one row per ID (wide).
  dataset <- dataset_long %>%
    select(ID, second, warmth) %>%
    mutate(second = str_c(variable_prefix, ".", second)) %>%
    pivot_wider(
      id_cols = ID,
      names_from = second,
      values_from = warmth
    )

  # Identify all second-by-second warmth variables created for this condition.
  warmth_vars <- names(dataset)[str_starts(names(dataset), paste0(variable_prefix, "."))]

  # Calculate each ID's average warmth across all available coded seconds.
  dataset[[average_name]] <- rowMeans(dataset[warmth_vars], na.rm = TRUE)

  # If all second-level warmth values are missing, rowMeans(..., na.rm = TRUE)
    # returns NaN because there are no numeric values left to average.
    # Convert NaN to NA so the video-level average is stored as missing.
  dataset[[average_name]][is.nan(dataset[[average_name]])] <- NA_real_

  dataset
}
```

## Validate a finished wide warmth dataset {#sec-validation-function}

This function is the final data-quality stop point for the wide production datasets. It checks that there is one row per ID and that all nonmissing second-by-second warmth values are within **1–7** after code 0 has been converted to missing.

```{r}
#| label: validation-function
validate_warmth_dataset <- function(dataset, dataset_name) {

  # Stop if the finished dataset contains duplicate couple IDs.
  duplicate_ids <- dataset %>%
    count(ID) %>%
    filter(n > 1)

  if (nrow(duplicate_ids) > 0) {
    stop(
      dataset_name, " contains duplicate IDs: ",
      paste(duplicate_ids$ID, collapse = ", ")
    )
  }

  # Select second-by-second warmth variables while excluding ID and average variables.
  warmth_vars <- names(dataset)[
    names(dataset) != "ID" &
      !str_detect(names(dataset), "\\.avg$")
  ]

  # Stop if any nonmissing warmth value falls outside 1-7 after zero-to-missing conversion.
  invalid_values <- dataset %>%
    select(all_of(warmth_vars)) %>%
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "warmth"
    ) %>%
    filter(!is.na(warmth), warmth < 1 | warmth > 7)

  if (nrow(invalid_values) > 0) {
    stop(dataset_name, " contains warmth values outside the expected 1-7 range.")
  }

  message(dataset_name, ": data-quality checks passed.")
}
```

# 3. Process the Original Baseline Warmth Files

Each section below calls the same [processing function](#sec-single-file-function) with the file pattern and variable prefix for one warmth condition. The four condition-specific datasets remain in R and are not saved separately.

## Baseline Mom Negative Warmth

This chunk imports all baseline Mom Negative Warmth files (`BL_M_NEG`) and creates `BL_M_NEG_WH`. Variables are prefixed `blmnwh`, and the average score is saved as `blmnwh.avg`.

```{r}
#| label: baseline-mom-negative
BL_M_NEG_WH <- process_warmth_files(
  path = PATH_CARMA_BASELINE_WARMTH,
  pattern = "BL_M_NEG",
  variable_prefix = "blmnwh",
  average_name = "blmnwh.avg"
)
```

## Baseline Partner Negative Warmth

This chunk imports all baseline Partner Negative Warmth files (`BL_P_NEG`) and creates `BL_P_NEG_WH`. Variables retain the existing `blfnwh` prefix, and the average score is saved as `blfnwh.avg`.

```{r}
#| label: baseline-partner-negative
BL_P_NEG_WH <- process_warmth_files(
  path = PATH_CARMA_BASELINE_WARMTH,
  pattern = "BL_P_NEG",
  variable_prefix = "blfnwh",
  average_name = "blfnwh.avg"
)
```

## Baseline Mom Positive Warmth

This chunk imports all baseline Mom Positive Warmth files (`BL_M_POS`) and creates `BL_M_POS_WH`. Variables are prefixed `blmpwh`, and the average score is saved as `blmpwh.avg`.

```{r}
#| label: baseline-mom-positive
BL_M_POS_WH <- process_warmth_files(
  path = PATH_CARMA_BASELINE_WARMTH,
  pattern = "BL_M_POS",
  variable_prefix = "blmpwh",
  average_name = "blmpwh.avg"
)
```

## Baseline Partner Positive Warmth

This chunk imports all baseline Partner Positive Warmth files (`BL_P_POS`) and creates `BL_P_POS_WH`. Variables retain the existing `blfpwh` prefix, and the average score is saved as `blfpwh.avg`.

```{r}
#| label: baseline-partner-positive
BL_P_POS_WH <- process_warmth_files(
  path = PATH_CARMA_BASELINE_WARMTH,
  pattern = "BL_P_POS",
  variable_prefix = "blfpwh",
  average_name = "blfpwh.avg"
)
```

# 4. Validate and Merge the Production Datasets

Before merging, each condition-specific R dataset is checked for duplicate IDs and invalid warmth values using the [validation function](#sec-validation-function) defined above. The four datasets are then merged by `ID` using full joins so that every couple appearing in at least one condition is retained.

```{r}
#| label: validate-and-merge-production
# Run data-quality checks on each condition-specific R dataset before merging.
validate_warmth_dataset(BL_M_NEG_WH, "BL_M_NEG_WH")
validate_warmth_dataset(BL_P_NEG_WH, "BL_P_NEG_WH")
validate_warmth_dataset(BL_M_POS_WH, "BL_M_POS_WH")
validate_warmth_dataset(BL_P_POS_WH, "BL_P_POS_WH")

# Merge all four baseline warmth datasets by couple ID.
BL_WH <- list(
  BL_M_NEG_WH,
  BL_P_NEG_WH,
  BL_M_POS_WH,
  BL_P_POS_WH
) %>%
  reduce(full_join, by = "ID") %>%
  arrange(ID)

# Run the final quality check on the complete merged production dataset.
validate_warmth_dataset(BL_WH, "BL_WH")
```

## Preview the final baseline warmth dataset

If the final checks pass, the merged production dataset is ready to be saved as **`BL_WH.sav`** in the source project (see [Setup](#sec-file-paths) for `PATH_OUTPUT_BL_WH`). This portfolio page does not write that file — it only previews the merged dataset so a reader can see the shape of the result without an export step running against Box.

```{r}
#| label: preview-merged-baseline-warmth
#| echo: false
#| eval: false

# Row-level preview suppressed for the public site since it would print
# real per-couple values even with the code hidden. Flip eval to true to
# preview locally.
knitr::kable(head(BL_WH, n = 5)[1:6], caption = "Merged Baseline Warmth Data (BL_WH)")
```

# 5. Prepare the Original-Coder Data for Reliability Analysis {#sec-original-coder-data}

The production datasets are intentionally wide because that is the useful structure for the final SPSS handoff. Reliability analysis is easier to audit in **long form**, where each row identifies one couple, one second, one condition, one target person, and one coder's warmth value.

The function below converts one already-processed production dataset (`BL_M_NEG_WH`, `BL_P_NEG_WH`, `BL_M_POS_WH`, `BL_P_POS_WH`, built above) back to long form for reliability matching. It uses the processed R data rather than re-importing the original CARMA files, so the reliability coder is compared directly against the values that entered the computed production datase.

```{r}
#| label: original-to-long-function
production_to_long <- function(dataset, variable_prefix, condition, person) {

  # Convert the second-by-second production variables back to one row per second.
  dataset %>%
    select(ID, starts_with(paste0(variable_prefix, "."))) %>%
    pivot_longer(
      cols = -ID,
      names_to = "variable",
      values_to = "warmth"
    ) %>%
    mutate(
      second = suppressWarnings(as.numeric(str_remove(variable, paste0("^", variable_prefix, "\\."))))
    ) %>%
    filter(!is.na(second)) %>%
    transmute(
      ID = ID,
      condition = condition,
      person = person,
      second = second,
      OG = warmth
    )
}
```

The four original-coded condition datasets are stacked into one long reference table. This table is temporary and is used only for matching the reliability files to the original coded values.

```{r}
#| label: original-reliability-reference
# Convert each processed production condition to long form and label its condition and target person.
ORIGINAL_WARMTH_LONG <- bind_rows(
  production_to_long(BL_M_NEG_WH, "blmnwh", "Negative", "Mom"),
  production_to_long(BL_P_NEG_WH, "blfnwh", "Negative", "Partner"),
  production_to_long(BL_M_POS_WH, "blmpwh", "Positive", "Mom"),
  production_to_long(BL_P_POS_WH, "blfpwh", "Positive", "Partner")
)
```

# 6. Import Reliability-Coded Warmth Files {#sec-import-reliability-files}

**Identify Mom versus Partner from reliability file names**

Reliability files need to be matched to the correct original Mom or Partner coding stream. Positive reliability files use names such as `124BL_M_POS_WH_RM.csv` and `133BL_P_POS_WH_RM.csv`. Negative reliability naming is less standardized, so this function recognizes several explicit Mom/Partner markers while avoiding assumptions based on fixed character positions.

The script stops and prints the unclassified file name if it cannot determine Mom versus Partner. That stop is intentional: guessing the target person would create an invalid reliability comparison.

```{r}
#| label: reliability-person-function
classify_reliability_person <- function(filename_full) {

  # Standardize the full path to uppercase for case-insensitive pattern matching.
  text <- str_to_upper(filename_full)

  # Identify explicit Mom/Mother or standalone M markers.
  is_mom <- str_detect(text, "MOM|MOTHER|(^|[/_ -])M([/_ -]|$)")

  # Identify explicit Partner/Father/Dad or standalone P markers.
  is_partner <- str_detect(text, "PARTNER|FATHER|DAD|(^|[/_ -])P([/_ -]|$)")

  # Stop if the file is ambiguous or cannot be classified.
  if (is_mom && is_partner) {
    stop("Reliability file contains both Mom and Partner markers: ", basename(filename_full))
  }

  if (!is_mom && !is_partner) {
    stop(
      "Could not determine Mom versus Partner from reliability file: ",
      basename(filename_full)
    )
  }

  if (is_mom) "Mom" else "Partner"
}
```

**Process a set of reliability files**

This function sends each selected reliability CSV through the same [`process_single_warmth_file()`](#sec-single-file-function) function used for the original CARMA files. It then labels the condition, target person, and reliability coder.

Before returning the data, it checks that there is only one reliability source file for each `ID × condition × person` combination.

```{r}
#| label: reliability-processing-function
process_reliability_files <- function(files_full, condition) {

  # Stop if no reliability CSV files were found for this condition.
  if (length(files_full) == 0) {
    stop("No ", condition, " reliability files were found.")
  }

  # Build a file-level index before processing the second-by-second data.
  file_index <- tibble(
    source_path = files_full,
    ID = map_chr(files_full, extract_couple_id),
    person = map_chr(files_full, classify_reliability_person),
    condition = condition
  )

  # Stop if more than one reliability file exists for the same video/coder target.
  duplicate_files <- file_index %>%
    count(ID, condition, person) %>%
    filter(n > 1)

  if (nrow(duplicate_files) > 0) {
    stop(
      "Duplicate reliability source files found for: ",
      paste(
        str_c(duplicate_files$ID, duplicate_files$condition, duplicate_files$person, sep = " / "),
        collapse = "; "
      )
    )
  }

  # Process every reliability file with the same CARMA cleaning rules as production files.
  map_dfr(seq_len(nrow(file_index)), function(i) {

    processed <- process_single_warmth_file(file_index$source_path[i])

    processed %>%
      transmute(
        ID = ID,
        condition = file_index$condition[i],
        person = file_index$person[i],
        second = second,
        RC = warmth,
        source_file = source_file
      )
  })
}
```

**Select and import Positive reliability files**

The Positive reliability folder is searched for CSV files. Files do not need to occupy fixed character positions; the three-digit ID is read from the beginning of the base file name, and Mom versus Partner is determined from the explicit `M` or `P` marker.

```{r}
#| label: import-positive-reliability
  # Confirm that the Positive reliability folder exists before searching it.
  if (!dir.exists(PATH_WARMTH_POS_RELIABILITY)) {
    stop("Positive reliability folder does not exist: ", PATH_WARMTH_POS_RELIABILITY)
  }


  # Find all Positive reliability CSV files in the folder.
  positive_reliability_files <- list.files(
    path = PATH_WARMTH_POS_RELIABILITY,
    pattern = "\\.csv$",
    full.names = TRUE,
    recursive = FALSE,
    ignore.case = TRUE
  )

  # Import and process all Positive reliability files.
  RELIABILITY_POSITIVE_LONG <- process_reliability_files(
    files_full = positive_reliability_files,
    condition = "Positive"
  )

```

**Select and import Negative reliability files**

The Positive reliability folder is searched for CSV files. Files do not need to occupy fixed character positions; the three-digit ID is read from the beginning of the base file name, and Mom versus Partner is determined from the explicit `M` or `P` marker.

```{r}
#| label: import-negative-reliability
  # Confirm that the Negative reliability folder exists before searching it.
    if (!dir.exists(PATH_WARMTH_NEG_RELIABILITY)) {
      stop("Negative reliability folder does not exist: ", PATH_WARMTH_NEG_RELIABILITY)
    }


    # Find all Negative reliability CSV files in the folder.
    negative_reliability_files <- list.files(
      path = PATH_WARMTH_NEG_RELIABILITY,
      pattern = "\\.csv$",
      full.names = TRUE,
      recursive = FALSE,
      ignore.case = TRUE
    )

    # Import and process all Negative reliability files.
    RELIABILITY_NEGATIVE_LONG <- process_reliability_files(
      files_full = negative_reliability_files,
      condition = "Negative"
    )
```

**Combine all reliability-coded observations**

The Positive RC and Negative RC observations are stacked into a single temporary long-form reliability dataset. One row represents one coded second from one reliability coder.

```{r}
#| label: combine-reliability
  # Combine Positive and Negative reliability observations into one long dataset.
  RELIABILITY_LONG <- bind_rows(
    RELIABILITY_POSITIVE_LONG,
    RELIABILITY_NEGATIVE_LONG
  )

  duplicate_reliability <- RELIABILITY_LONG %>%
    count(ID, condition, person, second) %>%
    filter(n > 1)
```

# 7. Match Reliability Codes to the Original Coder {#sec-match-reliability}

ICC should compare the **same video at the same second** for the [original](#sec-original-coder-data) and [reliability](#sec-import-reliability-files) coders. This section therefore matches data on `ID + condition + person + second` before calculating reliability.

The quality checks are restricted to videos that actually have a reliability file. The script stops if a selected reliability video is absent from the original processed data or if the original and reliability versions contain different sets of seconds. This prevents an ICC from being calculated on accidentally misaligned time windows.

```{r}
#| label: match-reliability
  # Identify the exact videos that have been selected for reliability analysis.
  reliability_video_keys <- RELIABILITY_LONG %>%
    distinct(ID, condition, person)

  # Restrict the original reference data to those same reliability-coded videos.
  ORIGINAL_FOR_RELIABILITY <- ORIGINAL_WARMTH_LONG %>%
    semi_join(
      reliability_video_keys,
      by = c("ID", "condition", "person")
    )

  # Identify reliability-coded seconds that have no matching original-coded second.
  reliability_without_original <- RELIABILITY_LONG %>%
    anti_join(
      ORIGINAL_FOR_RELIABILITY,
      by = c("ID", "condition", "person", "second")
    )

  # Identify original-coded seconds that are absent from the corresponding reliability file.
  original_without_reliability <- ORIGINAL_FOR_RELIABILITY %>%
    anti_join(
      RELIABILITY_LONG,
      by = c("ID", "condition", "person", "second")
    )

  # Stop rather than calculate ICC on mismatched second-by-second windows.
  if (nrow(reliability_without_original) > 0) {
    stop(
      "Reliability-coded observations were found without matching original ID/person/second values. ",
      "Review object: reliability_without_original"
    )
  }
  # Stop only if a reliability-coded observation cannot be matched to its
  # corresponding original-coded observation.
  reliability_without_original <- RELIABILITY_LONG %>%
    anti_join(
      ORIGINAL_WARMTH_LONG,
      by = c("ID", "condition", "person", "second")
    )

  if (nrow(reliability_without_original) > 0) {
    stop(
      "Reliability-coded observations were found without matching original ",
      "ID/condition/person/second values. Review object: reliability_without_original"
    )
  }

  # Keep only observations that were coded by both the original coder and
  # the reliability coder.
  RELIABILITY_MERGED <- ORIGINAL_WARMTH_LONG %>%
    inner_join(
      RELIABILITY_LONG,
      by = c("ID", "condition", "person", "second")
    )

  # Join the original and reliability values for exact ID/person/second comparisons.
  RELIABILITY_MATCHED <- ORIGINAL_FOR_RELIABILITY %>%
    inner_join(
      RELIABILITY_LONG,
      by = c("ID", "condition", "person", "second")
    ) %>%
    rename(warmth_original = OG, warmth_reliability = RC) %>%
    arrange(condition, person, ID, second)
```

**Review the reliability samples**

This table shows the number of reliability-coded videos, the number of matched seconds, and the number of seconds with nonmissing warmth values from **both** coders. ICC uses only rows where both coder values are observed.

```{r}
#| label: reliability-sample-review
# Summarize the reliability sample that will feed each ICC estimate.
RELIABILITY_SAMPLE_SUMMARY <- RELIABILITY_MATCHED %>%
  group_by(condition, person) %>%
  summarise(
    videos = n_distinct(ID),
    matched_seconds = n(),
    complete_pairs = sum(!is.na(warmth_original) & !is.na(warmth_reliability)),
    .groups = "drop"
  )

knitr::kable(RELIABILITY_SAMPLE_SUMMARY, caption = "Reliability Sample Summary")
```

# 8. Calculate Inter-Rater Reliability (Gwet's AC2)

The primary reliability statistic is **Gwet's AC2 with linear weighting**, calculated from the [matched](#sec-match-reliability) second-by-second warmth codes. Gwet's AC2 is a chance-corrected agreement statistic for ordinal ratings and is less sensitive than Cohen's kappa to highly concentrated or imbalanced category distributions [@gwet2008; @demirhan2023; @quarfoot2016].

Linear weighting is used because warmth is coded on an ordered 1–7 scale. Ratings that are closer together receive more agreement "credit" than ratings that are farther apart, and disagreements are penalized proportionally to the distance between ratings. This provides a more conservative weighting approach than quadratic weighting.

Gwet's AC2 is calculated separately for **Mom Positive, Partner Positive, Mom Negative, and Partner Negative**.

Rows where either coder has a missing warmth value are excluded from that condition's AC2 calculation. The output includes the AC2 estimate, weighted observed agreement, expected agreement, 95% confidence interval, *p*-value, number of videos, and number of complete second-by-second coder pairs used.

```{r}
#| label: ac2-function
calculate_warmth_ac2 <- function(data) {

  # Stop with a clear instruction if the irrCAC package is not installed.
  if (!requireNamespace("irrCAC", quietly = TRUE)) {
    stop(
      "The 'irrCAC' package is required for Gwet's AC2. ",
      "Install it once with install.packages('irrCAC'), ",
      "then rerun this section.")
  }

  # Keep only second-level observations with a nonmissing score from both coders.
  ac2_data <- data %>%
    filter(!is.na(warmth_original), !is.na(warmth_reliability)) %>%
    select(warmth_original, warmth_reliability)

  # Gwet's AC2 requires at least two complete observations.
  if (nrow(ac2_data) < 2) {
    return(
      tibble(
        AC2 = NA_real_,
        observed_agreement = NA_real_,
        expected_agreement = NA_real_,
        CI_lower = NA_real_,
        CI_upper = NA_real_,
        p_value = NA_real_,
        complete_pairs = nrow(ac2_data)) )
  }

  # Calculate linear-weighted Gwet's AC2 for the ordered 1-7 warmth scale.
  ac2_result <- irrCAC::gwet.ac1.raw(
    ac2_data,
    weights = "linear",
    categ.labels = 1:7)

  # Extract the AC2 results returned by irrCAC.
  ac2_estimate <- ac2_result$est

  # Return the main AC2 statistics in a compact one-row table.
  tibble(
    AC2 = ac2_estimate$coeff.val,
    observed_agreement = ac2_estimate$pa,
    expected_agreement = ac2_estimate$pe,
    CI_lower = as.numeric(
      sub(
        "^\\(([-0-9.]+),.*$",
        "\\1",
        ac2_estimate$conf.int)),
    CI_upper = as.numeric(
      sub(
        "^\\([-0-9.]+,([-0-9.]+)\\)$",
        "\\1",
        ac2_estimate$conf.int)),
    p_value = ac2_estimate$p.value,
    complete_pairs = nrow(ac2_data))
}
```

```{r}
#| label: calculate-ac2
# Calculate Gwet's AC2 separately for each condition and target person.
AC2_RESULTS <- RELIABILITY_MATCHED %>%
  group_by(condition, person) %>%
  group_modify(~ calculate_warmth_ac2(.x)) %>%
  ungroup() %>%
  left_join(
    RELIABILITY_MATCHED %>%
      group_by(condition, person) %>%
      summarise(
        videos = n_distinct(ID),
        .groups = "drop"),
    by = c("condition", "person")) %>%
  select(
    condition,
    person,
    videos,
    complete_pairs,
    AC2,
    observed_agreement,
    expected_agreement,
    CI_lower,
    CI_upper,
    p_valu)

knitr::kable(
  AC2_RESULTS,
  digits = 3,
  caption = "Inter-Rater Reliability (Linear-Weighted Gwet's AC2)")
```

# 9. Percent Agreement Reliability

As a complementary reliability metric to the [matched](#sec-match-reliability) ICC estimates above, this section calculates percent agreement: the share of seconds where the reliability coder's score falls within ±1 point of the original coder's score.

```{r}
#| label: percent-agreement
  # Calculate whether each second is within ±1 point between the original coder
  # and reliability coder.
  RELIABILITY_MERGED <- RELIABILITY_MERGED %>%
    mutate(
      abs_difference = abs(OG - RC),
      reliable = case_when(
        abs_difference <= 1 ~ 1,
        abs_difference >= 2 ~ 0,
        TRUE ~ NA_real_  ))

  # Calculate percent agreement separately for Mom Positive, Partner Positive,
  # Mom Negative, and Partner Negative.
  #
  # Percent agreement is the percentage of seconds for which the reliability
  # coder's score is within ±1 point of the original coder's score.

  PERCENT_AGREEMENT <- RELIABILITY_MERGED %>%
    filter(!is.na(reliable)) %>%
    group_by(condition) %>%
    summarise(
      reliable_seconds = sum(reliable),
      total_seconds = n(),
      percent_agreement = reliable_seconds / total_seconds * 100,
      .groups = "drop" ) %>%
    arrange(condition)

  knitr::kable(PERCENT_AGREEMENT, caption = "Percent Agreement (±1 point) by Condition")
```

# 10. Final Review

If the production section runs through the checks above without stopping, all four original condition datasets and the final merged `BL_WH` dataset passed the automated duplicate-ID and warmth-range checks. If the reliability sections above (5–9) also run through the ICC calculation, every selected reliability file was processed from second 6 using the same rules as the original CARMA files and matched exactly to the corresponding original `ID + condition + person + second` observations.

In the source project, only `BL_WH.sav` is written to Box. Reliability-coded observations and ICC results remain as R objects and are never exported.

```{r}
#| label: final-review
# Summarize the number of unique couple IDs in each production dataset.
PRODUCTION_DATASET_SUMMARY <- tibble(
  dataset = c(
    "BL_M_NEG_WH",
    "BL_P_NEG_WH",
    "BL_M_POS_WH",
    "BL_P_POS_WH",
    "BL_WH"),
  unique_IDs = c(
    n_distinct(BL_M_NEG_WH$ID),
    n_distinct(BL_P_NEG_WH$ID),
    n_distinct(BL_M_POS_WH$ID),
    n_distinct(BL_P_POS_WH$ID),
    n_distinct(BL_WH$ID) ))

knitr::kable(PRODUCTION_DATASET_SUMMARY, caption = "Unique Couple IDs by Dataset")
```
