---
title: "Resilience Analysis: Structural Disadvantage and Young-Adult Outcomes"
subtitle: "Reproducible secondary analysis of the National Longitudinal Study of Adolescent to Adult Health (Add Health)"
author: "Rachel C. Marcus, Ph.D."
bibliography: references.bib
format:
  html:
    code-fold: true
    code-summary: "Show code"
    toc: true
    toc-depth: 3
editor:
  markdown:
    wrap: 72
---

# Project Overview

This quantitative work sample demonstrates an end-to-end secondary-data
analysis using the National Longitudinal Study of Adolescent to Adult
Health (Add Health) [@harris2013]. The analysis was originally developed
for the 21st Annual Sloboda and Bukoski SPR Cup and was selected as one
of four national finalist projects presented at the Society for
Prevention Research 34th Annual Meeting [@marcus2026].

::: {.callout-note appearance="simple" icon="false"}
### Project at a Glance

**Research question:** Which individual, family, school, and community
factors are associated with better-than-expected young-adult outcomes
given structural disadvantage in adolescence?

**Data:** Add Health Waves I and III; analytic *N* = 4,882.

**Methods:** Secondary-data cleaning and harmonization; complex
survey-weighted regression; outcome-specific beta-weighted structural
disadvantage composites; residual-based continuous resilience scores;
multivariable regression; exploratory subgroup comparisons and formal
interaction tests.

**Tools:** R, SPSS, Quarto, Git/GitHub.

**My role:** Led the analytic design, operationalization, statistical
analysis, interpretation, visualization, and reproducible documentation
as part of a multidisciplinary research team.
:::

The analysis proceeds in five steps:

[Step 0](#sec-step0): Wave I structural disadvantage indicators are used to build
**outcome-specific disadvantage composites** [@marquez2023].

[Step 1](#sec-step1): Wave III outcomes are **regressed on the corresponding
disadvantage composites** [@miller-lewis2013].

[Step 2](#sec-step2): Standardized residuals from those models become **continuous
resilience scores**: the degree to which a young adult is doing better
or worse than expected given their adolescent disadvantage exposure.

[Step 3](#sec-step3): Multilevel resilience factors (individual, family, school,
and community) are tested as **predictors of resilience** in the full
sample.

[Step 4](#sec-step4): Exploratory low- and high-disadvantage subgroup estimates
are paired with **formal interaction tests** to evaluate whether
associations differ as disadvantage increases.

## Reproducible Analysis Environment

All primary models use survey-weighted regression (`survey`) to account
for Add Health's sampling weights. `tidyverse`, `broom`, `gtsummary`,
`flextable`, and `gt` support data preparation, model extraction, and
publication-ready reporting. Code is collapsed by default so the page
can be read as a research report while retaining the full reproducible
workflow.

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(gtsummary)   # Regression summary tables
library(flextable)   # Table formatting and export
library(haven)       # Read SPSS .sav files
library(tidyverse)   # Data wrangling and iteration
library(survey)      # Complex survey / weighted regression
library(broom)       # Tidy model output
library(gt)          # Formatted tables
```

The analytic dataset is the merged Add Health public-use survey data
(`AddHealth_AllData.sav`), imported with `haven::read_sav()`. The local
file path is intentionally hidden from the rendered work sample.

```{r}
#| label: load-data
#| include: false
# Import the cleaned SPSS analytic file into R while preserving variable labels and SPSS metadata
dat <- read_sav("C:/Users/rcm5550/Desktop/PSU OD Copy/Rachel - The Pennsylvania State University/SPR Cup 2026/Analysis/AddHealth_AllData.sav")
```

```{r}
#| label: save-tables
#| include: false
# Create `fig_dir` for use in the subsequent analysis/reporting steps
fig_dir <- "../../figures/spr-cup/"
if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
```

## Data Preparation and Quality Assurance

Data preparation was completed in SPSS before import into R. Four
domain-specific syntax files covered demographics, structural
disadvantage, young-adult outcomes, and hypothesized resilience factors.
The cleaning workflow was intentionally documented before modeling so
that transformations could be audited independently of the R analysis.

Across the four domains, preparation included:

- **Demographics:** recoding and QA for biological sex, age, and
  race/ethnicity variables used as covariates or structural-disadvantage
  indicators.
- **Structural disadvantage:** construction of minority-status, school,
  economic-hardship, family socioeconomic, subjective-neighborhood, and
  census-neighborhood measures, including direction harmonization and
  standardization where needed.
- **Young-adult outcomes:** construction of Wave III depression,
  substance-use, educational-attainment, and general-health measures.
- **Resilience factors:** construction of community connectedness,
  community integration, optimism/hopefulness, self-esteem,
  self-efficacy, parent-child relationship, and school connectedness.
- **Quality assurance:** inspection of source-variable frequencies,
  explicit data cleaning of special missing-value codes, reverse
  scoring, dummy coding, scale/composite construction, reliability
  checks, post-transformation frequency checks, and standardization
  within the valid weighted analytic sample.

The excerpts below show representative data cleaning code to make the
preprocessing workflow visible while keeping the page readable.

### Explicit Data Cleaning and Missing-Data Handling Example

Across the SPSS preprocessing files, source-variable frequencies were
inspected before recoding so that valid responses, study-specific
special codes, and legitimate skips could be handled explicitly rather
than treated as equivalent values.

Demographic covariates, including biological sex, age, and
race/ethnicity, were cleaned separately in `Demographics.sps` using the
same pattern of source-frequency checks, explicit recoding, variable
labels, and post-transformation validation.

The parental-employment component of family socioeconomic disadvantage
provides an example of this process.

For both the responding parent and their partner, the source variables
distinguish full-time employment (`1`), not employed full time (`0`),
refusal (`6`), and a legitimate skip for respondents who do not work
outside the home (`7`). Refusals are converted to system missing,
whereas legitimate skips are coded as not employed full time because
that response carries substantive information for the employment
indicator. Frequencies are checked before and after the transformation.
The two cleaned indicators are then averaged across available parents,
allowing the composite to retain information when employment data are
available for only one parent.

``` spss
*****************************PARENT EMPLOYMENT - FULL TIME.
*Q: Are you employed full time? (PA17).
*Inspect the original response distribution before recoding.
frequencies variables = PA17 /statistics = all.
*Answer codes: 0 = no, 1 = yes, 6 = refused,
*7 = legitimate skip [does not work outside home].

*Recode valid yes/no responses to a binary full-time-employment indicator.
if (PA17 eq 1) parentfulltimeD = 1.
if (PA17 eq 0) parentfulltimeD = 0.

*Treat refusal as missing rather than assigning it a substantive value.
if (PA17 eq 6) parentfulltimeD = -99.

*Code the legitimate skip as not employed full time because the skip
*indicates that the respondent does not work outside the home.
if (PA17 eq 7) parentfulltimeD = 0.

*Convert the temporary missing-data code to SPSS system missing.
recode parentfulltimeD (-99 = sysmis).

variable labels parentfulltimeD
    "Parent employed full time (PA17, 1=yes)".
value labels parentfulltimeD
    0 "not employed full time"
    1 "employed full time".
execute.

*Verify the distribution after recoding.
frequencies variables = parentfulltimeD /statistics = all.


*****************************PARTNER EMPLOYMENT - FULL TIME.
*Q: Is partner employed full time? (PB13).
*Inspect the original partner-employment response distribution.
frequencies variables = PB13 /statistics = all.
*Answer codes: 0 = no, 1 = yes, 6 = refused,
*7 = legitimate skip [does not work outside home].

*Apply the same coding rules to the partner variable.
if (PB13 eq 1) partnerfulltimeD = 1.
if (PB13 eq 0) partnerfulltimeD = 0.
if (PB13 eq 6) partnerfulltimeD = -99.
if (PB13 eq 7) partnerfulltimeD = 0.

*Convert refusal to SPSS system missing.
recode partnerfulltimeD (-99 = sysmis).

variable labels partnerfulltimeD
    "Partner employed full time (PB13, 1=yes)".
value labels partnerfulltimeD
    0 "not employed full time"
    1 "employed full time".
execute.

*Verify the cleaned partner-employment distribution.
frequencies variables = partnerfulltimeD /statistics = all.


*****************************PARENTAL EMPLOYMENT COMPOSITE.
*Average full-time employment across available parent reports.
*SPSS MEAN() uses the nonmissing component when only one is available.
compute parentemploy = mean(parentfulltimeD, partnerfulltimeD).

variable labels parentemploy
    "Average parental employment 0-1 scale (mean of available parents, higher = employed)".
execute.

*Check the final composite before it is incorporated into family SES.
frequencies variables = parentemploy /statistics = all.
```

### Structural Disadvantage

`Disadvantage.sps` constructs the six Wave I disadvantage domains used
in the outcome-specific cumulative-disadvantage composites. The family
socioeconomic-disadvantage example illustrates the broader workflow:
standardize component measures within the weighted analytic sample,
orient them to a common direction, combine them, and then verify the
resulting distribution and internal consistency.

``` spss
*****************************FAMILY SOCIOECONOMIC DISADVANTAGE COMPOSITE.
*Components: famincomecat (continuous $, higher = more income),
*pareduc_avg (0-8, higher = more education),
*parentemploy (0-1, higher = more employment).
*Standardizing each component to z-score then reverse code so higher is
*more disadvantage, then take the average of all variables.

*standardize based on weight filtered dataset.
delete variables Zfamincomecat Zpareduc_avg Zparentemploy.
    *delete variable if previously standardized with whole sample.
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0).
descriptives variables = famincomecat pareduc_avg parentemploy /save.
execute.

compute Zfamincomecat_r = Zfamincomecat * -1.
variable labels Zfamincomecat_r
    "Family income z-scored reversed (higher = lower income = more disadvantage)".
execute.

compute Zpareduc_avg_r = Zpareduc_avg * -1.
variable labels Zpareduc_avg_r
    "Average parental education z-scored reversed (higher = less education = more disadvantage)".
execute.

compute Zparentemploy_r = Zparentemploy * -1.
variable labels Zparentemploy_r
    "Average parental employment z-scored reversed (higher = less employed = more disadvantage)".
execute.

compute famSES_disadv =
    mean(Zfamincomecat_r, Zpareduc_avg_r, Zparentemploy_r).
variable labels famSES_disadv
    "Family SES disadvantage composite: mean of z-scored reversed income, education, employment (higher = more disadvantage)".
execute.

frequencies variables = famSES_disadv /statistics = all.

reliability variables = famincomecat pareduc_avg parentemploy
    /scale('famSES') all /statistics = all.

*standardize based on weight filtered dataset.
temporary.
select if (GSWGT3_2 ge 0).
descriptives variables = famSES_disadv /save.
execute.
```

### Young-Adult Outcomes

`Outcomes.sps` constructs the four Wave III outcomes used in the
resilience models: depression, substance use, educational attainment,
and general health. Substance use combines cleaned alcohol, cigarette,
and marijuana-use indicators; the remaining outcomes are similarly
recoded and checked before standardization.

Depressive symptoms provide a useful example of the full
item-to-analysis-variable workflow. The Center for Epidemiologic Studies
Depression Scale (CES-D) was originally developed as a self-report
measure of depressive symptomatology in the general population
[@radloff1977]. The project uses the modified nine-item CES-D available
in Add Health Wave III, consistent with prior Add Health work using the
CES-D-9 [@desch2023].

The source items use a 0–3 response scale. Special response codes (6 =
refused, 8 = don't know, 9 = not applicable) are converted to system
missing. Two positively worded items are reverse scored so that all nine
items point in the same direction before the 0–27 sum score is created.

``` spss
*****************************CES-D DEPRESSION SCALE (WAVE 3).
*Items H3SP5-H3SP13: 9-item CES-D scale from Wave 3 Section 12.
*Answer codes all items: 0=never or rarely, 1=sometimes,
*2=a lot of the time, 3=most/all of the time,
*6=refused, 8=don't know, 9=not applicable.
*H3SP7 and H3SP11 are positive items - reverse before summing.

***check source frequencies.
frequencies variables =
    H3SP5 H3SP6 H3SP7 H3SP8 H3SP9 H3SP10 H3SP11 H3SP12 H3SP13
    /statistics = all.

*Negative items: pass through directly (higher = more depressive symptoms).
if (H3SP5 ge 0 and H3SP5 le 3) dep_bothered = H3SP5.
if (H3SP5 eq 6 or H3SP5 eq 8 or H3SP5 eq 9) dep_bothered = -99.
recode dep_bothered (-99 = sysmis).
variable labels dep_bothered
    "Bothered by things (H3SP5, 0-3, higher = more depressive symptoms)".
execute.

if (H3SP6 ge 0 and H3SP6 le 3) dep_blues = H3SP6.
if (H3SP6 eq 6 or H3SP6 eq 8 or H3SP6 eq 9) dep_blues = -99.
recode dep_blues (-99 = sysmis).
variable labels dep_blues
    "Could not shake off blues (H3SP6, 0-3, higher = more depressive symptoms)".
execute.

*H3SP7: positive item - reverse before summing.
if (H3SP7 ge 0 and H3SP7 le 3) dep_asgood = (3 - H3SP7).
if (H3SP7 eq 6 or H3SP7 eq 8 or H3SP7 eq 9) dep_asgood = -99.
recode dep_asgood (-99 = sysmis).
variable labels dep_asgood
    "Did NOT feel as good as others (H3SP7 reversed, higher = more depressive symptoms)".
execute.

if (H3SP8 ge 0 and H3SP8 le 3) dep_concentrate = H3SP8.
if (H3SP8 eq 6 or H3SP8 eq 8 or H3SP8 eq 9) dep_concentrate = -99.
recode dep_concentrate (-99 = sysmis).
variable labels dep_concentrate
    "Trouble concentrating (H3SP8, 0-3, higher = more depressive symptoms)".
execute.

if (H3SP9 ge 0 and H3SP9 le 3) dep_depressed = H3SP9.
if (H3SP9 eq 6 or H3SP9 eq 8 or H3SP9 eq 9) dep_depressed = -99.
recode dep_depressed (-99 = sysmis).
variable labels dep_depressed
    "Felt depressed (H3SP9, 0-3, higher = more depressive symptoms)".
execute.

if (H3SP10 ge 0 and H3SP10 le 3) dep_tired = H3SP10.
if (H3SP10 eq 6 or H3SP10 eq 8 or H3SP10 eq 9) dep_tired = -99.
recode dep_tired (-99 = sysmis).
variable labels dep_tired
    "Too tired to do things (H3SP10, 0-3, higher = more depressive symptoms)".
execute.

*H3SP11: positive item - reverse before summing.
if (H3SP11 ge 0 and H3SP11 le 3) dep_enjoyed = (3 - H3SP11).
if (H3SP11 eq 6 or H3SP11 eq 8 or H3SP11 eq 9) dep_enjoyed = -99.
recode dep_enjoyed (-99 = sysmis).
variable labels dep_enjoyed
    "Did NOT enjoy life (H3SP11 reversed, higher = more depressive symptoms)".
execute.

if (H3SP12 ge 0 and H3SP12 le 3) dep_sad = H3SP12.
if (H3SP12 eq 6 or H3SP12 eq 8 or H3SP12 eq 9) dep_sad = -99.
recode dep_sad (-99 = sysmis).
variable labels dep_sad
    "Felt sad (H3SP12, 0-3, higher = more depressive symptoms)".
execute.

if (H3SP13 ge 0 and H3SP13 le 3) dep_disliked = H3SP13.
if (H3SP13 eq 6 or H3SP13 eq 8 or H3SP13 eq 9) dep_disliked = -99.
recode dep_disliked (-99 = sysmis).
variable labels dep_disliked
    "Felt people disliked them (H3SP13, 0-3, higher = more depressive symptoms)".
execute.

***check cleaned-item frequencies.
frequencies variables =
    dep_bothered dep_blues dep_asgood dep_concentrate dep_depressed
    dep_tired dep_enjoyed dep_sad dep_disliked
    /statistics = all.

*****************************CES-D DEPRESSION COMPOSITE (WAVE 3).
*9 items, all oriented so higher = more depressive symptoms.
*Sum score range = 0-27; scoring follows Desch et al. (2023).

compute cesD_sum =
    sum(dep_bothered, dep_blues, dep_asgood, dep_concentrate,
        dep_depressed, dep_tired, dep_enjoyed, dep_sad, dep_disliked).
variable labels cesD_sum
    "CES-D depression sum score Wave 3 (0-27, higher = more depressive symptoms)".
execute.

*Internal-consistency check.
reliability variables =
    dep_bothered dep_blues dep_asgood dep_concentrate dep_depressed
    dep_tired dep_enjoyed dep_sad dep_disliked
    /scale('CESD') all /statistics = all.

*Check final composite distribution.
frequencies variables = cesD_sum /statistics = all.

*Standardize based on the valid Wave III weighted analytic sample.
temporary.
select if (GSWGT3_2 ge 0).
descriptives variables = cesD_sum /save.
```

This sequence demonstrates the workflow from **data cleaning** to
**scale scoring**: invalid response codes are handled before the scale
is constructed, positive items are reverse scored, the cleaned items are
inspected again, the composite is validated with a reliability check,
and then the outcome is standardized for the analytic dataset.

### Resilience Factors

`Protective.sps` constructs the individual, family, school, and
community factors evaluated as predictors of residual resilience. The
community-connectedness excerpt illustrates item-level recoding,
composite construction, reliability assessment, and weighted-sample
standardization.

``` spss
***************************COMMUNITY RESOURCES - CONNECTEDNESS (ADOLESCENT REPORT).

*Example item-level cleaning.
if (H1NB1 eq 1) knowneighbR = 0.
if (H1NB1 eq 2) knowneighbR = 1.
if (H1NB1 eq 6 or H1NB1 eq 8 or H1NB1 eq 9) knowneighbR = -99.
recode knowneighbR (-99 = sysmis).
variable labels knowneighbR
    "Does NOT know most neighbors (H1NB1 reversed, 1=low connectedness/more disadvantage)".
value labels knowneighbR
    0 "knows most neighbors"
    1 "does not know most neighbors".
execute.

***************************COMMUNITY CONNECTEDNESS INDEX
*(Wickrama & Bryant 2003).
*Items coded in protective direction (1 = connected):
*knowneighb, talkneighb, neighbwatch, nbhdsafe.
*Sum 0-4; higher = more connectedness = more protective.

compute commconnect =
    sum(knowneighb, talkneighb, neighbwatch, nbhdsafe).
variable labels commconnect
    "Community connectedness (sum 0-4, higher = more connectedness)".
execute.

reliability variables =
    knowneighb talkneighb neighbwatch nbhdsafe
    /scale('commconnect') all /statistics = all.

frequencies variables = commconnect /statistics = all.

*standardize based on weight filtered dataset.
delete variables Zcommconnect.
temporary.
select if (GSWGT3_2 ge 0).
descriptives variables = commconnect /save.
```

After domain-specific cleaning and QA, the analysis ready variables were
merged into `AddHealth_AllData.sav` and imported into R.

## Complex Survey Design and Analytic Sample

Add Health uses a complex sampling design, and the Wave III grand sample
weight adjusts for unequal selection probabilities and nonresponse. For
this analysis, respondents without a valid Wave III weight were excluded
before model estimation, leaving the analytic sample used throughout the
work sample.

**Check and Restrict to Respondents with a Valid Wave III Weight**

```{r}
#| label: check missing sample weight

# How many cases missing
cat("Missing GSWGT3_2:", sum(is.na(dat$GSWGT3_2)), "\n")
# Print this diagnostic/result so the rendered analysis documents the value produced at this stage
cat("Non-missing GSWGT3_2:", sum(!is.na(dat$GSWGT3_2)), "\n")
```

**Create the Analytic Dataset**

```{r}
#| label: filter for analytic dataset
# ── Subset to respondents with valid Wave 3 weight ────────────────────────────
dat_w3 <- dat |> filter(!is.na(GSWGT3_2)) # Dat_w3 is name of filtered dataset 

# Print this diagnostic/result so the rendered analysis documents the value produced at this stage
cat("Analytic n after excluding missing weights:", nrow(dat_w3), "\n")
```

**Specify the Survey-Weighted Design Object**

```{r}
#| label: weighted dataset
# Wave III grand sample weight accounts for unequal selection probability
# And nonresponse. ids = ~1 reflects the weighting specification used in the
# Original SPR Cup analysis

# Define the survey-design object that carries the Wave III sampling weights into subsequent survey-weighted estimates
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)

# Print this diagnostic/result so the rendered analysis documents the value produced at this stage
cat("Sum of weights:", round(sum(dat_w3$GSWGT3_2), 0), "\n")
```

## Descriptive Statistics

Table @tbl-descriptives summarizes the disadvantage indicators,
resilience factors, and young-adult outcomes. Means and standard
deviations are survey-weighted; ranges are shown on the observed
unweighted scale.

```{r}
#| label: create descriptives table
# ── Variable lists with display labels ────────────────────────────────────────
disadvantage_vars <- c(
  "minorityD"          = "Minority Status",
  "school_disadv"     = "School Disadvantage",
  "econhardship"      = "Economic Hardship",
  "famSES_disadv"      = "Family SES Disadvantage",
  "nbhd_subj_disadv"  = "Neighborhood Disadvantage (Subjective)",
  "cens_nbhd_disadv"  = "Neighborhood Disadvantage (Census)")

# Store the variable names/labels in `protective_vars` so later code can iterate over them programmatically
protective_vars <- c(
  "commconnect"       = "Community Connectedness",
  "commintegr_disadv" = "Community Integration",
  "opt_hopeful"       = "Optimism / Hopefulness",
  "selfesteem"        = "Self-Esteem",
  "self_efficacy"     = "Self-Efficacy",
  "parentchild_rel"   = "Parent-Child Relationship",
  "schoolconnect"     = "School Connectedness")

# Store the variable names/labels in `outcome_vars` so later code can iterate over them programmatically
outcome_vars <- c(
  "cesD_sum"    = "Depression Symptoms",
  "subuse_W3"   = "Substance Use",
  "edattain_W3" = "Education Attainment",
  "genhealth_W3"= "General Health")

# ── Function: get weighted M and SD, and unweighted range ─────────────────────
get_desc <- function(var, design) {
  mn  <- as.numeric(svymean(as.formula(paste0("~", var)), design, na.rm = TRUE))
  vr  <- as.numeric(svyvar(as.formula(paste0("~", var)),  design, na.rm = TRUE))
  rng <- range(design$variables[[var]], na.rm = TRUE)
  data.frame(
    M     = formatC(mn,       digits = 2, format = "f"),
    SD    = formatC(sqrt(vr), digits = 2, format = "f"),
    Range = paste0(formatC(rng[1], digits = 2, format = "f"),
                   "–",
                   formatC(rng[2], digits = 2, format = "f")))
}

# ── Build rows for each variable group ────────────────────────────────────────
make_desc_rows <- function(var_named_vec, design, section_label) {
  rows <- lapply(names(var_named_vec), function(v) {
    d <- get_desc(v, design)
    data.frame(Variable = var_named_vec[[v]], M = d$M, SD = d$SD, Range = d$Range)
  })
  rows <- do.call(rbind, rows)
  rbind(data.frame(Variable = section_label, M = "", SD = "", Range = ""), rows)
}

# Create `desc_tbl` for use in the subsequent analysis/reporting steps
desc_tbl <- rbind(
  make_desc_rows(disadvantage_vars,  design_w3, "Disadvantage Predictors"),
  make_desc_rows(protective_vars, design_w3, "Protective Factors"),
  make_desc_rows(outcome_vars,    design_w3, "Outcomes"))

# ── Build flextable ───────────────────────────────────────────────────────────
header_rows <- which(desc_tbl$M == "")

# Build `descriptives` through a reproducible sequence of data-transformation steps
descriptives <- flextable(desc_tbl) |>
  set_header_labels(Variable = "Variable", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Weighted Descriptive Statistics"),
    colwidths = c(1, 3)                                        # Was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  bold(i = header_rows, part = "body") |>
  width(j = 1, width = 3.0) |>
  width(j = 2:4, width = 1.0) |>                              # Was 2:3
  align(j = 2:4, align = "center", part = "all") |>           # Was 2:3
  padding(i = setdiff(seq_len(nrow(desc_tbl)), header_rows),
          j = 1, padding.left = 20, part = "body") |>
  add_footer_lines("Note. Means and standard deviations are weighted using the Wave 3 grand sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-descriptives
#| echo: false
descriptives
```

The analytic sample reports moderate average levels of structural
disadvantage (e.g., *M* = 0.52 school disadvantage on a 0–3 scale)
alongside relatively high protective-factor scores (e.g., *M* = 4.15
parent-child relationship on a 1–5 scale), consistent with a normative,
not exclusively high-risk, longitudinal sample.

```{r}
#| label: save-descriptives-png
#| include: false

# Build `descriptives_png` through a reproducible sequence of data-transformation steps
descriptives_png <- descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 3.2) |>
  width(j = 2:4, width = 1.05)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(descriptives_png, path = paste0(fig_dir, "descriptives.png"), zoom = 3, expand = 0)
```

# Step 0: Build Outcome-Specific Disadvantage Composites {#sec-step0}

## Analytic Rationale

A simple cumulative-risk index would assign equal weight to each
disadvantage indicator. Here, the six indicators showed meaningfully
different associations with depression, substance use, educational
attainment, and general health. I therefore used **outcome-specific
beta-weighted composites**, allowing each indicator's contribution to
reflect its empirical association with the outcome being modeled.

Race/ethnicity is included here as a marker of exposure to structural
inequity within the original SPR Cup operationalization, not as an
individual deficit or intrinsic risk characteristic.

For each outcome, the same survey-weighted regression is fit separately
for minority status, school disadvantage, economic hardship, family
socioeconomic disadvantage, subjective neighborhood disadvantage, and
census-based neighborhood disadvantage. Rather than repeat 24 nearly
identical code blocks, the workflow is parameterized and iterated below.

```{r}
#| label: individual-disadvantage-regressions

# Create `disadvantage_specs` for use in the subsequent analysis/reporting steps
disadvantage_specs <- tribble(
  ~predictor,             ~predictor_label,
  "minorityD",            "Minority Status",
  "Zschool_disadv",       "School Disadvantage",
  "Zeconhardship",        "Economic Hardship",
  "famSES_disadv",        "Family SES",
  "Znbhd_subj_disadv",    "Neighborhood (Subjective)",
  "Zcens_nbhd_disadv",    "Neighborhood (Census)"
)

# Create `outcome_specs` for use in the subsequent analysis/reporting steps
outcome_specs <- tribble(
  ~outcome,          ~outcome_label,
  "ZcesD_sum",       "Depression",
  "Zsubuse_W3",      "Substance Use",
  "Zedattain_W3",    "Education",
  "Zgenhealth_W3",   "Health"
)

# Define `fit_disadvantage_model()` once so the same operation can be applied consistently across variables/models without duplicating code
fit_disadvantage_model <- function(outcome, predictor, design) {
  svyglm(
    reformulate(predictor, response = outcome),
    design = design,
    family = gaussian()
  )
}

# Generate every required outcome-by-predictor combination before fitting the repeated models programmatically
weight_results <- crossing(outcome_specs, disadvantage_specs) |>
  mutate(
    model = map2(
      outcome, predictor,
      ~fit_disadvantage_model(.x, .y, design_w3)
    ),
    b = map2_dbl(
      model, predictor,
      ~unname(coef(.x)[.y])
    ),
    p = map2_dbl(
      model, predictor,
      ~{
        coef_table <- summary(.x)$coefficients
        unname(coef_table[.y, "Pr(>|t|)"])
      }
    )
  )
```

This structure makes the repeated analysis auditable while reducing
copy-paste risk: the model specification is defined once and applied
consistently across all 24 predictor-outcome combinations.

## Disadvantage Weights

Table @tbl-disadvantage-weight-table displays the standardized
coefficient and *p*-value from each individual survey-weighted
regression. These coefficients become the empirical weights used in the
outcome-specific structural disadvantage composites.

```{r}
#| label: create-disadvantage-weight-table

# Build `weight_summary` through a reproducible sequence of data-transformation steps
weight_summary <- weight_results |>
  select(Predictor = predictor_label, Outcome = outcome_label, b, p) |>
  mutate(
    b = round(b, 2),
    p = case_when(
      p < .001 ~ "<.001",
      TRUE ~ sprintf("%.3f", p)
    )
  ) |>
  pivot_wider(
    names_from = Outcome,
    values_from = c(b, p),
    names_glue = "{Outcome}_{.value}"
  ) |>
  select(
    Predictor,
    Depression_b, Depression_p,
    `Substance Use_b`, `Substance Use_p`,
    Education_b, Education_p,
    Health_b, Health_p
  )

# Build `table_weights` through a reproducible sequence of data-transformation steps
table_weights <- flextable(weight_summary) |>
  set_header_labels(
    Predictor = "Predictor",
    Depression_b = "b", Depression_p = "p",
    `Substance Use_b` = "b", `Substance Use_p` = "p",
    Education_b = "b", Education_p = "p",
    Health_b = "b", Health_p = "p"
  ) |>
  add_header_row(
    values = c("", "Depression", "Substance Use", "Education", "Health"),
    colwidths = c(1, 2, 2, 2, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 1.5) |>
  width(j = 2:9, width = .6) |>
  align(j = 2:9, align = "center", part = "all") |>
  add_footer_lines(
    "Note. Continuous predictors and outcomes were standardized. Each disadvantage indicator was modeled separately for each outcome."
  ) |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-disadvantage-weight-table
#| echo: false
table_weights
```

Family socioeconomic disadvantage shows the largest absolute association
across the four outcomes. The magnitude and statistical strength of the
remaining indicator-outcome associations vary across outcomes, which is
the rationale for retaining outcome-specific weights rather than
assigning each indicator equal influence.

```{r}
#| label: save-disadvantage-weight-table-png
#| include: false
# Build `table_weights_png` through a reproducible sequence of data-transformation steps
table_weights_png <- table_weights |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 1.6) |>
  width(j = 2:9, width = 0.6)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(
  table_weights_png,
  path = paste0(fig_dir, "table_weights.png"),
  zoom = 3, expand = 0
)
```

## Structural Disadvantage Composites {#cumulative-adversity-composites}

For each outcome, the structural disadvantage composite is the weighted
sum of the six indicators, using the standardized coefficient from Step
0 as the weight. This approach retains the same conceptual domains
across outcomes while allowing their empirical contribution to differ.

To make interpretation consistent across outcomes, composites tied to
positively coded outcomes (educational attainment and general health)
are multiplied by `-1` after weighting. **Higher values therefore
indicate greater structural disadvantage for all four composites.**

::: {.callout-important appearance="minimal" icon="false"}
The weighting strategy is deliberately outcome-specific. A disadvantage
indicator with a stronger relationship to depression, for example,
contributes more to the depression composite than to an outcome for
which its association is weaker.
:::

::: {.callout-note appearance="simple" icon="false"}
**Interpretive note.** The beta weights are estimated in the same
analytic sample in which the composites are subsequently used. This
makes the approach useful for the present analysis, but the magnitude of
the weighted composite associations should not be treated as
out-of-sample predictive performance. A future extension could
cross-validate the weights or evaluate them in an independent sample.
:::

```{r}
#| label: cumulative-composites

# Define `get_weight_vector()` once so the same operation can be applied consistently across variables/models without duplicating code
get_weight_vector <- function(outcome_name) {
  weight_results |>
    filter(outcome == outcome_name) |>
    select(predictor, b) |>
    deframe()
}

# Create `w_dep` for use in the subsequent analysis/reporting steps
w_dep <- get_weight_vector("ZcesD_sum")
# Create `w_sub` for use in the subsequent analysis/reporting steps
w_sub <- get_weight_vector("Zsubuse_W3")
# Create `w_edu` for use in the subsequent analysis/reporting steps
w_edu <- get_weight_vector("Zedattain_W3")
# Create `w_hlt` for use in the subsequent analysis/reporting steps
w_hlt <- get_weight_vector("Zgenhealth_W3")

# Define `weighted_composite()` once so the same operation can be applied consistently across variables/models without duplicating code
weighted_composite <- function(data, weights) {
  x <- as.matrix(data[, names(weights), drop = FALSE])
  as.numeric(x %*% weights)
}

# Build `dat_w3` through a reproducible sequence of data-transformation steps
dat_w3 <- dat_w3 |>
  mutate(
    cumadv_depression = weighted_composite(cur_data(), w_dep),
    cumadv_substance  = weighted_composite(cur_data(), w_sub),
    cumadv_education  = -weighted_composite(cur_data(), w_edu),
    cumadv_health     = -weighted_composite(cur_data(), w_hlt),
    Zcumadv_depression = as.numeric(scale(cumadv_depression)),
    Zcumadv_substance  = as.numeric(scale(cumadv_substance)),
    Zcumadv_education  = as.numeric(scale(cumadv_education)),
    Zcumadv_health     = as.numeric(scale(cumadv_health))
  )

# Define the survey-design object that carries the Wave III sampling weights into subsequent survey-weighted estimates
design_w3 <- svydesign(
  ids = ~1,
  weights = ~GSWGT3_2,
  data = dat_w3
)
```

```{r}
#| label: create cumulative disadvantage descriptives
# Create table labels for each outcome
cumadv_vars <- c(
  "cumadv_depression" = "Depression",
  "cumadv_substance"  = "Substance Use",
  "cumadv_education"  = "Education Attainment",
  "cumadv_health"     = "General Health")

# Create function for disadvantage composite descriptives
cumadv_rows <- do.call(rbind, lapply(names(cumadv_vars), function(v) {
  d <- get_desc(v, design_w3)
  data.frame(Variable = cumadv_vars[[v]], M = d$M, SD = d$SD, Range = d$Range)
}))

# Create table for disadvantage composite descriptives
cumulative_descriptives <- flextable(cumadv_rows) |>
  set_header_labels(Variable = "Outcome", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Cumulative Disadvantage Composites"),
    colwidths = c(1, 3)                              # Was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 1.0) |>                   
  align(j = 2:4, align = "center", part = "all") |> 
  add_footer_lines("Note. Each structural disadvantage composite is outcome-specific, weighted by the association of each disadvantage indicator with that outcome. Descriptive statistics are survey-weighted using the Wave 3 sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer") |>
  fit_to_width(max_width = 6.5)
```

```{r}
#| label: tbl-cumulative
#| echo: false

cumulative_descriptives
```

```{r}
#| label: save-cumulative-png
#| include: false

# Build `cumulative_descriptives_png` through a reproducible sequence of data-transformation steps
cumulative_descriptives_png <- cumulative_descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 0.9)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(cumulative_descriptives_png, path = "cumulative_descriptives.png", zoom = 3, expand = 0)
```

# Step 1: Structural Disadvantage and Young-Adult Outcomes {#sec-step1}

Each standardized young-adult outcome is regressed on its corresponding
standardized structural disadvantage composite. These models establish
the expected outcome given adolescent disadvantage; their residuals are
used to operationalize resilience in Step 2. Results are summarized in
@tbl-disadvantage-regressions.

**Survey-Weighted Outcome Models**

```{r}
#| label: step1 regressions
# Model: Disadvantage → CES-D Depression 
model_depression <- svyglm(
  ZcesD_sum ~ Zcumadv_depression,
  design = design_w3,
  family = gaussian() )

# Model: Disadvantage → Substance Use
model_substance <- svyglm(
  Zsubuse_W3 ~ Zcumadv_substance,
  design = design_w3,
  family = gaussian() )

# Model: Disadvantage → Education Attainment
model_education <- svyglm(
  Zedattain_W3 ~ Zcumadv_education,
  design = design_w3,
  family = gaussian() )

# Model: Disadvantage → General Health
model_health <- svyglm(
  Zgenhealth_W3 ~ Zcumadv_health,
  design = design_w3,
  family = gaussian() )
```

**Step 1 Table Creation**

```{r}
#| label: create step1 table
# Create function to pull beta and p value from step 1 regressions
fmt_step1 <- function(model, var) {
  coefs <- summary(model)$coefficients
  b <- coefs[var, "Estimate"]
  p <- coefs[var, "Pr(>|t|)"]
  stars <- case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    p < .10  ~ "^",
    TRUE     ~ ""
  )
  paste0(formatC(b, digits = 2, format = "f"), stars)
}

# Create dataframe for regression table
step1_tbl <- data.frame(
  Predictor     = "Cumulative Disadvantage",
  Depression    = fmt_step1(model_depression, "Zcumadv_depression"),
  Substance_Use = fmt_step1(model_substance,  "Zcumadv_substance"),
  Education     = fmt_step1(model_education,  "Zcumadv_education"),
  Health        = fmt_step1(model_health,     "Zcumadv_health"))

# Create table for step 1 regressions
disadvantage_regressions <- flextable(step1_tbl) |>
  set_header_labels(
    Predictor     = "",
    Depression    = "Depression",
    Substance_Use = "Substance Use",
    Education     = "Education",
    Health        = "General Health"
  ) |>
  add_header_row(
    values    = c("", "Unstandardized Coefficients (b)"),
    colwidths = c(1, 4)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4) |>
  align(j = 2:5, align = "center", part = "all") |>
  add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-disadvantage-regressions
#| echo: false

disadvantage_regressions
```

Higher structural disadvantage is associated with more depressive
symptoms and substance use and with lower educational attainment and
poorer general health. Because all four composites are oriented so that
higher scores indicate greater disadvantage, the direction of the
coefficients is directly interpretable across outcomes. The residuals
from these models---the gap between predicted and observed
functioning---become the resilience scores used in the next step.

```{r}
#| label: save-disadvantageregressions-png
#| include: false

# Build `disadvantage_regressions_png` through a reproducible sequence of data-transformation steps
disadvantage_regressions_png <- disadvantage_regressions |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(disadvantage_regressions_png, path = "disadvantage_regressions.png", zoom = 2, expand = 0)
```

# Step 2: Residuals {#sec-step2}

For each outcome, the residual from its Step 1 model captures how much
better or worse a young adult is doing than expected given their
adolescent disadvantage exposure. Treating resilience continuously
preserves meaningful variation in adaptation and avoids imposing an
arbitrary threshold for classifying individuals as "resilient" or
"vulnerable."

Depression and substance-use residuals are reverse-coded (`* -1`) so
that, consistent with education and health, **higher values always
indicate greater resilience**.

```{r}
#| label: residuals
# ── Depression ─────────────────────────────────────────────────────────────────
dep_int    <- coef(model_depression)[1]
# Extract the fitted regression coefficient for reuse in the next analytic step
dep_Zcumadv <- coef(model_depression)[2]

dat_w3$predicted_depression <- dep_int + dep_Zcumadv * dat_w3$Zcumadv_depression
dat_w3$residual_depression  <- dat_w3$ZcesD_sum - dat_w3$predicted_depression
# Flip: higher = more resilient
dat_w3$residual_depressionR <- dat_w3$residual_depression * -1  

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_int    <- coef(model_substance)[1]
# Extract the fitted regression coefficient for reuse in the next analytic step
substance_Zcumadv <- coef(model_substance)[2]

dat_w3$predicted_substance  <- substance_int + substance_Zcumadv * dat_w3$Zcumadv_substance
dat_w3$residual_substance   <- dat_w3$Zsubuse_W3 - dat_w3$predicted_substance
# Flip: higher = more resilient
dat_w3$residual_substanceR  <- dat_w3$residual_substance * -1    

# ── Education ─────────────────────────────────────────────────────────────────
education_int    <- coef(model_education)[1]
# Extract the fitted regression coefficient for reuse in the next analytic step
education_Zcumadv <- coef(model_education)[2]

dat_w3$predicted_education <- education_int + education_Zcumadv * dat_w3$Zcumadv_education
dat_w3$residual_education  <- dat_w3$Zedattain_W3 - dat_w3$predicted_education

# ── Health ────────────────────────────────────────────────────────────────────
health_int    <- coef(model_health)[1]
# Extract the fitted regression coefficient for reuse in the next analytic step
health_Zcumadv <- coef(model_health)[2]

dat_w3$predicted_health <- health_int + health_Zcumadv * dat_w3$Zcumadv_health
dat_w3$residual_health  <- dat_w3$Zgenhealth_W3 - dat_w3$predicted_health

# Rerun design with residual variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

# Step 3: Multilevel Factors Associated with Resilience {#sec-step3}

The directionally harmonized resilience scores are regressed on
standardized individual, family, school, and community factors while
adjusting for sex and age. Separate survey-weighted models are estimated
for depression (@tbl-step3-depression), substance use
(@tbl-step3-substance), education (@tbl-step3-education), and general
health (@tbl-step3-health).

**Survey-Weighted Multivariable Models**

```{r}
#| label: step3 regressions
# ── Depression ─────────────────────────────────────────────────────────────────
depression_protect <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_protect <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
education_protect <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
health_protect <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )
```

**Create Step 3 Tables**

```{r}
#| label: create step3 tables
# Create variable labels for protective factors
var_labels_step3 <- list(
  Zcommconnect       ~ "Community Connectedness",
  Zcommintegr_disadv ~ "Community Integration",
  Zopt_hopeful       ~ "Optimism / Hopefulness",
  Zselfesteem        ~ "Self-Esteem",
  Zself_efficacy     ~ "Self-Efficacy",
  Zparentchild_rel   ~ "Parent-Child Relationship",
  Zschoolconnect     ~ "School Connectedness",
  bio_sexD           ~ "Sex (Female = 1)",
  Zwave3age          ~ "Age")

# Create function for table
fmt_tbl_regression <- function(model) {
  tbl_regression(model,
                 exponentiate = FALSE,
                 label = var_labels_step3) |>
    bold_p(t = .05) |>
    bold_labels() |>
    italicize_levels() |>
    modify_header(label = "**Variable**") |>
    modify_footnote(everything() ~ NA) |>
    as_flex_table() |>
    flextable::font(fontname = "Times New Roman", part = "all") |>
    flextable::fontsize(size = 12, part = "all") |>
    flextable::theme_apa()
}

# Run function for each outcome to generate four tables
tbl_step3_depression <- fmt_tbl_regression(depression_protect)
# Create `tbl_step3_substance` for use in the subsequent analysis/reporting steps
tbl_step3_substance  <- fmt_tbl_regression(substance_protect)
# Create `tbl_step3_education` for use in the subsequent analysis/reporting steps
tbl_step3_education  <- fmt_tbl_regression(education_protect)
# Create `tbl_step3_health` for use in the subsequent analysis/reporting steps
tbl_step3_health     <- fmt_tbl_regression(health_protect)
```

```{r}
#| label: tbl-step3-depression
#| echo: false

tbl_step3_depression
```

Self-esteem shows the strongest association with depression resilience
(*b* = 0.12, *p* \< .001), followed by parent-child relationship (*b* =
0.07, *p* = .003) and optimism/hopefulness (*b* = 0.06, *p* = .002) —
supporting individual- and family-level factors as key predictors .
Community integration is the only factor with no significant
association.

```{r}
#| label: save-step3-depression-png
#| include: false

# Build `tbl_step3_depression_png` through a reproducible sequence of data-transformation steps
tbl_step3_depression_png <- tbl_step3_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_step3_depression_png, path = paste0(fig_dir, "tbl_step3_depression.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-substance
#| echo: false

tbl_step3_substance
```

Parent-child relationship (*b* = 0.10, *p* \< .001) and sex (*b* = 0.43,
*p* \< .001) are the strongest predictors of substance use resilience.

```{r}
#| label: save-step3-substance-png
#| include: false

# Build `tbl_step3_substance_png` through a reproducible sequence of data-transformation steps
tbl_step3_substance_png <- tbl_step3_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_step3_substance_png, path = paste0(fig_dir, "tbl_step3_substance.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-education
#| echo: false

tbl_step3_education
```

School connectedness (*b* = 0.12, *p* \< .001) and age (*b* = 0.27, *p*
\< .001) are the strongest predictors of educational resilience, with
community integration (*b* = 0.07, *p* \< .001) and self-efficacy (*b* =
0.08, *p* \< .001) also significant. Self-esteem and parent-child
relationship are the only factors not significantly associated with this
outcome.

```{r}
#| label: save-step3-education-png
#| include: false

# Build `tbl_step3_education_png` through a reproducible sequence of data-transformation steps
tbl_step3_education_png <- tbl_step3_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_step3_education_png, path = paste0(fig_dir, "tbl_step3_education.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-health
#| echo: false

tbl_step3_health
```

Self-esteem (*b* = 0.14, *p* \< .001) and school connectedness (*b* =
0.11, *p* \< .001) are the only individual- and school-level factors
significant for health resilience; the community-level factors
(connectedness, integration) are not significant here, unlike their role
in the depression and education models.

```{r}
#| label: save-step3-health-png
#| include: false

# Build `tbl_step3_health_png` through a reproducible sequence of data-transformation steps
tbl_step3_health_png <- tbl_step3_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_step3_health_png, path = paste0(fig_dir, "tbl_step3_health.png"), zoom = 2, expand = 0)
```

## Cross-Outcome Interpretation

The pattern is outcome-specific rather than uniformly protective across
domains. Self-esteem is most prominent for depression and general-health
resilience; parent-child relationship is strongest for substance-use
resilience; and school connectedness is especially important for
educational resilience. This heterogeneity is substantively useful:
factors associated with doing better than expected in one developmental
domain are not necessarily the strongest correlates in another.

# Step 4: Does the Association Differ by Disadvantage? {#sec-step4}

Step 4 compared estimates in the bottom and top thirds of structural
disadvantage to describe whether associations appeared broadly
**promotive** or stronger under high disadvantage, which would be
**consistent with a protective pattern**.

::: {.callout-tip appearance="simple" icon="false"}
Results can be found in the following tables: depression
(@tbl-compare-depression), substance use (@tbl-compare-substance),
education (@tbl-compare-education), and health (@tbl-compare-health).
:::

**Create survey-weighted low-, middle-, and high-disadvantage groups**

For subgroup estimates, tertile cut points are calculated from the
survey-weighted distribution rather than from unweighted sample
quantiles.

```{r}
#| label: disadvantage-splits

# Define `weighted_tertiles()` once so the same operation can be applied consistently across variables/models without duplicating code
weighted_tertiles <- function(var, design) {
  cut_points <- as.numeric(
    coef(
      svyquantile(
        as.formula(paste0("~", var)),
        design,
        quantiles = c(.33, .66),
        ci = FALSE,
        na.rm = TRUE
      )
    )
  )

  cut(
    design$variables[[var]],
    breaks = c(-Inf, cut_points, Inf),
    labels = c("lo", "mid", "hi"),
    include.lowest = TRUE
  )
}

dat_w3$disadvantage_depression <-
  weighted_tertiles("Zcumadv_depression", design_w3)
dat_w3$disadvantage_substance <-
  weighted_tertiles("Zcumadv_substance", design_w3)
dat_w3$disadvantage_education <-
  weighted_tertiles("Zcumadv_education", design_w3)
dat_w3$disadvantage_health <-
  weighted_tertiles("Zcumadv_health", design_w3)

# Define the survey-design object that carries the Wave III sampling weights into subsequent survey-weighted estimates
design_w3 <- svydesign(
  ids = ~1,
  weights = ~GSWGT3_2,
  data = dat_w3
)
```

**Run residual regressions with each sub sample, for each outcome.**

```{r}
#| label: step4 regressions

# ── Depression ─────────────────────────────────────────────────────────────────
model_depression_lo <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_depression == "lo"),
  family = gaussian() )

# Fit a survey-weighted Gaussian regression so coefficient estimates incorporate the specified sampling weights
model_depression_hi <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_depression == "hi"),
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
model_substance_lo <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_substance == "lo"),
  family = gaussian() )

# Fit a survey-weighted Gaussian regression so coefficient estimates incorporate the specified sampling weights
model_substance_hi <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_substance == "hi"),
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
model_education_lo <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_education == "lo"),
  family = gaussian() )

# Fit a survey-weighted Gaussian regression so coefficient estimates incorporate the specified sampling weights
model_education_hi <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_education == "hi"),
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
model_health_lo <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_health == "lo"),
  family = gaussian() )

# Fit a survey-weighted Gaussian regression so coefficient estimates incorporate the specified sampling weights
model_health_hi <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, disadvantage_health == "hi"),
  family = gaussian() )
```

**Create Step 4 Tables**

```{r}
#| label: table helper function
# Create function to make comparison table
make_protect_table <- function(model_full, model_lo, model_hi, outcome_label) {
  
  protective_vars <- c(
    "Zcommconnect", "Zcommintegr_disadv",
    "Zopt_hopeful", "Zselfesteem", "Zself_efficacy",
    "Zparentchild_rel",
    "Zschoolconnect"
  )
  
  var_labels <- c(
    "Zcommconnect"       = "Community Connectedness",
    "Zcommintegr_disadv" = "Community Integration",
    "Zopt_hopeful"       = "Optimism / Hopefulness",
    "Zselfesteem"        = "Self-Esteem",
    "Zself_efficacy"     = "Self-Efficacy",
    "Zparentchild_rel"   = "Parent-Child Relationship",
    "Zschoolconnect"     = "School Connectedness"
  )
  
  fmt_coef <- function(model, var) {
    coefs <- summary(model)$coefficients
    if (!var %in% rownames(coefs)) return("—")
    b <- coefs[var, "Estimate"]
    p <- coefs[var, "Pr(>|t|)"]
    stars <- case_when(
      p < .001 ~ "***",
      p < .01  ~ "**",
      p < .05  ~ "*",
      p < .10  ~ "^",
      TRUE     ~ ""
    )
    paste0(formatC(b, digits = 2, format = "f"), stars)
  }
  
  tbl <- data.frame(
    Variable       = var_labels[protective_vars],
    Full_Sample    = sapply(protective_vars, \(v) fmt_coef(model_full, v)),
    Low_Disadvantage  = sapply(protective_vars, \(v) fmt_coef(model_lo,   v)),
    High_Disadvantage = sapply(protective_vars, \(v) fmt_coef(model_hi,   v)),
    row.names = NULL
  )
  
  flextable(tbl) |>
    set_header_labels(
      Variable       = "Protective Factor",
      Full_Sample    = "Full Sample",
      Low_Disadvantage  = "Low Disadvantage",
      High_Disadvantage = "High Disadvantage"
    ) |>
    add_header_row(
      values    = c("", outcome_label),
      colwidths = c(1, 3)
    ) |>
    theme_apa() |>
    font(fontname = "Times New Roman", part = "all") |>
    fontsize(size = 11, part = "all") |>
    bold(part = "header") |>
    width(j = 1, width = 2.2) |>
    width(j = 2:4, width = 1.5) |>
    align(j = 2:4, align = "center", part = "all") |>
    add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
    fontsize(size = 9, part = "footer") |>
    italic(part = "footer")
}

# Run function for each outcome
tbl_compare_depression <- make_protect_table(
  depression_protect, model_depression_lo, model_depression_hi, "Depression")

# Create `tbl_compare_substance` for use in the subsequent analysis/reporting steps
tbl_compare_substance <- make_protect_table(
  substance_protect, model_substance_lo,  model_substance_hi, "Substance Use")

# Create `tbl_compare_education` for use in the subsequent analysis/reporting steps
tbl_compare_education <- make_protect_table(
  education_protect, model_education_lo, model_education_hi, "Education Attainment")

# Create `tbl_compare_health` for use in the subsequent analysis/reporting steps
tbl_compare_health <- make_protect_table(
  health_protect, model_health_lo, model_health_hi,   "General Health")
```

```{r}
#| label: tbl-compare-depression
#| echo: false

tbl_compare_depression
```

The subgroup estimates are **consistent with stronger associations under
high disadvantage** for community connectedness and self-efficacy,
whereas self-esteem and parent-child relationship show associations in
both disadvantage groups.

```{r}
#| label: save-compare-depression-png
#| include: false

# Build `tbl_compare_depression_png` through a reproducible sequence of data-transformation steps
tbl_compare_depression_png <- tbl_compare_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_compare_depression_png, path = paste0(fig_dir, "tbl_compare_depression.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-substance
#| echo: false

tbl_compare_substance
```

Community connectedness shows a similar association with substance-use
resilience in both disadvantage groups. Parent-child relationship shows
a larger coefficient in the high-disadvantage group than in the
low-disadvantage group, a pattern **consistent with moderation** that is
evaluated directly in the interaction models below.

```{r}
#| label: save-compare-substance-png
#| include: false

# Build `tbl_compare_substance_png` through a reproducible sequence of data-transformation steps
tbl_compare_substance_png <- tbl_compare_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_compare_substance_png, path = paste0(fig_dir, "tbl_compare_substance.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-education
#| echo: false

tbl_compare_education
```

Community integration, self-efficacy, and school connectedness show
associations with educational resilience in both disadvantage groups.
School connectedness has a somewhat larger coefficient in the
high-disadvantage group; the interaction analysis below provides the
appropriate test of whether that difference is statistically meaningful.

```{r}
#| label: save-compare-education-png
#| include: false

# Build `tbl_compare_education_png` through a reproducible sequence of data-transformation steps
tbl_compare_education_png <- tbl_compare_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_compare_education_png, path = paste0(fig_dir, "tbl_compare_education.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-health
#| echo: false

tbl_compare_health
```

Optimism/hopefulness and community connectedness show associations with
general-health resilience primarily in the high-disadvantage subsample,
while self-esteem is associated with resilience in both groups. These
patterns motivate the formal moderation tests below rather than serving
as stand-alone evidence that subgroup effects differ.

```{r}
#| label: save-compare-health-png
#| include: false

# Build `tbl_compare_health_png` through a reproducible sequence of data-transformation steps
tbl_compare_health_png <- tbl_compare_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

# Save the completed output to the project output directory; this chunk is hidden from the rendered portfolio
save_as_image(tbl_compare_health_png, path = paste0(fig_dir, "tbl_compare_health.png"), zoom = 3, expand = 0)
```

# Key Findings and Analytic Takeaways

This analysis demonstrates an end-to-end quantitative workflow: cleaning
and harmonizing a large longitudinal secondary dataset, documenting
measurement decisions, incorporating survey weights, constructing
outcome-specific measures, fitting and summarizing multivariable models,
and translating technical results into substantive developmental
interpretation.

Two findings are especially important:

1.  **Structural disadvantage is multidimensional and
    outcome-specific.** The relative contribution of school, family,
    economic, and neighborhood disadvantage differs across young-adult
    outcomes.

2.  **Resilience is also domain-specific.** Individual, family, school,
    and community factors do not show the same pattern across
    depression, substance use, education, and general health.
