---
title: "SPR Cup Analysis"
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

# Set Up

## Overview

This analysis was originally developed for the 21st Annual Sloboda and
Bukoski SPR Cup, presented at the Society for Prevention Research 34th
Annual Meeting [@marcus2026]. It tests a residual-based model of
resilience using the Add Health dataset ([@harris2013], specifically
Waves I and III (*N* = 4,882) and can be broken down into the following
steps:

-   indicators are used to build **outcome-specific adversity
    composites** [@marquez2023]**.**
-   **regressed** on those composites [@miller-lewis2013].
-   become **continuous** **resilience scores** (better- or
    worse-than-expected adaptation given adversity exposure).
-   [Step 3: Protective Factors]: Multilevel resilience factors
    (individual, family, school, community) are tested as **predictors
    of resilience**, first in the full sample.
-   and high-adversity subsamples to distinguish between **promotive
    effects** (help regardless of adversity level) and **protective
    effects** (buffer specifically under high adversity).

**Load Packages**

All analyses use survey-weighted regression (`survey`) to account for
Add Health's complex sampling design, with `gtsummary`/`flextable`/`gt`
for APA-style output tables.

```{r}
#| label: load packages
#| output: false
library(gtsummary)   # regression summary tables
library(flextable)   # table formatting and export
library(gtools)      # quantile splitting 
library(haven)       # read SPSS .sav files
library(tidyverse)   # data wrangling
library(survey)      # complex survey / weighted regression
library(broom)       # tidy model output
library(gt)          # formatted tables
```

The analytic dataset is the merged Add Health restricted-use survey data
(`AddHealth_AllData.sav`), read in with `haven::read_sav()`. Import code
is included but not rendered below.

```{r}
#| label: load data
#| include: false
library(haven)
dat <- read_sav("C:/Users/rcm5550/Desktop/PSU OD Copy/Rachel - The Pennsylvania State University/SPR Cup 2026/Analysis/AddHealth_AllData.sav")
```

```{r}
#| label: save tables
#| include: false
# ── Output directory for saved table images ──────────────────────────────────
fig_dir <- "../../figures/spr-cup/"
if (!dir.exists(fig_dir)) dir.create(fig_dir, recursive = TRUE)
```

## Data Cleaning (SPSS)

Variable recoding, dummy coding, and composite construction for this
project were done in SPSS prior to import into R, across four syntax
files by domain (demographics, structural disadvantage, outcomes,
protective factors). Each is excerpted below to document the approach.

### Demographics

`Demographics.sps` recodes core demographic variables (child sex, age,
race/ethnicity) into analysis-ready measures, checking frequencies
before and after each recode.

``` spss
*****************************CHILD SEX.
*Q: What is your biological sex?
***check frequencies for zeros.
frequencies variables = bio_sex /statistics = all.
*male = 1, female = 2.

*create dummy code for child biological sex (female = 1).
if(bio_sex eq 1) bio_sexD = 0.
if(bio_sex eq 2) bio_sexD = 1.
variable labels bio_sexD
    "child biological sex".
value labels bio_sexD
    0 "male"
    1 "female".
execute.

*check frequencies of new variable.
frequencies variables = bio_sexD /statistics = all.
```

### Structural Disadvantage

`Disadvantage.sps` (672 lines) builds individual- and household-level
disadvantage indicators, including race/ethnicity dummy codes with
missing-data recodes.

``` spss

*****************************FAMILY SOCIOECONOMIC DISADVANTAGE COMPOSITE.
*Components: famincomecat (continuous $, higher = more income), pareduc_avg (0-8, higher = more education), parentemploy (0-1, higher = more employment).
**Standardizing each component to z-score then reverse code so higher is more disadvantage, then take the average of all variables.

*standardize based on weight filtered dataset.
delete variables  Zfamincomecat  Zpareduc_avg Zparentemploy.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = famincomecat pareduc_avg parentemploy /save.
*Creates Zfamincomecat, Zpareduc_avg, Zparentemploy.
execute.

compute Zfamincomecat_r = Zfamincomecat * -1.
variable labels Zfamincomecat_r
    "Family income z-scored reversed (higher = lower income = more disadvantage)".
execute.

compute Zpareduc_avg_r = Zpareduc_avg * -1.
variable labels Zpareduc_avg_r
    "Average parental education z-scored reversed (higher = less education = more disadvantage)".
execute.

compute Zparentemploy_r = Zparentemploy * -1.
variable labels Zparentemploy_r
    "Average parental employment z-scored reversed (higher = less employed = more disadvantage)".
execute.

compute famSES_disadv = mean(Zfamincomecat_r, Zpareduc_avg_r, Zparentemploy_r).
variable labels famSES_disadv
    "Family SES disadvantage composite: mean of z-scored reversed income, education, employment (higher = more disadvantage)".
execute.

frequencies variables = famSES_disadv /statistics = all.

reliability variables = famincomecat pareduc_avg parentemploy
    /scale('famSES') all /statistics = all.

*standardize based on weight filtered dataset.
temporary.
select if (GSWGT3_2 ge 0). 
descriptives variables = famSES_disadv /save.
*Creates Zfamincomecat, Zpareduc_avg, Zparentemploy.
execute.
```

### Outcomes

`Outcomes.sps` (473 lines) constructs the substance-use outcome variable
(alcohol, cigarette, and tobacco use composite), depression, general
health, and educational attainment. Below is a exert from the final
section of scoring the depression outcome.

``` spss
*****************************CES-D DEPRESSION COMPOSITE (WAVE 3).
*9 items (all in same 0-3 direction after reverse coding H3SP7 and H3SP11 above).
*Sum score (0-27 range); Desch et al. (2023).

compute cesD_sum = sum(dep_bothered, dep_blues, dep_asgood, dep_concentrate, dep_depressed,
                       dep_tired, dep_enjoyed, dep_sad, dep_disliked).
variable labels cesD_sum
    "CES-D depression sum score Wave 3 (0-27, higher = more depressive symptoms)".
execute.

reliability variables = dep_bothered dep_blues dep_asgood dep_concentrate dep_depressed dep_tired dep_enjoyed dep_sad dep_disliked
    /scale('CESD') all /statistics = all.

frequencies variables = cesD_sum /statistics = all.

*standardize based on weight filtered dataset.
delete variables  ZcesD_sum.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = cesD_sum / save.
```

### Protective Factors

`Protective.sps` (887 lines) builds the protective-factor composites,
several of which are reverse-scored so that higher values consistently
indicate greater disadvantage.Below is an example of the data cleaning
of one variable and then the creation of one protective factor
composite.

``` spss
***************************COMMUNITY RESOURCES - CONNECTEDNESS (ADOLESCENT REPORT).

*Note: Items reverse-scored here so higher = LOWER community resources (disadvantage direction),
*consistent with the structural disadvantage framing of this composite.
*Community Connectedness (Wickrama & Bryant 2003): 4 items, yes/no (0/1), summed 0-4.
*Reverse scoring: recode 1->0 and 0->1 so high score = LOW connectedness = more disadvantage.

if (H1NB1 eq 1) knowneighbR = 0.
    *knows neighbors = connected (reverse to 0=low disadvantage).
if (H1NB1 eq 2) knowneighbR = 1.
    *doesn't know = disconnected = higher disadvantage.
if (H1NB1 eq 6 or H1NB1 eq 8 or H1NB1 eq 9) knowneighbR = -99.
recode knowneighbR (-99 = sysmis).
variable labels knowneighbR
    "Does NOT know most neighbors (H1NB1 reversed, 1=low connectedness/more disadvantage)".
value labels knowneighbR 0 "knows most neighbors" 1 "does not know most neighbors".
execute.
    
***************************COMMUNITY CONNECTEDNESS INDEX (Wickrama & Bryant 2003).
*Items coded in protective direction (1 = connected):
*knowneighb, talkneighb, neighbwatch, nbhdsafe.
*Sum 0-4; higher = more connectedness = more protective.

compute commconnect = sum(knowneighb, talkneighb, neighbwatch, nbhdsafe).
variable labels commconnect
    "Community connectedness (Wickrama & Bryant 2003, sum 0-4, higher = more connectedness)".
execute.

reliability variables = knowneighb talkneighb neighbwatch nbhdsafe
    /scale('commconnect') all /statistics = all.
*alpha = .51.

frequencies variables = commconnect /statistics = all.


*standardize based on weight filtered dataset.
delete variables  Zcommconnect.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = commconnect / save.
*******************************************************************************
```

The cleaned file was exported and merged into `AddHealth_AllData.sav`
for the R analysis above.

## Survey Design Specification

Add Health uses a complex, clustered, stratified sampling design. All
regressions use the Wave 3 grand sample weight (`GSWGT3_2_2`) via the
`survey` package to produce design-consistent estimates.

**Missing Sample Weight at Wave 3**

Respondents missing GSWGT3_2 were not part of the Wave 3 sample (e.g.,
lost to follow-up, deceased, ineligible) and should be excluded from
analysis

```{r}
#| label: check missing sample weight

#how many cases missing
cat("Missing GSWGT3_2:", sum(is.na(dat$GSWGT3_2)), "\n")
cat("Non-missing GSWGT3_2:", sum(!is.na(dat$GSWGT3_2)), "\n")
```

**Filtered Dataset**

```{r}
#| label: filter for analytic dataset
# ── Subset to respondents with valid Wave 3 weight ────────────────────────────
dat_w3 <- dat |> filter(!is.na(GSWGT3_2)) #dat_w3 is name of filtered dataset 

cat("Analytic n after excluding missing weights:", nrow(dat_w3), "\n")
```

**Sample Weighted Dataset**

```{r}
#| label: weighted dataset
# ── Specify survey design on the subsetted data ───────────────────────────────
# Use design_w3 as survey weight design in regressions
# svydesign() with ids = ~1 specifies no clustering beyond the given weights
# Add Health's Wave 3 grand sample weight (GSWGT3_2) adjusts for unequal selection probability and non-response so estimates generalize to the population

design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)

cat("Sum of weights:", round(sum(dat_w3$GSWGT3_2), 0), "\n")
```

## Descriptive Statistics

Create @tbl-descriptives to display the *M*, *SD*, and range of each
analytic variable.

```{r}
#| label: create descriptives table
# ── Variable lists with display labels ────────────────────────────────────────
adversity_vars <- c(
  "minorityD"          = "Minority Status",
  "school_disadv"     = "School Disadvantage",
  "econhardship"      = "Economic Hardship",
  "famSES_disadv"      = "Family SES Disadvantage",
  "nbhd_subj_disadv"  = "Neighborhood Disadvantage (Subjective)",
  "cens_nbhd_disadv"  = "Neighborhood Disadvantage (Census)")

protective_vars <- c(
  "commconnect"       = "Community Connectedness",
  "commintegr_disadv" = "Community Integration",
  "opt_hopeful"       = "Optimism / Hopefulness",
  "selfesteem"        = "Self-Esteem",
  "self_efficacy"     = "Self-Efficacy",
  "parentchild_rel"   = "Parent-Child Relationship",
  "schoolconnect"     = "School Connectedness")

outcome_vars <- c(
  "cesD_sum"    = "Depression Symptoms",
  "subuse_W3"   = "Substance Use",
  "edattain_W3" = "Education Attainment",
  "genhealth_W3"= "General Health")

# ── Function: get weighted M and SD, and unweighted range ─────────────────────
get_desc <- function(var, design) {
  mn  <- as.numeric(svymean(as.formula(paste0("~", var)), design, na.rm = TRUE))
  vr  <- as.numeric(svyvar(as.formula(paste0("~", var)),  design, na.rm = TRUE))
  rng <- range(design$variables[[var]], na.rm = TRUE)
  data.frame(
    M     = formatC(mn,       digits = 2, format = "f"),
    SD    = formatC(sqrt(vr), digits = 2, format = "f"),
    Range = paste0(formatC(rng[1], digits = 2, format = "f"),
                   "–",
                   formatC(rng[2], digits = 2, format = "f")))
}

# ── Build rows for each variable group ────────────────────────────────────────
make_desc_rows <- function(var_named_vec, design, section_label) {
  rows <- lapply(names(var_named_vec), function(v) {
    d <- get_desc(v, design)
    data.frame(Variable = var_named_vec[[v]], M = d$M, SD = d$SD, Range = d$Range)
  })
  rows <- do.call(rbind, rows)
  rbind(data.frame(Variable = section_label, M = "", SD = "", Range = ""), rows)
}

desc_tbl <- rbind(
  make_desc_rows(adversity_vars,  design_w3, "Adversity Predictors"),
  make_desc_rows(protective_vars, design_w3, "Protective Factors"),
  make_desc_rows(outcome_vars,    design_w3, "Outcomes"))

# ── Build flextable ───────────────────────────────────────────────────────────
header_rows <- which(desc_tbl$M == "")

header_rows <- which(desc_tbl$M == "")

descriptives <- flextable(desc_tbl) |>
  set_header_labels(Variable = "Variable", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Weighted Descriptive Statistics"),
    colwidths = c(1, 3)                                        # was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  bold(i = header_rows, part = "body") |>
  width(j = 1, width = 3.0) |>
  width(j = 2:4, width = 1.0) |>                              # was 2:3
  align(j = 2:4, align = "center", part = "all") |>           # was 2:3
  padding(i = setdiff(seq_len(nrow(desc_tbl)), header_rows),
          j = 1, padding.left = 20, part = "body") |>
  add_footer_lines("Note. Means and standard deviations are weighted using the Wave 3 grand sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-descriptives
#| echo: false
descriptives
```

The analytic sample reports moderate average levels of structural
disadvantage (e.g., *M* = 0.52 school disadvantage on a 0–3 scale)
alongside relatively high protective-factor scores (e.g., *M* = 4.15
parent-child relationship on a 1–5 scale), consistent with a normative,
not exclusively high-risk, longitudinal sample.

```{r}
#| label: save-descriptives-png
#| include: false

descriptives_png <- descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 3.2) |>
  width(j = 2:4, width = 1.05)

save_as_image(descriptives_png, path = paste0(fig_dir, "descriptives.png"), zoom = 3, expand = 0)
```

# Step 0: Adversity Composites

## Individual Disadvantage Regressions

The next several chunks repeat the same regression for each structural
disadvantage indicator (minority status, school, economic, family SES,
neighborhood subjective/census) predicting each outcome (depression,
substance use, educational attainment, general health). Each regression
uses the standardized outcome and standardized predictor variable, and
each chunk stores the beta and p-value, which become the weights used to
build the cumulative adversity composite in the
@cumulative-adversity-composites section below.

::: {.callout-note appearance="simple" icon="false"}
**Regress outcome on structural disadvantage indicator.**

`outcome_disadvantage <-`

`svyglm(outcome ~ disadvantage, design = design_w3, family = gaussian())`
:::

::: {.callout-note appearance="simple" icon="false"}
**Extra standardized betas for later beta-weighted composites.**

`dat_w3$outcome_disadvantage <- coef(outcome_disadvantage)[2]`
:::

::: {.callout-note appearance="simple" icon="false"}
**Extra standardized betas and p values for predictor weight table.**

`outcome_disadvantageB<- coef(outcome_disadvantage[2]`

`outcome_disadvantageP`
`<-summary(outcome_disadvantage)$coefficients["disadvantage","Pr(\>\|t\|)"\]`
:::

::: {.callout-note appearance="minimal" icon="false"}
Jump to @tbl-adversity-weight-table to see the results of the individual
disadvantage regressions.
:::

**Depression**

Depression Weights - Minority Status

```{r}
#| label: depression weights - minority
# run regression
depression_minorityD <- svyglm(
  ZcesD_sum ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_minorityD <- coef(depression_minorityD)[2]
dep_minorityD_B <- coef(depression_minorityD)[2]
dep_minorityD_P <-summary(depression_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

Depression Weights - School Disadvantage

```{r}
#| label: depression weights - school

# run regression
depression_school_disadv <- svyglm(
  ZcesD_sum ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_school <- coef(depression_school_disadv)[2]
dep_school_B <- coef(depression_school_disadv)[2]
dep_school_P <-summary(depression_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

Depression Weights - Economic Hardship

```{r}
#| label: depression weights - econ

# run regression
depression_econhardship <- svyglm(
  ZcesD_sum ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_econ <- coef(depression_econhardship)[2]
dep_econ_B <- coef(depression_econhardship)[2]
dep_econ_P <-summary(depression_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

Depression Weights - Family SES

```{r}
#| label: depression weights - fam

# run regression
depression_famSES_disadv <- svyglm(
  ZcesD_sum ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_fam <- coef(depression_famSES_disadv)[2]
dep_famSES_B <- coef(depression_famSES_disadv)[2]
dep_famSES_P <-summary(depression_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

Depression Weights - Neighborhood Disadvantage, Subjective

```{r}
#| label: depression weights - nbhdsub

# run regression
depression_nbhd_subj_disadv <- svyglm(
  ZcesD_sum ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_nbhdsub <- coef(depression_nbhd_subj_disadv)[2]
dep_nbhdsub_B <- coef(depression_nbhd_subj_disadv)[2]
dep_nbhdsub_P <-summary(depression_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

Depression Weights - Neighborhood Disadvantage, Census

```{r}
#| label: depression weights - cens nbhd

# run regression
depression_cens_nbhd_disadv <- svyglm(
  ZcesD_sum ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_censnbhd <- coef(depression_cens_nbhd_disadv)[2]
dep_censnbhd_B <- coef(depression_cens_nbhd_disadv)[2]
dep_censnbhd_P <-summary(depression_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

**Substance Use**

Substance Use Weights - Minority Status

```{r}
#| label: substance use weights - minority

# run regression
substance_minorityD <- svyglm(
  Zsubuse_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_minorityD <- coef(substance_minorityD)[2]
substance_minorityD_B <- coef(substance_minorityD)[2]
substance_minorityD_P <-summary(substance_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

Substance Use Weights - School Disadvantage

```{r}
#| label: substance use weights - school

# run regression
substance_school_disadv <- svyglm(
  Zsubuse_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_school <- coef(substance_school_disadv)[2]
substance_school_B <- coef(substance_school_disadv)[2]
substance_school_P <-summary(substance_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

Substance Use Weights - Economic Hardship

```{r}
#| label: substance use weights - econ

# run regression
substance_econhardship <- svyglm(
  Zsubuse_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_econ <- coef(substance_econhardship)[2]
substance_econ_B <- coef(substance_econhardship)[2]
substance_econ_P <-summary(substance_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

Substance Use Weights - Family SES

```{r}
#| label: substance use weights - fam

# run regression
substance_famSES_disadv <- svyglm(
  Zsubuse_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_fam <- coef(substance_famSES_disadv)[2]
substance_famSES_B <- coef(substance_famSES_disadv)[2]
substance_famSES_P <-summary(substance_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

Substance Use Weights - Neighborhood Disadvantage, Subjective

```{r}
#| label: substance use weights - nbhdsub

# run regression
substance_nbhd_subj_disadv <- svyglm(
  Zsubuse_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_nbhdsub <- coef(substance_nbhd_subj_disadv)[2]
substance_nbhdsub_B <- coef(substance_nbhd_subj_disadv)[2]
substance_nbhdsub_P <-summary(substance_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

Substance Use Weights - Neighborhood Disadvantage, Census

```{r}
#| label: substance use weights - cens nbhd

# run regression
substance_cens_nbhd_disadv <- svyglm(
  Zsubuse_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_censnbhd <- coef(substance_cens_nbhd_disadv)[2]
substance_censnbhd_B <- coef(substance_cens_nbhd_disadv)[2]
substance_censnbhd_P <-summary(substance_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

**Education Attainment**

Education Weights - Minority Status

```{r}
#| label: education weights - minority

# run regression
education_minorityD <- svyglm(
  Zedattain_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_minorityD <- coef(education_minorityD)[2]
education_minorityD_B <- coef(education_minorityD)[2]
education_minorityD_P <- summary(education_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

Education Weights - School Disadvantage

```{r}
#| label: education weights - school

# run regression
education_school_disadv <- svyglm(
  Zedattain_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_school <- coef(education_school_disadv)[2]
education_school_B <- coef(education_school_disadv)[2]
education_school_P <- summary(education_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

Education Weights - Economic Hardship

```{r}
#| label: education weights - econ

# run regression
education_econhardship <- svyglm(
  Zedattain_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_econ <- coef(education_econhardship)[2]
education_econ_B <- coef(education_econhardship)[2]
education_econ_P <- summary(education_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

Education Weights - Family SES

```{r}
#| label: education weights - fam

# run regression
education_famSES_disadv <- svyglm(
  Zedattain_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_fam <- coef(education_famSES_disadv)[2]
education_famSES_B <- coef(education_famSES_disadv)[2]
education_famSES_P <- summary(education_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

Education Weights - Neighborhood Disadvantage, Subjective

```{r}
#| label: education weights - nbhdsub

# run regression
education_nbhd_subj_disadv <- svyglm(
  Zedattain_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_nbhdsub <- coef(education_nbhd_subj_disadv)[2]
education_nbhdsub_B <- coef(education_nbhd_subj_disadv)[2]
education_nbhdsub_P <- summary(education_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

Education Weights - Neighborhood Disadvantage, Census

```{r}
#| label: education weights - cens nbhd

# run regression
education_cens_nbhd_disadv <- svyglm(
  Zedattain_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_censnbhd <- coef(education_cens_nbhd_disadv)[2]
education_censnbhd_B <- coef(education_cens_nbhd_disadv)[2]
education_censnbhd_P <- summary(education_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

**General Health**

General Health Weights - Minority Status

```{r}
#| label: health weights - minority

# run regression
health_minorityD <- svyglm(
  Zgenhealth_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_minorityD <- coef(health_minorityD)[2]
health_minorityD_B <- coef(health_minorityD)[2]
health_minorityD_P <- summary(health_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

General Health Weights - School Disadvantage

```{r}
#| label: health weights - school

# run regression
health_school_disadv <- svyglm(
  Zgenhealth_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_school <- coef(health_school_disadv)[2]
health_school_B <- coef(health_school_disadv)[2]
health_school_P <- summary(health_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

General Health Weights - Economic Hardship

```{r}
#| label: health weights - econ

# run regression
health_econhardship <- svyglm(
  Zgenhealth_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_econ <- coef(health_econhardship)[2]
health_econ_B <- coef(health_econhardship)[2]
health_econ_P <- summary(health_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

General Health Weights - Family SES

```{r}
#| label: health weights - fam

# run regression
health_famSES_disadv <- svyglm(
  Zgenhealth_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_fam <- coef(health_famSES_disadv)[2]
health_famSES_B <- coef(health_famSES_disadv)[2]
health_famSES_P <- summary(health_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

General Health Weights - Neighborhood Disadvantage, Subjective

```{r}
#| label: health weights - nbhdsub

# run regression
health_nbhd_subj_disadv <- svyglm(
  Zgenhealth_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_nbhdsub <- coef(health_nbhd_subj_disadv)[2]
health_nbhdsub_B <- coef(health_nbhd_subj_disadv)[2]
health_nbhdsub_P <- summary(health_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

General Health Weights - Neighborhood Disadvantage, Census

```{r}
#| label: health weights - cens nbhd

# run regression
health_cens_nbhd_disadv <- svyglm(
  Zgenhealth_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_censnbhd <- coef(health_cens_nbhd_disadv)[2]
health_censnbhd_B <- coef(health_cens_nbhd_disadv)[2]
health_censnbhd_P <- summary(health_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

## Adversity Weights

The code below utilizes the extracted coefficients and *p*-values and
creates a presentation-ready table (@tbl-adversity-weight-table).

```{r}
#| label: create adversity weight table

weight_summary <- data.frame(
  Predictor = c("Minority Status", "School Disadvantage", "Economic Hardship",
                "Family SES", "Neighborhood (Subj.)", "Neighborhood (Census)"),

  # coefficient and p values for depression
  dep_B  = round(c(dep_minorityD_B, dep_school_B, dep_econ_B,
                   dep_famSES_B, dep_nbhdsub_B, dep_censnbhd_B), 2),
  dep_P  = round(c(dep_minorityD_P, dep_school_P, dep_econ_P,
                   dep_famSES_P, dep_nbhdsub_P, dep_censnbhd_P), 2),
  
  # coefficient and p values for substance use
  sub_B  = round(c(substance_minorityD_B, substance_school_B, 
                   substance_econ_B, substance_famSES_B, substance_nbhdsub_B,
                   substance_censnbhd_B), 2),
  sub_P  = round(c(substance_minorityD_P, substance_school_P, substance_econ_P,
                   substance_famSES_P, substance_nbhdsub_P,
                   substance_censnbhd_P), 2),

 # coefficient and p values for education
  edu_B  = round(c(education_minorityD_B, education_school_B, education_econ_B,
                   education_famSES_B, education_nbhdsub_B,
                   education_censnbhd_B), 2),
  edu_P  = round(c(education_minorityD_P, education_school_P, education_econ_P,
                   education_famSES_P, education_nbhdsub_P,
                   education_censnbhd_P), 2),

  # coefficient and p values for general health
  hlt_B  = round(c(health_minorityD_B, health_school_B, health_econ_B,
                   health_famSES_B, health_nbhdsub_B, health_censnbhd_B), 2),
  hlt_P  = round(c(health_minorityD_P, health_school_P, health_econ_P,
                   health_famSES_P, health_nbhdsub_P, health_censnbhd_P), 2))

# ── Build table ───────────────────────────────────────────────────────────
table_weights <- flextable(weight_summary) |>
  set_header_labels(
    Predictor = "Predictor",
    dep_B = "b", dep_P = "p",
    sub_B = "b", sub_P = "p",
    edu_B = "b", edu_P = "p",
    hlt_B = "b", hlt_P = "p"
  ) |>
  add_header_row(
    values     = c("", "Depression", "Substance Use", "Education", "Health"),
    colwidths  = c(1, 2, 2, 2, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 1.5) |>
  width(j = 2:9, width = .6) |>
  align(j = 2:9, align = "center", part = "all") |>
  add_footer_lines("Note. All continuous predictors and outocmes were standardized. Each predictor was standardized and then regressed on the outcome separately.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-adversity-weight-table
#| echo: false

table_weights
```

Family SES disadvantage is the strongest predictor across all four
outcomes (\|*b*\| = 0.13–0.58), and minority status shows the largest
effect specifically for substance use (*b* = -0.40) and education (*b*
=-0.25). All predictor-outcome associations are significant at *p*
\<.001, supporting the use of each as a weighted component of its
outcome-specific composite below.

```{r}
#| label: save-adversity-weight-table-png
#| include: false
table_weights_png <- table_weights |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  set_header_labels(dep_B = "Β", dep_P = "p", sub_B = "Β", sub_P = "p",
                    edu_B = "Β", edu_P = "p", hlt_B = "Β", hlt_P = "p") |>
  italic(i = 2, j = c(3, 5, 7, 9), part = "header") |>
  width(j = 1, width = 1.6) |>
  width(j = 2:9, width = 0.6)

save_as_image(table_weights_png, path = paste0(fig_dir, "table_weights.png"), zoom = 3, expand = 0)
```

## Structural Disadvantage Composites {#cumulative-adversity-composites}

Each outcome's composite is a weighted average of the six structural
disadvantage indicators, where the weight for each indicator is the beta
extracted above (indicator value × its beta for this specific outcome).
This means the same six indicators contribute differently to each of the
four composites, based on each indicator's empirical association with
that outcome.

::: {.callout-important appearance="minimal" icon="false"}
A predictor with a stronger relationship to depression, for example, is
weighted more heavily in the depression composite than in the substance
use composite. The descriptives of the structural disadvantage
composites can be found in @tbl-cumulative.
:::

**Compute Composites & Add to Dataset**

```{r}
#| label: cumulative composites
# compute weighted composites for all four outcomes
dat_w3 <- dat_w3 %>%
  mutate(
    
    # depression composite
    cumadv_depression = (
      (minorityD)         * dep_minorityD +
      (Zschool_disadv)    * dep_school +
      (Zeconhardship)     * dep_econ +
      (famSES_disadv)     * dep_fam +
      (Znbhd_subj_disadv) * dep_nbhdsub +
      (Zcens_nbhd_disadv) * dep_censnbhd),
    
    # substance use composite
    cumadv_substance = (
      (minorityD)         * substance_minorityD +
      (Zschool_disadv)    * substance_school +
      (Zeconhardship)     * substance_econ +
      (famSES_disadv)     * substance_fam +
      (Znbhd_subj_disadv) * substance_nbhdsub +
      (Zcens_nbhd_disadv) * substance_censnbhd),

    # education composite
    cumadv_education = (
      (minorityD)         * education_minorityD +
      (Zschool_disadv)    * education_school +
      (Zeconhardship)     * education_econ +
      (famSES_disadv)     * education_fam +
      (Znbhd_subj_disadv) * education_nbhdsub +
      (Zcens_nbhd_disadv) * education_censnbhd),
    
    #health composite
    cumadv_health = (
      (minorityD)         * health_minorityD +
      (Zschool_disadv)    * health_school +
      (Zeconhardship)     * health_econ +
      (famSES_disadv)     * health_fam +
      (Znbhd_subj_disadv) * health_nbhdsub +
      (Zcens_nbhd_disadv) * health_censnbhd))

# standardize composites and add into dataset 
dat_w3 <- dat_w3 %>%
  mutate(
    Zcumadv_depression = as.numeric(scale(cumadv_depression)),
    Zcumadv_substance  = as.numeric(scale(cumadv_substance)),
    Zcumadv_education  = as.numeric(scale(cumadv_education)),
    Zcumadv_health     = as.numeric(scale(cumadv_health)))

# rerun design with composite variables now added to dataset
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

```{r}
#| label: create cumulative disadvantage descriptives
# create table labels for each outcome
cumadv_vars <- c(
  "cumadv_depression" = "Depression",
  "cumadv_substance"  = "Substance Use",
  "cumadv_education"  = "Education Attainment",
  "cumadv_health"     = "General Health")

# create function for disadvantage composite descriptives
cumadv_rows <- do.call(rbind, lapply(names(cumadv_vars), function(v) {
  d <- get_desc(v, design_w3)
  data.frame(Variable = cumadv_vars[[v]], M = d$M, SD = d$SD, Range = d$Range)
}))

# create table for disadvantage composite descriptives
cumulative_descriptives <- flextable(cumadv_rows) |>
  set_header_labels(Variable = "Outcome", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Cumulative Adversity Composites"),
    colwidths = c(1, 3)                              # was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 1.0) |>                   
  align(j = 2:4, align = "center", part = "all") |> 
  add_footer_lines("Note. Each structural disadvantage composite is outcome-specific, weighted by the association of each disadvantage indicator with that outcome. Descriptive statistics are survey-weighted using the Wave 3 sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer") |>
  fit_to_width(max_width = 6.5)
```

```{r}
#| label: tbl-cumulative
#| echo: false

cumulative_descriptives
```

```{r}
#| label: save-cumulative-png
#| include: false

cumulative_descriptives_png <- cumulative_descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 0.9)

save_as_image(cumulative_descriptives_png, path = "cumulative_descriptives.png", zoom = 3, expand = 0)
```

# Step 1: Cumulative Adversity to Outcome

Now, regress each outcome onto its respective (standardized) structural
disadvantage composite, the results can be found in
@tbl-adversityregressions.

**Step 1 Regressions**

```{r}
#| label: step1 regressions
# Model: Adversity → CES-D Depression 
model_depression <- svyglm(
  ZcesD_sum ~ Zcumadv_depression,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → Substance Use
model_substance <- svyglm(
  Zsubuse_W3 ~ Zcumadv_substance,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → Education Attainment
model_education <- svyglm(
  Zedattain_W3 ~ Zcumadv_education,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → General Health
model_health <- svyglm(
  Zgenhealth_W3 ~ Zcumadv_health,
  design = design_w3,
  family = gaussian() )
```

**Step 1 Table Creation**

```{r}
#| label: create step1 table
# create function to pull beta and p value from step 1 regressions
fmt_step1 <- function(model, var) {
  coefs <- summary(model)$coefficients
  b <- coefs[var, "Estimate"]
  p <- coefs[var, "Pr(>|t|)"]
  stars <- case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    p < .10  ~ "^",
    TRUE     ~ ""
  )
  paste0(formatC(b, digits = 2, format = "f"), stars)
}

# create dataframe for regression table
step1_tbl <- data.frame(
  Predictor     = "Cumulative Adversity",
  Depression    = fmt_step1(model_depression, "Zcumadv_depression"),
  Substance_Use = fmt_step1(model_substance,  "Zcumadv_substance"),
  Education     = fmt_step1(model_education,  "Zcumadv_education"),
  Health        = fmt_step1(model_health,     "Zcumadv_health"))

# create table for step 1 regressions
adversity_regressions <- flextable(step1_tbl) |>
  set_header_labels(
    Predictor     = "",
    Depression    = "Depression",
    Substance_Use = "Substance Use",
    Education     = "Education",
    Health        = "General Health"
  ) |>
  add_header_row(
    values    = c("", "Unstandardized Coefficients (b)"),
    colwidths = c(1, 4)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4) |>
  align(j = 2:5, align = "center", part = "all") |>
  add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

```{r}
#| label: tbl-adversityregressions
#| echo: false

adversity_regressions
```

Structural disadvantage significantly predicts all four Wave III
outcomes (*p* \< .001): higher disadvantage is associated with more
depressive symptoms (*b* = 0.15), more substance use (*b* = 0.22), lower
educational attainment (*b* = 0.48), and worse general health (*b* =
0.12). These results support the hypothesis higher disadvantage in
adolescence is associated with worse outcomes in young adulthood.. The
residuals from these models (the gap between predicted and observed
outcomes) become the resilience scores used in Step 3.

```{r}
#| label: save-adversityregressions-png
#| include: false

adversity_regressions_png <- adversity_regressions |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4)

save_as_image(adversity_regressions_png, path = "adversity_regressions.png", zoom = 2, expand = 0)
```

# Step 2: Residuals

For each outcome, the residual from its Step 1 model reflects how much
better or worse an young adult is doing than their adolescent
disadvantage level would predict. A positive residual is "better than
expected" (resilient) and a negative residual is "worse than expected"
(vulnerable). Depression and substance use residuals are reverse-coded
(`* -1`) so that, consistent with education and health, higher values
always indicate greater resilience.

```{r}
#| label: residuals
# ── Depression ─────────────────────────────────────────────────────────────────
dep_int    <- coef(model_depression)[1]
dep_Zcumadv <- coef(model_depression)[2]

dat_w3$predicted_depression <- dep_int + dep_Zcumadv * dat_w3$Zcumadv_depression
dat_w3$residual_depression  <- dat_w3$ZcesD_sum - dat_w3$predicted_depression
# flip: higher = more resilient
dat_w3$residual_depressionR <- dat_w3$residual_depression * -1  

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_int    <- coef(model_substance)[1]
substance_Zcumadv <- coef(model_substance)[2]

dat_w3$predicted_substance  <- substance_int + substance_Zcumadv * dat_w3$Zcumadv_substance
dat_w3$residual_substance   <- dat_w3$Zsubuse_W3 - dat_w3$predicted_substance
# flip: higher = more resilient
dat_w3$residual_substanceR  <- dat_w3$residual_substance * -1    

# ── Education ─────────────────────────────────────────────────────────────────
education_int    <- coef(model_education)[1]
education_Zcumadv <- coef(model_education)[2]

dat_w3$predicted_education <- education_int + education_Zcumadv * dat_w3$Zcumadv_education
dat_w3$residual_education  <- dat_w3$Zedattain_W3 - dat_w3$predicted_education

# ── Health ────────────────────────────────────────────────────────────────────
health_int    <- coef(model_health)[1]
health_Zcumadv <- coef(model_health)[2]

dat_w3$predicted_health <- health_int + health_Zcumadv * dat_w3$Zcumadv_health
dat_w3$residual_health  <- dat_w3$Zgenhealth_W3 - dat_w3$predicted_health

# rerun design with residual variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

# Step 3: Protective Factors

Now, regress the residual scores on each (standardized) protective
factor, resulting in results for depression (@tbl-step3-depression),
substance use (@tbl-step3-substance), education (@tbl-step3-education),
and health (@tbl-step3-health).

**Residual Regressions**

```{r}
#| label: step3 regressions
# ── Depression ─────────────────────────────────────────────────────────────────
depression_protect <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_protect <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
education_protect <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
health_protect <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )
```

**Create Step 3 Tables**

```{r}
#| label: create step3 tables
# create variable labels for protective factors
var_labels_step3 <- list(
  Zcommconnect       ~ "Community Connectedness",
  Zcommintegr_disadv ~ "Community Integration",
  Zopt_hopeful       ~ "Optimism / Hopefulness",
  Zselfesteem        ~ "Self-Esteem",
  Zself_efficacy     ~ "Self-Efficacy",
  Zparentchild_rel   ~ "Parent-Child Relationship",
  Zschoolconnect     ~ "School Connectedness",
  bio_sexD           ~ "Sex (0 = Male)",
  Zwave3age          ~ "Age")

# create function for table
fmt_tbl_regression <- function(model) {
  tbl_regression(model,
                 exponentiate = FALSE,
                 label = var_labels_step3) |>
    bold_p(t = .05) |>
    bold_labels() |>
    italicize_levels() |>
    modify_header(label = "**Variable**") |>
    modify_footnote(everything() ~ NA) |>
    as_flex_table() |>
    flextable::font(fontname = "Times New Roman", part = "all") |>
    flextable::fontsize(size = 12, part = "all") |>
    flextable::theme_apa()
}

# run function for each outcome to generate four tables
tbl_step3_depression <- fmt_tbl_regression(depression_protect)
tbl_step3_substance  <- fmt_tbl_regression(substance_protect)
tbl_step3_education  <- fmt_tbl_regression(education_protect)
tbl_step3_health     <- fmt_tbl_regression(health_protect)
```

```{r}
#| label: tbl-step3-depression
#| echo: false

tbl_step3_depression
```

Self-esteem shows the strongest association with depression resilience
(β = 0.12, *p* \< .001), followed by parent-child relationship (β =
0.07, *p* = .003) and optimism/hopefulness (β = 0.06, *p* = .002) —
supporting individual- and family-level factors as key predictors
(Hypothesis 1b). Community integration is the only factor with no
significant association.

```{r}
#| label: save-step3-depression-png
#| include: false

tbl_step3_depression_png <- tbl_step3_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_depression_png, path = paste0(fig_dir, "tbl_step3_depression.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-substance
#| echo: false

tbl_step3_substance
```

Parent-child relationship (β = 0.10, *p* \< .001) and sex (β = 0.43, *p*
\< .001) are the strongest predictors of substance use resilience.

```{r}
#| label: save-step3-substance-png
#| include: false

tbl_step3_substance_png <- tbl_step3_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_substance_png, path = paste0(fig_dir, "tbl_step3_substance.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-education
#| echo: false

tbl_step3_education
```

School connectedness (β = 0.12, *p* \< .001) and age (β = 0.27, *p* \<
.001) are the strongest predictors of educational resilience, with
community integration (β = 0.07, *p* \< .001) and self-efficacy (β =
0.08, *p* \< .001) also significant. Self-esteem and parent-child
relationship are the only factors not significantly associated with this
outcome.

```{r}
#| label: save-step3-education-png
#| include: false

tbl_step3_education_png <- tbl_step3_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_education_png, path = paste0(fig_dir, "tbl_step3_education.png"), zoom = 2, expand = 0)
```

```{r}
#| label: tbl-step3-health
#| echo: false

tbl_step3_health
```

Self-esteem (β = 0.14, *p* \< .001) and school connectedness (β = 0.11,
*p* \< .001) are the only individual- and school-level factors
significant for health resilience; the community-level factors
(connectedness, integration) are not significant here, unlike their role
in the depression and education models.

```{r}
#| label: save-step3-health-png
#| include: false

tbl_step3_health_png <- tbl_step3_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_health_png, path = paste0(fig_dir, "tbl_step3_health.png"), zoom = 2, expand = 0)
```

# Step 4: Protective vs. Promotive

Comparing the low- and high-disadvantage subsamples distinguishes two
types of effects: **promotive** factors are significant regardless of
disadvantage level (help everyone similarly), while **protective**
factors are significant specifically, or more strongly, in the
high-disadvantage subsample (buffer against disadvantages' effects). A
factor with a larger, more significant coefficient in the High
Disadvantage column than the Low Disadvantage column is acting
protectively; one with similar, significant coefficients across both
columns is acting promotively.

::: {.callout-tip appearance="simple" icon="false"}
Results can be found in the following tables: depression
(@tbl-compare-depression), substance use (@tbl-compare-substance),
education (@tbl-compare-education), and health (@tbl-compare-health).
:::

**Split sample into bottom third and top third of structural
disadvantage**

```{r}
#| label: disadvantage splits
# save outcome-specific adversity splits as variables in dat_w3
dat_w3$adversity_depression <- quantcut(dat_w3$Zcumadv_depression, c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_substance  <- quantcut(dat_w3$Zcumadv_substance,  c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_education  <- quantcut(dat_w3$Zcumadv_education,  c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_health     <- quantcut(dat_w3$Zcumadv_health,     c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))

# rerun design with split variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

**Run residual regressions with each sub sample, for each outcome.**

```{r}
#| label: step4 regressions

# ── Depression ─────────────────────────────────────────────────────────────────
model_depression_lo <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_depression == "lo"),
  family = gaussian() )

model_depression_hi <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_depression == "hi"),
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
model_substance_lo <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_substance == "lo"),
  family = gaussian() )

model_substance_hi <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_substance == "hi"),
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
model_education_lo <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_education == "lo"),
  family = gaussian() )

model_education_hi <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_education == "hi"),
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
model_health_lo <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_health == "lo"),
  family = gaussian() )

model_health_hi <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_health == "hi"),
  family = gaussian() )
```

**Create Step 4 Tables**

```{r}
#| label: table helper function
# create function to make comparison table
make_protect_table <- function(model_full, model_lo, model_hi, outcome_label) {
  
  protective_vars <- c(
    "Zcommconnect", "Zcommintegr_disadv",
    "Zopt_hopeful", "Zselfesteem", "Zself_efficacy",
    "Zparentchild_rel",
    "Zschoolconnect"
  )
  
  var_labels <- c(
    "Zcommconnect"       = "Community Connectedness",
    "Zcommintegr_disadv" = "Community Integration",
    "Zopt_hopeful"       = "Optimism / Hopefulness",
    "Zselfesteem"        = "Self-Esteem",
    "Zself_efficacy"     = "Self-Efficacy",
    "Zparentchild_rel"   = "Parent-Child Relationship",
    "Zschoolconnect"     = "School Connectedness"
  )
  
  fmt_coef <- function(model, var) {
    coefs <- summary(model)$coefficients
    if (!var %in% rownames(coefs)) return("—")
    b <- coefs[var, "Estimate"]
    p <- coefs[var, "Pr(>|t|)"]
    stars <- case_when(
      p < .001 ~ "***",
      p < .01  ~ "**",
      p < .05  ~ "*",
      p < .10  ~ "^",
      TRUE     ~ ""
    )
    paste0(formatC(b, digits = 2, format = "f"), stars)
  }
  
  tbl <- data.frame(
    Variable       = var_labels[protective_vars],
    Full_Sample    = sapply(protective_vars, \(v) fmt_coef(model_full, v)),
    Low_Adversity  = sapply(protective_vars, \(v) fmt_coef(model_lo,   v)),
    High_Adversity = sapply(protective_vars, \(v) fmt_coef(model_hi,   v)),
    row.names = NULL
  )
  
  flextable(tbl) |>
    set_header_labels(
      Variable       = "Protective Factor",
      Full_Sample    = "Full Sample",
      Low_Adversity  = "Low Adversity",
      High_Adversity = "High Adversity"
    ) |>
    add_header_row(
      values    = c("", outcome_label),
      colwidths = c(1, 3)
    ) |>
    theme_apa() |>
    font(fontname = "Times New Roman", part = "all") |>
    fontsize(size = 11, part = "all") |>
    bold(part = "header") |>
    width(j = 1, width = 2.2) |>
    width(j = 2:4, width = 1.5) |>
    align(j = 2:4, align = "center", part = "all") |>
    add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
    fontsize(size = 9, part = "footer") |>
    italic(part = "footer")
}

# run function for each outcome
tbl_compare_depression <- make_protect_table(
  depression_protect, model_depression_lo, model_depression_hi, "Depression")

tbl_compare_substance <- make_protect_table(
  substance_protect, model_substance_lo,  model_substance_hi, "Substance Use")

tbl_compare_education <- make_protect_table(
  education_protect, model_education_lo, model_education_hi, "Education Attainment")

tbl_compare_health <- make_protect_table(
  health_protect, model_health_lo, model_health_hi,   "General Health")
```

```{r}
#| label: tbl-compare-depression
#| echo: false

tbl_compare_depression
```

Community connectedness and self-efficacy show classic protective
patterns for depression: non-significant or weaker in the low-adversity
group but significant in the high-adversity group (β = 0.08 and 0.07,
respectively). Self-esteem and parent-child relationship are promotive,
remaining significant across both subsamples.

```{r}
#| label: save-compare-depression-png
#| include: false

tbl_compare_depression_png <- tbl_compare_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_depression_png, path = paste0(fig_dir, "tbl_compare_depression.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-substance
#| echo: false

tbl_compare_substance
```

Community connectedness is promotive for substance use, remaining
significant in both the low- and high-adversity groups (β = -0.07 and
-0.08). Parent-child relationship, by contrast, is protective —
non-significant in the low-adversity group (β = 0.02) but the strongest
predictor in the high-adversity group (β = 0.16, *p* \< .001).

```{r}
#| label: save-compare-substance-png
#| include: false

tbl_compare_substance_png <- tbl_compare_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_substance_png, path = paste0(fig_dir, "tbl_compare_substance.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-education
#| echo: false

tbl_compare_education
```

Community integration, self-efficacy, and school connectedness are
promotive for educational attainment, remaining significant across both
low- and high-adversity groups. School connectedness's effect is notably
stronger under high adversity (β = 0.15 vs. 0.10), suggesting it also
carries a protective component alongside its promotive one.

```{r}
#| label: save-compare-education-png
#| include: false

tbl_compare_education_png <- tbl_compare_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_education_png, path = paste0(fig_dir, "tbl_compare_education.png"), zoom = 3, expand = 0)
```

```{r}
#| label: tbl-compare-health
#| echo: false

tbl_compare_health
```

Optimism/hopefulness and community connectedness show protective
patterns for general health: both are non-significant in the
low-adversity group but emerge as significant in the high-adversity
group (β = 0.10 and 0.06, respectively). Self-esteem is promotive,
remaining significant and substantial in both groups (β = 0.21 and
0.09).

```{r}
#| label: save-compare-health-png
#| include: false

tbl_compare_health_png <- tbl_compare_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_health_png, path = paste0(fig_dir, "tbl_compare_health.png"), zoom = 3, expand = 0)
```
