---
title: "SPR Cup Analysis"
---

# Set Up

```{r}
#| label: load packages
library(gtsummary)   # regression summary tables
library(flextable)   # table formatting and export
library(gtools)      # quantile splitting 
library(haven)       # read SPSS .sav files
library(tidyverse)   # data wrangling
library(survey)      # complex survey / weighted regression
library(broom)       # tidy model output
library(gt)          # formatted tables
library(ggplot2)     # plots
library(patchwork)   # combine plots
```

The analytic dataset is the merged Add Health restricted-use survey data (`AddHealth_AllData.sav`), read in with `haven::read_sav()`. Import code is included but not rendered below.

```{r}
#| label: load data
#| echo: false
library(haven)
dat <- read_sav("C:/Users/rcm5550/OneDrive - The Pennsylvania State University/SPR Cup 2026/Analysis/AddHealth_AllData.sav")
```

## Data Cleaning (SPSS)

Variable recoding, dummy coding, and composite construction for this project were done in SPSS prior to import into R, across four syntax files by domain (demographics, structural disadvantage, outcomes, protective factors). Each is excerpted below to document the approach.

### Demographics

`Demographics.sps` recodes core demographic variables (child sex, age, race/ethnicity) into analysis-ready measures, checking frequencies before and after each recode.

```spss
*****************************CHILD SEX.
*Q: What is your biological sex?
***check frequencies for zeros.
frequencies variables = bio_sex /statistics = all.
*male = 1, female = 2.

*create dummy code for child biological sex (female = 1).
if(bio_sex eq 1) bio_sexD = 0.
if(bio_sex eq 2) bio_sexD = 1.
variable labels bio_sexD
    "child biological sex".
value labels bio_sexD
    0 "male"
    1 "female".
execute.

*check frequencies of new variable.
frequencies variables = bio_sexD /statistics = all.
```

### Structural Disadvantage

`Disadvantage.sps` (672 lines) builds individual- and household-level disadvantage indicators, including race/ethnicity dummy codes with missing-data recodes.

```spss

*****************************FAMILY SOCIOECONOMIC DISADVANTAGE COMPOSITE.
*Components: famincomecat (continuous $, higher = more income), pareduc_avg (0-8, higher = more education),
*            parentemploy (0-1, higher = more employment).
*Standardizing each component to z-score then reverse code so higher is more disadvantage, then average.

*standardize based on weight filtered dataset.
delete variables  Zfamincomecat  Zpareduc_avg Zparentemploy.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = famincomecat pareduc_avg parentemploy /save.
*Creates Zfamincomecat, Zpareduc_avg, Zparentemploy.
execute.

compute Zfamincomecat_r = Zfamincomecat * -1.
variable labels Zfamincomecat_r
    "Family income z-scored reversed (higher = lower income = more disadvantage)".
execute.

compute Zpareduc_avg_r = Zpareduc_avg * -1.
variable labels Zpareduc_avg_r
    "Average parental education z-scored reversed (higher = less education = more disadvantage)".
execute.

compute Zparentemploy_r = Zparentemploy * -1.
variable labels Zparentemploy_r
    "Average parental employment z-scored reversed (higher = less employed = more disadvantage)".
execute.

compute famSES_disadv = mean(Zfamincomecat_r, Zpareduc_avg_r, Zparentemploy_r).
variable labels famSES_disadv
    "Family SES disadvantage composite: mean of z-scored reversed income, education, employment (higher = more disadvantage)".
execute.

frequencies variables = famSES_disadv /statistics = all.

reliability variables = famincomecat pareduc_avg parentemploy
    /scale('famSES') all /statistics = all.

*standardize based on weight filtered dataset.
temporary.
select if (GSWGT3_2 ge 0). 
descriptives variables = famSES_disadv /save.
*Creates Zfamincomecat, Zpareduc_avg, Zparentemploy.
execute.
```

### Outcomes

`Outcomes.sps` (473 lines) constructs the substance-use outcome variable (alcohol, cigarette, and tobacco use composite), depression, general health, and educational attainment. Below is a exert from the final section of scoring the depression outcome.

```spss
*****************************CES-D DEPRESSION COMPOSITE (WAVE 3).
*9 items (all in same 0-3 direction after reverse coding H3SP7 and H3SP11 above).
*Sum score (0-27 range); Desch et al. (2023).

compute cesD_sum = sum(dep_bothered, dep_blues, dep_asgood, dep_concentrate, dep_depressed,
                       dep_tired, dep_enjoyed, dep_sad, dep_disliked).
variable labels cesD_sum
    "CES-D depression sum score Wave 3 (0-27, higher = more depressive symptoms)".
execute.

reliability variables = dep_bothered dep_blues dep_asgood dep_concentrate dep_depressed dep_tired dep_enjoyed dep_sad dep_disliked
    /scale('CESD') all /statistics = all.

frequencies variables = cesD_sum /statistics = all.

*standardize based on weight filtered dataset.
delete variables  ZcesD_sum.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = cesD_sum / save.
```

### Protective Factors

`Protective.sps` (887 lines) builds the protective-factor composites, several of which are reverse-scored so that higher values consistently indicate greater disadvantage.Below is an example of the data cleaning of one variable and then the creation of one protective factor composite.

```spss
*****************************COMMUNITY RESOURCES - CONNECTEDNESS (ADOLESCENT REPORT).

*Note: Items reverse-scored here so higher = LOWER community resources (disadvantage direction),
*consistent with the structural disadvantage framing of this composite.
*Community Connectedness (Wickrama & Bryant 2003): 4 items, yes/no (0/1), summed 0-4.
*Reverse scoring: recode 1->0 and 0->1 so high score = LOW connectedness = more disadvantage.

if (H1NB1 eq 1) knowneighbR = 0.
    *knows neighbors = connected (reverse to 0=low disadvantage).
if (H1NB1 eq 2) knowneighbR = 1.
    *doesn't know = disconnected = higher disadvantage.
if (H1NB1 eq 6 or H1NB1 eq 8 or H1NB1 eq 9) knowneighbR = -99.
recode knowneighbR (-99 = sysmis).
variable labels knowneighbR
    "Does NOT know most neighbors (H1NB1 reversed, 1=low connectedness/more disadvantage)".
value labels knowneighbR 0 "knows most neighbors" 1 "does not know most neighbors".
execute.
    
*****************************COMMUNITY CONNECTEDNESS INDEX (Wickrama & Bryant 2003).
*Items coded in protective direction (1 = connected):
*knowneighb, talkneighb, neighbwatch, nbhdsafe.
*Sum 0-4; higher = more connectedness = more protective.

compute commconnect = sum(knowneighb, talkneighb, neighbwatch, nbhdsafe).
variable labels commconnect
    "Community connectedness (Wickrama & Bryant 2003, sum 0-4, higher = more connectedness)".
execute.

reliability variables = knowneighb talkneighb neighbwatch nbhdsafe
    /scale('commconnect') all /statistics = all.
*alpha = .51.

frequencies variables = commconnect /statistics = all.


*standardize based on weight filtered dataset.
delete variables  Zcommconnect.
    *delete variable if previously standardized with whole sample. 
temporary.
    *temporary can't be used with delete.
select if (GSWGT3_2 ge 0). 
descriptives variables = commconnect / save.
*******************************************************************************
```

The cleaned file was exported and merged into `AddHealth_AllData.sav` for the R analysis above.

## Overview

-   **Step 0: Create cumulative risk variable.**
    -   Approach: regress each outcome on all individual adversity indicators simultaneously. Extract standardized betas. Use betas as weights to construct an outcome-specific adversity composite.
        -   Citation: Utilize predictor weighting — meaning the contribution of each adversity indicator to the cumulative composite is weighted by its empirical association with the outcome. Marquez et al. (2023) utilized this method; Miller-Lewis et al. (2013) examined bivariate correlations with each outcome and found them all small, so created one composite adversity index.
-   **Step 1: Regress cumulative risk on outcome**
-   **Step 2: Extract residuals**
-   **Step 3: Regress protective factors on residuals**
-   **Step 4: Repeat step 2-4 with high and low adversity sample split**

## Survey Design Specification

Add Health uses a complex, clustered, stratified sampling design. All regressions use the Wave 3 grand sample weight (`GSWGT3_2_2`) via the `survey` package to produce design-consistent estimates.

```{r}
#| label: check missing sample weight

#how many cases missing
cat("Missing GSWGT3_2:", sum(is.na(dat$GSWGT3_2)), "\n")
cat("Non-missing GSWGT3_2:", sum(!is.na(dat$GSWGT3_2)), "\n")

# ── Subset to respondents with valid Wave 3 weight ────────────────────────────
# Respondents missing GSWGT3_2 were not part of the Wave 3 sample
# (e.g., lost to follow-up, deceased, ineligible) and should be excluded
dat_w3 <- dat |> filter(!is.na(GSWGT3_2))  ###dat_w3 is name of new dataset for analysis

cat("Analytic n after excluding missing weights:", nrow(dat_w3), "\n")

# ── Specify survey design on the subsetted data ───────────────────────────────
# Use design_w3 as survey weight design in regressions
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3  
)

cat("Sum of weights:", round(sum(dat_w3$GSWGT3_2), 0), "\n")
```

## Descriptive Statistics

### Code

```{r}
#| label: create descriptives table

# ── Variable lists with display labels ────────────────────────────────────────
adversity_vars <- c(
  "minorityD"          = "Minority Status",
  "school_disadv"     = "School Disadvantage",
  "econhardship"      = "Economic Hardship",
  "famSES_disadv"      = "Family SES Disadvantage",
  "nbhd_subj_disadv"  = "Neighborhood Disadvantage (Subjective)",
  "cens_nbhd_disadv"  = "Neighborhood Disadvantage (Census)"
)

protective_vars <- c(
  "commconnect"       = "Community Connectedness",
  "commintegr_disadv" = "Community Integration",
  "opt_hopeful"       = "Optimism / Hopefulness",
  "selfesteem"        = "Self-Esteem",
  "self_efficacy"     = "Self-Efficacy",
  "parentchild_rel"   = "Parent-Child Relationship",
  "schoolconnect"     = "School Connectedness"
)

outcome_vars <- c(
  "cesD_sum"    = "Depression Symptoms",
  "subuse_W3"   = "Substance Use",
  "edattain_W3" = "Education Attainment",
  "genhealth_W3"= "General Health"
)

# ── Function: get weighted M and SD, and unweighted range ─────────────────────
get_desc <- function(var, design) {
  mn  <- as.numeric(svymean(as.formula(paste0("~", var)), design, na.rm = TRUE))
  vr  <- as.numeric(svyvar(as.formula(paste0("~", var)),  design, na.rm = TRUE))
  rng <- range(design$variables[[var]], na.rm = TRUE)
  data.frame(
    M     = formatC(mn,       digits = 2, format = "f"),
    SD    = formatC(sqrt(vr), digits = 2, format = "f"),
    Range = paste0(formatC(rng[1], digits = 2, format = "f"),
                   "–",
                   formatC(rng[2], digits = 2, format = "f"))
  )
}

# ── Build rows for each variable group ────────────────────────────────────────
make_desc_rows <- function(var_named_vec, design, section_label) {
  rows <- lapply(names(var_named_vec), function(v) {
    d <- get_desc(v, design)
    data.frame(Variable = var_named_vec[[v]], M = d$M, SD = d$SD, Range = d$Range)
  })
  rows <- do.call(rbind, rows)
  rbind(data.frame(Variable = section_label, M = "", SD = "", Range = ""), rows)
}

desc_tbl <- rbind(
  make_desc_rows(adversity_vars,  design_w3, "Adversity Predictors"),
  make_desc_rows(protective_vars, design_w3, "Protective Factors"),
  make_desc_rows(outcome_vars,    design_w3, "Outcomes")
)

# ── Build flextable ───────────────────────────────────────────────────────────
header_rows <- which(desc_tbl$M == "")

header_rows <- which(desc_tbl$M == "")

descriptives <- flextable(desc_tbl) |>
  set_header_labels(Variable = "Variable", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Weighted Descriptive Statistics"),
    colwidths = c(1, 3)                                        # was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  bold(i = header_rows, part = "body") |>
  width(j = 1, width = 3.0) |>
  width(j = 2:4, width = 1.0) |>                              # was 2:3
  align(j = 2:4, align = "center", part = "all") |>           # was 2:3
  padding(i = setdiff(seq_len(nrow(desc_tbl)), header_rows),
          j = 1, padding.left = 20, part = "body") |>
  add_footer_lines("Note. Means and standard deviations are weighted using the Wave 3 grand sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

## Table: Descriptive Statistics

@tbl-descriptives

```{r}
#| label: tbl-descriptives

descriptives
```

```{r}
#| label: save-descriptives-png

descriptives_png <- descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 3.2) |>
  width(j = 2:4, width = 1.05)

save_as_image(descriptives_png, path = "descriptives.png", zoom = 3, expand = 0)
```

# Step 0: Adversity Composites

## Individual Disadvantage Regressions

### Depression

```{r}
#| label: depression weights - minority

# run regression
depression_minorityD <- svyglm(
  ZcesD_sum ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_minorityD <- coef(depression_minorityD)[2]
dep_minorityD_B <- coef(depression_minorityD)[2]
dep_minorityD_P <-summary(depression_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

```{r}
#| label: depression weights - school

# run regression
depression_school_disadv <- svyglm(
  ZcesD_sum ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_school <- coef(depression_school_disadv)[2]
dep_school_B <- coef(depression_school_disadv)[2]
dep_school_P <-summary(depression_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

```{r}
#| label: depression weights - econ

# run regression
depression_econhardship <- svyglm(
  ZcesD_sum ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_econ <- coef(depression_econhardship)[2]
dep_econ_B <- coef(depression_econhardship)[2]
dep_econ_P <-summary(depression_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

```{r}
#| label: depression weights - fam

# run regression
depression_famSES_disadv <- svyglm(
  ZcesD_sum ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_fam <- coef(depression_famSES_disadv)[2]
dep_famSES_B <- coef(depression_famSES_disadv)[2]
dep_famSES_P <-summary(depression_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

```{r}
#| label: depression weights - nbhdsub

# run regression
depression_nbhd_subj_disadv <- svyglm(
  ZcesD_sum ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_nbhdsub <- coef(depression_nbhd_subj_disadv)[2]
dep_nbhdsub_B <- coef(depression_nbhd_subj_disadv)[2]
dep_nbhdsub_P <-summary(depression_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

```{r}
#| label: depression weights - cens nbhd

# run regression
depression_cens_nbhd_disadv <- svyglm(
  ZcesD_sum ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$dep_censnbhd <- coef(depression_cens_nbhd_disadv)[2]
dep_censnbhd_B <- coef(depression_cens_nbhd_disadv)[2]
dep_censnbhd_P <-summary(depression_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

### Substance Use

```{r}
#| label: substance use weights - minority

# run regression
substance_minorityD <- svyglm(
  Zsubuse_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_minorityD <- coef(substance_minorityD)[2]
substance_minorityD_B <- coef(substance_minorityD)[2]
substance_minorityD_P <-summary(substance_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

```{r}
#| label: substance use weights - school

# run regression
substance_school_disadv <- svyglm(
  Zsubuse_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_school <- coef(substance_school_disadv)[2]
substance_school_B <- coef(substance_school_disadv)[2]
substance_school_P <-summary(substance_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

```{r}
#| label: substance use weights - econ

# run regression
substance_econhardship <- svyglm(
  Zsubuse_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_econ <- coef(substance_econhardship)[2]
substance_econ_B <- coef(substance_econhardship)[2]
substance_econ_P <-summary(substance_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

```{r}
#| label: substance use weights - fam

# run regression
substance_famSES_disadv <- svyglm(
  Zsubuse_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_fam <- coef(substance_famSES_disadv)[2]
substance_famSES_B <- coef(substance_famSES_disadv)[2]
substance_famSES_P <-summary(substance_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

```{r}
#| label: substance use weights - nbhdsub

# run regression
substance_nbhd_subj_disadv <- svyglm(
  Zsubuse_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_nbhdsub <- coef(substance_nbhd_subj_disadv)[2]
substance_nbhdsub_B <- coef(substance_nbhd_subj_disadv)[2]
substance_nbhdsub_P <-summary(substance_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

```{r}
#| label: substance use weights - cens nbhd

# run regression
substance_cens_nbhd_disadv <- svyglm(
  Zsubuse_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$substance_censnbhd <- coef(substance_cens_nbhd_disadv)[2]
substance_censnbhd_B <- coef(substance_cens_nbhd_disadv)[2]
substance_censnbhd_P <-summary(substance_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

### Education Attainment

```{r}
#| label: education weights - minority

# run regression
education_minorityD <- svyglm(
  Zedattain_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_minorityD <- coef(education_minorityD)[2]
education_minorityD_B <- coef(education_minorityD)[2]
education_minorityD_P <- summary(education_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

```{r}
#| label: education weights - school

# run regression
education_school_disadv <- svyglm(
  Zedattain_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_school <- coef(education_school_disadv)[2]
education_school_B <- coef(education_school_disadv)[2]
education_school_P <- summary(education_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

```{r}
#| label: education weights - econ

# run regression
education_econhardship <- svyglm(
  Zedattain_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_econ <- coef(education_econhardship)[2]
education_econ_B <- coef(education_econhardship)[2]
education_econ_P <- summary(education_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

```{r}
#| label: education weights - fam

# run regression
education_famSES_disadv <- svyglm(
  Zedattain_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_fam <- coef(education_famSES_disadv)[2]
education_famSES_B <- coef(education_famSES_disadv)[2]
education_famSES_P <- summary(education_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

```{r}
#| label: education weights - nbhdsub

# run regression
education_nbhd_subj_disadv <- svyglm(
  Zedattain_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_nbhdsub <- coef(education_nbhd_subj_disadv)[2]
education_nbhdsub_B <- coef(education_nbhd_subj_disadv)[2]
education_nbhdsub_P <- summary(education_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

```{r}
#| label: education weights - cens nbhd

# run regression
education_cens_nbhd_disadv <- svyglm(
  Zedattain_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$education_censnbhd <- coef(education_cens_nbhd_disadv)[2]
education_censnbhd_B <- coef(education_cens_nbhd_disadv)[2]
education_censnbhd_P <- summary(education_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

### General Health

```{r}
#| label: health weights - minority

# run regression
health_minorityD <- svyglm(
  Zgenhealth_W3 ~ minorityD,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_minorityD <- coef(health_minorityD)[2]
health_minorityD_B <- coef(health_minorityD)[2]
health_minorityD_P <- summary(health_minorityD)$coefficients["minorityD", "Pr(>|t|)"]
```

```{r}
#| label: health weights - school

# run regression
health_school_disadv <- svyglm(
  Zgenhealth_W3 ~ Zschool_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_school <- coef(health_school_disadv)[2]
health_school_B <- coef(health_school_disadv)[2]
health_school_P <- summary(health_school_disadv)$coefficients["Zschool_disadv", "Pr(>|t|)"]
```

```{r}
#| label: health weights - econ

# run regression
health_econhardship <- svyglm(
  Zgenhealth_W3 ~ Zeconhardship,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_econ <- coef(health_econhardship)[2]
health_econ_B <- coef(health_econhardship)[2]
health_econ_P <- summary(health_econhardship)$coefficients["Zeconhardship", "Pr(>|t|)"]
```

```{r}
#| label: health weights - fam

# run regression
health_famSES_disadv <- svyglm(
  Zgenhealth_W3 ~ famSES_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_fam <- coef(health_famSES_disadv)[2]
health_famSES_B <- coef(health_famSES_disadv)[2]
health_famSES_P <- summary(health_famSES_disadv)$coefficients["famSES_disadv", "Pr(>|t|)"]
```

```{r}
#| label: health weights - nbhdsub

# run regression
health_nbhd_subj_disadv <- svyglm(
  Zgenhealth_W3 ~ Znbhd_subj_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_nbhdsub <- coef(health_nbhd_subj_disadv)[2]
health_nbhdsub_B <- coef(health_nbhd_subj_disadv)[2]
health_nbhdsub_P <- summary(health_nbhd_subj_disadv)$coefficients["Znbhd_subj_disadv", "Pr(>|t|)"]
```

```{r}
#| label: health weights - cens nbhd

# run regression
health_cens_nbhd_disadv <- svyglm(
  Zgenhealth_W3 ~ Zcens_nbhd_disadv,
  design = design_w3,
  family = gaussian() )

#save beta and p value
dat_w3$health_censnbhd <- coef(health_cens_nbhd_disadv)[2]
health_censnbhd_B <- coef(health_cens_nbhd_disadv)[2]
health_censnbhd_P <- summary(health_cens_nbhd_disadv)$coefficients["Zcens_nbhd_disadv", "Pr(>|t|)"]
```

## Adversity Weights

### Code

```{r}
#| label: create adversity weight table

weight_summary <- data.frame(
  Predictor = c("Minority Status", "School Disadvantage", "Economic Hardship",
                "Family SES", "Neighborhood (Subj.)", "Neighborhood (Census)"),

  dep_B  = round(c(dep_minorityD_B, dep_school_B, dep_econ_B,
                   dep_famSES_B, dep_nbhdsub_B, dep_censnbhd_B), 2),
  dep_P  = round(c(dep_minorityD_P, dep_school_P, dep_econ_P,
                   dep_famSES_P, dep_nbhdsub_P, dep_censnbhd_P), 2),

  sub_B  = round(c(substance_minorityD_B, substance_school_B, substance_econ_B,
                   substance_famSES_B, substance_nbhdsub_B, substance_censnbhd_B), 2),
  sub_P  = round(c(substance_minorityD_P, substance_school_P, substance_econ_P,
                   substance_famSES_P, substance_nbhdsub_P, substance_censnbhd_P), 2),

  edu_B  = round(c(education_minorityD_B, education_school_B, education_econ_B,
                   education_famSES_B, education_nbhdsub_B, education_censnbhd_B), 2),
  edu_P  = round(c(education_minorityD_P, education_school_P, education_econ_P,
                   education_famSES_P, education_nbhdsub_P, education_censnbhd_P), 2),

  hlt_B  = round(c(health_minorityD_B, health_school_B, health_econ_B,
                   health_famSES_B, health_nbhdsub_B, health_censnbhd_B), 2),
  hlt_P  = round(c(health_minorityD_P, health_school_P, health_econ_P,
                   health_famSES_P, health_nbhdsub_P, health_censnbhd_P), 2)
)

table_weights <- flextable(weight_summary) |>
  set_header_labels(
    Predictor = "Predictor",
    dep_B = "b", dep_P = "p",
    sub_B = "b", sub_P = "p",
    edu_B = "b", edu_P = "p",
    hlt_B = "b", hlt_P = "p"
  ) |>
  add_header_row(
    values     = c("", "Depression", "Substance Use", "Education", "Health"),
    colwidths  = c(1, 2, 2, 2, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 1.5) |>
  width(j = 2:9, width = .6) |>
  align(j = 2:9, align = "center", part = "all") |>
  add_footer_lines("Note. All continuous predictors and outocmes were standardized. Each predictor was standardized and then regressed on the outcome separately.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

## Table: Adversity Predictor Weights

@tbl-adversity-weight-table

```{r}
#| label: tbl-adversity-weight-table

table_weights
```

```{r}
#| label: save-adversity-weight-table-png

table_weights_png <- table_weights |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  set_header_labels(dep_B = "Β", dep_P = "p", sub_B = "Β", sub_P = "p",
                    edu_B = "Β", edu_P = "p", hlt_B = "Β", hlt_P = "p") |>
  italic(i = 2, j = c(3, 5, 7, 9), part = "header") |>
  width(j = 1, width = 1.6) |>
  width(j = 2:9, width = 0.6)

save_as_image(table_weights_png, path = "table_weights.png", zoom = 3, expand = 0)
```

## Cumulative Adversity Composites

### Code

```{r}
#| label: cumulative composites

#compute weighted composites for all four outcomes
dat_w3 <- dat_w3 %>%
  mutate(
    cumadv_depression = (
      (minorityD)         * dep_minorityD +
      (Zschool_disadv)    * dep_school +
      (Zeconhardship)     * dep_econ +
      (famSES_disadv)     * dep_fam +
      (Znbhd_subj_disadv) * dep_nbhdsub +
      (Zcens_nbhd_disadv) * dep_censnbhd),

    cumadv_substance = (
      (minorityD)         * substance_minorityD +
      (Zschool_disadv)    * substance_school +
      (Zeconhardship)     * substance_econ +
      (famSES_disadv)     * substance_fam +
      (Znbhd_subj_disadv) * substance_nbhdsub +
      (Zcens_nbhd_disadv) * substance_censnbhd),

    cumadv_education = (
      (minorityD)         * education_minorityD +
      (Zschool_disadv)    * education_school +
      (Zeconhardship)     * education_econ +
      (famSES_disadv)     * education_fam +
      (Znbhd_subj_disadv) * education_nbhdsub +
      (Zcens_nbhd_disadv) * education_censnbhd),

    cumadv_health = (
      (minorityD)         * health_minorityD +
      (Zschool_disadv)    * health_school +
      (Zeconhardship)     * health_econ +
      (famSES_disadv)     * health_fam +
      (Znbhd_subj_disadv) * health_nbhdsub +
      (Zcens_nbhd_disadv) * health_censnbhd)
  )

dat_w3 <- dat_w3 %>%
  mutate(
    Zcumadv_depression = as.numeric(scale(cumadv_depression)),
    Zcumadv_substance  = as.numeric(scale(cumadv_substance)),
    Zcumadv_education  = as.numeric(scale(cumadv_education)),
    Zcumadv_health     = as.numeric(scale(cumadv_health))
  )

# rerun design with composite variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

```{r}
#| label: create cumulative adversity descriptives

cumadv_vars <- c(
  "cumadv_depression" = "Depression",
  "cumadv_substance"  = "Substance Use",
  "cumadv_education"  = "Education Attainment",
  "cumadv_health"     = "General Health"
)

cumadv_rows <- do.call(rbind, lapply(names(cumadv_vars), function(v) {
  d <- get_desc(v, design_w3)
  data.frame(Variable = cumadv_vars[[v]], M = d$M, SD = d$SD, Range = d$Range)
}))

cumulative_descriptives <- flextable(cumadv_rows) |>
  set_header_labels(Variable = "Outcome", M = "M", SD = "SD", Range = "Range") |>
  add_header_row(
    values    = c("", "Cumulative Adversity Composites"),
    colwidths = c(1, 3)                              # was c(1, 2)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 1.0) |>                    # was 2:3
  align(j = 2:4, align = "center", part = "all") |> # was 2:3
  add_footer_lines("Note. Each cumulative adversity composite is outcome-specific, weighted by the association of each adversity indicator with that outcome. Descriptive statistics are survey-weighted using the Wave 3 sample weight (GSWGT3_2). Range is unweighted.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer") |>
  fit_to_width(max_width = 6.5)
```

## Table: Cumulative Adversity Descriptives

@tbl-cumulative

```{r}
#| label: tbl-cumulative

cumulative_descriptives
```

```{r}
#| label: save-cumulative-png

cumulative_descriptives_png <- cumulative_descriptives |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  italic(i = 2, j = 2:3, part = "header") |>
  width(j = 1, width = 2.5) |>
  width(j = 2:4, width = 0.9)

save_as_image(cumulative_descriptives_png, path = "cumulative_descriptives.png", zoom = 3, expand = 0)
```

# Step 1: Cumulative Adversity to Outcome

### Code

```{r}
#| label: step1 regressions

# Model: Adversity → CES-D Depression 
model_depression <- svyglm(
  ZcesD_sum ~ Zcumadv_depression,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → Substance Use
model_substance <- svyglm(
  Zsubuse_W3 ~ Zcumadv_substance,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → Education Attainment
model_education <- svyglm(
  Zedattain_W3 ~ Zcumadv_education,
  design = design_w3,
  family = gaussian() )

# Model: Adversity → General Health
model_health <- svyglm(
  Zgenhealth_W3 ~ Zcumadv_health,
  design = design_w3,
  family = gaussian() )
```

```{r}
#| label: create step1 table

fmt_step1 <- function(model, var) {
  coefs <- summary(model)$coefficients
  b <- coefs[var, "Estimate"]
  p <- coefs[var, "Pr(>|t|)"]
  stars <- case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    p < .10  ~ "^",
    TRUE     ~ ""
  )
  paste0(formatC(b, digits = 2, format = "f"), stars)
}

step1_tbl <- data.frame(
  Predictor     = "Cumulative Adversity",
  Depression    = fmt_step1(model_depression, "Zcumadv_depression"),
  Substance_Use = fmt_step1(model_substance,  "Zcumadv_substance"),
  Education     = fmt_step1(model_education,  "Zcumadv_education"),
  Health        = fmt_step1(model_health,     "Zcumadv_health")
)

adversity_regressions <- flextable(step1_tbl) |>
  set_header_labels(
    Predictor     = "",
    Depression    = "Depression",
    Substance_Use = "Substance Use",
    Education     = "Education",
    Health        = "General Health"
  ) |>
  add_header_row(
    values    = c("", "Unstandardized Coefficients (b)"),
    colwidths = c(1, 4)
  ) |>
  theme_apa() |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 11, part = "all") |>
  bold(part = "header") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4) |>
  align(j = 2:5, align = "center", part = "all") |>
  add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

## Table: Cumulative Adversity Predicting Outcomes

@tbl-adversityregressions

```{r}
#| label: tbl-adversityregressions

adversity_regressions
```

```{r}
#| label: save-adversityregressions-png

adversity_regressions_png <- adversity_regressions |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:5, width = 1.4)

save_as_image(adversity_regressions_png, path = "adversity_regressions.png", zoom = 2, expand = 0)
```

# Step 2: Residuals

```{r}
#| label: residuals

# ── Depression ─────────────────────────────────────────────────────────────────
dep_int    <- coef(model_depression)[1]
dep_Zcumadv <- coef(model_depression)[2]

dat_w3$predicted_depression <- dep_int + dep_Zcumadv * dat_w3$Zcumadv_depression
dat_w3$residual_depression  <- dat_w3$ZcesD_sum - dat_w3$predicted_depression
dat_w3$residual_depressionR <- dat_w3$residual_depression * -1   # flip: higher = more resilient

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_int    <- coef(model_substance)[1]
substance_Zcumadv <- coef(model_substance)[2]

dat_w3$predicted_substance  <- substance_int + substance_Zcumadv * dat_w3$Zcumadv_substance
dat_w3$residual_substance   <- dat_w3$Zsubuse_W3 - dat_w3$predicted_substance
dat_w3$residual_substanceR  <- dat_w3$residual_substance * -1    # flip: higher = more resilient

# ── Education (no flip: higher score = better outcome) ────────────────────────
education_int    <- coef(model_education)[1]
education_Zcumadv <- coef(model_education)[2]

dat_w3$predicted_education <- education_int + education_Zcumadv * dat_w3$Zcumadv_education
dat_w3$residual_education  <- dat_w3$Zedattain_W3 - dat_w3$predicted_education

# ── Health (verify scale direction before deciding whether to flip) ────────────
health_int    <- coef(model_health)[1]
health_Zcumadv <- coef(model_health)[2]

dat_w3$predicted_health <- health_int + health_Zcumadv * dat_w3$Zcumadv_health
dat_w3$residual_health  <- dat_w3$Zgenhealth_W3 - dat_w3$predicted_health

# rerun design with residual variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

# Step 3: Protective Factors

### Code

```{r}
#| label: step3 regressions

# ── Depression ─────────────────────────────────────────────────────────────────
depression_protect <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
substance_protect <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
education_protect <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect +
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
health_protect <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel +
    Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = design_w3,
  family = gaussian() )
```

```{r}
#| label: create step3 tables

var_labels_step3 <- list(
  Zcommconnect       ~ "Community Connectedness",
  Zcommintegr_disadv ~ "Community Integration",
  Zopt_hopeful       ~ "Optimism / Hopefulness",
  Zselfesteem        ~ "Self-Esteem",
  Zself_efficacy     ~ "Self-Efficacy",
  Zparentchild_rel   ~ "Parent-Child Relationship",
  Zschoolconnect     ~ "School Connectedness",
  bio_sexD           ~ "Sex (0 = Male)",
  Zwave3age          ~ "Age"
)

fmt_tbl_regression <- function(model) {
  tbl_regression(model,
                 exponentiate = FALSE,
                 label = var_labels_step3) |>
    bold_p(t = .05) |>
    bold_labels() |>
    italicize_levels() |>
    modify_header(label = "**Variable**") |>
    modify_footnote(everything() ~ NA) |>
    as_flex_table() |>
    flextable::font(fontname = "Times New Roman", part = "all") |>
    flextable::fontsize(size = 12, part = "all") |>
    flextable::theme_apa()
}

tbl_step3_depression <- fmt_tbl_regression(depression_protect)
tbl_step3_substance  <- fmt_tbl_regression(substance_protect)
tbl_step3_education  <- fmt_tbl_regression(education_protect)
tbl_step3_health     <- fmt_tbl_regression(health_protect)
```

## Table: Protective Factors — Depression

@tbl-step3-depression

```{r}
#| label: tbl-step3-depression

tbl_step3_depression
```

```{r}
#| label: save-step3-depression-png

tbl_step3_depression_png <- tbl_step3_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_depression_png, path = "tbl_step3_depression.png", zoom = 2, expand = 0)
```

## Table: Protective Factors — Substance Use

@tbl-step3-substance

```{r}
#| label: tbl-step3-substance

tbl_step3_substance
```

```{r}
#| label: save-step3-substance-png

tbl_step3_substance_png <- tbl_step3_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_substance_png, path = "tbl_step3_substance.png", zoom = 2, expand = 0)
```

## Table: Protective Factors — Education

@tbl-step3-education

```{r}
#| label: tbl-step3-education

tbl_step3_education
```

```{r}
#| label: save-step3-education-png

tbl_step3_education_png <- tbl_step3_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_education_png, path = "tbl_step3_education.png", zoom = 2, expand = 0)
```

## Table: Protective Factors — General Health

@tbl-step3-health

```{r}
#| label: tbl-step3-health

tbl_step3_health
```

```{r}
#| label: save-step3-health-png

tbl_step3_health_png <- tbl_step3_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 18, part = "body") |>
  fontsize(size = 20, part = "header") |>
  fontsize(size = 13, part = "footer") |>
  fit_to_width(max_width = 7.3)

save_as_image(tbl_step3_health_png, path = "tbl_step3_health.png", zoom = 2, expand = 0)
```

# Step 4: Protective vs. Promotive

### Code

```{r}
#| label: adversity splits

# save outcome-specific adversity splits as variables in dat_w3
dat_w3$adversity_depression <- quantcut(dat_w3$Zcumadv_depression, c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_substance  <- quantcut(dat_w3$Zcumadv_substance,  c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_education  <- quantcut(dat_w3$Zcumadv_education,  c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))
dat_w3$adversity_health     <- quantcut(dat_w3$Zcumadv_health,     c(0, .33, .66, 1), lab = c("lo", "mid", "hi"))

# rerun design with split variables added
design_w3 <- svydesign(
  ids     = ~1,
  weights = ~GSWGT3_2,
  data    = dat_w3)
```

```{r}
#| label: step4 regressions

# ── Depression ─────────────────────────────────────────────────────────────────
model_depression_lo <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_depression == "lo"),
  family = gaussian() )

model_depression_hi <- svyglm(
  residual_depressionR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_depression == "hi"),
  family = gaussian() )

# ── Substance Use ──────────────────────────────────────────────────────────────
model_substance_lo <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_substance == "lo"),
  family = gaussian() )

model_substance_hi <- svyglm(
  residual_substanceR ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_substance == "hi"),
  family = gaussian() )

# ── Education ──────────────────────────────────────────────────────────────────
model_education_lo <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect + 
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_education == "lo"),
  family = gaussian() )

model_education_hi <- svyglm(
  residual_education ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_education == "hi"),
  family = gaussian() )

# ── Health ─────────────────────────────────────────────────────────────────────
model_health_lo <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_health == "lo"),
  family = gaussian() )

model_health_hi <- svyglm(
  residual_health ~ 
    Zcommconnect + Zcommintegr_disadv +
    Zopt_hopeful + Zselfesteem + Zself_efficacy +
    Zparentchild_rel + Zschoolconnect +
    bio_sexD + Zwave3age,
  design = subset(design_w3, adversity_health == "hi"),
  family = gaussian() )
```

```{r}
#| label: table helper function

make_protect_table <- function(model_full, model_lo, model_hi, outcome_label) {
  
  protective_vars <- c(
    "Zcommconnect", "Zcommintegr_disadv",
    "Zopt_hopeful", "Zselfesteem", "Zself_efficacy",
    "Zparentchild_rel",
    "Zschoolconnect"
  )
  
  var_labels <- c(
    "Zcommconnect"       = "Community Connectedness",
    "Zcommintegr_disadv" = "Community Integration",
    "Zopt_hopeful"       = "Optimism / Hopefulness",
    "Zselfesteem"        = "Self-Esteem",
    "Zself_efficacy"     = "Self-Efficacy",
    "Zparentchild_rel"   = "Parent-Child Relationship",
    "Zschoolconnect"     = "School Connectedness"
  )
  
  fmt_coef <- function(model, var) {
    coefs <- summary(model)$coefficients
    if (!var %in% rownames(coefs)) return("—")
    b <- coefs[var, "Estimate"]
    p <- coefs[var, "Pr(>|t|)"]
    stars <- case_when(
      p < .001 ~ "***",
      p < .01  ~ "**",
      p < .05  ~ "*",
      p < .10  ~ "^",
      TRUE     ~ ""
    )
    paste0(formatC(b, digits = 2, format = "f"), stars)
  }
  
  tbl <- data.frame(
    Variable       = var_labels[protective_vars],
    Full_Sample    = sapply(protective_vars, \(v) fmt_coef(model_full, v)),
    Low_Adversity  = sapply(protective_vars, \(v) fmt_coef(model_lo,   v)),
    High_Adversity = sapply(protective_vars, \(v) fmt_coef(model_hi,   v)),
    row.names = NULL
  )
  
  flextable(tbl) |>
    set_header_labels(
      Variable       = "Protective Factor",
      Full_Sample    = "Full Sample",
      Low_Adversity  = "Low Adversity",
      High_Adversity = "High Adversity"
    ) |>
    add_header_row(
      values    = c("", outcome_label),
      colwidths = c(1, 3)
    ) |>
    theme_apa() |>
    font(fontname = "Times New Roman", part = "all") |>
    fontsize(size = 11, part = "all") |>
    bold(part = "header") |>
    width(j = 1, width = 2.2) |>
    width(j = 2:4, width = 1.5) |>
    align(j = 2:4, align = "center", part = "all") |>
    add_footer_lines("^ p < .10. * p < .05. ** p < .01. *** p < .001.") |>
    fontsize(size = 9, part = "footer") |>
    italic(part = "footer")
}

tbl_compare_depression <- make_protect_table(depression_protect, model_depression_lo, model_depression_hi, "Depression")
tbl_compare_substance  <- make_protect_table(substance_protect,  model_substance_lo,  model_substance_hi,  "Substance Use")
tbl_compare_education  <- make_protect_table(education_protect,  model_education_lo,  model_education_hi,  "Education Attainment")
tbl_compare_health     <- make_protect_table(health_protect,     model_health_lo,     model_health_hi,     "General Health")
```

## Table: Protective vs. Promotive — Depression

```{r}
#| label: tbl-compare-depression

tbl_compare_depression
```

```{r}
#| label: save-compare-depression-png

tbl_compare_depression_png <- tbl_compare_depression |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_depression_png, path = "tbl_compare_depression.png", zoom = 3, expand = 0)
```

## Table: Protective vs. Promotive — Substance Use

@tbl-compare-substance

```{r}
#| label: tbl-compare-substance

tbl_compare_substance
```

```{r}
#| label: save-compare-substance-png

tbl_compare_substance_png <- tbl_compare_substance |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_substance_png, path = "tbl_compare_substance.png", zoom = 3, expand = 0)
```

## Table: Protective vs. Promotive — Education

@tbl-compare-education

```{r}
#| label: tbl-compare-education

tbl_compare_education
```

```{r}
#| label: save-compare-education-png

tbl_compare_education_png <- tbl_compare_education |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_education_png, path = "tbl_compare_education.png", zoom = 3, expand = 0)
```

## Table: Protective vs. Promotive — General Health

@tbl-compare-health

```{r}
#| label: tbl-compare-health

tbl_compare_health
```

```{r}
#| label: save-compare-health-png

tbl_compare_health_png <- tbl_compare_health |>
  font(fontname = "Calibri", part = "all") |>
  fontsize(size = 20, part = "body") |>
  fontsize(size = 22, part = "header") |>
  fontsize(size = 14, part = "footer") |>
  width(j = 1, width = 2.0) |>
  width(j = 2:4, width = 1.4)

save_as_image(tbl_compare_health_png, path = "tbl_compare_health.png", zoom = 3, expand = 0)
```
