---
title: "Indicator Profiles — Standardized (Z-Score)"
code-fold: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(haven)   # read_sav
```

# Standardized Profile Panels

I also standardized based on the within class average for each indicator, allowing for a visualization of how each profile compares to the mean. The code used to create the figures is below.


# Variable Labels

```{r}
#| label: variable-labels

#Variable labels for plots
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# "Negative" = negative risk indicators; "Positive" = positive risk indicators
indicator_risk_panel <- c(
  WSDalc               = "Negative",
  cigwhole             = "Negative",
  skipschool           = "Negative",
  liewhere             = "Negative",
  awayhome             = "Negative",
  dare_dum             = "Negative",
  seatbelt             = "Negative",
  drvdrunk             = "Negative",
  BC_dum               = "Negative",
  steal12M             = "Negative",
  phyfight             = "Negative",
  WSDsport             = "Positive",
  WSDactivity          = "Positive",
  avoidproblems        = "Positive",
  disclose             = "Positive",
  intitaterelationship = "Positive")

# Desired x-axis order within each panel — edit to reorder bars
neg_order <- c(
  "WSDalc", "cigwhole", "skipschool", "liewhere", "awayhome",
  "dare_dum", "seatbelt", "drvdrunk", "BC_dum", "steal12M", "phyfight")

pos_order <- c(
  "WSDsport", "WSDactivity", "avoidproblems", "disclose", "intitaterelationship")
```

# Standardization

```{r}
#| label: import
#| echo: false

sav_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/LCAwASP_4CProbabilities.sav"

sav_raw <- read_sav(sav_path)
```

Class probabilities and indicator values are read in from the dissertation's saved SPSS output (a 4-class LCA solution with an auxiliary ASP scale). Each indicator is then z-scored so the plotted classes are on a common scale.

```{r}
#| label: standardize

# Variable names present in the indicator map
indicator_vars <- names(indicator_label_map)

# Standardize each indicator across the full sample
sav_scaled <- sav_raw %>%
  mutate(across(all_of(indicator_vars), ~ as.numeric(scale(.x))))

# Compute within-class means of z-scores
# Class variable assumed to be named "class" — update if different in your file
zscore_long <- sav_scaled %>%
  select(`clu#`, all_of(indicator_vars)) %>%
  pivot_longer(
    cols      = all_of(indicator_vars),
    names_to  = "variable",
    values_to = "z_score"
  ) %>%
  group_by(`clu#`, variable) %>%
  summarise(mean_z = mean(z_score, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    cluster    = paste0("C", `clu#`),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    label      = indicator_label_map[variable],
    Risk       = indicator_risk_panel[variable],
    label      = factor(label, levels = indicator_label_map[c(neg_order, pos_order)])
  )

```

# Legend Labels

```{r}
#| label: legend-labels

# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Edit the description for each cluster here
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR"
)

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions, " (", cluster_sizes_z, "%)"),
  names(cluster_sizes_z)
)
cluster_colors <- c(
  "C1" = "#44AA99",
  "C2" = "#88CCEE",   
  "C3" = "#882255",
  "C4" = "#DDCC77"   
)

# Colors — pulls from cluster_colors defined in _setup.R
zscore_colors <- cluster_colors
```

# Z-Score Profile Panels

A horizontal reference line at z = 0 marks the sample average so it is immediately clear which profiles fall above vs. below the mean on each indicator. 

```{r}
#| label: make-zscore-panel

# Function to create panel for profiles
make_zscore_panel <- function(risk_label) {

  # Set x-axis variable order
  panel_order <- if (risk_label == "Negative") {
    indicator_label_map[neg_order]
  } else {
    indicator_label_map[pos_order]
  }

  # Select columns from dataset
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order))
  
  # Recode for graph 
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order)) %>%
    mutate(mean_z = case_when(
      mean_z >  1.0 ~  1.1,
      mean_z < -0.5 ~ -0.6,
      TRUE          ~  mean_z
    ))

  # Create plot
  ggplot(panel_data, aes(x = label, y = mean_z, fill = cluster)) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.6, linetype = "solid") +
   geom_vline(
    xintercept = seq(1.5, length(panel_order) - 0.5, by = 1),
    color      = "gray63",
    linewidth  = .8
  ) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width
    ) +
    scale_fill_manual(
      values = zscore_colors,
      labels = cluster_labels_z,
      name   = NULL
    ) +
    scale_y_continuous(
      limits = c(-.6, 1.1),
     # breaks = seq(-.5, 1.0, by = 0.5),
    #  labels = number_format(accuracy = 0.1),
    breaks = c(-0.5,  0, 0.5, 1.0),
  labels = c("Below -0.5", "0.0", "0.5", "Above 1.0"),
      expand = expansion(mult = c(0.02, 0.02))
    ) + 
    labs(
      title   = paste(risk_label, "Risk-Taking Indicators"),
      y       = "Mean (Standardized)",
      caption = "Mean standardized (z-scored) indicator values by latent class. Values below −0.5 or above 1.0 are capped for display."
    ) +
    base_theme +
    theme(
      axis.text.x     = element_text(angle = 30, hjust = 1,
                                     size = axis_text_size, family = plot_font),
      legend.position = "none", 
       plot.margin       = margin(40, 0, 40, 100)
    )
}

# Run function for each indicator set
p_zneg <- make_zscore_panel("Negative")
p_zpos <- make_zscore_panel("Positive")

# Save negative panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_neg.png"),
  device   = agg_png,
  plot     = p_zneg,
  width    = 14, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")

# Save positive panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_pos.png"),
  device   = agg_png,
  plot     = p_zpos,
  width    = 14, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")
```

```{r}
#| label: panel-standardized-neg
#| echo: false
print(p_zneg)  
```

```{r}
#| label: panel-standardized-pos
#| echo: false
print(p_zpos)
```