---
title: "Predictor OR Line Graphs"
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
```

## Data Import

Odds-ratio parameters and paired comparisons for each predictor, referenced against Class 1 and Class 3 respectively, are read in from the project's Excel results workbook (line-graph version of the odds-ratio figures). Import code is included but not rendered below.

```{r}
#| label: import
#| echo: false
c1_params_raw <- read_excel(data_path, sheet = "c1 as reference parameters")
c3_params_raw <- read_excel(data_path, sheet = "c3 as reference parameters")
c1_pairs_raw  <- read_excel(data_path, sheet = "predictors paired comparisons")
c3_pairs_raw  <- read_excel(data_path, sheet = "predictors paired (C3) ")
```

## Wrangle Parameter Sheets

```{r}
#| label: wrangle-params
pred_labels <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing")
  #"Zasp"         = "Autonomy Supportive Parenting"


tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds"
    ) %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds)
    ) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}

c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
c3_params <- tidy_params(c3_params_raw, c("Cluster3", "Cluster4", "Cluster1", "Cluster2"))
```

## BH Correction on Pairwise Comparisons

```{r}
#| label: bh-pairs
apply_bh <- function(raw) {
  raw %>%
    rename(predictor = covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05
    )
}

c1_pairs <- apply_bh(c1_pairs_raw)
c3_pairs <- apply_bh(c3_pairs_raw)
```

## Significance Color Settings

```{r}
#| label: sig-settings

# Colors for OR points — edit hex codes to change
color_sig_positive <- "#228833"   # green  — OR > 1, sig vs. reference
color_sig_negative <- "#EE6677"   # purple — OR < 1, sig vs. reference
color_nonsig       <- "#BBBBBB"   # grey   — not sig vs. reference
color_reference    <- "black"     # black  — reference cluster (OR = 1.0 by definition)

# Reference line at OR = 1.0
or_refline_color <- "grey20"
or_refline_type  <- "dashed"

# Point shape (size and line width set in appearance-settings below)
or_point_shape <- 21    # filled circle with border
```

## Figure Appearance Settings

```{r}
#| label: appearance-settings

# Axis labels
ylab_or_c1 <- "Odds Ratio (vs. C1)"
ylab_or_c3 <- "Odds Ratio (vs. C3)"
xlab_or_C1 <- "Reference: C1"
xlab_or_C3 <- "Reference: C3"

# Font and text sizes — inherited from _setup.R:
#   plot_font, title_size, axis_title_size, axis_text_size, legend_text_size

# Point and line appearance — word-doc scale
or_point_shape <- 21   # filled circle with border
or_point_size  <- 3
or_line_size   <- 0.8

# Legend color lookup (used by both line and bar figures)
or_legend_colors <- c(
  "Higher odds (sig.)"   = color_sig_positive,
  "Lower odds (sig.)"    = color_sig_negative,
  "Not significant"      = color_nonsig,
  "Reference (OR = 1.0)" = color_reference
)

# Shared legend panel — appended to combined figures via patchwork
p_or_legend <- ggplot(
  data.frame(
    x     = 1:4,
    y     = 1,
    label = factor(names(or_legend_colors), levels = names(or_legend_colors))
  ),
  aes(x = x, y = y, fill = label, color = label)
) +
  geom_point(shape = or_point_shape, size = or_point_size + 1, stroke = 0.5) +
  scale_fill_manual(values  = or_legend_colors, name = NULL) +
  scale_color_manual(values = or_legend_colors, name = NULL) +
  guides(
    fill  = guide_legend(nrow = 1, keywidth = 0.6, keyheight = 0.6,
                         override.aes = list(shape = or_point_shape,
                                             size  = or_point_size)),
    color = "none"
  ) +
  coord_cartesian(ylim = c(100, 101)) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = legend_text_size, family = plot_font),
    plot.background = element_rect(fill = NA, color = NA)
  )
```

## Build OR Data: Helper Function

```{r}
#| label: build-or-data

# build_or_long():
#   params     = output of tidy_params() for a given reference class
#   pairs      = output of apply_bh() for a given reference class
#   ref_cluster = "C1" or "C3" — the reference cluster label
#   ref_num    = 1 or 3 — the numeric cluster number used in pairs columns
#
# Returns a long data frame with one row per predictor × cluster,
# with columns: predictor_var, predictor_label, cluster, log_odds, OR, or_color
# The reference cluster row is added with OR = 1.0 and or_color = color_reference.

build_or_long <- function(params, pairs, ref_cluster, ref_num) {

  pred_var_to_pairs_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing")
    #"Zasp"        = "Zasp")

  or_colored <- params %>%
    filter(cluster != ref_cluster) %>%
    rowwise() %>%
    mutate(
      pairs_pred = pred_var_to_pairs_name[predictor_var],
      is_sig = {
        hit <- pairs %>%
          filter(
            predictor == pairs_pred,
            (cluster_a == ref_num | cluster_b == ref_num),
            sig == TRUE
          ) %>%
          mutate(other = ifelse(cluster_a == ref_num, cluster_b, cluster_a)) %>%
          filter(other == as.integer(gsub("C", "", as.character(cluster))))
        nrow(hit) > 0
      },
      or_color  = case_when(
        !is_sig             ~ color_nonsig,
        is_sig & OR > 1     ~ color_sig_positive,
        is_sig & OR < 1     ~ color_sig_negative,
        TRUE                ~ color_nonsig
      ),
      sig_label = case_when(
        !is_sig             ~ "Not significant",
        is_sig & OR > 1     ~ "Higher odds (sig.)",
        is_sig & OR < 1     ~ "Lower odds (sig.)",
        TRUE                ~ "Not significant"
      )
    ) %>%
    ungroup() %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Reference cluster row: OR = 1.0 by definition
  ref_rows <- params %>%
    filter(cluster == ref_cluster) %>%
    mutate(
      or_color  = color_reference,
      OR        = 1.0,
      log_odds  = 0,
      sig_label = "Reference (OR = 1.0)"
    ) %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  bind_rows(or_colored, ref_rows) %>%
    mutate(
      cluster   = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      sig_label = factor(sig_label, levels = names(or_legend_colors))
    )
}

c1_or_long <- build_or_long(c1_params, c1_pairs, ref_cluster = "C1", ref_num = 1)
c3_or_long <- build_or_long(c3_params, c3_pairs, ref_cluster = "C3", ref_num = 3)
```

## Figure: C1 as Reference — One Panel per Predictor (Line)

```{r}
#| label: fig-or-c1-panels
#| fig-width: 1.75
#| fig-height: 4.0

# Predictor display order and titles
pred_order  <- c("female", "Zwave1age", "Zpeer_subuse", "pub_timing")
pred_titles <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing")
  # "Zasp"        = "Autonomy Supportive Parenting")

# x-axis cluster label colors to match cluster colors
x_axis_colors_c1 <- cluster_colors[c("C1", "C2", "C3", "C4")]

# Build and print one line graph per predictor (C1 reference)
plots_or_c1 <- lapply(pred_order, function(pred_var) {

  pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  other_sig_c1 <- c1_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 1, cluster_b != 1,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  fig_caption_c1 <- if (length(other_sig_c1) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c1, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.4) +
    geom_line(aes(group = 1),
              color     = "#BBBBBB",
              linewidth = or_line_size,
              linetype  = "solid") +
    geom_point(aes(fill = sig_label, color = sig_label),
               shape  = or_point_shape,
               size   = or_point_size,
               stroke = 0.5) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_log10(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.15, 0.15))
    ) +
    labs(title = pred_title, caption = fig_caption_c1) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c1),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA)
    )

  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c1, "\n\n")

  ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c1) <- pred_order
```

------------------------------------------------------------------------

## Figure: C3 as Reference — One Panel per Predictor (Line)

```{r}
#| label: fig-or-c3-panels
#| fig-width: 1.75
#| fig-height: 4.0

# x-axis cluster label colors for C3 reference panels
x_axis_colors_c3 <- cluster_colors[c("C1", "C2", "C3", "C4")]

# Build and print one line graph per predictor (C3 reference)
plots_or_c3 <- lapply(pred_order, function(pred_var) {

  pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  other_sig_c3 <- c3_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 3, cluster_b != 3,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  fig_caption_c3 <- if (length(other_sig_c3) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c3, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.4) +
    geom_line(aes(group = 1),
              color     = "grey65",
              linewidth = or_line_size,
              linetype  = "solid") +
    geom_point(aes(fill = sig_label, color = sig_label),
               shape  = or_point_shape,
               size   = or_point_size,
               stroke = 0.5) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_log10(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.15, 0.15))
    ) +
    labs(title = pred_title, caption = fig_caption_c3) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c3),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA)
    )

  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c3, "\n\n")

  ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c3) <- pred_order
```

------------------------------------------------------------------------

## Bar Plots: C1 as Reference — One Panel per Predictor

```{r}
#| label: fig-or-c1-bar-panels
#| fig-width: 1.75
#| fig-height: 4.0

plots_or_c1_bar <- lapply(pred_order, function(pred_var) {

  pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  other_sig_c1 <- c1_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 1, cluster_b != 1,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  fig_caption_c1 <- if (length(other_sig_c1) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c1, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR, fill = sig_label)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.5) +
    geom_col(width = 0.65, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_continuous(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.05, 0.15))
    ) +
    labs(title = pred_title, caption = fig_caption_c1) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c1),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA)
    )

  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c1, "\n\n")

  ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_bar.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c1_bar) <- pred_order
```

------------------------------------------------------------------------

## Bar Plots: C3 as Reference — One Panel per Predictor

```{r}
#| label: fig-or-c3-bar-panels
#| fig-width: 1.75
#| fig-height: 4.0

plots_or_c3_bar <- lapply(pred_order, function(pred_var) {

  pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  other_sig_c3 <- c3_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 3, cluster_b != 3,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  fig_caption_c3 <- if (length(other_sig_c3) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c3, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR, fill = sig_label)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.5) +
    geom_col(width = 0.65, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_continuous(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.05, 0.15))
    ) +
    labs(title = pred_title, caption = fig_caption_c3) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c3),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA)
    )

  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c3, "\n\n")

  ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_bar.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c3_bar) <- pred_order
```

------------------------------------------------------------------------

## Legend Panel (standalone inspect / export)

```{r}
#| label: fig-or-legend
#| fig-width: 6.5
#| fig-height: 1.0

print(p_or_legend)

ggsave(file.path(fig_or_c1, "legend_or.png"),
       plot = p_or_legend,
       width = 6.5, height = 1.0,
       dpi = fig_dpi, units = "in",
       bg  = "transparent")
```

------------------------------------------------------------------------

## Draft Figure Notes

**C1 as Reference Group**

*Figure X.* Odds ratios for \[predictor\] predicting profile membership relative to C1 (reference group, OR = 1.0). Sex was dummy-coded (0 = male, 1 = female); all continuous predictors were standardized (*M* = 0, *SD* = 1). The y-axis is log-scaled. Points are colored by direction and significance: green = significantly higher odds than C1; purple = significantly lower odds than C1; grey = not significant vs. C1; black = C1 reference (OR = 1.0 by definition). Significance determined by Benjamini-Hochberg false discovery rate correction (FDR \< .05).

**C3 as Reference Group**

*Figure X.* Odds ratios for \[predictor\] predicting profile membership relative to C3 (reference group, OR = 1.0). Sex was dummy-coded (0 = male, 1 = female); all continuous predictors were standardized (*M* = 0, *SD* = 1). The y-axis is log-scaled. Points are colored by direction and significance: green = significantly higher odds than C3; purple = significantly lower odds than C3; grey = not significant vs. C3; black = C3 reference (OR = 1.0 by definition). Significance determined by Benjamini-Hochberg false discovery rate correction (FDR \< .05).
