---
title: "4-Class LCA: Benjamini-Hochberg Adjustment for Pairwise Comparisons"
author: "Rachel C. Marcus"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
---


```{r}
#| label: setup
#| include: false
source("_setup.R")
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(patchwork)
library(knitr)
library(ragg)
library(readxl)
```

## Data Import

Predictor and paired-comparison results for the BCH (Bolck-Croon-Hagenaars) three-step auxiliary variable analysis are read in from the project's Excel results workbook. Import code is included but not rendered below.

```{r}
#| label: import
#| echo: false

pred_raw <- read_excel(data_path, sheet = "predictors paired comparisons")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")
```


# BH Corrections
```{r}
#| label: bh-pairs

pairs_6c <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(clustera),
    cluster_b = as.integer(clusterb),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05
  )

pred_pairs <- pred_raw %>%
  filter(covariate != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(cluster_a),
    cluster_b = as.integer(cluster_b),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05
  )

```

## Table: Non-Significant Indicator Comparisons
@tbl-nonsig-6c: List of non-significant indicator comparisons
```{r}
#| label: tbl-nonsig-6c
nonsig_6c <- pairs_6c %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")

```

## Table: Significant Indicator Comparisons
@tbl-sig-6c: List of significant indicator comparisons
```{r}
#| label: tbl-sig-6c
sig_6c <- pairs_6c %>%
  filter(sig == TRUE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(sig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")
```


## Table: Non-Significant Predictor Comparisons
@tbl-nonsig-6pred: List of non-significant Predictor comparisons
```{r}
#| label: tbl-nonsig-6pred
nonsig_pred <- pred_pairs %>%
  filter(sig == FALSE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(nonsig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```


## Table: Significant Predictor Comparisons
@tbl-sig-6pred: List of significant Predictor comparisons
```{r}
#| label: tbl-sig-6pred
sig_pred <- pred_pairs %>%
  filter(sig == TRUE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(sig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

## Significant Indicator Comparisons by Profile
@tbl-sig-cluster-6c: Total Significant Pairwise Indicator Comparisons by Cluster
```{r}
#| label: tbl-sig-cluster-6c
sig_pairs_6c <- pairs_6c %>% filter(sig == TRUE)

by_cluster_6c <- bind_rows(
  sig_pairs_6c %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6c,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")


```

## Table: Class 1 vs 2 Filtered
```{r}
#| label: tbl-sig-cluster-6c12
sig_pairs_6c12 <- pairs_6c %>% filter(sig == TRUE & clustera == 1 & clusterb == 2)

sig_pairs_6c12 <- bind_rows(
  sig_pairs_6c12 %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c12 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c12,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")



nonsig_pairs_6c12 <- pairs_6c %>% filter(sig == FALSE & clustera == 1 & clusterb == 2)

nonsig_pairs_6c12 <- bind_rows(
  nonsig_pairs_6c12 %>% select(cluster = cluster_a, Indicator),
  nonsig_pairs_6c12 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c12,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```


## Table: Class 3 vs 4 Filtered
```{r}
#| label: tbl-sig-cluster-6c34
sig_pairs_6c34 <- pairs_6c %>% filter(sig == TRUE & clustera == 3 & clusterb == 4)

sig_pairs_6c34 <- bind_rows(
  sig_pairs_6c34 %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c34 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c34,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")



nonsig_pairs_6c34 <- pairs_6c %>% filter(sig == FALSE & clustera == 3 & clusterb == 4)

nonsig_pairs_6c34 <- bind_rows(
  nonsig_pairs_6c34 %>% select(cluster = cluster_a, Indicator),
  nonsig_pairs_6c34 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c34,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```

## Significant Demograpnic Comparisons by Profile
@tbl-sig-cluster-6pred: Total Significant Pairwise Predictor Comparisons by Cluster
```{r}
#| label: tbl-sig-cluster-6pred

sig_pred <- pred_pairs %>% filter(sig == TRUE)

by_cluster_6pred <- bind_rows(
  sig_pred %>% select(cluster = cluster_a, covariate),
  sig_pred %>% select(cluster = cluster_b, covariate)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6pred,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant Predictor comparisons (BH-corrected, FDR < .05) involving each profile.")
```

## Significant Comparisons by Indicator
@tbl-sig-indicator-6c: Total Significant Pairwise Comparisons by Indicator
```{r}
#| label: tbl-sig-indicator-6c
by_indicator_6c <- pairs_6c %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6c,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```



## Significant Comparisons by Predictors
@tbl-sig-indicator-6pred: Total Significant Pairwise Comparisons by Predictors
```{r}
#| label: tbl-sig-indicator-6pred
by_indicator_6pred <- pred_pairs %>%
  group_by(covariate) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6pred,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by Predictor")
```