---
title: "Interaction Effects & BH-Corrected Comparisons"
toc: true
toc-depth: 3
code-fold: true
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(gt)      # used for the data-view tables added in this consolidation
library(stringr) #for wrap of captions
```

# Overview

This page covers two robustness checks on the profile-membership findings from [Odds Ratios](Odds-Ratios.qmd):

1.  whether the autonomy supportive parenting to profile association is moderated by biological sex or child age, and
2.  a side-by-side view of BH-corrected indicator- and predictor-level comparisons

In @sec-interaction, the ASP-predictor association is tested for moderation by sex and age. In @sec-bh-comparisons, Benjamini-Hochberg (BH) corrected indicator- and predictor-level comparisons are shown side by side.

# Interaction Effects {#sec-interaction}

Tests whether the ASP-predictor association shown in [Odds Ratios](Odds-Ratios.qmd) differs by biological sex or age -- i.e., whether autonomy-supportive parenting predicts class membership differently for boys vs. girls, or younger vs. older adolescents, rather than having a uniform effect across the sample. A regression with the four demographic covariates and autonomy suppportive parenting (ASP) was run first, then the two interaction terms (ASP x sex, ASP x age) were added.

Interaction-effect parameters and paired comparisons, by reference class, are read in from the project's Excel results workbook. Import code is included but not rendered below.

::: {.callout-note appearance="simple" icon="false"}
The code for this section is collapsed by default, but can be expanded individually for each section or all together with the toggle on the upper right.
:::

```{r}
#| label: import-interaction
#| include: false
c1_params_raw <- read_excel(data_path, sheet = "interaction parameters (c1)")
c3_params_raw <- read_excel(data_path, sheet = "interaction parameters (c3)")
c1_pairs_raw  <- read_excel(data_path, sheet = "int paired comparisons (c1)")
c3_pairs_raw  <- read_excel(data_path, sheet = "int paired comparisons (c3)")
```

# Results Import 

**Raw Results Import**
```{r}
#| label: data-view-int-raw
# Preview raw import
gt(c1_params_raw) |> tab_header(title = "interaction parameters (c1)")
```

**Clean Results Import**
```{r}
#| label: wrangle-params-interaction
# Add labels for predictors
pred_labels <- c(
  "bio_sexD"       = "Sex (Female vs. Male)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"         = "Autonomy Supportive Parenting",
  "ZaspXbio_sexD" =	"Autonomy Supportive Parenting * Child Sex",
  "ZaspXZwave1age" =	"Autonomy Supportive Parenting * Child Age")

# Create function to clean data import
tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds" ) %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds) ) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}

# Apply the cleaning function to the C1-reference parameter sheet
c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
```

## BH Correction on Pairwise Comparisons {#sec-int-bh}

```{r}
#| label: bh-pairs-interaction
# Create function to do BH correction
apply_bh_cap <- function(raw) {
  raw %>%
    rename(predictor = Covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05 )
}

# Apply BH correction to the C1-reference paired comparisons
c1_pairs <- apply_bh_cap(c1_pairs_raw)
```

**Cleaned Data**

```{r}
#| label: data-view-int-tidy
gt(c1_params) |> tab_header(title = "data after custom function to clean output")
```

Neither interaction reached significance: ASP x sex, Wald = 1.45, *p* = .69; ASP x age, Wald = 4.32, *p* = .23. This means the ASP-profile association shown on the Odds Ratios page holds equally for males and females, and equally across the adolescent age range studied -- higher ASP shifts profile membership toward C1 (High PR-Low NR) to the same degree regardless of the adolescent's sex or age. This suggests parenting's role in shaping risk-taking profiles is not confined to a particular developmental window within adolescence, nor does it operate through sex-differentiated pathways [@branje2018].

## Autonomy-Supportive Parenting × Sex

```{r}
#| label: c1-aspsex
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_sex <- c1_params %>% filter(predictor_var == "ZaspXbio_sexD")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "ZaspXbio_sexD",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "ZaspXbio_sexD",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_sex <- c1_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct )

# Plot text: title, axis labels, caption
plot_title   <- "Autonomy Supportive Parenting * Child Sex"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_sex <- ggplot(c1_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label,
       caption = str_wrap(fig_caption, width = 90)) +
 or_theme + theme(legend.position = "none")
```

```{r}
#| label: c1-aspsex-save
#| include: false

ggsave(file.path(fig_interactions, "aspXsex_OR_C1.png"),
      plot = p_c1_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-c1-aspsex
#| echo: false
#| fig-width: 14
#| fig-height: 8
print(p_c1_sex)
```

## Autonomy-Supportive Parenting × Age

```{r}
#| label: c1-aspage
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_age <- c1_params %>% filter(predictor_var == "ZaspXZwave1age")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "ZaspXZwave1age",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "ZaspXZwave1age",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_age <- c1_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct
  )

# Plot text: title, axis labels, caption
plot_title   <- "Autonomy Supportive Parenting * Child Age"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_age <- ggplot(c1_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label,
       caption = str_wrap(fig_caption, width = 90)) +
 or_theme + theme(legend.position = "none")
```

```{r}
#| label: c1-aspage-save
#| include: false

ggsave(file.path(fig_interactions, "aspXage_OR_C1.png"),
      plot = p_c1_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-c1-aspage
#| echo: false
#| fig-width: 14
#| fig-height: 8
print(p_c1_age)
```

# BH-Corrected Comparisons {#sec-bh-comparisons}

Benjamini-Hochberg correction applied to the indicator- and predictor-level pairwise comparisons already introduced on [Profiles](Profiles.qmd) and [Odds Ratios](Odds-Ratios.qmd), shown here together for direct comparison. Per Table 3 of the dissertation, negative risk indicators were significant on 57 of 65 possible comparisons and positive risk indicators on 21 of 29, while among predictors, peer substance use was the only one significant on all 6 comparisons -- sex was the least discriminating predictor at 4 of 6.

These indicator- and predictor-level pairwise comparisons were already computed against posterior class probabilities from the BCH three-step model estimated in LatentGOLD, which corrects for measurement error in profile assignment [@bolck2004]. What happens on this page is Benjamini-Hochberg correction of those already BCH-corrected comparisons for the multiple-testing problem created by six pairwise class comparisons per indicator/predictor [@benjamini1995a; @benjamini2000a] -- no additional BCH-specific computation happens in this R code. The indicator-level comparisons here are the same `paired comparisons` sheet used on the [Profiles](Profiles.qmd) page; the predictor-level comparisons are the same `predictors paired comparisons` sheet used on the [Odds Ratios](Odds-Ratios.qmd) page, shown together here for direct comparison.

::: {.callout-note appearance="simple" icon="false"}
The code for this section is collapsed by default, but can be expanded individually for each section or all together with the toggle on the upper right.
:::

```{r}
#| label: import-bch
#| include: false

# Import predictor- and indicator-level paired comparisons
pred_raw <- read_excel(data_path, sheet = "predictors paired comparisons")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")
```

**Data Import**

```{r}
#| label: data-view-bch-raw
gt(pred_raw) |> tab_header(title = "predictors paired comparisons")
```

**Variable Labels**

```{r}
#| label: indicator-labels
# Label Indicators
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# "Negative" = negative risk indicators; "Positive" = positive risk indicators
# (not currently used in a table below, kept here for future grouping)
indicator_risk_panel <- c(
  WSDalc               = "Negative",
  cigwhole             = "Negative",
  skipschool           = "Negative",
  liewhere             = "Negative",
  awayhome             = "Negative",
  dare_dum             = "Negative",
  seatbelt             = "Negative",
  drvdrunk             = "Negative",
  BC_dum               = "Negative",
  steal12M             = "Negative",
  phyfight             = "Negative",
  WSDsport             = "Positive",
  WSDactivity          = "Positive",
  avoidproblems        = "Positive",
  disclose             = "Positive",
  intitaterelationship = "Positive")
```

## BH Corrections {#sec-comparisons-bh}

The indicator- and predictor-level patterns agree in direction: both sets of comparisons separate most sharply along the same alcohol-problem-status divide (C1/C2 vs. C3/C4) that structures the indicator profiles themselves, and peer substance use -- the predictor with the most consistent indicator-level analog (negative risk indicators broadly) -- is also the most consistently significant predictor. BH correction on top of the BCH-corrected estimates does not change this overall pattern; it affects which specific pairwise comparisons clear the FDR < .05 threshold; see @tbl-nonsig-6c and @tbl-nonsig-6pred for the individual comparisons that did not.

```{r}
#| label: bh-pairs-bch

# Clean, BH-correct, and relabel the indicator-level paired comparisons
pairs_6c <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    clustera = as.integer(clustera),
    clusterb = as.integer(clusterb),
    pair      = paste0("C", clustera, " vs. C", clusterb),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05,
    Indicator = recode(Indicator, !!!indicator_label_map))

# Clean, BH-correct, and relabel the predictor-level paired comparisons
pred_pairs <- pred_raw %>%
  filter(covariate != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    class = as.integer(cluster_a),
    class = as.integer(cluster_b),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05,
    covariate = recode(covariate, !!!pred_labels))

```

**Non-Significant Indicator Comparisons**

```{r}
#| label: tbl-nonsig-6c
# Filter to non-significant indicator comparisons only
nonsig_6c <- pairs_6c %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")

```

**Significant Indicator Comparisons**

```{r}
#| label: tbl-sig-6c
# Filter to significant indicator comparisons only
sig_6c <- pairs_6c %>%
  filter(sig == TRUE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(sig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

**Non-Significant Predictor Comparisons**

```{r}
#| label: tbl-nonsig-6pred
# Filter to non-significant predictor comparisons only
nonsig_pred <- pred_pairs %>%
  filter(sig == FALSE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(nonsig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

**Significant Predictor Comparisons**

```{r}
#| label: tbl-sig-6pred
# Filter to significant predictor comparisons only
sig_pred <- pred_pairs %>%
  filter(sig == TRUE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(sig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

**Significant Indicator Comparisons by Profile**

```{r}
#| label: tbl-sig-cluster-6c
# Count how many significant comparisons involve each profile
sig_pairs_6c <- pairs_6c %>% filter(sig == TRUE)

by_cluster_6c <- bind_rows(
  sig_pairs_6c %>% select(cluster = clustera, Indicator),
  sig_pairs_6c %>% select(cluster = clusterb, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6c,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")


```

**Indicator Comparisons by Profile: C1 vs. C2**

```{r}
#| label: tbl-sig-cluster-6c12
# Count significant comparisons specific to the C1 vs. C2 pair
sig_pairs_6c12 <- pairs_6c %>% filter(sig == TRUE & clustera == 1 & clusterb == 2)

sig_pairs_6c12 <- bind_rows(
  sig_pairs_6c12 %>% select(cluster = clustera, Indicator),
  sig_pairs_6c12 %>% select(cluster = clusterb, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c12,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")


# Count non-significant comparisons specific to the C1 vs. C2 pair
nonsig_pairs_6c12 <- pairs_6c %>% filter(sig == FALSE & clustera == 1 & clusterb == 2)

nonsig_pairs_6c12 <- bind_rows(
  nonsig_pairs_6c12 %>% select(cluster = clustera, Indicator),
  nonsig_pairs_6c12 %>% select(cluster = clusterb, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c12,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```

**Indicator Comparisons by Profile: C3 vs. C4**

```{r}
#| label: tbl-sig-cluster-6c34
# Count significant comparisons specific to the C3 vs. C4 pair
sig_pairs_6c34 <- pairs_6c %>% filter(sig == TRUE & clustera == 3 & clusterb == 4)

sig_pairs_6c34 <- bind_rows(
  sig_pairs_6c34 %>% select(cluster = clustera, Indicator),
  sig_pairs_6c34 %>% select(cluster = clusterb, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c34,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")


# Count non-significant comparisons specific to the C3 vs. C4 pair
nonsig_pairs_6c34 <- pairs_6c %>% filter(sig == FALSE & clustera == 3 & clusterb == 4)

nonsig_pairs_6c34 <- bind_rows(
  nonsig_pairs_6c34 %>% select(cluster = clustera, Indicator),
  nonsig_pairs_6c34 %>% select(cluster = clusterb, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c34,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```

**Significant Demographic Comparisons by Profile**

```{r}
#| label: tbl-sig-cluster-6pred
# Count how many significant predictor comparisons involve each profile
sig_pred <- pred_pairs %>% filter(sig == TRUE)

by_cluster_6pred <- bind_rows(
  sig_pred %>% select(cluster = cluster_a, covariate),
  sig_pred %>% select(cluster = cluster_b, covariate)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6pred,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant Predictor comparisons (BH-corrected, FDR < .05) involving each profile.")
```

**Significant Comparisons by Indicator**

```{r}
#| label: tbl-sig-indicator-6c
# Total significant vs. non-significant comparisons, by indicator
by_indicator_6c <- pairs_6c %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6c,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```

**Cleaned Data**

```{r}
#| label: data-view-bch-tidy
gt(pred_pairs) |> tab_header(title = "After: BH-corrected predictor pairwise comparisons")
```


**Significant Comparisons by Predictor**

```{r}
#| label: tbl-sig-indicator-6pred
# Total significant vs. non-significant comparisons, by predictor
by_indicator_6pred <- pred_pairs %>%
  group_by(covariate) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6pred,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by Predictor")
```