---
title: "Interaction Effects & BH-Corrected Comparisons"
toc: true
toc-depth: 3
code-fold: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(gt)      # used for the data-view tables added in this consolidation
```

# Overview

This page covers two robustness checks on the profile-membership findings from [Odds Ratios](Odds-Ratios.qmd): whether the ASP-profile association is moderated by sex or age, and a side-by-side view of BH-corrected indicator- and predictor-level comparisons. Each section below defines its own local `tidy_params()`/`apply_bh_cap()` (a Quarto document each render independently, so these can't be imported from Odds-Ratios.qmd's Helper Functions the way they can within that single file) -- the logic is identical to the versions there.

**A bug fix made in this consolidation:** the original `06_interaction.qmd` computed which comparisons were significant against the reference class using `ifelse(clustera == 1, cluster_b, cluster_a)` -- but `cluster_b`/`cluster_a` (with an underscore) were never defined anywhere in that file; the actual columns are `clustera`/`clusterb` (no underscore), as used two lines earlier in the same chunk. This would have thrown an "object not found" error on render. Fixed to `ifelse(clustera == 1, clusterb, clustera)` in both the ASP\\*Sex and ASP\\*Age figure chunks below.

# 1. Interaction Effects

## Overview

Tests whether the ASP-predictor association shown in [Odds Ratios](Odds-Ratios.qmd) differs by biological sex or age -- i.e., whether autonomy-supportive parenting predicts class membership differently for boys vs. girls, or younger vs. older adolescents, rather than having a uniform effect across the sample. A regression with the four demographic covariates and ASP was run first, then the two interaction terms (ASP x sex, ASP x age) were added.

## Data Import

Interaction-effect parameters and paired comparisons, by reference class, are read in from the project's Excel results workbook. Import code is included but not rendered below.

```{r}
#| label: import-interaction
#| echo: false
c1_params_raw <- read_excel(data_path, sheet = "interaction parameters (c1)")
c3_params_raw <- read_excel(data_path, sheet = "interaction parameters (c3)")
c1_pairs_raw  <- read_excel(data_path, sheet = "int paired comparisons (c1)")
c3_pairs_raw  <- read_excel(data_path, sheet = "int paired comparisons (c3)")
```

```{r}
#| label: data-view-int-raw
gt(head(c1_params_raw, 6)) |> tab_header(title = "interaction parameters (c1)")
```


## Wrangle Parameter Sheets

```{r}
#| label: wrangle-params-interaction
pred_labels <- c(
  "female"       = "Sex (Female vs. Male)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"         = "Autonomy Supportive Parenting",
  "ZaspXbio_sexD" =	"Autonomy Supportive Parenting * Child Sex",
  "ZaspXZwave1age" =	"Autonomy Supportive Parenting * Child Age"
)

tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds"
    ) %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds)
    ) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}

c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
```


## BH Correction on Pairwise Comparisons

```{r}
#| label: bh-pairs-interaction
apply_bh_cap <- function(raw) {
  raw %>%
    rename(predictor = Covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05
    )
}

c1_pairs <- apply_bh_cap(c1_pairs_raw)
```

```{r}
#| label: data-view-int-tidy
gt(head(c1_params, 8)) |> tab_header(title = "After: tidy_params() output, interaction parameters")
```

Neither interaction reached significance: ASP x sex, Wald = 1.45, *p* = .69; ASP x age, Wald = 4.32, *p* = .23. This means the ASP-profile association shown on the Odds Ratios page holds equally for males and females, and equally across the adolescent age range studied -- higher ASP shifts profile membership toward C1 (High PR-Low NR) to the same degree regardless of the adolescent's sex or age. This suggests parenting's role in shaping risk-taking profiles is not confined to a particular developmental window within adolescence, nor does it operate through sex-differentiated pathways (Branje, 2018).

## Figure: ASP*Sex - C1 Reference
```{r}
#| label: fig-c1-aspsex

c1_sex <- c1_params %>% filter(predictor_var == "ZaspXbio_sexD")

ref_sig <- c1_pairs %>%
  filter(predictor == "ZaspXbio_sexD",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

other_sig <- c1_pairs %>%
  filter(predictor == "ZaspXbio_sexD",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

c1_sex <- c1_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct
  )

plot_title   <- "Autonomy Supportive Parenting * Child Sex"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }
p <- ggplot(c1_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme

ggsave(file.path(fig_interactions, "aspXsex_OR_C1.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```


### Figure: ASP*Age - C1 Reference
```{r}
#| label: fig-c1-aspage

c1_age <- c1_params %>% filter(predictor_var == "ZaspXZwave1age")

ref_sig <- c1_pairs %>%
  filter(predictor == "ZaspXZwave1age",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

other_sig <- c1_pairs %>%
  filter(predictor == "ZaspXZwave1age",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

c1_age <- c1_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct
  )

plot_title   <- "Autonomy Supportive Parenting * Child Age"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }
p <- ggplot(c1_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme

ggsave(file.path(fig_interactions, "aspXage_OR_C1.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

# 2. BH-Corrected Comparisons

### Overview

Benjamini-Hochberg correction applied to the indicator- and predictor-level pairwise comparisons already introduced on [Profiles](Profiles.qmd) and [Odds Ratios](Odds-Ratios.qmd), shown here together for direct comparison. Per Table 3 of the dissertation, negative risk indicators were significant on 57 of 65 possible comparisons and positive risk indicators on 21 of 29, while among predictors, peer substance use was the only one significant on all 6 comparisons -- sex was the least discriminating predictor at 4 of 6.

### Data Import

These indicator- and predictor-level pairwise comparisons were already computed against posterior class probabilities from the BCH three-step model estimated in LatentGOLD, which corrects for measurement error in profile assignment (Bolck et al., 2004). What happens on this page is Benjamini-Hochberg correction of those already BCH-corrected comparisons for the multiple-testing problem created by six pairwise class comparisons per indicator/predictor (Benjamini & Hochberg, 1995, 2000) -- no additional BCH-specific computation happens in this R code. (The original file's title said "Benjamini-Hochberg Adjustment," correctly, but this import paragraph said BCH -- fixed here to match what the code actually does.) The indicator-level comparisons here are the same `paired comparisons` sheet used on the Profiles page; the predictor-level comparisons are the same `predictors paired comparisons` sheet used on the Odds Ratios page, shown together here for direct comparison.

```{r}
#| label: import-bch
#| echo: false

pred_raw <- read_excel(data_path, sheet = "predictors paired comparisons")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")
```

```{r}
#| label: data-view-bch-raw
gt(head(pred_raw, 5)) |> tab_header(title = "predictors paired comparisons")
```



## BH Corrections
```{r}
#| label: bh-pairs-bch

pairs_6c <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(clustera),
    cluster_b = as.integer(clusterb),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05
  )

pred_pairs <- pred_raw %>%
  filter(covariate != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(cluster_a),
    cluster_b = as.integer(cluster_b),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05
  )

```

### Table: Non-Significant Indicator Comparisons
@tbl-nonsig-6c: List of non-significant indicator comparisons
```{r}
#| label: tbl-nonsig-6c
nonsig_6c <- pairs_6c %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")

```

### Table: Significant Indicator Comparisons
@tbl-sig-6c: List of significant indicator comparisons
```{r}
#| label: tbl-sig-6c
sig_6c <- pairs_6c %>%
  filter(sig == TRUE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(sig_6c,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant indicator comparisons after Benjamini-Hochberg correction (FDR < .05).")
```


### Table: Non-Significant Predictor Comparisons
@tbl-nonsig-6pred: List of non-significant Predictor comparisons
```{r}
#| label: tbl-nonsig-6pred
nonsig_pred <- pred_pairs %>%
  filter(sig == FALSE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(nonsig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```


### Table: Significant Predictor Comparisons
@tbl-sig-6pred: List of significant Predictor comparisons
```{r}
#| label: tbl-sig-6pred
sig_pred <- pred_pairs %>%
  filter(sig == TRUE) %>%
  select(covariate, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(covariate, pair)

kable(sig_pred,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Significant Predictor comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

### Significant Indicator Comparisons by Profile
@tbl-sig-cluster-6c: Total Significant Pairwise Indicator Comparisons by Cluster
```{r}
#| label: tbl-sig-cluster-6c
sig_pairs_6c <- pairs_6c %>% filter(sig == TRUE)

by_cluster_6c <- bind_rows(
  sig_pairs_6c %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6c,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")


```

### Table: Class 1 vs 2 Filtered
```{r}
#| label: tbl-sig-cluster-6c12
sig_pairs_6c12 <- pairs_6c %>% filter(sig == TRUE & clustera == 1 & clusterb == 2)

sig_pairs_6c12 <- bind_rows(
  sig_pairs_6c12 %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c12 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c12,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")



nonsig_pairs_6c12 <- pairs_6c %>% filter(sig == FALSE & clustera == 1 & clusterb == 2)

nonsig_pairs_6c12 <- bind_rows(
  nonsig_pairs_6c12 %>% select(cluster = cluster_a, Indicator),
  nonsig_pairs_6c12 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c12,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```


### Table: Class 3 vs 4 Filtered
```{r}
#| label: tbl-sig-cluster-6c34
sig_pairs_6c34 <- pairs_6c %>% filter(sig == TRUE & clustera == 3 & clusterb == 4)

sig_pairs_6c34 <- bind_rows(
  sig_pairs_6c34 %>% select(cluster = cluster_a, Indicator),
  sig_pairs_6c34 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(sig_pairs_6c34,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")



nonsig_pairs_6c34 <- pairs_6c %>% filter(sig == FALSE & clustera == 3 & clusterb == 4)

nonsig_pairs_6c34 <- bind_rows(
  nonsig_pairs_6c34 %>% select(cluster = cluster_a, Indicator),
  nonsig_pairs_6c34 %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(nonsig_pairs_6c34,
      col.names = c("Profile", "Number of Non-Significant Pairwise Comparisons"),
      caption   = "Count of non-significant indicator comparisons (BH-corrected, FDR < .05) involving each profile.")
```

### Significant Demograpnic Comparisons by Profile
@tbl-sig-cluster-6pred: Total Significant Pairwise Predictor Comparisons by Cluster
```{r}
#| label: tbl-sig-cluster-6pred

sig_pred <- pred_pairs %>% filter(sig == TRUE)

by_cluster_6pred <- bind_rows(
  sig_pred %>% select(cluster = cluster_a, covariate),
  sig_pred %>% select(cluster = cluster_b, covariate)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster_6pred,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant Predictor comparisons (BH-corrected, FDR < .05) involving each profile.")
```

### Significant Comparisons by Indicator
@tbl-sig-indicator-6c: Total Significant Pairwise Comparisons by Indicator
```{r}
#| label: tbl-sig-indicator-6c
by_indicator_6c <- pairs_6c %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6c,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```

```{r}
#| label: data-view-bch-tidy
gt(head(pred_pairs, 6)) |> tab_header(title = "After: BH-corrected predictor pairwise comparisons")
```

The indicator- and predictor-level patterns agree in direction: both sets of comparisons separate most sharply along the same alcohol-problem-status divide (C1/C2 vs. C3/C4) that structures the indicator profiles themselves, and peer substance use -- the predictor with the most consistent indicator-level analog (negative risk indicators broadly) -- is also the most consistently significant predictor. BH correction on top of the BCH-corrected estimates does not change this overall pattern; it affects which specific pairwise comparisons clear the FDR < .05 threshold; see the non-significant tables above for the individual comparisons that did not.




### Significant Comparisons by Predictors
@tbl-sig-indicator-6pred: Total Significant Pairwise Comparisons by Predictors
```{r}
#| label: tbl-sig-indicator-6pred
by_indicator_6pred <- pred_pairs %>%
  group_by(covariate) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator_6pred,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by Predictor")
```