---
title: "Odds Ratios: Predictors & ASP Scales"
toc: true
toc-depth: 3
code-fold: true
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(gt)      # used for the data-view tables added in this consolidation
library(stringr)   # str_wrap(), used to wrap figure captions
```

# Overview

This page tests whether the demographic, parenting, and peer predictors
described on [Profiles](Profiles.qmd) actually distinguish class
membership, using posterior-probability-weighted multinomial logistic
regression via the BCH three-step method to correctly account for
measurement error in profile assignment [@bolck2004], with all
continuous predictors standardized prior to analysis.

In @sec-BH, Benjamini-Hochberg correction is applied in R to the
resulting pairwise comparisons [@benjamini1995a; @benjamini2000a].
Running six pairwise cluster comparisons per predictor raises the odds
that some comparisons look significant purely by chance, so
Benjamini-Hochberg correction adjusts the p-values across that full set
of comparisons and keeps the false discovery rate at an acceptable level
before any pairwise difference gets flagged as significant.

In @sec-OR, predictor odds ratios (bar graphs and written summaries) are
shown. Alternative line graphs of the odd ratios can be found in
@sec-ORline

```{r}
#| label: helper-tidy-params
# Shared by every OR section below to clean data import
tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds") %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds)) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}
```

```{r}
#| label: helper-apply-bh
#| include: false

# For sheets with lowercase covariate/cluster_a/cluster_b (predictors paired comparisons).
apply_bh <- function(raw) {
  raw %>%
    rename(predictor = covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}

# For sheets with capitalized Covariate/clustera/clusterb (ASP, interaction,
# indicator-pairs). Reused on the Interaction & Comparisons page.
apply_bh_cap <- function(raw) {
  raw %>%
    rename(predictor = Covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}
```

# Data Import and Cleaning

Odds-ratio parameters and paired comparisons for each predictor,
referenced against Class 1 and Class 3 respectively, are read in from
the project's Excel results workbook. Import code is included but not
rendered below.

::: {.callout-note appearance="simple" icon="false"}
The code for this section is collapsed by default, but can be expanded
individually for each section or all together with the toggle on the
upper right.
:::

```{r}
#| label: import-predictor-or
#| include: false
c1_params_raw <- read_excel(data_path, sheet = "c1 as reference parameters")
c3_params_raw <- read_excel(data_path, sheet = "c3 as reference parameters")
c1_pairs_raw  <- read_excel(data_path, sheet = "predictors paired comparisons")
c3_pairs_raw  <- read_excel(data_path, sheet = "predictors paired (C3) ")
```

**Data Import of C1 as Reference**

```{r}
#| label: data-view-or-raw
# View first rows of c1 reference as data import example 
gt(c1_params_raw) |> tab_header(title = "c1 as reference parameters")
```

**Data Import of Paired Comparisons Between Classes (clusters)**

```{r}
#| label: data-view-or-raw2
# View last rows of paired comparisons as data import example 
gt(c1_pairs_raw) |>
  tab_header(title = "predictors paired comparisons") |>
  cols_label(
    cluster_a = "Class",
    cluster_b = "Class" )
```

**Clean Data Import**

```{r}
#| label: wrangle-params-predictor-or
# Display labels for each predictor
pred_labels <- c(
  "female"       = "Sex (Male = 0)",
  "bio_sexD"     = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"         = "ASP")

# Tidy each reference-class parameter sheet
c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
c3_params <- tidy_params(c3_params_raw, c("Cluster3", "Cluster4", "Cluster1", "Cluster2"))
```

## BH Correction on Pairwise Comparisons {#sec-BH}

```{r}
#| label: bh-pairs-predictor-or
#| include: false
# Add a labeled column alongside the raw "predictor" column -- the raw
# codes (e.g. "bio_sexD") stay in place since every figure build chunk
# filters on them; predictor_label is only for display in tables.
c1_pairs <- apply_bh(c1_pairs_raw) %>%
  mutate(predictor_label = recode(predictor, !!!pred_labels))
c3_pairs <- apply_bh(c3_pairs_raw) %>%
  mutate(predictor_label = recode(predictor, !!!pred_labels))
```

**Cleaned Data with C1 as Reference**

```{r}
#| label: data-view-or-tidy
# Hide predictor_var (raw code) from the preview, keep predictor_label
gt(c1_params %>% select(-predictor_var)) |>
  tab_header(title = "data after custom function to clean output")
```

**Cleaned Data of Paired Comparisons Between Classes**

```{r}
#| label: data-view-or-tidy2
# Show predictor_label instead of the raw "predictor" code
gt(c1_pairs %>% select(predictor_label, everything(), -predictor)) |>
  tab_header(title = "data after custom function to clean output") |>
  cols_label(
    predictor_label = "Predictor",
    cluster_a = "Class",
    cluster_b = "Class" )
```

# Odds Ratios {#sec-OR}

Odds ratios test whether each predictor distinguishes class membership,
computed twice: once with C1 (High PR-Low NR) as the reference class,
once with C3 (Low PR-High NR), so both alcohol-problem-status groupings
can anchor the comparison.

::: {.callout-tip appearance="simple" icon="false"}
**OR Interpretation**

An OR above 1 means higher odds of membership in the comparison class
relative to the reference, for a one-unit (or one-SD, for standardized
predictors) increase in that predictor.
:::

## Autonomy-Supportive Parenting

```{r}
#| label: c1-asp
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_asp <- c1_params %>% filter(predictor_var == "Zasp")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_asp <- c1_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "* Significantly different than C1"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
              } else {
                 "No additional pairwise differences were significant after BH correction."
}

# Build the plot
p_c1_asp <- ggplot(c1_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-asp-save
#| include: false

ggsave(file.path(fig_or_c1, "Zasp_OR_C1ref.png"),
       plot = p_c1_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

```{r}
#| label: c3-asp
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
c3_asp <- c3_params %>% filter(predictor_var == "Zasp")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_asp <- c3_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "* Significantly different than C3"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) { 
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_asp <- ggplot(c3_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-asp-save
#| include: false

ggsave(file.path(fig_or_c3, "Zasp_OR_C3ref.png"),
       plot = p_c3_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

```{r}
#| label: fig-asp-combined
#| echo: false
#| layout-ncol: 2



print(p_c1_asp)
print(p_c3_asp)
```

**Autonomy-Supportive Parenting.** As ASP increases, the odds of being
in any other profile compared to C1 (High PR–Low NR) significantly
decrease. Compared to C3 (Low PR–High NR), as ASP increases the odds of
being in either of the two zero-alcohol-problem profiles (C1, C2)
decrease; C2 vs. C3 and C2 vs. C4 were also significant. This is
consistent with self-determination theory [@deci2013; @ryan2022]:
autonomy-supportive parenting predicted profile membership even within
the two no-alcohol-problem profiles, which a variable-centered approach
combining them would have obscured [@bergman1997].

## Biological Sex

```{r}
#| label: c1-sex
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_sex <- c1_params %>% filter(predictor_var == "female")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_sex <- c1_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Sex (Female vs. Male)"
x_label      <- "Reference: C1"
y_label      <- "Odds Ratio - C1 as Reference"    
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_sex <- ggplot(c1_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-sex-save
#| include: false

ggsave(file.path(fig_or_c1, "bio_sexD_OR_C1ref.png"),
       plot = p_c1_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-sex
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
c3_sex <- c3_params %>% filter(predictor_var == "female")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_sex <- c3_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Sex (Female vs. Male)"
#x_label     <- "Reference: C3"
#y_label     <- "Odds Ratio - C3 as Reference"
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_sex <- ggplot(c3_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-sex-save
#| include: false

ggsave(file.path(fig_or_c3, "bio_sexD_OR_C3ref.png"),
       plot = p_c3_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-sex-combined
#| echo: false
#| layout-ncol: 2



print(p_c1_sex)
print(p_c3_sex)
```

**Sex.** As the proportion of females in a profile increases, compared
to C1 the odds of being in C2 or C3 significantly decrease; compared to
C3, the odds of being in C1 or C4 increase (C2 vs. C4 and C3 vs. C4 also
significant) — consistent with prior findings of higher male delinquency
rates [@miller2010].

## Child Age

```{r}
#| label: c1-age
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_age <- c1_params %>% filter(predictor_var == "Zwave1age")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_age <- c1_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Age"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
               }

# Build the plot
p_c1_age <- ggplot(c1_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-age-save
#| include: false

ggsave(file.path(fig_or_c1, "Zwave1age_OR_C1ref.png"),
       plot = p_c1_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-age
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
c3_age <- c3_params %>% filter(predictor_var == "Zwave1age")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_age <- c3_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Age"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_age<- ggplot(c3_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-age-save
#| include: false

ggsave(file.path(fig_or_c3, "Zwave1age_OR_C3ref.png"),
       plot = p_c3_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-age-combined
#| echo: false
#| layout-ncol: 2



print(p_c1_age)
print(p_c3_age)
```

**Child Age.** As age increases, the odds of being in any profile
compared to C1 increase significantly; compared to C3, the odds of being
in either zero-alcohol-problem profile decrease (C2 vs. C3 and C2 vs. C4
also significant) — consistent with developmental increases in negative
risk-taking across adolescence [@duell2018; @moffitt2017].

## Peer Substance Use

```{r}
#| label: c1-peer
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_peer <- c1_params %>% filter(predictor_var == "Zpeer_subuse")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_peer <- c1_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_peer <- ggplot(c1_peer, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-peer-save
#| include: false

ggsave(file.path(fig_or_c1, "Zpeer_subuse_OR_C1ref.png"),
       plot = p_c1_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-peer
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
c3_peer <- c3_params %>% filter(predictor_var == "Zpeer_subuse")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_peer <- c3_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_peer <- ggplot(c3_peer,aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-peer-save
#| include: false

ggsave(file.path(fig_or_c3, "Zpeer_subuse_OR_C3ref.png"),
       plot = p_c3_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-peer-combined
#| echo: false
#| layout-ncol: 2


print(p_c1_peer)
print(p_c3_peer)
```

**Peer Substance Use.** As peer substance use increases, the odds of
being in any profile compared to C1 increase significantly; compared to
C3, the odds of being in either zero-alcohol-problem profile or C4
decrease (all six pairwise comparisons significant — the only predictor
with every comparison reaching significance), matching literature
identifying peer behavior as a robust proximal predictor of adolescent
risk [@dishion1999; @dishion2011a; @henneberger2021].

## Pubertal Timing

```{r}
#| label: c1-pub
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
c1_pub <- c1_params %>% filter(predictor_var == "pub_timing")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_pub <- c1_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
             } else { 
                 "No additional pairwise differences were significant after BH correction."
}

# Build the plot
p_c1_pub <- ggplot(c1_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-pub-save
#| include: false

ggsave(file.path(fig_or_c1, "pub_timing_OR_C1ref.png"),
       plot = p_c1_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-pub
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
c3_pub <- c3_params %>% filter(predictor_var == "pub_timing")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_pub <- c3_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_pub <- ggplot(c3_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-pub-save
#| include: false

ggsave(file.path(fig_or_c3, "pub_timing_OR_C3ref.png"),
       plot = p_c3_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-pub-combined
#| echo: false
#| layout-ncol: 2



print(p_c1_pub)
print(p_c3_pub)
```

**Pubertal Timing.** As pubertal timing gets earlier, the odds of being
in C2 decrease compared to C1, while the odds of being in either
alcohol-problem profile increase compared to C1; compared to C3, the
odds of being in either zero-alcohol-problem profile decrease (C2 vs. C3
and C2 vs. C4 also significant) — reinforcing that biological maturation
shapes risk propensity above and beyond chronological age
[@natsuaki2009].

# Odds Ratios - Lines {#sec-ORline}

::: {.callout-tip appearance="minimal" icon="false"}
The following chunks of code create the foundation for the odd ratio
line graphs, or you can skip to the figures directly: @sec-aspline.
:::

**Significance Color Settings**

```{r}
#| label: sig-settings
# Colors for OR data points
color_sig_positive <- "#228833"   # green  — OR > 1, sig vs. reference
color_sig_negative <- "#EE6677"   # purple — OR < 1, sig vs. reference
color_nonsig       <- "#BBBBBB"   # grey   — not sig vs. reference
color_reference    <- "black"     # black  — reference cluster (OR = 1.0 by definition)

# Reference line at OR = 1.0
or_refline_color <- "grey20"
or_refline_type  <- "dashed"

# Point shape (size and line width set in appearance-settings below)
or_point_shape <- 21    # filled circle with border
```

**Figure Appearance Settings**

```{r}
#| label: appearance-settings
# Axis labels
ylab_or_c1 <- "Odds Ratio (vs. C1)"
ylab_or_c3 <- "Odds Ratio (vs. C3)"
xlab_or_C1 <- "Reference: C1"
xlab_or_C3 <- "Reference: C3"

# Font and text sizes — inherited from _setup.R:
#   plot_font, title_size, axis_title_size, axis_text_size, legend_text_size

# Point and line appearance — word-doc scale
or_point_shape <- 21   # filled circle with border
or_point_size  <- 6
or_line_size   <- 3

# Legend color lookup (used by both line and bar figures)
or_legend_colors <- c(
  "Higher odds (sig.)"   = color_sig_positive,
  "Lower odds (sig.)"    = color_sig_negative,
  "Not significant"      = color_nonsig,
  "Reference (OR = 1.0)" = color_reference)

# Shared legend panel — appended to combined figures via patchwork
p_or_legend <- ggplot(
  data.frame(
    x     = 1:4,
    y     = 1,
    label = factor(names(or_legend_colors), levels = names(or_legend_colors))),
  aes(x = x, y = y, fill = label, color = label)) +
  geom_point(shape = or_point_shape, size = or_point_size + 1, stroke = 0.5) +
  scale_fill_manual(values  = or_legend_colors, name = NULL) +
  scale_color_manual(values = or_legend_colors, name = NULL) +
  guides(
    fill  = guide_legend(nrow = 1, keywidth = 0.6, keyheight = 0.6,
                         override.aes = list(shape = or_point_shape,
                                             size  = or_point_size)),
    color = "none") +
  coord_cartesian(ylim = c(100, 101)) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = legend_text_size, family = plot_font),
    plot.background = element_rect(fill = NA, color = NA))
```

**Build OR Data: Helper Function**

```{r}
#| label: build-or-data
# build_or_long():
#   params     = output of tidy_params() for a given reference class
#   pairs      = output of apply_bh() for a given reference class
#   ref_cluster = "C1" or "C3" — the reference cluster label
#   ref_num    = 1 or 3 — the numeric cluster number used in pairs columns
#
# Returns a long data frame with one row per predictor × cluster,
# with columns: predictor_var, predictor_label, cluster, log_odds, OR, or_color
# The reference cluster row is added with OR = 1.0 and or_color = color_reference.

build_or_long <- function(params, pairs, ref_cluster, ref_num) {

  # Map each predictor's tidy_params() name to its name in the pairs sheet
  pred_var_to_pairs_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing",
    "Zasp"        = "Zasp")

  # Non-reference clusters: flag significance vs. reference and assign color
  or_colored <- params %>%
    filter(cluster != ref_cluster) %>%
    rowwise() %>%
    mutate(
      pairs_pred = pred_var_to_pairs_name[predictor_var],
      is_sig = {
        hit <- pairs %>%
          filter(
            predictor == pairs_pred,
            (cluster_a == ref_num | cluster_b == ref_num),
            sig == TRUE) %>%
          mutate(other = ifelse(cluster_a == ref_num, cluster_b, cluster_a)) %>%
          filter(other == as.integer(gsub("C", "", as.character(cluster))))
        nrow(hit) > 0
      },
      or_color  = case_when(
        !is_sig             ~ color_nonsig,
        is_sig & OR > 1     ~ color_sig_positive,
        is_sig & OR < 1     ~ color_sig_negative,
        TRUE                ~ color_nonsig),
      sig_label = case_when(
        !is_sig             ~ "Not significant",
        is_sig & OR > 1     ~ "Higher odds (sig.)",
        is_sig & OR < 1     ~ "Lower odds (sig.)",
        TRUE                ~ "Not significant")) %>%
    ungroup() %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Reference cluster row: OR = 1.0 by definition
  ref_rows <- params %>%
    filter(cluster == ref_cluster) %>%
    mutate(
      or_color  = color_reference,
      OR        = 1.0,
      log_odds  = 0,
      sig_label = "Reference (OR = 1.0)") %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Combine and set factor levels for consistent ordering
  bind_rows(or_colored, ref_rows) %>%
    mutate(
      cluster   = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      sig_label = factor(sig_label, levels = names(or_legend_colors)))
}

# Build the long-format OR data for each reference class
c1_or_long <- build_or_long(c1_params, c1_pairs, ref_cluster = "C1", ref_num = 1)
c3_or_long <- build_or_long(c3_params, c3_pairs, ref_cluster = "C3", ref_num = 3)

# Predictor display order and titles (shared across all four predictor sections below)
pred_order  <- c("female", "Zwave1age", "Zpeer_subuse", "pub_timing")
pred_titles <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"        = "Autonomy Supportive Parenting")

# x-axis cluster label colors to match cluster colors (same for C1 and C3 panels)
x_axis_colors_c1 <- cluster_colors[c("C1", "C2", "C3", "C4")]
x_axis_colors_c3 <- cluster_colors[c("C1", "C2", "C3", "C4")]
```

## Autonomy-Supportive Parenting {#sec-aspline}

```{r}
#| label: c1-line-asp
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
pred_var   <- "Zasp"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zasp"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_asp <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-asp-save
#| include: false

ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_asp,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-asp
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
pred_var   <- "Zasp"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zasp"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_asp <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-asp-save
#| include: false

ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_asp,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-asp-combined
#| echo: false
#| layout-ncol: 2



print(p_line_c1_asp)
print(p_line_c3_asp)
```

## Biological Sex

```{r}
#| label: c1-line-sex
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
pred_var   <- "female"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
pairs_pred_name <- "bio_sexD"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_female <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-sex-save
#| include: false

ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_female,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-sex
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
pred_var   <- "female"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
pairs_pred_name <- "bio_sexD"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_female <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-sex-save
#| include: false

ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_female,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-sex-combined
#| echo: false
#| layout-ncol: 2



print(p_line_c1_female)
print(p_line_c3_female)
```

## Child Age

```{r}
#| label: c1-line-age
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
pred_var   <- "Zwave1age"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zwave1age"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_age <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-age-save
#| include: false

ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_age,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-age
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
pred_var   <- "Zwave1age"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zwave1age"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_age <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-age-save
#| include: false

ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_age,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-age-combined
#| echo: false
#| layout-ncol: 2



print(p_line_c1_age)
print(p_line_c3_age)
```

## Peer Substance Use

```{r}
#| label: c1-line-peer
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
pred_var   <- "Zpeer_subuse"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zpeer_subuse"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_peer <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-peer-save
#| include: false

ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_peer,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-peer
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
pred_var   <- "Zpeer_subuse"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zpeer_subuse"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_peer <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-peer-save
#| include: false

ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_peer,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-peer-combined
#| echo: false
#| layout-ncol: 2



print(p_line_c1_peer)
print(p_line_c3_peer)
```

## Pubertal Timing

```{r}
#| label: c1-line-pub
#| code-fold: true
#| code-summary: "C1 graph code"

# Filter to this predictor
pred_var   <- "pub_timing"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "pub_timing"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_pub <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-pub-save
#| include: false

ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_pub,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-pub
#| code-fold: true
#| code-summary: "C3 graph code"

# Filter to this predictor
pred_var   <- "pub_timing"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "pub_timing"

# Significant pairwise differences not involving the reference (for caption)
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_pub <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-pub-save
#| include: false

ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_pub,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-pub-combined
#| echo: false
#| layout-ncol: 2



print(p_line_c1_pub)
print(p_line_c3_pub)
```
