---
title: "Odds Ratios: Predictors & ASP Scales"
toc: true
toc-depth: 3
code-fold: false
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

# Load shared figure settings and packages used for SPSS import, formatted tables, and caption wrapping
source("_setup.R")
library(gt)  # Used for the data-view tables added in this consolidation
library(stringr)  # Str_wrap(), used to wrap figure captions

# Figure output directories
fig_dir <- "../../figures/add-health"
fig_or <- file.path(fig_dir, "odds-ratios")
fig_or_c1 <- file.path(fig_or, "C1-reference")
fig_or_c3 <- file.path(fig_or, "C3-reference")

# Figure export settings
fig_dpi <- 300

# Create figure directories if needed
dir.create(fig_or_c1, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_or_c3, recursive = TRUE, showWarnings = FALSE)
```

# Overview

This is the second page of the Dissertation Study 1 workflow. The
[Profiles](Profiles.qmd) page established a four-class latent profile solution
capturing distinct combinations of positive and negative adolescent
risk-taking. This page asks the next question: **which demographic, peer, and
parenting characteristics distinguish membership in those profiles?**

Posterior-probability-weighted multinomial logistic regression was estimated
using the BCH three-step method so predictor associations account for
classification uncertainty in the latent profiles [@bolck2004]. Continuous
predictors were standardized before analysis.

The inferential workflow proceeds in three steps:

- **1. Import and clean model output.** Parameter estimates and pairwise class
  comparisons are imported from the saved model-results workbook and reshaped
  into consistent long-format objects for plotting and interpretation
- **2. Correct pairwise comparisons for multiple testing.** Six class-pair
  comparisons are possible for each predictor, so Benjamini-Hochberg (BH)
  correction is applied within each predictor's comparison set to control the
  false discovery rate [@benjamini1995a; @benjamini2000a]
- **3. Converted log-odds to odds ratios.** Odds ratios show how a one-unit
  increase—or one standard deviation for standardized predictors—changes the
  relative odds of membership in one class versus the selected reference class

Models are shown with both **C1 (High PR–Low NR)** and **C3 (Low PR–High NR)**
as reference classes so comparisons can be interpreted from both the
low-negative-risk and high-negative-risk sides of the profile solution.

::: {.callout-note appearance="simple" icon="false"}
### Analysis Workflow

**1. Profiles** → **2. Odds Ratios** → **3. Interaction & Comparisons**

[Profiles](Profiles.qmd) · [Odds Ratios](Odds-Ratios.qmd) · [Interaction & Comparisons](Interaction-and-Comparisons.qmd)
:::


The next page, [Interaction & Comparisons](Interaction-and-Comparisons.qmd), tests whether the
autonomy-supportive parenting association varies by adolescent sex or age and
places BH-corrected profile comparisons side by side.

```{r}
#| label: helper-tidy-params
# Shared by every OR section below to clean data import
tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds") %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds)) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}
```

```{r}
#| label: helper-apply-bh
#| include: false

# For sheets with lowercase covariate/cluster_a/cluster_b (predictors paired comparisons)
apply_bh <- function(raw) {
  raw %>%
    rename(predictor = covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}

# For sheets with capitalized Covariate/clustera/clusterb (ASP, interaction,
# Indicator-pairs). Reused on the Interaction & Comparisons page
apply_bh_cap <- function(raw) {
  raw %>%
    rename(predictor = Covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}
```

# Data Import and Cleaning

Odds-ratio parameters and paired comparisons for each predictor,
referenced against Class 1 and Class 3 respectively, are read in from
the project's Excel results workbook. Import code is included but not
rendered below.

```{r}
#| label: import-predictor-or
#| include: false

# Import multinomial-regression parameters and pairwise comparisons for both reference-class specifications
c1_params_raw <- read_excel(data_path, sheet = "c1 as reference parameters")
c3_params_raw <- read_excel(data_path, sheet = "c3 as reference parameters")
c1_pairs_raw  <- read_excel(data_path, sheet = "predictors paired comparisons")
c3_pairs_raw  <- read_excel(data_path, sheet = "predictors paired (C3) ")
```

**Data Import of C1 as Reference**

```{r}
#| label: data-view-or-raw
# View first rows of c1 reference as data import example
gt(c1_params_raw) |> tab_header(title = "c1 as reference parameters")
```

**Data Import of Paired Comparisons Between Classes (clusters)**

```{r}
#| label: data-view-or-raw2
# View last rows of paired comparisons as data import example
gt(c1_pairs_raw) |>
  tab_header(title = "predictors paired comparisons") |>
  cols_label(
    cluster_a = "Class",
    cluster_b = "Class" )
```

**Clean Data Import**

```{r}
#| label: wrangle-params-predictor-or
# ── Define Predictor Labels ───────────────────────────────────────────────────────
pred_labels <- c(
  "female"       = "Sex (Male = 0)",
  "bio_sexD"     = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"         = "ASP")

# ── Reshape Reference-Class Parameters ────────────────────────────────────────────
c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
c3_params <- tidy_params(c3_params_raw, c("Cluster3", "Cluster4", "Cluster1", "Cluster2"))
```

## BH Correction on Pairwise Comparisons {#sec-BH}

```{r}
#| label: bh-pairs-predictor-or
#| include: false
# ── Apply BH Correction and Display Labels ────────────────────────────────────────
# Codes (e.g. "bio_sexD") stay in place since every figure build chunk
# Filters on them; predictor_label is only for display in tables
c1_pairs <- apply_bh(c1_pairs_raw) %>%
  mutate(predictor_label = recode(predictor, !!!pred_labels))
c3_pairs <- apply_bh(c3_pairs_raw) %>%
  mutate(predictor_label = recode(predictor, !!!pred_labels))
```

**Cleaned Data with C1 as Reference**

```{r}
#| label: data-view-or-tidy
# Hide predictor_var (raw code) from the preview, keep predictor_label
gt(c1_params %>% select(-predictor_var)) |>
  tab_header(title = "data after custom function to clean output")
```

**Cleaned Data of Paired Comparisons Between Classes**

```{r}
#| label: data-view-or-tidy2
# Show predictor_label instead of the raw "predictor" code
gt(c1_pairs %>% select(predictor_label, everything(), -predictor)) |>
  tab_header(title = "data after custom function to clean output") |>
  cols_label(
    predictor_label = "Predictor",
    cluster_a = "Class",
    cluster_b = "Class" )
```

# Odds Ratios {#sec-OR}

Odds ratios test whether each predictor distinguishes class membership,
computed twice: once with C1 (High PR-Low NR) as the reference class,
once with C3 (Low PR-High NR), so both alcohol-problem-status groupings
can anchor the comparison.

::: {.callout-tip appearance="simple" icon="false"}
**OR Interpretation**

An OR above 1 means higher odds of membership in the comparison class
relative to the reference, for a one-unit (or one-SD, for standardized
predictors) increase in that predictor.
:::

## Autonomy-Supportive Parenting

```{r}
#| label: c1-asp
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c1_asp <- c1_params %>% filter(predictor_var == "Zasp")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c1_asp <- c1_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "* Significantly different than C1"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
              } else {
                 "No additional pairwise differences were significant after BH correction."
}

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c1_asp <- ggplot(c1_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-asp-save
#| include: false

# Save the C1-reference autonomy-supportive parenting odds-ratio figure
ggsave(file.path(fig_or_c1, "Zasp_OR_C1ref.png"),
       plot = p_c1_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

```{r}
#| label: c3-asp
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c3_asp <- c3_params %>% filter(predictor_var == "Zasp")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c3_asp <- c3_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "* Significantly different than C3"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) { 
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c3_asp <- ggplot(c3_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-asp-save
#| include: false

# Save the C3-reference autonomy-supportive parenting odds-ratio figure
ggsave(file.path(fig_or_c3, "Zasp_OR_C3ref.png"),
       plot = p_c3_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

```{r}
#| label: fig-asp-combined
#| echo: false
#| layout-ncol: 2

# Render the combined autonomy-supportive parenting odds-ratio comparison
print(p_c1_asp)
print(p_c3_asp)
```

**Autonomy-Supportive Parenting.** As ASP increases, the odds of being
in any other profile compared to C1 (High PR–Low NR) significantly
decrease. Compared to C3 (Low PR–High NR), as ASP increases the odds of
being in either of the two zero-alcohol-problem profiles (C1, C2)
decrease; C2 vs. C3 and C2 vs. C4 were also significant. This is
consistent with self-determination theory [@deci2013; @ryan2022]:
autonomy-supportive parenting predicted profile membership even within
the two no-alcohol-problem profiles, which a variable-centered approach
combining them would have obscured [@bergman1997].

## Biological Sex

```{r}
#| label: c1-sex
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c1_sex <- c1_params %>% filter(predictor_var == "female")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)  # Cluster numbers sig different from C1

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)  # Non-reference sig pairs for caption

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c1_sex <- c1_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title   <- "Sex (Female vs. Male)"
x_label      <- "Reference: C1"
y_label      <- "Odds Ratio - C1 as Reference"    
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c1_sex <- ggplot(c1_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-sex-save
#| include: false

# Save the C1-reference child-sex odds-ratio figure
ggsave(file.path(fig_or_c1, "bio_sexD_OR_C1ref.png"),
       plot = p_c1_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-sex
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c3_sex <- c3_params %>% filter(predictor_var == "female")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c3_sex <- c3_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Sex (Female vs. Male)"
# X_label     <- "Reference: C3"
# Y_label     <- "Odds Ratio - C3 as Reference"
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c3_sex <- ggplot(c3_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-sex-save
#| include: false

# Save the C3-reference child-sex odds-ratio figure
ggsave(file.path(fig_or_c3, "bio_sexD_OR_C3ref.png"),
       plot = p_c3_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-sex-combined
#| echo: false
#| layout-ncol: 2

# Render the combined child-sex odds-ratio comparison
print(p_c1_sex)
print(p_c3_sex)
```

**Sex.** As the proportion of females in a profile increases, compared
to C1 the odds of being in C2 or C3 significantly decrease; compared to
C3, the odds of being in C1 or C4 increase (C2 vs. C4 and C3 vs. C4 also
significant) — consistent with prior findings of higher male delinquency
rates [@miller2010].

## Child Age

```{r}
#| label: c1-age
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c1_age <- c1_params %>% filter(predictor_var == "Zwave1age")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c1_age <- c1_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Age"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
               }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c1_age <- ggplot(c1_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-age-save
#| include: false

# Save the C1-reference age odds-ratio figure
ggsave(file.path(fig_or_c1, "Zwave1age_OR_C1ref.png"),
       plot = p_c1_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-age
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c3_age <- c3_params %>% filter(predictor_var == "Zwave1age")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c3_age <- c3_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Age"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c3_age<- ggplot(c3_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-age-save
#| include: false

# Save the C3-reference age odds-ratio figure
ggsave(file.path(fig_or_c3, "Zwave1age_OR_C3ref.png"),
       plot = p_c3_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-age-combined
#| echo: false
#| layout-ncol: 2

# Render the combined age odds-ratio comparison
print(p_c1_age)
print(p_c3_age)
```

**Child Age.** As age increases, the odds of being in any profile
compared to C1 increase significantly; compared to C3, the odds of being
in either zero-alcohol-problem profile decrease (C2 vs. C3 and C2 vs. C4
also significant) — consistent with developmental increases in negative
risk-taking across adolescence [@duell2018; @moffitt2017].

## Peer Substance Use

```{r}
#| label: c1-peer
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c1_peer <- c1_params %>% filter(predictor_var == "Zpeer_subuse")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c1_peer <- c1_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c1_peer <- ggplot(c1_peer, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-peer-save
#| include: false

# Save the C1-reference peer-substance-use odds-ratio figure
ggsave(file.path(fig_or_c1, "Zpeer_subuse_OR_C1ref.png"),
       plot = p_c1_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-peer
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c3_peer <- c3_params %>% filter(predictor_var == "Zpeer_subuse")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c3_peer <- c3_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c3_peer <- ggplot(c3_peer,aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-peer-save
#| include: false

# Save the C3-reference peer-substance-use odds-ratio figure
ggsave(file.path(fig_or_c3, "Zpeer_subuse_OR_C3ref.png"),
       plot = p_c3_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-peer-combined
#| echo: false
#| layout-ncol: 2

# Render the combined peer-substance-use odds-ratio comparison
print(p_c1_peer)
print(p_c3_peer)
```

**Peer Substance Use.** As peer substance use increases, the odds of
being in any profile compared to C1 increase significantly; compared to
C3, the odds of being in either zero-alcohol-problem profile or C4
decrease (all six pairwise comparisons significant — the only predictor
with every comparison reaching significance), matching literature
identifying peer behavior as a robust proximal predictor of adolescent
risk [@dishion1999; @dishion2011a; @henneberger2021].

## Pubertal Timing

```{r}
#| label: c1-pub
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c1_pub <- c1_params %>% filter(predictor_var == "pub_timing")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c1_pub <- c1_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
             } else { 
                 "No additional pairwise differences were significant after BH correction."
}

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c1_pub <- ggplot(c1_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c1-pub-save
#| include: false

# Save the C1-reference pubertal-timing odds-ratio figure
ggsave(file.path(fig_or_c1, "pub_timing_OR_C1ref.png"),
       plot = p_c1_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: c3-pub
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
c3_pub <- c3_params %>% filter(predictor_var == "pub_timing")

# ── Identify Reference-Class Differences ──────────────────────────────────────────
ref_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# ── Prepare Plot Annotations ──────────────────────────────────────────────────────
c3_pub <- c3_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# ── Define Plot Labels ────────────────────────────────────────────────────────────
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# ── Build Odds-Ratio Figure ───────────────────────────────────────────────────────
p_c3_pub <- ggplot(c3_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, 
       caption = str_wrap(fig_caption, width = 60)) +
  or_theme +
  theme( legend.position    = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: c3-pub-save
#| include: false

# Save the C3-reference pubertal-timing odds-ratio figure
ggsave(file.path(fig_or_c3, "pub_timing_OR_C3ref.png"),
       plot = p_c3_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-pub-combined
#| echo: false
#| layout-ncol: 2

# Render the combined pubertal-timing odds-ratio comparison
print(p_c1_pub)
print(p_c3_pub)
```

**Pubertal Timing.** As pubertal timing gets earlier, the odds of being
in C2 decrease compared to C1, while the odds of being in either
alcohol-problem profile increase compared to C1; compared to C3, the
odds of being in either zero-alcohol-problem profile decrease (C2 vs. C3
and C2 vs. C4 also significant) — reinforcing that biological maturation
shapes risk propensity above and beyond chronological age
[@natsuaki2009].

# Odds Ratios - Lines {#sec-ORline}

::: {.callout-tip appearance="minimal" icon="false"}
The following chunks of code create the foundation for the odd ratio
line graphs, or you can skip to the figures directly: @sec-aspline.
:::

**Significance Color Settings**

```{r}
#| label: sig-settings
# Colors for OR data points
color_sig_positive <- "#228833"  # Green  — OR > 1, sig vs. reference
color_sig_negative <- "#EE6677"  # Purple — OR < 1, sig vs. reference
color_nonsig       <- "#BBBBBB"  # Grey   — not sig vs. reference
color_reference    <- "black"  # Black  — reference cluster (OR = 1.0 by definition)

# Reference line at OR = 1.0
or_refline_color <- "grey20"
or_refline_type  <- "dashed"

# Point shape (size and line width set in appearance-settings below)
or_point_shape <- 21  # Filled circle with border
```

**Figure Appearance Settings**

```{r}
#| label: appearance-settings
# Axis labels
ylab_or_c1 <- "Odds Ratio (vs. C1)"
ylab_or_c3 <- "Odds Ratio (vs. C3)"
xlab_or_C1 <- "Reference: C1"
xlab_or_C3 <- "Reference: C3"

# Font and text sizes — inherited from _setup.R:
# Plot_font, title_size, axis_title_size, axis_text_size, legend_text_size

# Point and line appearance — word-doc scale
or_point_shape <- 21  # Filled circle with border
or_point_size  <- 6
or_line_size   <- 3

# Legend color lookup (used by both line and bar figures)
or_legend_colors <- c(
  "Higher odds (sig.)"   = color_sig_positive,
  "Lower odds (sig.)"    = color_sig_negative,
  "Not significant"      = color_nonsig,
  "Reference (OR = 1.0)" = color_reference)

# Shared legend panel — appended to combined figures via patchwork
p_or_legend <- ggplot(
  data.frame(
    x     = 1:4,
    y     = 1,
    label = factor(names(or_legend_colors), levels = names(or_legend_colors))),
  aes(x = x, y = y, fill = label, color = label)) +
  geom_point(shape = or_point_shape, size = or_point_size + 1, stroke = 0.5) +
  scale_fill_manual(values  = or_legend_colors, name = NULL) +
  scale_color_manual(values = or_legend_colors, name = NULL) +
  guides(
    fill  = guide_legend(nrow = 1, keywidth = 0.6, keyheight = 0.6,
                         override.aes = list(shape = or_point_shape,
                                             size  = or_point_size)),
    color = "none") +
  coord_cartesian(ylim = c(100, 101)) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = legend_text_size, family = plot_font),
    plot.background = element_rect(fill = NA, color = NA))
```

**Build OR Data: Helper Function**

```{r}
#| label: build-or-data
# Build_or_long():
# Params     = output of tidy_params() for a given reference class
# Pairs      = output of apply_bh() for a given reference class
# Ref_cluster = "C1" or "C3" — the reference cluster label
# Ref_num    = 1 or 3 — the numeric cluster number used in pairs columns
# Returns a long data frame with one row per predictor × cluster,
# With columns: predictor_var, predictor_label, cluster, log_odds, OR, or_color
# The reference cluster row is added with OR = 1.0 and or_color = color_reference

build_or_long <- function(params, pairs, ref_cluster, ref_num) {

  # Map each predictor's tidy_params() name to its name in the pairs sheet
  pred_var_to_pairs_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing",
    "Zasp"        = "Zasp")

  # Non-reference clusters: flag significance vs. reference and assign color
  or_colored <- params %>%
    filter(cluster != ref_cluster) %>%
    rowwise() %>%
    mutate(
      pairs_pred = pred_var_to_pairs_name[predictor_var],
      is_sig = {
        hit <- pairs %>%
          filter(
            predictor == pairs_pred,
            (cluster_a == ref_num | cluster_b == ref_num),
            sig == TRUE) %>%
          mutate(other = ifelse(cluster_a == ref_num, cluster_b, cluster_a)) %>%
          filter(other == as.integer(gsub("C", "", as.character(cluster))))
        nrow(hit) > 0
      },
      or_color  = case_when(
        !is_sig             ~ color_nonsig,
        is_sig & OR > 1     ~ color_sig_positive,
        is_sig & OR < 1     ~ color_sig_negative,
        TRUE                ~ color_nonsig),
      sig_label = case_when(
        !is_sig             ~ "Not significant",
        is_sig & OR > 1     ~ "Higher odds (sig.)",
        is_sig & OR < 1     ~ "Lower odds (sig.)",
        TRUE                ~ "Not significant")) %>%
    ungroup() %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Reference cluster row: OR = 1.0 by definition
  ref_rows <- params %>%
    filter(cluster == ref_cluster) %>%
    mutate(
      or_color  = color_reference,
      OR        = 1.0,
      log_odds  = 0,
      sig_label = "Reference (OR = 1.0)") %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Combine and set factor levels for consistent ordering
  bind_rows(or_colored, ref_rows) %>%
    mutate(
      cluster   = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      sig_label = factor(sig_label, levels = names(or_legend_colors)))
}

# Build the long-format OR data for each reference class
c1_or_long <- build_or_long(c1_params, c1_pairs, ref_cluster = "C1", ref_num = 1)
c3_or_long <- build_or_long(c3_params, c3_pairs, ref_cluster = "C3", ref_num = 3)

# Predictor display order and titles (shared across all four predictor sections below)
pred_order  <- c("female", "Zwave1age", "Zpeer_subuse", "pub_timing")
pred_titles <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"        = "Autonomy Supportive Parenting")

# X-axis cluster label colors to match cluster colors (same for C1 and C3 panels)
x_axis_colors_c1 <- cluster_colors[c("C1", "C2", "C3", "C4")]
x_axis_colors_c3 <- cluster_colors[c("C1", "C2", "C3", "C4")]
```

## Autonomy-Supportive Parenting {#sec-aspline}

```{r}
#| label: c1-line-asp
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zasp"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zasp"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_asp <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-asp-save
#| include: false

# Save the alternative C1-reference ASP odds-ratio line figure
ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_asp,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-asp
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zasp"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zasp"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_asp <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-asp-save
#| include: false

# Save the alternative C3-reference ASP odds-ratio line figure
ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_asp,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-asp-combined
#| echo: false
#| layout-ncol: 2

# Render the combined ASP odds-ratio line figure
print(p_line_c1_asp)
print(p_line_c3_asp)
```

## Biological Sex

```{r}
#| label: c1-line-sex
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "female"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
pairs_pred_name <- "bio_sexD"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_female <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-sex-save
#| include: false

# Save the alternative C1-reference child-sex odds-ratio line figure
ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_female,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-sex
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "female"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
pairs_pred_name <- "bio_sexD"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_female <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-sex-save
#| include: false

# Save the alternative C3-reference child-sex odds-ratio line figure
ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_female,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-sex-combined
#| echo: false
#| layout-ncol: 2

# Render the combined child-sex odds-ratio line figure
print(p_line_c1_female)
print(p_line_c3_female)
```

## Child Age

```{r}
#| label: c1-line-age
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zwave1age"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zwave1age"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_age <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-age-save
#| include: false

# Save the alternative C1-reference age odds-ratio line figure
ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_age,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-age
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zwave1age"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zwave1age"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_age <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-age-save
#| include: false

# Save the alternative C3-reference age odds-ratio line figure
ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_age,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-age-combined
#| echo: false
#| layout-ncol: 2

# Render the combined age odds-ratio line figure
print(p_line_c1_age)
print(p_line_c3_age)
```

## Peer Substance Use

```{r}
#| label: c1-line-peer
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zpeer_subuse"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zpeer_subuse"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_peer <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-peer-save
#| include: false

# Save the alternative C1-reference peer-substance-use odds-ratio line figure
ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_peer,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-peer
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "Zpeer_subuse"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "Zpeer_subuse"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_peer <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-peer-save
#| include: false

# Save the alternative C3-reference peer-substance-use odds-ratio line figure
ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_peer,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-peer-combined
#| echo: false
#| layout-ncol: 2

# Render the combined peer-substance-use odds-ratio line figure
print(p_line_c1_peer)
print(p_line_c3_peer)
```

## Pubertal Timing

```{r}
#| label: c1-line-pub
#| code-fold: true
#| code-summary: "C1 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "pub_timing"
pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "pub_timing"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c1 <- c1_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c1 <- if (length(other_sig_c1) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c1, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c1_pub <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "#BBBBBB",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c1, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c1),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c1-line-pub-save
#| include: false

# Save the alternative C1-reference pubertal-timing odds-ratio line figure
ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
       plot   = p_line_c1_pub,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: c3-line-pub
#| code-fold: true
#| code-summary: "C3 graph code"

# ── Select Predictor and Significant Comparisons ──────────────────────────────────
pred_var   <- "pub_timing"
pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
pred_title <- pred_titles[pred_var]

# This predictor's name in the pairs sheet
pairs_pred_name <- "pub_timing"

# ── Identify Additional Pairwise Differences ──────────────────────────────────────
other_sig_c3 <- c3_pairs %>%
  filter(predictor == pairs_pred_name,
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair_label)

# Caption text summarizing those additional differences
fig_caption_c3 <- if (length(other_sig_c3) > 0) {
  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
         paste(other_sig_c3, collapse = "; "), ".")
} else {
  "No additional pairwise differences were significant after BH correction."
}

# Build this predictor's panel
p_line_c3_pub <- ggplot(pred_data, aes(x = cluster, y = OR)) +
  geom_hline(yintercept = 1.0,
             linetype   = or_refline_type,
             color      = or_refline_color,
             linewidth  = 0.4) +
  geom_line(aes(group = 1),
            color     = "grey65",
            linewidth = or_line_size,
            linetype  = "solid") +
  geom_point(aes(fill = sig_label, color = sig_label),
             shape  = or_point_shape,
             size   = or_point_size,
             stroke = 0.5) +
  scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
  scale_y_log10(
    labels = label_number(accuracy = 0.1),
    expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = pred_title,
       caption = str_wrap(fig_caption_c3, width = 60)) +
  theme_classic(base_size = axis_text_size, base_family = plot_font) +
  theme(
    legend.position    = "none",
    plot.caption    = element_text(size = 14, hjust = .5, face = "plain", 
                                   family = plot_font),
    plot.title         = element_text(size = axis_title_size, face = "plain",
                                      hjust = 0.5, family = plot_font),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                      family = plot_font, angle = 0,
                                      hjust = .5, vjust = 1,
                                      color = x_axis_colors_c3),
    axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                      face = "plain"),
    axis.ticks.x       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = NA, color = NA),
    panel.background   = element_rect(fill = NA, color = NA))
```

```{r}
#| label: c3-line-pub-save
#| include: false

# Save the alternative C3-reference pubertal-timing odds-ratio line figure
ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
       plot   = p_line_c3_pub,
       width  = 14,
       height = 8,
       dpi    = fig_dpi,
       units  = "in",
       bg     = "transparent")
```

```{r}
#| label: fig-line-pub-combined
#| echo: false
#| layout-ncol: 2

# Render the combined pubertal-timing odds-ratio line figure
print(p_line_c1_pub)
print(p_line_c3_pub)
```
