---
title: "Odds Ratios: Predictors & ASP Scales"
toc: true
toc-depth: 3
code-fold: true
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(gt)      # used for the data-view tables added in this consolidation
library(stringr)   # str_wrap(), used to wrap figure captions
```

# Overview

This page tests whether the demographic, parenting, and peer predictors described on [Profiles](Profiles.qmd) actually distinguish class membership, using posterior-probability-weighted multinomial logistic regression via the BCH three-step method to correctly account for measurement error in profile assignment (Bolck et al., 2004), with all continuous predictors standardized prior to analysis. Benjamini-Hochberg correction is then applied in R to the resulting pairwise comparisons (Benjamini & Hochberg, 1995, 2000). Predictor odds ratios (bar and line versions) are shown first, followed by the ASP-subscale-level breakdown.

# Helper Functions

Both `tidy_params()` and `apply_bh()`/`apply_bh_cap()` were previously defined separately in each of the three original files (`03_OR_graphs.qmd`, `04_OR_graphs_asp_scales.qmd`, `07_OR_graphsline.qmd`) — identically or near-identically each time. Defined once here, they're reused by every section below. Two BH-correction variants are kept (not one) because the underlying Excel sheets genuinely use two different column-naming conventions: the predictor-pairs sheet uses lowercase `covariate`/`cluster_a`/`cluster_b`, while the ASP, interaction, and indicator-pairs sheets all use `Covariate`/`clustera`/`clusterb`. `apply_bh_cap()` is also reused later on the [Interaction & Comparisons](Interaction-and-Comparisons.qmd) page, since those sheets share the same capitalized convention.

```{r}
#| label: helper-tidy-params

# Shared by every OR section below
tidy_params <- function(raw, cluster_cols) {
  raw %>%
    rename_with(~ "predictor_var", 1) %>%
    rename(omnibus_p = `p-value`) %>%
    pivot_longer(
      cols      = all_of(cluster_cols),
      names_to  = "cluster_raw",
      values_to = "log_odds") %>%
    mutate(
      predictor_label = recode(predictor_var, !!!pred_labels),
      log_odds        = as.numeric(log_odds),
      cluster         = recode(cluster_raw,
                          "Cluster1" = "C1", "Cluster2" = "C2",
                          "Cluster3" = "C3", "Cluster4" = "C4"),
      cluster         = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      OR              = exp(log_odds)) %>%
    select(predictor_label, predictor_var, cluster, log_odds, OR, omnibus_p)
}
```

```{r}
#| label: helper-apply-bh

# For sheets with lowercase covariate/cluster_a/cluster_b (predictors paired comparisons).
apply_bh <- function(raw) {
  raw %>%
    rename(predictor = covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}

# For sheets with capitalized Covariate/clustera/clusterb (ASP, interaction,
# indicator-pairs). Reused on the Interaction & Comparisons page.
apply_bh_cap <- function(raw) {
  raw %>%
    rename(predictor = Covariate, p_value = `p-value`) %>%
    filter(predictor != "Intercept") %>%
    mutate(
      p_value = as.numeric(p_value),
      p_adj   = p.adjust(p_value, method = "BH"),
      sig     = p_adj < .05)
}
```

# 1. Predictor Odds Ratios

Odds-ratio parameters and paired comparisons for each predictor, referenced against Class 1 and Class 3 respectively, are read in from the project's Excel results workbook. Import code is included but not rendered below.

::: {.callout-note}
The code for this section is collapsed by default, but can be expanded individually for each section or all together with the toggle on the upper right.
:::

```{r}
#| label: import-predictor-or
#| include: false
c1_params_raw <- read_excel(data_path, sheet = "c1 as reference parameters")
c3_params_raw <- read_excel(data_path, sheet = "c3 as reference parameters")
c1_pairs_raw  <- read_excel(data_path, sheet = "predictors paired comparisons")
c3_pairs_raw  <- read_excel(data_path, sheet = "predictors paired (C3) ")
```

```{r}
#| label: data-view-or-raw
gt(head(c1_params_raw, 5)) |> tab_header(title = "c1 as reference parameters")
gt(head(c1_pairs_raw, 5)) |> tab_header(title = "predictors paired comparisons")
```


## Wrangle Parameter Sheets

```{r}
#| label: wrangle-params-predictor-or

# Display labels for each predictor
pred_labels <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing",
  "Zasp"         = "ASP")

# Tidy each reference-class parameter sheet
c1_params <- tidy_params(c1_params_raw, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
c3_params <- tidy_params(c3_params_raw, c("Cluster3", "Cluster4", "Cluster1", "Cluster2"))
```

## BH Correction on Pairwise Comparisons

```{r}
#| label: bh-pairs-predictor-or
#| include: false
c1_pairs <- apply_bh(c1_pairs_raw)
```

```{r}
#| label: data-view-or-tidy
gt(head(c1_params, 6)) |> tab_header(title = "After: tidy_params() output")
gt(head(c1_pairs, 6))  |> tab_header(title = "After: apply_bh() output")
```

Odds ratios test whether each predictor distinguishes class membership, computed twice — once with C1 (High PR-Low NR) as the reference class, once with C3 (Low PR-High NR) — so both alcohol-problem-status groupings can anchor the comparison. An OR above 1 means higher odds of membership in the comparison class relative to the reference, for a one-unit (or one-SD, for standardized predictors) increase in that predictor.

**Autonomy-Supportive Parenting.** As ASP increases, the odds of being in any other profile compared to C1 (High PR–Low NR) significantly decrease. Compared to C3 (Low PR–High NR), as ASP increases the odds of being in either of the two zero-alcohol-problem profiles (C1, C2) decrease; C2 vs. C3 and C2 vs. C4 were also significant. This is consistent with self-determination theory (Deci & Ryan, 2013; Ryan & Deci, 2022): autonomy-supportive parenting predicted profile membership even within the two no-alcohol-problem profiles, which a variable-centered approach combining them would have obscured (Bergman & Magnusson, 1997).

**Sex.** As the proportion of females in a profile increases, compared to C1 the odds of being in C2 or C3 significantly decrease; compared to C3, the odds of being in C1 or C4 increase (C2 vs. C4 and C3 vs. C4 also significant) — consistent with prior findings of higher male delinquency rates (Miller et al., 2010).

**Age.** As age increases, the odds of being in any profile compared to C1 increase significantly; compared to C3, the odds of being in either zero-alcohol-problem profile decrease (C2 vs. C3 and C2 vs. C4 also significant) — consistent with developmental increases in negative risk-taking across adolescence (Duell et al., 2018; Moffitt, 2017).

**Peer Substance Use.** As peer substance use increases, the odds of being in any profile compared to C1 increase significantly; compared to C3, the odds of being in either zero-alcohol-problem profile or C4 decrease (all six pairwise comparisons significant — the only predictor with every comparison reaching significance), matching literature identifying peer behavior as a robust proximal predictor of adolescent risk (Dishion et al., 1999; Dishion & Tipsord, 2011; Henneberger et al., 2021).

**Pubertal Timing.** As pubertal timing gets earlier, the odds of being in C2 decrease compared to C1, while the odds of being in either alcohol-problem profile increase compared to C1; compared to C3, the odds of being in either zero-alcohol-problem profile decrease (C2 vs. C3 and C2 vs. C4 also significant) — reinforcing that biological maturation shapes risk propensity above and beyond chronological age (Natsuaki et al., 2009).

## Figures: C1 as Reference Group

### Figure: Sex - C1 Reference

```{r}
#| label: c1-sex

# Filter to this predictor
c1_sex <- c1_params %>% filter(predictor_var == "female")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_sex <- c1_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Sex (Female vs. Male)"
x_label      <- "Reference: C1"
y_label      <- "Odds Ratio - C1 as Reference"    
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_sex <- ggplot(c1_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)),
                    axis.title.x = element_blank() ,
                   axis.text.x = element_text(size = 32),
                   axis.title = element_text(size = 40),
                  axis.title.y = element_blank() )
```

```{r}
#| label: fig-c1-sex
#| echo: false

print(p_c1_sex)
```

```{r}
#| label: c1-sex-save
#| include: false

ggsave(file.path(fig_or_c1, "bio_sexD_OR_C1ref.png"),
       plot = p_c1_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

### Figure: Age - C1 Reference

```{r}
#| label: c1-age

# Filter to this predictor
c1_age <- c1_params %>% filter(predictor_var == "Zwave1age")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_age <- c1_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Age"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
               }

# Build the plot
p_c1_age <- ggplot(c1_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)), 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                 axis.text.x = element_text(size = 32),
                   axis.title = element_text(size = 40) )
```

```{r}
#| label: fig-c1-age
#| echo: false

print(p_c1_age)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c1-age-save
#| include: false

ggsave(file.path(fig_or_c1, "Zwave1age_OR_C1ref.png"),
       plot = p_c1_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Peer Substance Use - C1 Reference

```{r}
#| label: c1-peer

# Filter to this predictor
c1_peer <- c1_params %>% filter(predictor_var == "Zpeer_subuse")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_peer <- c1_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c1_peer <- ggplot(c1_peer, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme +  theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) , 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                 axis.text.x = element_text(size = 32),
                   axis.title = element_text(size = 40) )
```

```{r}
#| label: fig-c1-peer
#| echo: false

print(p_c1_peer)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c1-peer-save
#| include: false

ggsave(file.path(fig_or_c1, "Zpeer_subuse_OR_C1ref.png"),
       plot = p_c1_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Pubertal Timing - C1 Reference

```{r}
#| label: c1-pub

# Filter to this predictor
c1_pub <- c1_params %>% filter(predictor_var == "pub_timing")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_pub <- c1_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
             } else { 
                 "No additional pairwise differences were significant after BH correction."
}

# Build the plot
p_c1_pub <- ggplot(c1_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) , 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                 axis.text.x = element_text(size = 32),
                   axis.title = element_text(size = 40) )
```

```{r}
#| label: fig-c1-pub
#| echo: false

print(p_c1_pub)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c1-pub-save
#| include: false

ggsave(file.path(fig_or_c1, "pub_timing_OR_C1ref.png"),
       plot = p_c1_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Autonomy Supportive Parenting - C1 Reference

```{r}
#| label: c1-asp

# Filter to this predictor
c1_asp <- c1_params %>% filter(predictor_var == "Zasp")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 1 | cluster_b == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 1, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 1, cluster_b != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c1_asp <- c1_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "* Significantly different than C1"
x_label     <- "Reference: C1"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
              } else {
                 "No additional pairwise differences were significant after BH correction."
}

# Build the plot
p_c1_asp <- ggplot(c1_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: fig-c1-asp
#| echo: false

print(p_c1_asp)
cat("\n\n*", fig_caption, "*\n\n")
```

```{r}
#| label: c1-asp-save
#| include: false

ggsave(file.path(fig_or_c1, "Zasp_OR_C1ref.png"),
       plot = p_c1_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

------------------------------------------------------------------------

## Figures: C3 as Reference Group

### Figure: Sex - C3 Reference

```{r}
#| label: c3-sex

# Filter to this predictor
c3_sex <- c3_params %>% filter(predictor_var == "female")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "bio_sexD",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_sex <- c3_sex %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Sex (Female vs. Male)"
#x_label     <- "Reference: C3"
#y_label     <- "Odds Ratio - C3 as Reference"
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_sex <- ggplot(c3_sex, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
           linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) ,
                    axis.title.x = element_blank() ,
                   axis.text.x = element_text(size = 28) ,axis.title.y = element_blank() )
```

```{r}
#| label: fig-c3-sex
#| echo: false

print(p_c3_sex)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c3-sex-save
#| include: false

ggsave(file.path(fig_or_c3, "bio_sexD_OR_C3ref.png"),
       plot = p_c3_sex, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Age - C3 Reference

```{r}
#| label: c3-age

# Filter to this predictor
c3_age <- c3_params %>% filter(predictor_var == "Zwave1age")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zwave1age",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_age <- c3_age %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Age"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_age<- ggplot(c3_age, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) , 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                   axis.text.x = element_text(size = 28) )
```

```{r}
#| label: fig-c3-age
#| echo: false

print(p_c3_age)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c3-age-save
#| include: false

ggsave(file.path(fig_or_c3, "Zwave1age_OR_C3ref.png"),
       plot = p_c3_age, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Peer Substance Use - C3 Reference

```{r}
#| label: c3-peer

# Filter to this predictor
c3_peer <- c3_params %>% filter(predictor_var == "Zpeer_subuse")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zpeer_subuse",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_peer <- c3_peer %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Peer Substance Use"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_peer <- ggplot(c3_peer, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) , 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                   axis.text.x = element_text(size = 28) )
```

```{r}
#| label: fig-c3-peer
#| echo: false

print(p_c3_peer)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c3-peer-save
#| include: false

ggsave(file.path(fig_or_c3, "Zpeer_subuse_OR_C3ref.png"),
       plot = p_c3_peer, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Pubertal Timing - C3 Reference

```{r}
#| label: c3-pub

# Filter to this predictor
c3_pub <- c3_params %>% filter(predictor_var == "pub_timing")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "pub_timing",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_pub <- c3_pub %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "Pubertal Timing"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) {
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_pub <- ggplot(c3_pub, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme + theme(plot.caption    = element_text(size = 14, hjust = .5, face = "plain", family = plot_font,
                                   margin = margin(t = 8)) , 
                   axis.title.y = element_blank(),
                   axis.title.x = element_blank() ,
                   axis.text.x = element_text(size = 28) )
```

```{r}
#| label: fig-c3-pub
#| echo: false

print(p_c3_pub)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c3-pub-save
#| include: false

ggsave(file.path(fig_or_c3, "pub_timing_OR_C3ref.png"),
       plot = p_c3_pub, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

------------------------------------------------------------------------

### Figure: Autonomy Supportive Parenting - C3 Reference

```{r}
#| label: c3-asp

# Filter to this predictor
c3_asp <- c3_params %>% filter(predictor_var == "Zasp")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a == 3 | cluster_b == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(cluster_a == 3, cluster_b, cluster_a)) %>%
  pull(other)

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairs %>%
  filter(predictor == "Zasp",
         cluster_a != 3, cluster_b != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
  pull(pair)

# Add asterisk labels and their y-position
c3_asp <- c3_asp %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title  <- "* Significantly different than C3"
x_label     <- "Reference: C3"
y_label     <- y_lab
fig_caption <- if (length(other_sig) > 0) { 
                 paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                        paste(other_sig, collapse = "; "), ".")
               } else {
                 "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p_c3_asp <- ggplot(c3_asp, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = "gray20",
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank())
```

```{r}
#| label: fig-c3-asp
#| echo: false

print(p_c3_asp)
cat("\n*Caption:*", fig_caption, "\n")
```

```{r}
#| label: c3-asp-save
#| include: false

ggsave(file.path(fig_or_c3, "Zasp_OR_C3ref.png"),
       plot = p_c3_asp, width = 3.25, height = 8, dpi = fig_dpi, units = "in",
       bg = "transparent")
```

------------------------------------------------------------------------

## Figure Labels
```{r}
#| label: legend-labels-1

library(ggplot2)

# Legend text for each cluster
legend_labels <- c(
  "C1"      = "C1: High PR - Low NR (35.6%)",
  "C2"      = "C2: Low PR - Low NR (33.7%)",
  "C3"      = "C3: Low PR - High NR(15.8%)",
  "C4"      = "C4: High PR - High PR (14.9%)")

# Dummy data to drive the legend (bars pushed off-plot below)
dummy_df <- data.frame(
  x       = 1:4,
  y       = 1:4,
  cluster = factor(names(legend_labels), levels = names(legend_labels)))

# Build the legend-only plot
p <- ggplot(dummy_df, aes(x = x, y = y, fill = cluster)) +
  geom_col() +
  scale_fill_manual(values = cluster_colors, labels = legend_labels, name = NULL) +
  guides(fill = guide_legend(ncol = 1, keywidth = 1.2, keyheight = 1.2)) +
  coord_cartesian(ylim = c(100, 101)) +   # pushes bars completely out of view
  theme_void() +
  theme(
    legend.position.inside = c(0.5, 0.5),
    legend.position        = "inside",
    legend.text            = element_text(size = 14),
    legend.spacing.y       = unit(6, "pt"))
```

```{r}
#| label: legend-labels-1-save
#| include: false

ggsave(file.path(fig_or_c1, "legend_only.png"),
       plot = p, width = 14, height = 8, dpi = 300, units = "in")
```

```{r}
#| label: legend-labels-2

# Legend text for each cluster (alternate wording, with risk-taker labels)
legend_labels2 <- c(
  "C1"      = "C1: Adaptive Risk-Takers     | High PR-Low NR (35.6%)",
  "C2"      = "C2: Low Risk-Takers            | Low PR-Low NR (33.7%)",
  "C3"      = "C3: Maldaptive Risk-Takers | Low PR-High NR (15.8%)",
  "C4"      = "C4: High Risk-Takers           | High PR-High NR (14.9%)")

# Dummy data to drive the legend (bars pushed off-plot below)
dummy_df2 <- data.frame(
  x       = 1:4,
  y       = 1:4,
  cluster = factor(names(legend_labels2), levels = names(legend_labels2)))

# Build the legend-only plot, single column
p2 <- ggplot(dummy_df2, aes(x = x, y = y, fill = cluster)) +
  geom_col() +
  scale_fill_manual(values = cluster_colors, labels = legend_labels2, name = NULL) +
  guides(fill = guide_legend(ncol = 1, keywidth = 1.2, keyheight = 1.2)) +
  coord_cartesian(ylim = c(100, 101)) +   # pushes bars completely out of view
  theme_void() +
  theme(
    legend.position.inside = c(0.5, 0.5),
    legend.position        = "inside",
    legend.text            = element_text(size = 14),
    legend.spacing.y       = unit(6, "pt"))

# Build the legend-only plot, two columns
p3 <- ggplot(dummy_df2, aes(x = x, y = y, fill = cluster)) +
  geom_col() +
  scale_fill_manual(values = cluster_colors, labels = legend_labels2, name = NULL) +
  guides(fill = guide_legend(ncol = 2, keywidth = 1.2, keyheight = 1.2)) +
  coord_cartesian(ylim = c(100, 101)) +   # pushes bars completely out of view
  theme_void() +
  theme(
    legend.position.inside = c(0.5, 0.5),
    legend.position        = "inside",
    legend.text            = element_text(size = 14),
    legend.spacing.y       = unit(6, "pt"))
```

```{r}
#| label: legend-labels-2-save
#| include: false

ggsave(file.path(fig_or_c1, "legend_only2.png"),
       plot = p2, width = 8, height = 4, dpi = 300, units = "in")

ggsave(file.path(fig_or_c1, "legend_only3.png"),
       plot = p3, width = 14, height = 8, dpi = 300, units = "in")
```



------------------------------------------------------------------------

### Data

Line-graph version of the predictor odds ratios above — same underlying
`c1_params`, `c3_params`, `c1_pairs`, and `c3_pairs` objects built in
[Predictor Odds Ratios](#1-predictor-odds-ratios) (bar version), re-plotted
here as one line per cluster so the trend across all four classes is visible
at a glance, where the bar version emphasizes pairwise significance against
the reference class instead. Reusing those objects also means the line
panels now pick up the same, complete `pred_labels` mapping (including ASP)
that the bar panels use — the original standalone version of this file had
its own, incomplete copy of `pred_labels` that omitted ASP. ASP itself is
still excluded from the predictor-panel loop below by design (its
odds-ratio pattern is shown as its own dedicated figure in Predictor Odds
Ratios), not because the label is missing.

### Significance Color Settings

```{r}
#| label: sig-settings

# Colors for OR points — edit hex codes to change
color_sig_positive <- "#228833"   # green  — OR > 1, sig vs. reference
color_sig_negative <- "#EE6677"   # purple — OR < 1, sig vs. reference
color_nonsig       <- "#BBBBBB"   # grey   — not sig vs. reference
color_reference    <- "black"     # black  — reference cluster (OR = 1.0 by definition)

# Reference line at OR = 1.0
or_refline_color <- "grey20"
or_refline_type  <- "dashed"

# Point shape (size and line width set in appearance-settings below)
or_point_shape <- 21    # filled circle with border
```

### Figure Appearance Settings

```{r}
#| label: appearance-settings

# Axis labels
ylab_or_c1 <- "Odds Ratio (vs. C1)"
ylab_or_c3 <- "Odds Ratio (vs. C3)"
xlab_or_C1 <- "Reference: C1"
xlab_or_C3 <- "Reference: C3"

# Font and text sizes — inherited from _setup.R:
#   plot_font, title_size, axis_title_size, axis_text_size, legend_text_size

# Point and line appearance — word-doc scale
or_point_shape <- 21   # filled circle with border
or_point_size  <- 3
or_line_size   <- 0.8

# Legend color lookup (used by both line and bar figures)
or_legend_colors <- c(
  "Higher odds (sig.)"   = color_sig_positive,
  "Lower odds (sig.)"    = color_sig_negative,
  "Not significant"      = color_nonsig,
  "Reference (OR = 1.0)" = color_reference)

# Shared legend panel — appended to combined figures via patchwork
p_or_legend <- ggplot(
  data.frame(
    x     = 1:4,
    y     = 1,
    label = factor(names(or_legend_colors), levels = names(or_legend_colors))),
  aes(x = x, y = y, fill = label, color = label)) +
  geom_point(shape = or_point_shape, size = or_point_size + 1, stroke = 0.5) +
  scale_fill_manual(values  = or_legend_colors, name = NULL) +
  scale_color_manual(values = or_legend_colors, name = NULL) +
  guides(
    fill  = guide_legend(nrow = 1, keywidth = 0.6, keyheight = 0.6,
                         override.aes = list(shape = or_point_shape,
                                             size  = or_point_size)),
    color = "none") +
  coord_cartesian(ylim = c(100, 101)) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = legend_text_size, family = plot_font),
    plot.background = element_rect(fill = NA, color = NA))
```

### Build OR Data: Helper Function

```{r}
#| label: build-or-data

# build_or_long():
#   params     = output of tidy_params() for a given reference class
#   pairs      = output of apply_bh() for a given reference class
#   ref_cluster = "C1" or "C3" — the reference cluster label
#   ref_num    = 1 or 3 — the numeric cluster number used in pairs columns
#
# Returns a long data frame with one row per predictor × cluster,
# with columns: predictor_var, predictor_label, cluster, log_odds, OR, or_color
# The reference cluster row is added with OR = 1.0 and or_color = color_reference.

build_or_long <- function(params, pairs, ref_cluster, ref_num) {

  # Map each predictor's tidy_params() name to its name in the pairs sheet
  pred_var_to_pairs_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing")
    #"Zasp"        = "Zasp")

  # Non-reference clusters: flag significance vs. reference and assign color
  or_colored <- params %>%
    filter(cluster != ref_cluster) %>%
    rowwise() %>%
    mutate(
      pairs_pred = pred_var_to_pairs_name[predictor_var],
      is_sig = {
        hit <- pairs %>%
          filter(
            predictor == pairs_pred,
            (cluster_a == ref_num | cluster_b == ref_num),
            sig == TRUE) %>%
          mutate(other = ifelse(cluster_a == ref_num, cluster_b, cluster_a)) %>%
          filter(other == as.integer(gsub("C", "", as.character(cluster))))
        nrow(hit) > 0
      },
      or_color  = case_when(
        !is_sig             ~ color_nonsig,
        is_sig & OR > 1     ~ color_sig_positive,
        is_sig & OR < 1     ~ color_sig_negative,
        TRUE                ~ color_nonsig),
      sig_label = case_when(
        !is_sig             ~ "Not significant",
        is_sig & OR > 1     ~ "Higher odds (sig.)",
        is_sig & OR < 1     ~ "Lower odds (sig.)",
        TRUE                ~ "Not significant")) %>%
    ungroup() %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Reference cluster row: OR = 1.0 by definition
  ref_rows <- params %>%
    filter(cluster == ref_cluster) %>%
    mutate(
      or_color  = color_reference,
      OR        = 1.0,
      log_odds  = 0,
      sig_label = "Reference (OR = 1.0)") %>%
    select(predictor_var, predictor_label, cluster, log_odds, OR, or_color, sig_label)

  # Combine and set factor levels for consistent ordering
  bind_rows(or_colored, ref_rows) %>%
    mutate(
      cluster   = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
      sig_label = factor(sig_label, levels = names(or_legend_colors)))
}

# Build the long-format OR data for each reference class
c1_or_long <- build_or_long(c1_params, c1_pairs, ref_cluster = "C1", ref_num = 1)
c3_or_long <- build_or_long(c3_params, c3_pairs, ref_cluster = "C3", ref_num = 3)
```

### Figure: C1 as Reference — One Panel per Predictor (Line)

```{r}
#| label: fig-or-c1-panels
#| fig-width: 1.75
#| fig-height: 4.0

# Predictor display order and titles
pred_order  <- c("female", "Zwave1age", "Zpeer_subuse", "pub_timing")
pred_titles <- c(
  "female"       = "Sex (Male = 0)",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer Substance Use",
  "pub_timing"   = "Pubertal Timing")
  # "Zasp"        = "Autonomy Supportive Parenting")

# x-axis cluster label colors to match cluster colors
x_axis_colors_c1 <- cluster_colors[c("C1", "C2", "C3", "C4")]

# Build and print one line graph per predictor (C1 reference)
plots_or_c1 <- lapply(pred_order, function(pred_var) {

  # This predictor's long-format OR data and display title
  pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  # This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  # Significant pairwise differences not involving the reference (for caption)
  other_sig_c1 <- c1_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 1, cluster_b != 1,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  # Caption text summarizing those additional differences
  fig_caption_c1 <- if (length(other_sig_c1) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c1, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  # Build this predictor's panel
  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.4) +
    geom_line(aes(group = 1),
              color     = "#BBBBBB",
              linewidth = or_line_size,
              linetype  = "solid") +
    geom_point(aes(fill = sig_label, color = sig_label),
               shape  = or_point_shape,
               size   = or_point_size,
               stroke = 0.5) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_log10(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.15, 0.15))) +
    labs(title = pred_title, caption = fig_caption_c1) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c1),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA))

  # Display and save this panel
  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c1, "\n\n")

  ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_line.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c1) <- pred_order
```

------------------------------------------------------------------------

### Figure: C3 as Reference — One Panel per Predictor (Line)

```{r}
#| label: fig-or-c3-panels
#| fig-width: 1.75
#| fig-height: 4.0

# x-axis cluster label colors for C3 reference panels
x_axis_colors_c3 <- cluster_colors[c("C1", "C2", "C3", "C4")]

# Build and print one line graph per predictor (C3 reference)
plots_or_c3 <- lapply(pred_order, function(pred_var) {

  # This predictor's long-format OR data and display title
  pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  # This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  # Significant pairwise differences not involving the reference (for caption)
  other_sig_c3 <- c3_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 3, cluster_b != 3,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  # Caption text summarizing those additional differences
  fig_caption_c3 <- if (length(other_sig_c3) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c3, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  # Build this predictor's panel
  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.4) +
    geom_line(aes(group = 1),
              color     = "grey65",
              linewidth = or_line_size,
              linetype  = "solid") +
    geom_point(aes(fill = sig_label, color = sig_label),
               shape  = or_point_shape,
               size   = or_point_size,
               stroke = 0.5) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_color_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_log10(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.15, 0.15))) +
    labs(title = pred_title, caption = fig_caption_c3) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c3),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA))

  # Display and save this panel
  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c3, "\n\n")

  ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_line.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c3) <- pred_order
```

------------------------------------------------------------------------

### Bar Plots: C1 as Reference — One Panel per Predictor

```{r}
#| label: fig-or-c1-bar-panels
#| fig-width: 1.75
#| fig-height: 4.0

plots_or_c1_bar <- lapply(pred_order, function(pred_var) {

  # This predictor's long-format OR data and display title
  pred_data  <- c1_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  # This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  # Significant pairwise differences not involving the reference (for caption)
  other_sig_c1 <- c1_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 1, cluster_b != 1,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  # Caption text summarizing those additional differences
  fig_caption_c1 <- if (length(other_sig_c1) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c1, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  # Build this predictor's panel
  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR, fill = sig_label)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.5) +
    geom_col(width = 0.65, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_continuous(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.05, 0.15))) +
    labs(title = pred_title, caption = fig_caption_c1) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c1),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA))

  # Display and save this panel
  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c1, "\n\n")

  ggsave(file.path(fig_or_c1, paste0(pred_var, "_OR_C1ref_bar.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c1_bar) <- pred_order
```

------------------------------------------------------------------------

### Bar Plots: C3 as Reference — One Panel per Predictor

```{r}
#| label: fig-or-c3-bar-panels
#| fig-width: 1.75
#| fig-height: 4.0

plots_or_c3_bar <- lapply(pred_order, function(pred_var) {

  # This predictor's long-format OR data and display title
  pred_data  <- c3_or_long %>% filter(predictor_var == pred_var)
  pred_title <- pred_titles[pred_var]

  # This predictor's name in the pairs sheet (differs from tidy_params() name for sex)
  pairs_pred_name <- c(
    "female"       = "bio_sexD",
    "Zwave1age"    = "Zwave1age",
    "Zpeer_subuse" = "Zpeer_subuse",
    "pub_timing"   = "pub_timing"
  )[pred_var]

  # Significant pairwise differences not involving the reference (for caption)
  other_sig_c3 <- c3_pairs %>%
    filter(predictor == pairs_pred_name,
           cluster_a != 3, cluster_b != 3,
           sig == TRUE) %>%
    mutate(pair_label = paste0("C", cluster_a, " vs. C", cluster_b)) %>%
    pull(pair_label)

  # Caption text summarizing those additional differences
  fig_caption_c3 <- if (length(other_sig_c3) > 0) {
    paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
           paste(other_sig_c3, collapse = "; "), ".")
  } else {
    "No additional pairwise differences were significant after BH correction."
  }

  # Build this predictor's panel
  plot_obj <- ggplot(pred_data, aes(x = cluster, y = OR, fill = sig_label)) +
    geom_hline(yintercept = 1.0,
               linetype   = or_refline_type,
               color      = or_refline_color,
               linewidth  = 0.5) +
    geom_col(width = 0.65, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = or_legend_colors, name = NULL, drop = FALSE) +
    scale_y_continuous(
      labels = label_number(accuracy = 0.1),
      expand = expansion(mult = c(0.05, 0.15))) +
    labs(title = pred_title, caption = fig_caption_c3) +
    theme_classic(base_size = axis_text_size, base_family = plot_font) +
    theme(
      legend.position    = "none",
      plot.title         = element_text(size = axis_title_size, face = "plain",
                                        hjust = 0.5, family = plot_font),
      axis.title.y       = element_blank(),
      axis.title.x       = element_blank(),
      axis.text.x        = element_text(size = axis_text_size, face = "plain",
                                        family = plot_font, angle = 0,
                                        hjust = .5, vjust = 1,
                                        color = x_axis_colors_c3),
      axis.text.y        = element_text(size = axis_text_size, family = plot_font,
                                        face = "plain"),
      axis.ticks.x       = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.background    = element_rect(fill = NA, color = NA),
      panel.background   = element_rect(fill = NA, color = NA))

  # Display and save this panel
  print(plot_obj)
  cat("\n*Caption:*", fig_caption_c3, "\n\n")

  ggsave(file.path(fig_or_c3, paste0(pred_var, "_OR_C3ref_bar.png")),
         plot   = plot_obj,
         width  = 14,
         height = 8,
         dpi    = fig_dpi,
         units  = "in",
         bg     = "transparent")

  plot_obj
})
names(plots_or_c3_bar) <- pred_order
```

------------------------------------------------------------------------

### Legend Panel (standalone inspect / export)

```{r}
#| label: fig-or-legend
#| fig-width: 6.5
#| fig-height: 1.0
#| echo: false

print(p_or_legend)
```

```{r}
#| label: or-legend-save
#| include: false

ggsave(file.path(fig_or_c1, "legend_or.png"),
       plot = p_or_legend,
       width = 6.5, height = 1.0,
       dpi = fig_dpi, units = "in",
       bg  = "transparent")
```

------------------------------------------------------------------------

# 2. ASP Scale Odds Ratios

## Overview

Same odds-ratio approach as Predictor Odds Ratios above, restricted to the three Autonomy Supportive Parenting (ASP) subscales — support, involvement, and warmth — tested individually rather than as the single composite ASP score used elsewhere. The dissertation reports ASP odds ratios only at the composite level; the subscale-level breakdown here is exploratory and not something the dissertation text interprets directly. SDT treats involvement (parental attention/engagement) and autonomy support proper (perspective-taking, minimizing pressure) as conceptually separable but interrelated facets (Grolnick, 2009; Teuber et al., 2022), so if one subscale predicts profile membership more strongly than the others, that would suggest not all forms of autonomy support are equally proximal to positive and negative risk-taking.

## Data Import

ASP scale odds-ratio parameters and paired comparisons, by reference class, are read in from the project's Excel results workbook. Import code is included but not rendered below.

::: {.callout-note}
The code for this section is collapsed by default, but can be expanded individually for each section or all together with the toggle on the upper right.
:::
```{r}
#| label: import-asp-or
#| echo: false
c1_params_asp <- read_excel(data_path, sheet = "asp scale parameters")
c3_params_asp <- read_excel(data_path, sheet = "asp scale parameters C3")
c1_pairs_asp <- read_excel(data_path, sheet = "asp paired C1")
c3_pairs_asp  <- read_excel(data_path, sheet = "asp paired C3")
```

```{r}
#| label: data-view-asp-or-raw
gt(head(c1_params_asp, 5)) |> tab_header(title = "asp scale parameters")
gt(head(c1_pairs_asp, 5)) |> tab_header(title = "asp paired C1")
```

## Wrangle Parameter Sheets
```{r}
#| label: wrangle-params-asp-or

# Display labels for each ASP subscale
pred_labels <- c(
  "parentwarmth"       = "Parental Warmth",
  "parentinvolve"    = "Parental Involvement",
  "autonomysupport" = "Autonomy Support")

# Tidy each reference-class parameter sheet
c1_asp_or <- tidy_params(c1_params_asp, c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))
c3_asp_or <- tidy_params(c3_params_asp, c("Cluster3", "Cluster4", "Cluster1", "Cluster2"))
```

### BH Correction on Pairwise Comparisons
```{r}
#| label: bh-pairs-asp-or
# apply_bh_cap() is the shared helper defined in Helper Functions above (handles
# the "Covariate"/clustera/clusterb sheet layout used by the ASP, interaction,
# and indicator-pairs sheets). The dead apply_b2h()/c1_pairsasp2/c3_pairsasp2
# from the original file are removed here -- that second correction was
# computed but never referenced by any downstream figure or table.
c1_pairsasp <- apply_bh_cap(c1_pairs_asp)
```

```{r}
#| label: data-view-asp-or-tidy
gt(head(c1_pairsasp, 6)) |> tab_header(title = "After: BH-corrected ASP pairwise comparisons")
```

*RCM: 2-3 sentences on which subscale (support, involvement, or warmth) most strongly predicts profile membership, once you've looked at the figures below — the dissertation doesn't report this breakdown, so this is the one place in the portfolio without dissertation text to draw the interpretation from directly.

## Figures: C1 as Reference Group

### Figure: Parental Warmth - C1 Reference

```{r}
#| label: c1-warmth

# Filter to this predictor
c1_warmth <- c1_asp_or %>% filter(predictor_var == "parentwarmth")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairsasp %>%
  filter(predictor == "parentwarmth",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairsasp %>%
  filter(predictor == "parentwarmth",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_warmth <- c1_warmth %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Parental Warmth"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c1_warmth, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c1-warmth
#| echo: false

print(p)
```

```{r}
#| label: c1-warmth-save
#| include: false

ggsave(file.path(fig_asp_or_c1, "warmth_OR_C1.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```


### Figure: Parental Involvement - C1 Reference

```{r}
#| label: c1-involve

# Filter to this predictor
c1_involve <- c1_asp_or %>% filter(predictor_var == "parentinvolve")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairsasp %>%
  filter(predictor == "parentinvolve",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairsasp %>%
  filter(predictor == "parentinvolve",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_involve <- c1_involve %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Parental Involvement"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c1_involve, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c1-involve
#| fig-width: 6
#| fig-height: 5
#| echo: false

print(p)
```

```{r}
#| label: c1-involve-save
#| include: false

ggsave(file.path(fig_asp_or_c1, "involve_OR_C1.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```


### Figure: Autonomy Support  - C1 Reference

```{r}
#| label: c1-support

# Filter to this predictor
c1_support <- c1_asp_or %>% filter(predictor_var == "autonomysupport")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c1_pairsasp %>%
  filter(predictor == "autonomysupport",
         clustera == 1 | clusterb == 1,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 1, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C1

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c1_pairsasp %>%
  filter(predictor == "autonomysupport",
         clustera != 1, clusterb != 1,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c1_support <- c1_support %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Autonomy Support"
x_label      <- "Reference: C1"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c1_involve, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c1-support
#| fig-width: 6
#| fig-height: 5
#| echo: false

print(p)
```

```{r}
#| label: c1-support-save
#| include: false

ggsave(file.path(fig_asp_or_c1, "support_OR_C1.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```

## Figures: C3 as Reference Group

### Figure: Parental Warmth - C3 Reference

```{r}
#| label: c3-warmth

# Filter to this predictor
c3_warmth <- c3_asp_or %>% filter(predictor_var == "parentwarmth")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairsasp %>%
  filter(predictor == "parentwarmth",
         clustera == 3 | clusterb == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 3, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C3

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairsasp %>%
  filter(predictor == "parentwarmth",
         clustera != 3, clusterb != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c3_warmth <- c3_warmth %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Parental Warmth"
x_label      <- "Reference: C3"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c3_warmth, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c3-warmth
#| fig-width: 6
#| fig-height: 5
#| echo: false

print(p)
```

```{r}
#| label: c3-warmth-save
#| include: false

ggsave(file.path(fig_asp_or_c3, "warmth_OR_C3.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```


### Figure: Parental Involvement - C3 Reference

```{r}
#| label: c3-involve

# Filter to this predictor
c3_involve <- c3_asp_or %>% filter(predictor_var == "parentinvolve")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairsasp %>%
  filter(predictor == "parentinvolve",
         clustera == 3 | clusterb == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 3, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C3

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairsasp %>%
  filter(predictor == "parentinvolve",
         clustera != 3, clusterb != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c3_involve <- c3_involve %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Parental Involvement"
x_label      <- "Reference: C3"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c3_involve, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c3-involve
#| fig-width: 6
#| fig-height: 5
#| echo: false

print(p)
```

```{r}
#| label: c3-involve-save
#| include: false

ggsave(file.path(fig_asp_or_c3, "involve_OR_C3.png"),
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```


### Figure: Autonomy Support  - C3 Reference

```{r}
#| label: c3-support

# Filter to this predictor
c3_support <- c3_asp_or %>% filter(predictor_var == "autonomysupport")

# Cluster(s) significantly different from the reference (for asterisk labels)
ref_sig <- c3_pairsasp %>%
  filter(predictor == "autonomysupport",
         clustera == 3 | clusterb == 3,
         sig == TRUE) %>%
  mutate(other = ifelse(clustera == 3, clusterb, clustera)) %>%
  pull(other)                          # cluster numbers sig different from C3

# Significant pairwise differences not involving the reference (for caption)
other_sig <- c3_pairsasp %>%
  filter(predictor == "autonomysupport",
         clustera != 3, clusterb != 3,
         sig == TRUE) %>%
  mutate(pair = paste0("C", clustera, " vs. C", clusterb)) %>%
  pull(pair)                           # non-reference sig pairs for caption

# Add asterisk labels and their y-position
c3_support <- c3_support %>%
  mutate(
    ast_label = ifelse(as.integer(cluster) %in% ref_sig, "*", ""),
    ast_y     = OR + max(OR) * ast_nudge_pct)

# Plot text: title, axis labels, caption
plot_title   <- "Autonomy Support"
x_label      <- "Reference: C3"
y_label      <- y_lab    # from formatting chunk
fig_caption  <- if (length(other_sig) > 0) { 
                  paste0("Additional significant pairwise differences after BH correction (FDR < .05): ",
                         paste(other_sig, collapse = "; "), ".")
              }  else {
                  "No additional pairwise differences were significant after BH correction."
 }

# Build the plot
p <- ggplot(c3_involve, aes(x = cluster, y = OR, fill = cluster)) +
  geom_col(width = bar_width, alpha = bar_alpha,
            linewidth = bar_outline_lwd) +
  geom_hline(yintercept = 1, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  geom_text(aes(y = OR, label = round(OR, 2)),
            color = or_label_color, size = or_label_size, vjust = or_label_vjust) +
  geom_text(aes(y = ast_y, label = ast_label),
            size = ast_text_size, vjust = 0.5) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = number_format(accuracy = 0.1)) +
  labs(title = plot_title, x = x_label, y = y_label, caption = fig_caption) +
  or_theme
```

```{r}
#| label: fig-c3-support
#| fig-width: 6
#| fig-height: 5
#| echo: false

print(p)
```

```{r}
#| label: c3-support-save
#| include: false

ggsave(file.path(fig_asp_or_c3, "support_OR_C3.png"), 
      plot = p, width = 14, height = 8, dpi = fig_dpi, units = "in")
```
