---
title: "Indicator, Predictor & ASP Profiles"
toc: true
toc-depth: 3
code-fold: true
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(haven)   # read_sav, used in the z-score section
library(gt)      # used for the data-view tables added in this consolidation
```

# Overview

This page covers the descriptive class profiles from the 4-class latent
profile analysis of Add Health Wave I adolescents (*N* = 6,469): the
negative and positive risk-taking indicators that defined the classes,
the demographic and parenting predictors tested as correlates of class
membership, and the parenting-scale-level breakdown of the
autonomy-supportive parenting (ASP) composite used throughout this
project. The inferential odds-ratio work built on these profiles lives
on the [Odds Ratios](Odds-Ratios.qmd) page; effect-modification and
BH-corrected robustness checks live on [Interaction Effects &
Comparisons](Interaction-and-Comparisons.qmd).

# 1. Indicator Profiles

### Overview

These panels show the 4-class latent profile solution's indicator
profiles — the risk and protective indicators that defined the classes
themselves, as distinct from the predictor and covariate profiles in the
sections that follow. Fit indices favored the four-profile solution:
AIC/BIC improvements became marginal beyond four classes, the
five-profile solution failed to replicate (a local maximum), and the
four-profile solution showed adequate entropy (\> .80) with
theoretically coherent classes (LPA conducted in LatentGOLD; Vermunt &
Magidson, 2005).

The four classes are C1 (High PR–Low NR, 35.6%), C2 (Low PR–Low NR,
33.7%), C3 (Low PR–High NR, 15.8%), and C4 (High PR–High NR, 14.9%) — PR
= positive risk-taking, NR = negative risk-taking. Alcohol problems
functioned as a near-binary separator: both C1 and C2 reported
essentially no alcohol problems, while C3 and C4 reported substantially
elevated problem counts. Within each of those two broad groupings,
positive risk indicators (sports, school activities, friend disclosure,
romantic involvement) were the primary source of further differentiation
— adolescents with comparably low negative risk nonetheless occupied
meaningfully distinct profiles with respect to positive risk engagement
([@duell2019], [@duell2021],[@fryt2021]).

**Data Import**

Indicator profile probabilities/means and paired-comparison significance
tests across the latent class solution are read in from a local Excel
results workbook from saved LatentGOLD results. Import code is not
rendered.

```{r}
#| label: import-indicator
#| include: false

data_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/Study1_R/data/Tables.xlsx"

profile_raw <- read_excel(data_path, sheet = "profile R")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")
```

```{r}
#| label: data-view-profile-raw
gt(head(profile_raw, 6)) |> tab_header(title = "profile R — indicator probabilities/means by class")
```

### Wrangle Profile Data

Binary indicators (probabilities, 0–1) and continuous indicators (means,
0–3) are plotted together on a shared primary axis. Continuous values
are divided by `scale_factor` before plotting so they fit the 0–1
primary axis; the right-hand secondary axis reverses that transformation
to show the true 0–3 scale. `scale_factor` is defined in one place —
adjust if indicator ranges change.

```{r}
#| label: wrangle-profile

# Legend labels with cluster size in parentheses
cluster_sizes <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

cluster_labels <- c(
  "C1"      = paste0("C1 (", cluster_sizes["C1"], "%)"),
  "C2"      = paste0("C2 (", cluster_sizes["C2"], "%)"),
  "C3"      = paste0("C3 (", cluster_sizes["C3"], "%)"),
  "C4"      = paste0("C4 (", cluster_sizes["C4"], "%)")
)

# Fill colors — no outlines
indicator_colors <- cluster_colors

# Scale factor: right axis max / left axis max (continuous max ~3, binary max 1)
# Continuous values are divided by this before plotting, then right axis
# multiplies back to show true scale. Edit here if any continuous value > 3.
scale_factor <- 3

# Continuous indicators — values can exceed 1 and read off the right axis
continuous_indicators <- c(
  "Problems from alcohol",
  "Sport Teams",
  "Extracurriculars",
  "Don't avoid problems"
)

# Negative panel: binary on left (0-1), continuous on right (0-3)
#   → continuous values divided by scale_factor to fit primary axis
# Positive panel: continuous on left (0-3), binary on right (0-1)
#   → binary values multiplied by scale_factor to fit primary axis
# This arrangement puts the 0-3 axes adjacent in the middle of the combined figure.

neg_continuous <- c("Problems from alcohol")
pos_continuous <- c("Sport Teams", "Extracurriculars", "Don't avoid problems")

profile_long <- profile_raw %>%
  pivot_longer(
    cols      = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to  = "cluster_raw",
    values_to = "value"
  ) %>%
  mutate(
    cluster    = recode(cluster_raw,
                   "Cluster1" = "C1", "Cluster2" = "C2",
                   "Cluster3" = "C3", "Cluster4" = "C4"),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    Indicators = factor(Indicators, levels = unique(profile_raw$Indicators)),
    # Rescale for plotting:
    #   Negative: continuous / scale_factor (plots on 0-1 axis, reads off right 0-3)
    #   Positive: binary * scale_factor     (plots on 0-3 axis, reads off right 0-1)
    value_plot = case_when(
      Risk == "Negative" & Indicators %in% neg_continuous ~ value / scale_factor,
      Risk == "Positive" & !Indicators %in% pos_continuous ~ value * scale_factor,
      TRUE ~ value
    )
  )
```

### BH Correction on Pairwise Comparisons

Non-estimable rows (p = ".") and the Intercept block are dropped before
correction is applied across all remaining tests simultaneously.

```{r}
#| label: bh-pairs
pairs <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(clustera),
    cluster_b = as.integer(clusterb),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05
  )
```

Benjamini-Hochberg correction is applied here because every indicator is
compared across all six class-pairs simultaneously; without correction,
running this many pairwise tests inflates the false-positive rate well
past .05 (Benjamini & Hochberg, 1995, 2000). Overall, the negative risk
indicators contributed about equally to differentiating classes, while
positive risk indicators were the primary driver of separation within
each alcohol-problem grouping (see Overview above).

```{r}
#| label: data-view-pairs-bh
gt(head(pairs, 6)) |> tab_header(title = "After: BH-corrected pairwise comparisons")
```

### Profile Figure

@fig-indicator-profileneg: Negative Indicators Left y-axis (0–1): binary
indicators, read as probabilities. Right y-axis (0–3): continuous
indicators, read as means.

```{r}
#| label: fig-indicator-profileneg

fig_caption <- paste(
  "Note. Left y-axis (0\u20131) = binary indicators (probability of behavior).",
  "Right y-axis (0\u20133) = continuous indicators (within-class mean).",
  "Continuous bars are rescaled to the primary axis for display;"
)

# make_panel() takes explicit axis config so Negative and Positive can differ.
#
# Negative: left = binary 0-1, right = continuous 0-3
#   primary axis shows 0-1; sec.axis multiplies by scale_factor to show 0-3
#
# Positive: left = continuous 0-3, right = binary 0-1
#   primary axis shows 0-3; sec.axis divides by scale_factor to show 0-1

make_panel <- function(risk_label,
                       left_name = "" ,left_lim, left_breaks,
                       sec_transform, right_name = "", right_breaks) {

  df <- profile_long %>% filter(Risk == risk_label)

  ggplot(df, aes(x = Indicators, y = value_plot, fill = cluster)) +
     geom_vline(
  xintercept = seq(1.5, length(unique(df$Indicators)) - 0.5, by = 1),
  color      = "#DDDDDD",
  linewidth  = 0.4
) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width
    ) +
    scale_fill_manual(values = indicator_colors, labels = cluster_labels,
                      name = NULL) +
    scale_y_continuous(
      #name   = left_name,
      limits = left_lim,
      breaks = left_breaks,
      labels = number_format(accuracy = 0.1),
      expand = expansion(mult = c(0, 0.03)),
      sec.axis = sec_axis(
        transform = sec_transform,
        name      = right_name,
        breaks    = right_breaks,
        labels    = number_format(accuracy = 0.1)
      )
    ) +
    labs(title = paste(risk_label, "Risk Indicators"), y = NULL, caption = fig_caption) +
    base_theme +
    theme(
      axis.text.x        = element_text(angle = 35, hjust = 1,
                                        size = axis_text_size),
      axis.title.y.right = element_text(size = axis_title_size,
                                        color = "gray40"),
      legend.position    = "none",
      plot.title         = element_text(size = title_size, face = "bold",
                                        hjust = 0.5),
          plot.margin       = margin(20, 0, 40, 100)
    )
}

# Negative: left=binary(0-1), right=continuous(0-3)
p_neg <- make_panel(
  risk_label    = "Negative",
  #left_name     = "Probability (binary indicators)",
  left_lim      = c(0, 1.08),
  left_breaks   = seq(0, 1, by = 0.2),
  sec_transform = ~ . * scale_factor,
 # right_name    = "Within-class mean (continuous indicators)",
  right_breaks  = seq(0, scale_factor, by = 1)
)

# Positive: left=continuous(0-3), right=binary(0-1)
p_pos <- make_panel(
  risk_label    = "Positive",
 # left_name     = "Within-class mean (continuous indicators)",
  left_lim      = c(0, scale_factor * 1.08),
  left_breaks   = seq(0, scale_factor, by = 1),
  sec_transform = ~ . / scale_factor,
 # right_name    = "Probability (binary indicators)",
  right_breaks  = seq(0, 1, by = 0.2)
)

print(p_neg)


ggsave(
  filename = file.path(fig_indicator, "profile_figure_neg.png"),
     device = agg_png,
  plot     = p_neg,
  width    = 14, height = 6, dpi = fig_dpi, units = "in"
)
```

@fig-indicator-profilepos: Positive Indicators Left y-axis (0–3):
continuous indicators, read as means. Right y-axis (0–1): binary
indicators, read as probabilities.

```{r}
#| label: fig-indicator-profilepos

print(p_pos)
ggsave(
  filename = file.path(fig_indicator, "profile_figure_pos.png"),
  plot     = p_pos,    
  device = agg_png,
  width    = 14, height = 6, dpi = fig_dpi, units = "in"
)
```

@fig-indicator-profile: Profiles The two panels (Negative risk /
Positive risk) share dual-axis and are combined into one figure.

```{r}
#| label: fig-indicator-profile
#| fig-width: 14
#| fig-height: 6
# Remove left axis from positive panel
# so the shared 0-3 middle axis appears only once
p_neg_trim <- p_neg +
  theme(
    axis.text.y.right = element_text(margin = margin(l = 12)),
    plot.margin       = margin(12, 0, 20, 85)
  )
p_pos_trim <- p_pos +
  theme(
    axis.title.y.left = element_blank(),
    axis.text.y.left  = element_blank(),
   # axis.ticks.y.left = element_blank(),
    plot.margin       = margin(12, 12, 12, 0)    # no left margin only
  )
combined <- (p_neg_trim + p_pos_trim) +
  plot_layout(guides = "collect") &
  theme(legend.position = "none")
combined

ggsave(
  filename = file.path(fig_indicator, "profile_figures_combined.png"),
  plot     = combined,
  width    = 40, height = 6, dpi = fig_dpi, units = "in"
)
```

Sports participation, school activity involvement, and friend disclosure
most clearly distinguish C1 (High PR–Low NR) and C4 (High PR–High NR)
from C2 and C3, while alcohol problems and the broader negative-risk
indicators most clearly separate the two alcohol-problem classes (C3,
C4) from the two non-problem classes (C1, C2). The one indicator with
limited discriminating power across all four classes was problem
approach (class means ranged narrowly from 2.80–2.95, with only 4 of 6
pairwise comparisons significant).

### Table: Non-Significant Comparisons

@tbl-nonsig: List of non-significant pairwise comparisons

```{r}
#| label: tbl-nonsig
nonsig <- pairs %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant pairwise comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

### Significant Comparisons by Profile

@tbl-sig-cluster: Total Significant Pairwise Comparisons by Cluster

```{r}
#| label: tbl-sig-cluster
sig_pairs <- pairs %>% filter(sig == TRUE)

by_cluster <- bind_rows(
  sig_pairs %>% select(cluster = cluster_a, Indicator),
  sig_pairs %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant pairwise comparisons (BH-corrected, FDR < .05) involving each profile.")
```

### Significant Comparisons by Indicator

@tbl-sig-indicator: Total Significant Pairwise Comparisons by Indicator

```{r}
#| label: tbl-sig-indicator
by_indicator <- pairs %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```

### Line Profile Figure

```{r}
#| label: fig-indicator-profileneg-line
fig_caption_line <- paste(
  "Note. Left y-axis (0\u20131) = binary indicators (probability of behavior).",
  "Right y-axis (0\u20133) = continuous indicators (within-class mean).",
  "Continuous values are rescaled to the primary axis for display;"
)

make_panel_line <- function(risk_label,
                            left_name  = "", left_lim, left_breaks,
                            sec_transform, right_name = "", right_breaks) {
  df_line <- profile_long %>% filter(Risk == risk_label)
  ggplot(df_line, aes(x = Indicators, y = value_plot,
                      color = cluster, group = cluster)) +
 geom_vline(
  xintercept = seq(1.5, length(unique(df_line$Indicators)) - 0.5, by = 1),
  color      = "gray90",
  linewidth  = 0.4
) +
    geom_line(linewidth = 3) +
    geom_point(size = 4) + 
    scale_color_manual(values = indicator_colors, labels = cluster_labels,
                       name = NULL) +
    scale_y_continuous(
      limits = left_lim,
      breaks = left_breaks,
      labels = number_format(accuracy = 0.1),
      expand = expansion(mult = c(0, 0.03)),
      sec.axis = sec_axis(
        transform = sec_transform,
        name      = right_name,
        breaks    = right_breaks,
        labels    = number_format(accuracy = 0.1)
      )
    ) +
    labs(title = paste(risk_label, "Risk Indicators"), y = NULL, caption = fig_caption_line) +
    base_theme +
    theme(
      axis.text.x        = element_text(angle = 35, hjust = 1,
                                        size = axis_text_size),
      axis.title.y.right = element_text(size = axis_title_size,
                                        color = "gray40"),
      legend.position    = "none",
      plot.title         = element_text(size = title_size, face = "bold",
                                        hjust = 0.5),
      plot.margin       = margin(12, 0, 20, 85)
    )
}

# Negative: left=binary(0-1), right=continuous(0-3)
p_line_neg <- make_panel_line(
  risk_label    = "Negative",
  left_lim      = c(0, 1.08),
  left_breaks   = seq(0, 1, by = 0.2),
  sec_transform = ~ . * scale_factor,
  right_breaks  = seq(0, scale_factor, by = 1)
)

# Positive: left=continuous(0-3), right=binary(0-1)
p_line_pos <- make_panel_line(
  risk_label    = "Positive",
  left_lim      = c(0, scale_factor * 1.08),
  left_breaks   = seq(0, scale_factor, by = 1),
  sec_transform = ~ . / scale_factor,
  right_breaks  = seq(0, 1, by = 0.2)
)


print(p_line_neg)
ggsave(
  filename = file.path(fig_indicator, "profile_figure_line_neg.png"),
  device   = agg_png,
  plot     = p_line_neg,
  width    = 24, height = 8, dpi = fig_dpi, units = "in"
)

print(p_line_pos)
ggsave(
  filename = file.path(fig_indicator, "profile_figure_line_pos.png"),
  device   = agg_png,
  plot     = p_line_pos,
  width    = 24, height = 8, dpi = fig_dpi, units = "in"
)
```

```{r}
# Remove left axis from positive panel so the shared 0-3 middle axis appears only once
p_neg_trim_line <- p_line_neg +
  theme(
    axis.text.y.right = element_text(margin = margin(l = 12)),
    plot.margin       = margin(12, 0, 20, 85)
  )
p_pos_trim_line <- p_line_pos +
  theme(
    axis.title.y.left = element_blank(),
    axis.text.y.left  = element_blank(),
   # axis.ticks.y.left = element_blank(),
    plot.margin       = margin(12, 12, 12, 0)    # no left margin only
  )
combined_line <- (p_neg_trim_line + p_pos_trim_line) +
  plot_layout(guides = "collect") &
  theme(legend.position = "none")
combined_line

ggsave(
  filename = file.path(fig_indicator, "profile_figures_combined_line.png"),
  plot     = combined_line,
  width    = 24, height = 14, dpi = fig_dpi, units = "in"
)
```

# 2. Indicator Profiles — Standardized (Z-Score)

### Standardized Profile Panels

I also standardized based on the within class average for each
indicator, allowing for a visualization of how each profile compares to
the mean. The code used to create the figures is below.

Raw indicator profiles above mix binary probabilities and continuous
means on different scales; z-scoring every indicator here puts all of
them on the same standardized axis so their relative separation across
classes can be compared directly, which the raw panels cannot show on
their own.

### Variable Labels

```{r}
#| label: variable-labels

#Variable labels for plots
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# "Negative" = negative risk indicators; "Positive" = positive risk indicators
indicator_risk_panel <- c(
  WSDalc               = "Negative",
  cigwhole             = "Negative",
  skipschool           = "Negative",
  liewhere             = "Negative",
  awayhome             = "Negative",
  dare_dum             = "Negative",
  seatbelt             = "Negative",
  drvdrunk             = "Negative",
  BC_dum               = "Negative",
  steal12M             = "Negative",
  phyfight             = "Negative",
  WSDsport             = "Positive",
  WSDactivity          = "Positive",
  avoidproblems        = "Positive",
  disclose             = "Positive",
  intitaterelationship = "Positive")

# Desired x-axis order within each panel — edit to reorder bars
neg_order <- c(
  "WSDalc", "cigwhole", "skipschool", "liewhere", "awayhome",
  "dare_dum", "seatbelt", "drvdrunk", "BC_dum", "steal12M", "phyfight")

pos_order <- c(
  "WSDsport", "WSDactivity", "avoidproblems", "disclose", "intitaterelationship")
```

### Standardization

```{r}
#| label: import-zscore
#| include: false

sav_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/LCAwASP_4CProbabilities.sav"

sav_raw <- read_sav(sav_path)
```

Class probabilities and indicator values are read in from the
dissertation's saved SPSS output (a 4-class LCA solution with an
auxiliary ASP scale). Each indicator is then z-scored so the plotted
classes are on a common scale.

```{r}
#| label: standardize

# Variable names present in the indicator map
indicator_vars <- names(indicator_label_map)

# Standardize each indicator across the full sample
sav_scaled <- sav_raw %>%
  mutate(across(all_of(indicator_vars), ~ as.numeric(scale(.x))))

# Compute within-class means of z-scores
# Class variable assumed to be named "class" — update if different in your file
zscore_long <- sav_scaled %>%
  select(`clu#`, all_of(indicator_vars)) %>%
  pivot_longer(
    cols      = all_of(indicator_vars),
    names_to  = "variable",
    values_to = "z_score"
  ) %>%
  group_by(`clu#`, variable) %>%
  summarise(mean_z = mean(z_score, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    cluster    = paste0("C", `clu#`),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    label      = indicator_label_map[variable],
    Risk       = indicator_risk_panel[variable],
    label      = factor(label, levels = indicator_label_map[c(neg_order, pos_order)])
  )

```

```{r}
#| label: data-view-zscore
gt(head(sav_scaled, 6)) |> tab_header(title = "After: z-scored indicators, by class")
```

### Legend Labels

```{r}
#| label: legend-labels

# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Edit the description for each cluster here
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR"
)

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions, " (", cluster_sizes_z, "%)"),
  names(cluster_sizes_z)
)
cluster_colors <- c(
  "C1" = "#44AA99",
  "C2" = "#88CCEE",   
  "C3" = "#882255",
  "C4" = "#DDCC77"   
)

# Colors — pulls from cluster_colors defined in _setup.R
zscore_colors <- cluster_colors
```

### Z-Score Profile Panels

A horizontal reference line at z = 0 marks the sample average so it is
immediately clear which profiles fall above vs. below the mean on each
indicator.

```{r}
#| label: make-zscore-panel

# Function to create panel for profiles
make_zscore_panel <- function(risk_label) {

  # Set x-axis variable order
  panel_order <- if (risk_label == "Negative") {
    indicator_label_map[neg_order]
  } else {
    indicator_label_map[pos_order]
  }

  # Select columns from dataset
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order))
  
  # Recode for graph 
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order)) %>%
    mutate(mean_z = case_when(
      mean_z >  1.0 ~  1.1,
      mean_z < -0.5 ~ -0.6,
      TRUE          ~  mean_z
    ))

  # Create plot
  ggplot(panel_data, aes(x = label, y = mean_z, fill = cluster)) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.6, linetype = "solid") +
   geom_vline(
    xintercept = seq(1.5, length(panel_order) - 0.5, by = 1),
    color      = "gray63",
    linewidth  = .8
  ) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width
    ) +
    scale_fill_manual(
      values = zscore_colors,
      labels = cluster_labels_z,
      name   = NULL
    ) +
    scale_y_continuous(
      limits = c(-.6, 1.1),
     # breaks = seq(-.5, 1.0, by = 0.5),
    #  labels = number_format(accuracy = 0.1),
    breaks = c(-0.5,  0, 0.5, 1.0),
  labels = c("Below -0.5", "0.0", "0.5", "Above 1.0"),
      expand = expansion(mult = c(0.02, 0.02))
    ) + 
    labs(
      title   = paste(risk_label, "Risk-Taking Indicators"),
      y       = "Mean (Standardized)",
      caption = "Mean standardized (z-scored) indicator values by latent class. Values below −0.5 or above 1.0 are capped for display."
    ) +
    base_theme +
    theme(
      axis.text.x     = element_text(angle = 30, hjust = 1,
                                     size = axis_text_size, family = plot_font),
      legend.position = "none", 
       plot.margin       = margin(40, 0, 40, 100)
    )
}

# Run function for each indicator set
p_zneg <- make_zscore_panel("Negative")
p_zpos <- make_zscore_panel("Positive")

# Save negative panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_neg.png"),
  device   = agg_png,
  plot     = p_zneg,
  width    = 14, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")

# Save positive panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_pos.png"),
  device   = agg_png,
  plot     = p_zpos,
  width    = 14, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")
```

```{r}
#| label: panel-standardized-neg
#| echo: false
print(p_zneg)  
```

```{r}
#| label: panel-standardized-pos
#| echo: false
print(p_zpos)
```

# 3. Predictor Profiles

#### Overview

Predictor profiles show where the four classes fall on the variables
being tested as correlates of class membership — biological sex, age,
peer substance use, pubertal timing, and autonomy-supportive parenting
(ASP) — as distinct from the indicators that defined the classes in the
first place (see Indicator Profiles above). Sex was a moderately
discriminating covariate (4 of 6 pairwise comparisons significant); age,
pubertal timing, and ASP each discriminated broadly (5 of 6
significant); peer substance use discriminated across all 6 comparisons
and was the single most consistent predictor of profile membership.

#### Data Import

Predictor variable profiles across latent classes are read in from the
same local Excel results workbook used throughout this project. Import
code is included but not rendered below.

```{r}
#| label: import-predictor
#| echo: false
pred_raw <- read_excel(data_path, sheet = "predictor profiles R")
```

```{r}
#| label: data-view-pred-raw
gt(head(pred_raw, 6)) |> tab_header(title = "predictor profiles R — predictor means/proportions by class")
```

#### Build Dataframes

Standardized profile dataframe (z-scored predictors + % male).

```{r}
#| label: build-profile-predictor
pred_order <- c("Zmale", "Zwave1age", "Zpeer_subuse", "Zub_timing", "Zasp")

display_labels <- c(
  "Zmale"        = "% Male",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer SU",
  "Zub_timing"   = "Puberty",
  "Zasp"         = "ASP"
)

profile_df <- pred_raw %>%
  filter(Covariate %in% pred_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value"
  ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = pred_order),
    display   = factor(display_labels[as.character(Covariate)],
                       levels = display_labels[pred_order])
  )
```

```{r}
#| label: data-view-pred-long
gt(head(profile_df, 8)) |> tab_header(title = "After: long format, one row per predictor x cluster")
```

Unstandardized dataframe (raw means per cluster).

```{r}
#| label: build-unstandardized
unstand_order <- c("male", "wave1age", "peer_subuse", "pub_timing", "asp")

unstand_df <- pred_raw %>%
  filter(Covariate %in% unstand_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value"
  ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = unstand_order)
  )
```

Age distribution dataframe. Oldest group (18–19) stacks at the top.

```{r}
#| label: build-age
age_order <- c("age_13", "age_14", "age_16", "age_17", "age_18")

age_df <- pred_raw %>%
  filter(Covariate %in% age_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "proportion"
  ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    age_group = factor(label, levels = c(
      "18-19 years old", "17 years old", "16 years old",
      "14-15 years old", "13 years old"
    ))
  )
```

#### Predictor Profile Graph (Standardized)

```{r}
#| label: fig-profile-standardized
#| fig-cap: "Note. Age, Peer Substance Use, Pubertal Timing, and Autonomy Supportive Parenting were standardized (M = 0, SD = 1). Sex dummy-coded (0 = male, 1 = female); % Male shown as raw proportion."
#| fig-width: 14
#| fig-height: 8
# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Create the description for each cluster
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR")


cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions),
  names(cluster_sizes_z))

# Colors — pulls from cluster_colors defined in _setup.R
zscore_colors <- cluster_colors

  
p_profile <- ggplot(profile_df, aes(x = display, y = value, fill = cluster)) +
  geom_vline(
    xintercept = seq(1.5, length(profile_df) - 0.5, by = 1),
    color      = "#DDDDDD",
    linewidth  = 0.4
  ) +
  geom_col(position = position_dodge(width = dodge_width), width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
 scale_fill_manual(values = cluster_colors, labels = cluster_labels_z) +
  scale_y_continuous(
    name   = "Mean (Standardized)",
    limits = c(-.5, 1.15),
    breaks = seq(-.5, 1, by = .5),
    labels = number_format(accuracy = 0.01)
  ) +  
  labs(
    title   = "ASP and Covariate Levels Across Profiles",
    caption = "Note. Age, Peer Substance Use, Pubertal Timing, and Autonomy Supportive Parenting were standardized (M = 0, SD = 1). Sex dummy-coded (0 = male, 1 = female); % Male shown as raw proportion."
  ) +
  base_theme +
  theme(legend.position = "top") +
  guides(fill = guide_legend(nrow = 1))

print(p_profile)

ggsave(file.path(fig_predictor, "Predictor_Profile_Standardized.png"),
       plot = p_profile, width = 13, height = 6.5,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```

Older adolescents and adolescents with earlier pubertal timing were
overrepresented in the two alcohol-problem classes (C3, C4), consistent
with developmental increases in negative risk-taking across adolescence
(Duell et al., 2018; Moffitt, 2017) and with the early-timing hypothesis
linking earlier physical maturation to elevated risk propensity
(Natsuaki et al., 2009). Male adolescents were more likely to fall in
the alcohol-problem classes, consistent with prior findings of higher
delinquency rates among males (Miller et al., 2010). Peer substance use
was the single strongest and most consistent predictor — the only
covariate with all pairwise comparisons against both C1 and C3
significant — matching developmental literature identifying peer
behavior as a robust predictor of adolescent delinquency (Dishion et
al., 1999; Dishion & Tipsord, 2011; Henneberger et al., 2021).

#### Age Distribution

Youngest (13) at bottom; oldest (18–19) at top.

```{r}
#| label: fig-age-distribution
#| fig-cap: "Note. Stacked bars show proportion within each age group per cluster (sum to 1.0). Ordered youngest (bottom) to oldest (top)."
#| fig-width: 14
#| fig-height: 7
p_age <- ggplot(age_df, aes(x = cluster, y = proportion, fill = age_group)) +
  geom_col(position = "stack", width = bar_width) +
  scale_fill_manual(
    values = age_colors, name = "Age Group",
    breaks = c("13 years old","14-15 years old","16 years old",
               "17 years old","18-19 years old")
  ) +
  scale_y_continuous(
    name   = "Proportion",
    limits = c(0, 1.01), expand = c(0, 0),
    breaks = seq(0, 1, by = 0.25),
    labels = number_format(accuracy = 0.01)
  ) +
  labs(
    title   = "Age Distribution by Cluster",
    caption = "Note. Stacked bars show proportion within each age group per cluster (sum to 1.0). Ordered youngest (bottom) to oldest (top)."
  ) +
    or_theme +
  theme(
    axis.title.x    = element_blank(),
    legend.position = "bottom"
  ) +
  guides(fill = guide_legend(nrow = 2))

print(p_age)

ggsave(file.path(fig_predictor, "Age_Distribution.png"),
       plot = p_age, width = 14, height = 7,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```

# 4. ASP Scale Profiles

This is the scale-composite counterpart to the ASP predictor shown above
— support, involvement, and warmth as separate parenting subscales,
rather than the single composite ASP score used elsewhere. The
dissertation's discussion notes this is worth examining at the subscale
level: SDT treats involvement (parental attention/engagement) and
autonomy support proper (perspective-taking, minimizing pressure) as
conceptually separable but interrelated facets (Grolnick, 2009; Teuber
et al., 2022), so subscale-level differences in how strongly each
predicts profile membership would be theoretically informative — a
question the composite-level odds ratios (below) can't answer on their
own.

#### Data Import

ASP scale profiles across latent classes are read in from the project's
Excel results workbook. Import code is included but not rendered below.

```{r}
#| label: import-asp
#| echo: false
asp_raw <- read_excel(data_path, sheet = "asp profiles R")
```

```{r}
#| label: data-view-asp-raw
gt(asp_raw) |> tab_header(title = "asp profiles R — scale means by class")
```

#### Build Dataframes

Standardized profile dataframe (z-scored predictors + % male).

```{r}
#| label: build-profile-asp
asp_pred_order <- c("support", "involve", "warmth", "Zsupport", "Zinvolve", "Zwarmth")

asp_display_labels <- c(
    "support" = "Autonomy Support",
  "involve"    = "Parental Involvement",
    "warmth"     = "Parental Warmth",
    "support" = "Autonomy Support",
  "involve"    = "Parental Involvement",
    "warmth"     = "Parental Warmth"
)

asp_df <- asp_raw %>%
  filter(Covariate %in% asp_pred_order) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value"
  ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = asp_pred_order),
    display   = factor(asp_display_labels[as.character(Covariate)],
                       levels = asp_display_labels[asp_pred_order])
  )
```

```{r}
#| label: data-view-asp-long
gt(head(asp_df, 8)) |> tab_header(title = "After: long format for plotting")
```

#### Profile Figures (Standardized)

```{r}
#| label: fig-asp-Zsupport

p_aspZ <- ggplot(asp_df %>% filter(Covariate == "Zsupport"),
                aes(x = cluster, y = value, fill = cluster)) +
  geom_col(width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(name = "Mean",
                     expand = expansion(mult = c(0.08, 0.08)),
                     labels = number_format(accuracy = 0.01)) +
  labs(title = "Autonomy Support",
       caption = "Mean standardized (z-scored) autonomy-supportive parenting score by latent class.") +
  or_theme +  theme(
    axis.title.x = element_blank())
print(p_aspZ)
ggsave(file.path(fig_asp_profiles, "stand_asp.png"),
       plot = p_aspZ, width = 5, height = 4, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-asp-Zinvolve

p_involveZ <- ggplot(asp_df %>% filter(Covariate == "Zinvolve"),
                aes(x = cluster, y = value, fill = cluster)) +
  geom_col(width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(name = "Mean",
                     expand = expansion(mult = c(0.08, 0.08)),
                     labels = number_format(accuracy = 0.01)) +
  labs(title = "Parental Involvement",
       caption = "Mean standardized (z-scored) parental involvement score by latent class.") +
  or_theme +  theme(
    axis.title.x = element_blank())
print(p_involveZ)
ggsave(file.path(fig_asp_profiles, "stand_involve.png"),
       plot = p_involveZ, width = 5, height = 4, dpi = fig_dpi, units = "in")
```

```{r}
#| label: fig-asp-Zwarmth

p_warmthZ <- ggplot(asp_df %>% filter(Covariate == "Zwarmth"),
                aes(x = cluster, y = value, fill = cluster)) +
  geom_col(width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(name = "Mean",
                     expand = expansion(mult = c(0.08, 0.08)),
                     labels = number_format(accuracy = 0.01)) +
  labs(title = "Parental Warmth",
       caption = "Mean standardized (z-scored) parental warmth score by latent class.") +
  or_theme +  theme(
    axis.title.x = element_blank())
print(p_warmthZ)
ggsave(file.path(fig_asp_profiles, "stand_warmth.png"),
       plot = p_warmthZ, width = 5, height = 4, dpi = fig_dpi, units = "in")
```
