---
title: "Indicator, Predictor & ASP Profiles"
toc: true
toc-depth: 3
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

# Load shared figure settings and packages used for SPSS import, formatted tables, and caption wrapping
source("_setup.R")
library(haven)  # Read_sav, used in the z-score section
library(gt)  # Used for the data-view tables added in this consolidation
library(stringr)  # Str_wrap(), used to wrap figure captions
```

# Overview

This three-page workflow presents analyses from **Dissertation Study
1**, which used Add Health Wave I data (*N* = 6,469) to examine
heterogeneity in adolescents' positive and negative risk-taking and the
developmental factors associated with different patterns of risk
behavior.

This first page establishes the descriptive profile solution. It shows
the indicators that defined each class and summarizes demographic and
developmental predictor patterns across classes.

The next page tests whether demographic, peer, and parenting predictors
distinguish class membership using posterior-probability-weighted
multinomial logistic regression with BCH adjustment for classification
uncertainty. The final page tests whether the ASP association differs by
adolescent sex or age and presents BH-corrected pairwise comparisons
across the profile analyses.

::: {.callout-note appearance="simple" icon="false"}
### Analysis Workflow

**1. Profiles** → **2. Odds Ratios** → **3. Interaction & Comparisons**

[Profiles](Profiles.qmd) · [Odds Ratios](Odds-Ratios.qmd) · [Interaction
& Comparisons](Interaction-and-Comparisons.qmd)
:::

Continue to [Odds Ratios](Odds-Ratios.qmd) for inferential comparisons
of class membership, or to [Interaction &
Comparisons](Interaction-and-Comparisons.qmd) for moderation and
BH-corrected robustness checks.

# Indicator Profiles

### Overview

These panels show the 4-class latent profile solution's indicator
profiles from the negative and positive risk-taking indicators. Fit
indices favored the four-profile solution: AIC/BIC improvements became
marginal beyond four classes, the five-profile solution failed to
replicate (a local maximum), and the four-profile solution showed
adequate entropy (\> .80) with theoretically coherent classes (LPA
conducted in LatentGOLD; [@vermunt2005]).

A four-class latent profile analysis identified groups that differed in
their joint levels of **positive risk-taking (PR)** and **negative
risk-taking (NR)**:

- **C1: High PR–Low NR** (35.6%)
- **C2: Low PR–Low NR** (33.7%)
- **C3: Low PR–High NR** (15.8%)
- **C4: High PR–High NR** (14.9%)

Alcohol problems functioned as a near-binary separator: both **C1 and C2
reported essentially no alcohol problems**, while **C3 and C4 reported
substantially elevated problem counts**. Within each of those two broad
groupings, positive risk indicators (sports, school activities, friend
disclosure, romantic involvement) were the primary source of further
differentiation. However, adolescents with comparably low negative risk
occupied meaningfully distinct profiles with respect to positive risk
engagement [@duell2019a; @duell2021a; @fryt2021a].

To jump straight to figures:

- indicator probabilities/mean: negative risk-taking
  [@fig-indicator-profileneg] and [@fig-indicator-profilepos]
- indicator within-class average: negative [@fig-standardized-neg] and
  positive [@fig-standardized-pos]
- predictor profiles: [@fig-profile-standardized]
- age distribution of each profile: [@fig-age-distribution]

**Data Import**

Indicator profile probabilities/means and paired-comparison significance
tests across the latent class solution are read in from a local Excel
results workbook from saved LatentGOLD results. Import code is not
rendered.

```{r}
#| label: import-indicator
#| include: false

data_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/Study1_R/data/Tables.xlsx"

profile_raw <- read_excel(data_path, sheet = "profile R")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")

# Relabel for visual ease
profile_indicator_recode <- c(
  "Had whole cigarrette"           = "Cigarette",
  "Skip school (no excuse)"        = "Skipped school",
  "Lied about whereabouts"         = "Lied about whereabouts",
  "Night out (no permission)"      = "Night out(unsanctioned)",
  "Done dangerous dare"            = "Dangerous dare",
  "Not always wear seatbelt"       = "No seatbelt",
  "Ever Drive drunk"               = "Drove Drunk",
  "No contraceptives used"         = "Risky Sex",
  "Stole Anything"                 = "Theft",
  "Serious Physical Fight"         = "Fight",
  "Problems from alcohol"          = "Alcohol Problems",
  "Sport Teams"                    = "Sport Teams",
  "Extracurriculars"               = "School Activities",
  "Don't avoid problems"           = "Problem Approach",
  "Disclosed problem to friend"    = "Friend Disclosure",
  "Initiate romantic relationship" = "Initiated Relationship")

profile_raw <- profile_raw %>%
  mutate(Indicators = recode(str_trim(Indicators), !!!profile_indicator_recode))

# Continuous indicators (within-class means, not probabilities/proportions)
# Moved up here (rather than defined down in indicator-profilecreate) so the
# Raw data-view table below can also flag them with an "(M)" suffix
indicator_continuous_fig <- c("Alcohol Problems", "Problem Approach",
                               "Sport Teams", "School Activities")

# Figure output directories
fig_dir <- "../../figures/add-health"
fig_indicator <- file.path(fig_dir, "indicator-profiles")
fig_predictor <- file.path(fig_dir, "predictor-profiles")

# Figure export settings
fig_dpi <- 300

# Create figure directories if needed
dir.create(fig_indicator, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_predictor, recursive = TRUE, showWarnings = FALSE)
```

```{r}
#| label: data-view-profile-raw
#| echo: false
# View first 6 rows of profiles — continuous indicators get an "(M)" suffix
# For display only; profile_raw itself keeps the plain labels everything
# Downstream matches on
gt(profile_raw %>%
     mutate(Indicators = if_else(Indicators %in% indicator_continuous_fig,
                                  paste0(Indicators, " (M)"),
                                  Indicators))) |>
  tab_header(title = "indicator probabilities/means by class")|>
  fmt_number(columns = where(is.numeric), decimals = 2)
```

### Wrangle Profile Data

Binary indicators (probabilities, 0–1) and continuous indicators (means,
0–3) are plotted together on a shared primary axis. Continuous values
are divided by `scale_factor` before plotting so they fit the 0–1
primary axis; the right-hand secondary axis reverses that transformation
to show the true 0–3 scale.

```{r}
#| label: wrangle-profile

# Legend labels with cluster size in parentheses
cluster_sizes <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

cluster_labels <- c(
  "C1"      = paste0("C1 (", cluster_sizes["C1"], "%)"),
  "C2"      = paste0("C2 (", cluster_sizes["C2"], "%)"),
  "C3"      = paste0("C3 (", cluster_sizes["C3"], "%)"),
  "C4"      = paste0("C4 (", cluster_sizes["C4"], "%)"))

# Fill colors — no outlines
indicator_colors <- cluster_colors

# Scale factor: right axis max / left axis max (continuous max ~3, binary max 1)
# Continuous values are divided by this before plotting, then right axis
# Multiplies back to show true scale. Edit here if any continuous value > 3
scale_factor <- 3

# Label indicators
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# Continuous indicators — values can exceed 1 and read off the right axis
continuous_indicators <- c(
  "Problems from alcohol",
  "Sport Teams",
  "Extracurriculars",
  "Don't avoid problems")

# Negative panel: binary on left (0-1), continuous on right (0-3)
# → Continuous values divided by scale_factor to fit primary axis
# Positive panel: continuous on left (0-3), binary on right (0-1)
# → Binary values multiplied by scale_factor to fit primary axis
# This arrangement puts the 0-3 axes adjacent in the middle of the combined figure

neg_continuous <- c("Alcohol Problems")
pos_continuous <- c("Sport Teams", "School Activities", "Problem Approach")

profile_long <- profile_raw %>%
  pivot_longer(
    cols      = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to  = "cluster_raw",
    values_to = "value"
  ) %>%
  mutate(
    cluster    = recode(cluster_raw,
                   "Cluster1" = "C1", "Cluster2" = "C2",
                   "Cluster3" = "C3", "Cluster4" = "C4"),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    Indicators = factor(Indicators, levels = unique(Indicators)),
    # Rescale for plotting:
    # Negative: continuous / scale_factor (plots on 0-1 axis, reads off right 0-3)
    # Positive: binary * scale_factor     (plots on 0-3 axis, reads off right 0-1)
    value_plot = case_when(
      Risk == "Negative" & Indicators %in% neg_continuous ~ value / scale_factor,
      Risk == "Positive" & !Indicators %in% pos_continuous ~ value * scale_factor,
      TRUE ~ value ) )
```

### BH Correction on Pairwise Comparisons

Non-estimable rows (p = ".") and the Intercept block (included in
LatentGOLD data export) are dropped before correction is applied across
all remaining tests simultaneously.

```{r}
#| label: bh-pairs

# Apply Benjamini-Hochberg correction to the pairwise profile comparisons before flagging significant differences
pairs <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(clustera),
    cluster_b = as.integer(clusterb),
    Indicator = recode(Indicator, !!!indicator_label_map),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05)
```

Benjamini-Hochberg correction is applied here because every indicator is
compared across all six class-pairs simultaneously; without correction,
running this many pairwise tests inflates the false-positive rate well
past .05 ([@benjamini1995], [@benjamini2000]). Overall, the negative
risk indicators contributed about equally to differentiating classes,
while positive risk indicators were the primary driver of separation
within each alcohol-problem grouping.

```{r}
#| label: data-view-pairs-bh

# Display the corrected pairwise-comparison results used to annotate the profile figures
gt(head(pairs)) |> tab_header(title = "After: BH-corrected pairwise comparisons")
```

### Profile Figure

```{r}
#| label: indicator-profilecreate

# Builds axis labels as plotmath expressions so only the "M" renders in
# Italics (the label text and parentheses stay plain). Named by the label
# Text itself so ggplot matches each break by name — not position — which
# Matters since p_neg/p_pos (and the line/z-score panels) each only show a
# Subset of the full indicator list
label_with_M <- function(labels_vec) {
  exprs <- lapply(labels_vec, function(lbl) {
    if (lbl %in% indicator_continuous_fig) {
      bquote(.(lbl) * " (" * italic(M) * ")")
    } else {
      bquote(.(lbl))
    }
  })
  out <- as.expression(exprs)
  names(out) <- labels_vec
  out
}

axis_labels_fig <- label_with_M(levels(factor(profile_long$Indicators)))

# Make_panel() takes explicit axis configuration so negative and positive can differ

make_panel <- function(risk_label,
                       left_name = "", left_lim, left_breaks,
                       sec_transform, right_name = "", right_breaks,
                       caption_text) {
  
  df <- profile_long %>% filter(Risk == risk_label)
  
  # Create panel
  ggplot(df, aes(x = Indicators, y = value_plot, fill = cluster)) +
    geom_vline(
      xintercept = seq(1.5, length(unique(df$Indicators)) - 0.5, by = 1),
      color = "#DDDDDD",
      linewidth = 0.4) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width = bar_width) +
    scale_fill_manual(
      values = indicator_colors,
      labels = cluster_labels,
      name = NULL) +
    scale_x_discrete(labels = axis_labels_fig) +
    scale_y_continuous(
      limits = left_lim,
      breaks = left_breaks,
      labels = number_format(accuracy = 0.1),
      expand = expansion(mult = c(0, 0.03)),
      sec.axis = sec_axis(
        transform = sec_transform,
        name = right_name,
        breaks = right_breaks,
        labels = number_format(accuracy = 0.1))) +
    labs(
      title = paste(risk_label, "Risk Indicators"),
      y = NULL,
      caption = str_wrap(caption_text, width = 150)) +
    base_theme +
    theme(
      # Adjust panel format from base theme
      axis.text.x = element_text(
        angle = 35,
        hjust = 1,
        size = axis_text_size),
      
      axis.title.y.right = element_text(
        size = axis_title_size,
        color = "gray40"),
      
      legend.position = "none",

      plot.caption = element_text(
        hjust = 0,
        lineheight = 1.2,
        size = 16),

      plot.title = element_text(
        size = title_size,
        face = "bold",
        hjust = 0.5))
}

# Negative: left = binary 0-1, right = continuous 0-3
p_neg <- make_panel(
  risk_label = "Negative",
  left_lim = c(0, 1.08),
  left_breaks = seq(0, 1, by = 0.2),
  sec_transform = ~ . * scale_factor,
  right_breaks = seq(0, scale_factor, by = 1),
  caption_text = paste(
    "Note. Left y-axis (0–1) = binary indicators (probability of behavior).",
    "Right y-axis (0–3) = continuous indicators (within-class mean).",
    "Continuous bars are rescaled to the primary axis for display."))

# Positive: left = continuous 0-3, right = binary 0-1
p_pos <- make_panel(
  risk_label = "Positive",
  left_lim = c(0, scale_factor * 1.08),
  left_breaks = seq(0, scale_factor, by = 1),
  sec_transform = ~ . / scale_factor,
  right_breaks = seq(0, 1, by = 0.2),
  caption_text = paste(
    "Note. Left y-axis (0–3) = continuous indicators (within-class mean).",
    "Right y-axis (0–1) = binary indicators (probability of behavior).",
    "Binary bars are rescaled to the primary axis for display."))
```

```{r}
#| label: fig-indicator-profileneg
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the negative-risk indicator profile figure created in the preceding plotting chunk
print(p_neg)
```

```{r}
#| label: fig-indicator-profilepos
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the positive-risk indicator profile figure created in the preceding plotting chunk
print(p_pos)
```

```{r}
#| label: indicator-profilesave
#| include: false

# Save the completed indicator-profile figures to the project figure directory
ggsave(
  filename = file.path(fig_indicator, "profile_figure_neg.png"),
     device = agg_png,
  plot     = p_neg,
  width    = 18, height = 8, dpi = fig_dpi, units = "in")

ggsave(
  filename = file.path(fig_indicator, "profile_figure_pos.png"),
  plot     = p_pos,    
  device = agg_png,
  width    = 18, height = 8, dpi = fig_dpi, units = "in")
```

Sports participation, school activity involvement, and friend disclosure
most clearly distinguish C1 (High PR–Low NR) and C4 (High PR–High NR)
from C2 and C3, while alcohol problems and the broader negative-risk
indicators most clearly separate the two alcohol-problem classes (C3,
C4) from the two non-problem classes (C1, C2). The one indicator with
limited discriminating power across all four classes was problem
approach (class means ranged narrowly from 2.80–2.95, with only 4 of 6
pairwise comparisons significant).

```{r}
#| label: tbl-nonsig
#| warning: false

# Display indicators with no BH-corrected pairwise class differences
nonsig <- pairs %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant pairwise comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

### Significant Comparisons by Profile

```{r}
#| label: tbl-sig-cluster

# Display significant BH-corrected comparisons organized by class pair
sig_pairs <- pairs %>% filter(sig == TRUE)

by_cluster <- bind_rows(
  sig_pairs %>% select(cluster = cluster_a, Indicator),
  sig_pairs %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant pairwise comparisons (BH-corrected, FDR < .05) involving each profile.")
```

### Significant Comparisons by Indicator

```{r}
#| label: tbl-sig-indicator
#| warning: false

# Display significant BH-corrected comparisons organized by indicator
by_indicator <- pairs %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```

# Indicator Profiles - Standardized

### Standardized Profile Panels

I also standardized based on the within class average for each
indicator, allowing for a visualization of how each profile compares to
the mean. The code used to create the figures is below.

Raw indicator profiles above mix binary probabilities and continuous
means on different scales; z-scoring every indicator here puts all of
them on the same standardized axis so their relative separation across
classes can be compared directly, which the raw panels cannot show on
their own.

### Variable Labels

```{r}
#| label: variable-labels

# Variable labels for plots
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# "Negative" = negative risk indicators; "Positive" = positive risk indicators
indicator_risk_panel <- c(
  WSDalc               = "Negative",
  cigwhole             = "Negative",
  skipschool           = "Negative",
  liewhere             = "Negative",
  awayhome             = "Negative",
  dare_dum             = "Negative",
  seatbelt             = "Negative",
  drvdrunk             = "Negative",
  BC_dum               = "Negative",
  steal12M             = "Negative",
  phyfight             = "Negative",
  WSDsport             = "Positive",
  WSDactivity          = "Positive",
  avoidproblems        = "Positive",
  disclose             = "Positive",
  intitaterelationship = "Positive")

# Desired x-axis order within each panel — edit to reorder bars
neg_order <- c(
  "WSDalc", "cigwhole", "skipschool", "liewhere", "awayhome",
  "dare_dum", "seatbelt", "drvdrunk", "BC_dum", "steal12M", "phyfight")

pos_order <- c(
  "WSDsport", "WSDactivity", "avoidproblems", "disclose", "intitaterelationship")
```

### Standardization

```{r}
#| label: import-zscore
#| include: false

# Import standardized indicator values used to compare within-class patterns on a common scale
sav_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/LCAwASP_4CProbabilities.sav"

sav_raw <- read_sav(sav_path)
```

Class probabilities and indicator values are read in from the
dissertation's saved SPSS output (a 4-class LCA solution with an
auxiliary ASP scale). Each indicator is then z-scored so the plotted
classes are on a common scale.

```{r}
#| label: standardize

# Variable names present in the indicator map
indicator_vars <- names(indicator_label_map)

# Standardize each indicator across the full sample
sav_scaled <- sav_raw %>%
  mutate(across(all_of(indicator_vars), ~ as.numeric(scale(.x))))

# Compute within-class means of standardized indicators
zscore_long <- sav_scaled %>%
  select(`clu#`, all_of(indicator_vars)) %>%
  pivot_longer(
    cols = all_of(indicator_vars),
    names_to = "variable",
    values_to = "z_score") %>%
  group_by(`clu#`, variable) %>%
  summarise(
    mean_z = mean(z_score, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(
    cluster = paste0("C", `clu#`),
    cluster = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    label = indicator_label_map[variable],
    Risk = indicator_risk_panel[variable],
    label = factor(
      label,
      levels = indicator_label_map[c(neg_order, pos_order)]))
```

### Legend Labels

```{r}
#| label: legend-labels

# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Edit the description for each cluster here
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR")

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions, " (", cluster_sizes_z, "%)"),
  names(cluster_sizes_z))

cluster_colors <- c(
  "C1" = "#44AA99",
  "C2" = "#88CCEE",   
  "C3" = "#882255",
  "C4" = "#DDCC77")

zscore_colors <- cluster_colors
```

### Z-Score Profile Panels

A horizontal reference line at z = 0 marks the sample average so it is
immediately clear which profiles fall above vs. below the mean on each
indicator.

```{r}
#| label: make-zscore-panel

# Function to create panel for profiles
make_zscore_panel <- function(risk_label) {

  # Set x-axis variable order
  panel_order <- if (risk_label == "Negative") {
    indicator_label_map[neg_order]
  } else {
    indicator_label_map[pos_order]
  }

  # Select columns from dataset
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order))
  
  # Recode for graph
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order)) %>%
    mutate(mean_z = case_when(
      mean_z >  1.0 ~  1.1,
      mean_z < -0.5 ~ -0.6,
      TRUE          ~  mean_z))

  # Create plot
  ggplot(panel_data, aes(x = label, y = mean_z, fill = cluster)) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.6, linetype = "solid") +
   geom_vline(
    xintercept = seq(1.5, length(panel_order) - 0.5, by = 1),
    color      = "gray63",
    linewidth  = .8) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width) +
    scale_fill_manual(
      values = zscore_colors,
      labels = cluster_labels_z,
      name   = NULL ) +
    scale_x_discrete(labels = label_with_M(panel_order)) +
    scale_y_continuous(
      limits = c(-.6, 1.1),
      breaks = c(-0.5,  0, 0.5, 1.0),
  labels = c("Below -0.5", "0.0", "0.5", "Above 1.0"),
      expand = expansion(mult = c(0.02, 0.02))) + 
    labs(
      title   = paste(risk_label, "Risk-Taking Indicators"),
      y       = "Mean (Standardized)",
      caption = str_wrap("Mean standardized (z-scored) indicator values by latent class. Values below −0.5 or above 1.0 are capped for display.", width = 150)) +
    base_theme +
    theme(
      axis.text.x     = element_text(angle = 30, hjust = 1,
                                     size = axis_text_size, family = plot_font),
       plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16),
      legend.position = "none")
}

# Run function for each indicator set
p_zneg <- make_zscore_panel("Negative")
p_zpos <- make_zscore_panel("Positive")
```

```{r}
#| label: fig-standardized-neg
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the standardized negative-risk indicator profile figure
print(p_zneg)  
```

```{r}
#| label: fig-standardized-pos
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the standardized positive-risk indicator profile figure
print(p_zpos)
```

```{r}
#| label: save-panel-standardized
#| include: false

# Save negative panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_neg.png"),
  device   = agg_png,
  plot     = p_zneg,
  width    = 18, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")

# Save positive panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_pos.png"),
  device   = agg_png,
  plot     = p_zpos,
  width    = 18, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")
```

# Predictor Profiles

#### Overview

Predictor profiles show where the four classes fall on the variables
being tested as correlates of class membership, biological sex, age,
peer substance use, pubertal timing, and autonomy-supportive parenting
(ASP).

- Biological sex was a moderately discriminating covariate (4 of 6
  pairwise comparisons significant),
- age, pubertal timing, and ASP each discriminated broadly (5 of 6
  significant), and
- peer substance use discriminated across all 6 comparisons and was the
  single most consistent predictor of profile membership.

```{r}
#| label: import-predictor
#| include: false

# Import class-level predictor summaries used for the descriptive predictor profiles
pred_raw <- read_excel(data_path, sheet = "predictor profiles R")
```

```{r}
#| label: data-view-pred-raw

# Display the raw predictor-profile import before reshaping
gt(head(pred_raw, 6)) |> tab_header(title = "predictor profiles R — predictor means/proportions by class")
```

#### Build Dataframes

Standardized profile dataframe (z-scored predictors + % male).

```{r}
#| label: build-profile-predictor

# Reshape and label the predictor-profile data for class-level plotting
pred_order <- c("Zmale", "Zwave1age", "Zpeer_subuse", "Zub_timing", "Zasp")

display_labels <- c(
  "Zmale"        = "% Male",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer SU",
  "Zub_timing"   = "Puberty",
  "Zasp"         = "ASP")

profile_df <- pred_raw %>%
  filter(Covariate %in% pred_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value") %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = pred_order),
    display   = factor(display_labels[as.character(Covariate)],
                       levels = display_labels[pred_order]))
```

```{r}
#| label: data-view-pred-long

# Display the cleaned long-format predictor data used in the profile figures
gt(head(profile_df, 8)) |> tab_header(title = "After: long format, one row per predictor x cluster")
```

Unstandardized dataframe (raw means per cluster).

```{r}
#| label: build-unstandardized

# Build the unstandardized predictor-profile plot for direct interpretation in original units
unstand_order <- c("male", "wave1age", "peer_subuse", "pub_timing", "asp")

unstand_df <- pred_raw %>%
  filter(Covariate %in% unstand_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value" ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = unstand_order) )
```

Age distribution dataframe. Oldest group (18–19) stacks at the top.

```{r}
#| label: build-age

# Prepare age-group distributions across the four latent profiles
age_order <- c("age_13", "age_14", "age_16", "age_17", "age_18")

age_df <- pred_raw %>%
  filter(Covariate %in% age_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "proportion" ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    age_group = factor(label, levels = c(
      "18-19 years old", "17 years old", "16 years old",
      "14-15 years old", "13 years old"  )))
```

#### Predictor Profile Graph (Standardized)

```{r}
#| label: profile-standardized
# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Create the description for each cluster
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR")

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions),
  names(cluster_sizes_z))

# Colors — pulls from cluster_colors defined in _setup.R
zscore_colors <- cluster_colors

# Create predictor profile plot
p_profile <- ggplot(profile_df, aes(x = display, y = value, fill = cluster)) +
  geom_vline(
    xintercept = seq(1.5, length(profile_df) - 0.5, by = 1),
    color      = "#DDDDDD",
    linewidth  = 0.4) +
  geom_col(position = position_dodge(width = dodge_width), width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
 scale_fill_manual(values = cluster_colors, labels = cluster_labels_z) +
  scale_y_continuous(
    name   = "Mean (Standardized)",
    limits = c(-.5, 1.15),
    breaks = seq(-.5, 1, by = .5),
    labels = number_format(accuracy = 0.01)) +  
  labs(
    title   = "ASP and Covariate Levels Across Profiles",
    caption = str_wrap("Note. Age, Peer Substance Use, Pubertal Timing, and Autonomy Supportive Parenting were standardized (M = 0, SD = 1). Sex dummy-coded (0 = male, 1 = female); % Male shown as raw proportion.", width = 150)) +
  base_theme +
  theme(legend.position = "top",    
        plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16)) +
  guides(fill = guide_legend(nrow = 1))
```

```{r}
#| label: fig-profile-standardized
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the standardized predictor-profile figure
print(p_profile)
```

```{r}
#| label: profile-standardized-save
#| include: false

# Save the completed standardized predictor-profile figure
ggsave(file.path(fig_predictor, "Predictor_Profile_Standardized.png"),
       plot = p_profile, width = 18, height = 8,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```

Older adolescents and adolescents with earlier pubertal timing were
overrepresented in the two alcohol-problem classes (C3, C4), consistent
with developmental increases in negative risk-taking across adolescence
([@duell2018], [@moffitt2017]) and with the early-timing hypothesis
linking earlier physical maturation to elevated risk propensity
([@natsuaki2009]). Male adolescents were more likely to fall in the
alcohol-problem classes, consistent with prior findings of higher
delinquency rates among males [@miller2010]. Peer substance use was the
single strongest and most consistent predictor, the only covariate with
all pairwise comparisons against both C1 and C3 significant, matching
developmental literature identifying peer behavior as a robust predictor
of adolescent delinquency ([@dishion2011; @dishion1999;
@henneberger2021]).

#### Age Distribution

Youngest (13) at bottom; oldest (18–19) at top.

```{r}
#| label: plot-age-distribution

# Create age distribution plot
p_age <- ggplot(age_df, aes(x = cluster, y = proportion, fill = age_group)) +
  geom_col(position = "stack", width = bar_width) +
  scale_fill_manual(
    values = age_colors, name = "Age Group",
    breaks = c("13 years old","14-15 years old","16 years old",
               "17 years old","18-19 years old") ) +
  scale_y_continuous(
    name   = "Proportion",
    limits = c(0, 1.01), expand = c(0, 0),
    breaks = seq(0, 1, by = 0.25),
    labels = number_format(accuracy = 0.01)) +
  labs(
    title   = "Age Distribution by Cluster",
    caption = str_wrap(
      "Note. Stacked bars show proportion within each age group per cluster (sum to 1.0). Ordered youngest (bottom) to oldest (top).",
      width = 150) ) +
    or_theme +
  theme(
    axis.title.x    = element_blank(),
    legend.position = "bottom" , 
    legend.text     = element_text(size = 14),
    legend.title    = element_text(size = 15, face = "bold"),
    legend.key.size = unit(1.1, "lines"),
     plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16)) +
  guides(fill = guide_legend(nrow =1))
```

```{r}
#| label: fig-age-distribution
#| fig-width: 18
#| fig-height: 8
#| echo: false

# Render the age-distribution figure across latent profiles
print(p_age)
```

```{r}
#| label: age-plot-save
#| include: false

# Save the completed age-distribution figure
ggsave(file.path(fig_predictor, "Age_Distribution.png"),
       plot = p_age, width = 18, height = 8,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```
