---
title: "Indicator, Predictor & ASP Profiles"
toc: true
toc-depth: 3
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
source("_setup.R")
library(haven)   # read_sav, used in the z-score section
library(gt)      # used for the data-view tables added in this consolidation
library(stringr)   # str_wrap(), used to wrap figure captions
```

# Overview

This page covers the descriptive class profiles from the 4-class latent
profile analysis of Add Health Wave I adolescents (*N* = 6,469): the
negative and positive risk-taking indicators that defined the classes,
the demographic and parenting predictors tested as correlates of class
membership, and the parenting-scale-level breakdown of the
autonomy-supportive parenting (ASP) composite used throughout this
project. The inferential odds-ratio work built on these profiles lives
on the [Odds Ratios](Odds-Ratios.qmd) page; effect-modification and
BH-corrected robustness checks live on [Interaction Effects &
Comparisons](Interaction-and-Comparisons.qmd).

# Indicator Profiles

### Overview

These panels show the 4-class latent profile solution's indicator
profiles from the negative and positive risk-taking indicators. Fit
indices favored the four-profile solution: AIC/BIC improvements became
marginal beyond four classes, the five-profile solution failed to
replicate (a local maximum), and the four-profile solution showed
adequate entropy (\> .80) with theoretically coherent classes (LPA
conducted in LatentGOLD; [@vermunt2005]).

::: {.callout-tip appearance="minimal" icon="false"}
**PR** = **p**ositive **r**isk-taking; **NR** = **n**egative
**r**isk-taking
:::

The four profiles are:

-   C1 (High PR–Low NR, 35.6%)
-   C2 (Low PR–Low NR, 33.7%)
-   C3 (Low PR–High NR, 15.8%)
-   C4 (High PR–High NR, 14.9%)

Alcohol problems functioned as a near-binary separator: both **C1 and C2
reported essentially no alcohol problems**, while **C3 and C4 reported
substantially elevated problem counts**. Within each of those two broad
groupings, positive risk indicators (sports, school activities, friend
disclosure, romantic involvement) were the primary source of further
differentiation. However, adolescents with comparably low negative risk
occupied meaningfully distinct profiles with respect to positive risk
engagement [@duell2019a; @duell2021a; @fryt2021a].

To jump straight to figures:

-   indicator probabilities/mean: negative risk-taking
    (@fig-indicator-profileneg) and (@fig-indicator-profilepos)
-   indicator within-class average: negative (@fig-standardized-neg) and
    positive (@fig-standardized-pos)
-   predictor profiles: @fig-profile-standardized
-   age distribution of each profile: @fig-age-distribution

**Data Import**

Indicator profile probabilities/means and paired-comparison significance
tests across the latent class solution are read in from a local Excel
results workbook from saved LatentGOLD results. Import code is not
rendered.

```{r}
#| label: import-indicator
#| include: false

data_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/Study1_R/data/Tables.xlsx"

profile_raw <- read_excel(data_path, sheet = "profile R")
pairs_raw   <- read_excel(data_path, sheet = "paired comparisons")

# Relabel for visual ease
profile_indicator_recode <- c(
  "Had whole cigarrette"           = "Cigarette",
  "Skip school (no excuse)"        = "Skipped school",
  "Lied about whereabouts"         = "Lied about whereabouts",
  "Night out (no permission)"      = "Night out(unsanctioned)",
  "Done dangerous dare"            = "Dangerous dare",
  "Not always wear seatbelt"       = "No seatbelt",
  "Ever Drive drunk"               = "Drove Drunk",
  "No contraceptives used"         = "Risky Sex",
  "Stole Anything"                 = "Theft",
  "Serious Physical Fight"         = "Fight",
  "Problems from alcohol"          = "Alcohol Problems",
  "Sport Teams"                    = "Sport Teams",
  "Extracurriculars"               = "School Activities",
  "Don't avoid problems"           = "Problem Approach",
  "Disclosed problem to friend"    = "Friend Disclosure",
  "Initiate romantic relationship" = "Initiated Relationship")

profile_raw <- profile_raw %>%
  mutate(Indicators = recode(str_trim(Indicators), !!!profile_indicator_recode))

# Continuous indicators (within-class means, not probabilities/proportions).
# Moved up here (rather than defined down in indicator-profilecreate) so the
# raw data-view table below can also flag them with an "(M)" suffix.
indicator_continuous_fig <- c("Alcohol Problems", "Problem Approach",
                               "Sport Teams", "School Activities")
```

```{r}
#| label: data-view-profile-raw
#| echo: false
# View first 6 rows of profiles — continuous indicators get an "(M)" suffix
# for display only; profile_raw itself keeps the plain labels everything
# downstream matches on.
gt(profile_raw %>%
     mutate(Indicators = if_else(Indicators %in% indicator_continuous_fig,
                                  paste0(Indicators, " (M)"),
                                  Indicators))) |>
  tab_header(title = "indicator probabilities/means by class")|>
  fmt_number(columns = where(is.numeric), decimals = 2)
```

### Wrangle Profile Data

Binary indicators (probabilities, 0–1) and continuous indicators (means,
0–3) are plotted together on a shared primary axis. Continuous values
are divided by `scale_factor` before plotting so they fit the 0–1
primary axis; the right-hand secondary axis reverses that transformation
to show the true 0–3 scale.

```{r}
#| label: wrangle-profile

# Legend labels with cluster size in parentheses
cluster_sizes <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

cluster_labels <- c(
  "C1"      = paste0("C1 (", cluster_sizes["C1"], "%)"),
  "C2"      = paste0("C2 (", cluster_sizes["C2"], "%)"),
  "C3"      = paste0("C3 (", cluster_sizes["C3"], "%)"),
  "C4"      = paste0("C4 (", cluster_sizes["C4"], "%)"))

# Fill colors — no outlines
indicator_colors <- cluster_colors

# Scale factor: right axis max / left axis max (continuous max ~3, binary max 1)
# Continuous values are divided by this before plotting, then right axis
# multiplies back to show true scale. Edit here if any continuous value > 3.
scale_factor <- 3

# Label indicators
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# Continuous indicators — values can exceed 1 and read off the right axis
continuous_indicators <- c(
  "Problems from alcohol",
  "Sport Teams",
  "Extracurriculars",
  "Don't avoid problems")

# Negative panel: binary on left (0-1), continuous on right (0-3)
#   → continuous values divided by scale_factor to fit primary axis
# Positive panel: continuous on left (0-3), binary on right (0-1)
#   → binary values multiplied by scale_factor to fit primary axis
# This arrangement puts the 0-3 axes adjacent in the middle of the combined figure.

neg_continuous <- c("Alcohol Problems")
pos_continuous <- c("Sport Teams", "School Activities", "Problem Approach")

profile_long <- profile_raw %>%
  pivot_longer(
    cols      = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to  = "cluster_raw",
    values_to = "value"
  ) %>%
  mutate(
    cluster    = recode(cluster_raw,
                   "Cluster1" = "C1", "Cluster2" = "C2",
                   "Cluster3" = "C3", "Cluster4" = "C4"),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    Indicators = factor(Indicators, levels = unique(Indicators)),
    # Rescale for plotting:
    #   Negative: continuous / scale_factor (plots on 0-1 axis, reads off right 0-3)
    #   Positive: binary * scale_factor     (plots on 0-3 axis, reads off right 0-1)
    value_plot = case_when(
      Risk == "Negative" & Indicators %in% neg_continuous ~ value / scale_factor,
      Risk == "Positive" & !Indicators %in% pos_continuous ~ value * scale_factor,
      TRUE ~ value ) )
```

### BH Correction on Pairwise Comparisons

Non-estimable rows (p = ".") and the Intercept block (included in
LatentGOLD data export) are dropped before correction is applied across
all remaining tests simultaneously.

```{r}
#| label: bh-pairs
pairs <- pairs_raw %>%
  filter(Indicator != "Intercept", `p-value` != ".") %>%
  mutate(
    p_value   = as.numeric(`p-value`),
    cluster_a = as.integer(clustera),
    cluster_b = as.integer(clusterb),
    Indicator = recode(Indicator, !!!indicator_label_map),
    pair      = paste0("C", cluster_a, " vs. C", cluster_b),
    p_adj     = p.adjust(p_value, method = "BH"),
    sig       = p_adj < .05)
```

Benjamini-Hochberg correction is applied here because every indicator is
compared across all six class-pairs simultaneously; without correction,
running this many pairwise tests inflates the false-positive rate well
past .05 ([@benjamini1995], [@benjamini2000]). Overall, the negative
risk indicators contributed about equally to differentiating classes,
while positive risk indicators were the primary driver of separation
within each alcohol-problem grouping.

```{r}
#| label: data-view-pairs-bh
gt(head(pairs)) |> tab_header(title = "After: BH-corrected pairwise comparisons")
```

### Profile Figure

```{r}
#| label: indicator-profilecreate

# Builds axis labels as plotmath expressions so only the "M" renders in
# italics (the label text and parentheses stay plain). Named by the label
# text itself so ggplot matches each break by name — not position — which
# matters since p_neg/p_pos (and the line/z-score panels) each only show a
# subset of the full indicator list.
label_with_M <- function(labels_vec) {
  exprs <- lapply(labels_vec, function(lbl) {
    if (lbl %in% indicator_continuous_fig) {
      bquote(.(lbl) * " (" * italic(M) * ")")
    } else {
      bquote(.(lbl))
    }
  })
  out <- as.expression(exprs)
  names(out) <- labels_vec
  out
}

axis_labels_fig <- label_with_M(levels(factor(profile_long$Indicators)))

# make_panel() takes explicit axis config so Negative and Positive can differ.

# Negative: left = binary 0-1, right = continuous 0-3
#   primary axis shows 0-1; sec.axis multiplies by scale_factor to show 0-3

# Positive: left = continuous 0-3, right = binary 0-1
#   primary axis shows 0-3; sec.axis divides by scale_factor to show 0-1

# Create a function to make panel
make_panel <- function(risk_label,
                       left_name = "" ,left_lim, left_breaks,
                       sec_transform, right_name = "", right_breaks) {

  df <- profile_long %>% filter(Risk == risk_label)

  # Create panel
  ggplot(df, aes(x = Indicators, y = value_plot, fill = cluster)) +
     geom_vline(
  xintercept = seq(1.5, length(unique(df$Indicators)) - 0.5, by = 1),
  color      = "#DDDDDD",
  linewidth  = 0.4) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width) +
    scale_fill_manual(values = indicator_colors, labels = cluster_labels,
                      name = NULL) + 
    scale_x_discrete(labels = axis_labels_fig) +
    scale_y_continuous(
      limits = left_lim,
      breaks = left_breaks,
      labels = number_format(accuracy = 0.1),
      expand = expansion(mult = c(0, 0.03)),
      sec.axis = sec_axis(
        transform = sec_transform,
        name      = right_name,
        breaks    = right_breaks,
        labels    = number_format(accuracy = 0.1)
     )
    ) +
    labs(title = paste(risk_label, "Risk Indicators"),
         y = NULL, 
         caption = str_wrap(
           "Note. Left y-axis (0\u20131) = binary indicators (probability of behavior).Right y-axis (0\u20133) = continuous indicators (within-class mean). Continuous bars are rescaled to the primary axis for display.", 
           width = 150)) +
    base_theme +
    theme(
      # adjust panel format from base theme
      axis.text.x = element_text(
        angle = 35, hjust = 1,size = axis_text_size),
      
      axis.title.y.right = element_text(
        size = axis_title_size, color = "gray40"),
      
      legend.position = "none",
      
      plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16),
      plot.title  = element_text(
        size = title_size, face = "bold",  hjust = 0.5))
      }

# Negative: left=binary(0-1), right=continuous(0-3)
p_neg <- make_panel(
  risk_label    = "Negative",
  #left_name     = "Probability (binary indicators)",
  left_lim      = c(0, 1.08),
  left_breaks   = seq(0, 1, by = 0.2),
  sec_transform = ~ . * scale_factor,
 # right_name    = "Within-class mean (continuous indicators)",
  right_breaks  = seq(0, scale_factor, by = 1))

# Positive: left=continuous(0-3), right=binary(0-1)
p_pos <- make_panel(
  risk_label    = "Positive",
 # left_name     = "Within-class mean (continuous indicators)",
  left_lim      = c(0, scale_factor * 1.08),
  left_breaks   = seq(0, scale_factor, by = 1),
  sec_transform = ~ . / scale_factor,
 # right_name    = "Probability (binary indicators)",
  right_breaks  = seq(0, 1, by = 0.2))
```

```{r}
#| label: fig-indicator-profileneg
#| fig-width: 18
#| fig-height: 8
#| echo: false
print(p_neg)
```

```{r}
#| label: fig-indicator-profilepos
#| fig-width: 18
#| fig-height: 8
#| echo: false
print(p_pos)
```

```{r}
#| label: indicator-profilesave
#| include: false

ggsave(
  filename = file.path(fig_indicator, "profile_figure_neg.png"),
     device = agg_png,
  plot     = p_neg,
  width    = 18, height = 8, dpi = fig_dpi, units = "in")

ggsave(
  filename = file.path(fig_indicator, "profile_figure_pos.png"),
  plot     = p_pos,    
  device = agg_png,
  width    = 18, height = 8, dpi = fig_dpi, units = "in")
```

Sports participation, school activity involvement, and friend disclosure
most clearly distinguish C1 (High PR–Low NR) and C4 (High PR–High NR)
from C2 and C3, while alcohol problems and the broader negative-risk
indicators most clearly separate the two alcohol-problem classes (C3,
C4) from the two non-problem classes (C1, C2). The one indicator with
limited discriminating power across all four classes was problem
approach (class means ranged narrowly from 2.80–2.95, with only 4 of 6
pairwise comparisons significant).

@tbl-nonsig: List of non-significant pairwise comparisons

```{r}
#| label: tbl-nonsig
#| warning: false
nonsig <- pairs %>%
  filter(sig == FALSE) %>%
  select(Indicator, pair, p_value, p_adj) %>%
  mutate(across(c(p_value, p_adj), ~ round(.x, 3))) %>%
  arrange(Indicator, pair)

kable(nonsig,
      col.names = c("Indicator", "Comparison", "Raw p", "BH-adjusted p"),
      caption   = "Non-significant pairwise comparisons after Benjamini-Hochberg correction (FDR < .05).")
```

### Significant Comparisons by Profile

```{r}
#| label: tbl-sig-cluster
sig_pairs <- pairs %>% filter(sig == TRUE)

by_cluster <- bind_rows(
  sig_pairs %>% select(cluster = cluster_a, Indicator),
  sig_pairs %>% select(cluster = cluster_b, Indicator)
) %>%
  mutate(cluster = paste0("C", cluster)) %>%
  count(cluster, name = "n_significant") %>%
  arrange(cluster)

kable(by_cluster,
      col.names = c("Profile", "Number of Significant Pairwise Comparisons"),
      caption   = "Count of significant pairwise comparisons (BH-corrected, FDR < .05) involving each profile.")
```

### Significant Comparisons by Indicator

```{r}
#| label: tbl-sig-indicator
#| warning: false
by_indicator <- pairs %>%
  group_by(Indicator) %>%
  summarise(
    n_total       = n(),
    n_significant = sum(sig),
    n_nonsig      = sum(!sig)
  ) %>%
  arrange(desc(n_significant))

kable(by_indicator,
      col.names = c("Indicator", "Total Comparisons", "Significant", "Non-Significant"),
      caption   = "Count of significant and non-significant pairwise comparisons (BH-corrected, FDR < .05) by indicator.")
```

# Indicator Profiles - Standardized

### Standardized Profile Panels

I also standardized based on the within class average for each
indicator, allowing for a visualization of how each profile compares to
the mean. The code used to create the figures is below.

Raw indicator profiles above mix binary probabilities and continuous
means on different scales; z-scoring every indicator here puts all of
them on the same standardized axis so their relative separation across
classes can be compared directly, which the raw panels cannot show on
their own.

### Variable Labels

```{r}
#| label: variable-labels

# Variable labels for plots
indicator_label_map <- c(
  WSDalc               = "Alcohol Problems",
  cigwhole             = "Cigarette",
  skipschool           = "Skipped school",
  liewhere             = "Lied about whereabouts",
  awayhome             = "Night out(unsanctioned)",
  dare_dum             = "Dangerous dare",
  seatbelt             = "No seatbelt",
  drvdrunk             = "Drove Drunk",
  BC_dum               = "Risky Sex",
  steal12M             = "Theft",
  phyfight             = "Fight",
  WSDsport             = "Sport Teams",
  WSDactivity          = "School Activities",
  avoidproblems        = "Problem Approach",
  disclose             = "Friend Disclosure",
  intitaterelationship = "Initiated Relationship")

# "Negative" = negative risk indicators; "Positive" = positive risk indicators
indicator_risk_panel <- c(
  WSDalc               = "Negative",
  cigwhole             = "Negative",
  skipschool           = "Negative",
  liewhere             = "Negative",
  awayhome             = "Negative",
  dare_dum             = "Negative",
  seatbelt             = "Negative",
  drvdrunk             = "Negative",
  BC_dum               = "Negative",
  steal12M             = "Negative",
  phyfight             = "Negative",
  WSDsport             = "Positive",
  WSDactivity          = "Positive",
  avoidproblems        = "Positive",
  disclose             = "Positive",
  intitaterelationship = "Positive")

# Desired x-axis order within each panel — edit to reorder bars
neg_order <- c(
  "WSDalc", "cigwhole", "skipschool", "liewhere", "awayhome",
  "dare_dum", "seatbelt", "drvdrunk", "BC_dum", "steal12M", "phyfight")

pos_order <- c(
  "WSDsport", "WSDactivity", "avoidproblems", "disclose", "intitaterelationship")
```

### Standardization

```{r}
#| label: import-zscore
#| include: false

sav_path <- "C:/Users/rcm5550/OneDrive - The Pennsylvania State University/Rachel M & Rina/Dissertation/Analysis/Study 1/LCAwASP_4CProbabilities.sav"

sav_raw <- read_sav(sav_path)
```

Class probabilities and indicator values are read in from the
dissertation's saved SPSS output (a 4-class LCA solution with an
auxiliary ASP scale). Each indicator is then z-scored so the plotted
classes are on a common scale.

```{r}
#| label: standardize

# Variable names present in the indicator map
indicator_vars <- names(indicator_label_map)

# Standardize each indicator across the full sample
sav_scaled <- sav_raw %>%
  mutate(across(all_of(indicator_vars), ~ as.numeric(scale(.x))))

# Compute within-class means of z-scores
# Class variable assumed to be named "class" — update if different in your file
zscore_long <- sav_scaled %>%
  select(`clu#`, all_of(indicator_vars)) %>%
  pivot_longer(
    cols      = all_of(indicator_vars),
    names_to  = "variable",
    values_to = "z_score") %>%
  group_by(`clu#`, variable) %>%
  summarise(mean_z = mean(z_score, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    cluster    = paste0("C", `clu#`),
    cluster    = factor(cluster, levels = c("C1", "C2", "C3", "C4")),
    label      = indicator_label_map[variable],
    Risk       = indicator_risk_panel[variable],
    label      = factor(label, levels = indicator_label_map[c(neg_order, pos_order)]) )
```

### Legend Labels

```{r}
#| label: legend-labels

# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Edit the description for each cluster here
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR")

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions, " (", cluster_sizes_z, "%)"),
  names(cluster_sizes_z))

cluster_colors <- c(
  "C1" = "#44AA99",
  "C2" = "#88CCEE",   
  "C3" = "#882255",
  "C4" = "#DDCC77")

zscore_colors <- cluster_colors
```

### Z-Score Profile Panels

A horizontal reference line at z = 0 marks the sample average so it is
immediately clear which profiles fall above vs. below the mean on each
indicator.

```{r}
#| label: make-zscore-panel

# Function to create panel for profiles
make_zscore_panel <- function(risk_label) {

  # Set x-axis variable order
  panel_order <- if (risk_label == "Negative") {
    indicator_label_map[neg_order]
  } else {
    indicator_label_map[pos_order]
  }

  # Select columns from dataset
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order))
  
  # Recode for graph 
  panel_data <- zscore_long %>%
    filter(Risk == risk_label) %>%
    mutate(label = factor(label, levels = panel_order)) %>%
    mutate(mean_z = case_when(
      mean_z >  1.0 ~  1.1,
      mean_z < -0.5 ~ -0.6,
      TRUE          ~  mean_z))

  # Create plot
  ggplot(panel_data, aes(x = label, y = mean_z, fill = cluster)) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.6, linetype = "solid") +
   geom_vline(
    xintercept = seq(1.5, length(panel_order) - 0.5, by = 1),
    color      = "gray63",
    linewidth  = .8) +
    geom_col(
      position = position_dodge(width = dodge_width),
      width    = bar_width) +
    scale_fill_manual(
      values = zscore_colors,
      labels = cluster_labels_z,
      name   = NULL ) +
    scale_x_discrete(labels = label_with_M(panel_order)) +
    scale_y_continuous(
      limits = c(-.6, 1.1),
      breaks = c(-0.5,  0, 0.5, 1.0),
  labels = c("Below -0.5", "0.0", "0.5", "Above 1.0"),
      expand = expansion(mult = c(0.02, 0.02))) + 
    labs(
      title   = paste(risk_label, "Risk-Taking Indicators"),
      y       = "Mean (Standardized)",
      caption = str_wrap("Mean standardized (z-scored) indicator values by latent class. Values below −0.5 or above 1.0 are capped for display.", width = 150)) +
    base_theme +
    theme(
      axis.text.x     = element_text(angle = 30, hjust = 1,
                                     size = axis_text_size, family = plot_font),
       plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16),
      legend.position = "none")
}

# Run function for each indicator set
p_zneg <- make_zscore_panel("Negative")
p_zpos <- make_zscore_panel("Positive")
```

```{r}
#| label: fig-standardized-neg
#| fig-width: 18
#| fig-height: 8
#| echo: false
print(p_zneg)  
```

```{r}
#| label: fig-standardized-pos
#| fig-width: 18
#| fig-height: 8
#| echo: false

print(p_zpos)
```

```{r}
#| label: save-panel-standardized
#| include: false

# Save negative panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_neg.png"),
  device   = agg_png,
  plot     = p_zneg,
  width    = 18, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")

# Save positive panel
ggsave(
  filename = file.path(fig_indicator, "zscore_profile_pos.png"),
  device   = agg_png,
  plot     = p_zpos,
  width    = 18, height = 8, dpi = fig_dpi, units = "in",
  bg       = "transparent")
```

# Predictor Profiles

#### Overview

Predictor profiles show where the four classes fall on the variables
being tested as correlates of class membership, biological sex, age,
peer substance use, pubertal timing, and autonomy-supportive parenting
(ASP). Biological sex was a moderately discriminating covariate (4 of 6
pairwise comparisons significant); age, pubertal timing, and ASP each
discriminated broadly (5 of 6 significant); peer substance use
discriminated across all 6 comparisons and was the single most
consistent predictor of profile membership.

```{r}
#| label: import-predictor
#| include: false
pred_raw <- read_excel(data_path, sheet = "predictor profiles R")
```

```{r}
#| label: data-view-pred-raw
gt(head(pred_raw, 6)) |> tab_header(title = "predictor profiles R — predictor means/proportions by class")
```

#### Build Dataframes

Standardized profile dataframe (z-scored predictors + % male).

```{r}
#| label: build-profile-predictor
pred_order <- c("Zmale", "Zwave1age", "Zpeer_subuse", "Zub_timing", "Zasp")

display_labels <- c(
  "Zmale"        = "% Male",
  "Zwave1age"    = "Age",
  "Zpeer_subuse" = "Peer SU",
  "Zub_timing"   = "Puberty",
  "Zasp"         = "ASP")

profile_df <- pred_raw %>%
  filter(Covariate %in% pred_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value") %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = pred_order),
    display   = factor(display_labels[as.character(Covariate)],
                       levels = display_labels[pred_order]))
```

```{r}
#| label: data-view-pred-long
gt(head(profile_df, 8)) |> tab_header(title = "After: long format, one row per predictor x cluster")
```

Unstandardized dataframe (raw means per cluster).

```{r}
#| label: build-unstandardized
unstand_order <- c("male", "wave1age", "peer_subuse", "pub_timing", "asp")

unstand_df <- pred_raw %>%
  filter(Covariate %in% unstand_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "value" ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    Covariate = factor(Covariate, levels = unstand_order) )
```

Age distribution dataframe. Oldest group (18–19) stacks at the top.

```{r}
#| label: build-age
age_order <- c("age_13", "age_14", "age_16", "age_17", "age_18")

age_df <- pred_raw %>%
  filter(Covariate %in% age_order) %>%
  select(-Overall) %>%
  pivot_longer(
    cols = c(Cluster1, Cluster2, Cluster3, Cluster4),
    names_to = "cluster", values_to = "proportion" ) %>%
  mutate(
    cluster   = factor(recode(cluster,
                   Cluster1 = "C1", Cluster2 = "C2",
                   Cluster3 = "C3", Cluster4 = "C4"),
                 levels = c("C1","C2","C3","C4")),
    age_group = factor(label, levels = c(
      "18-19 years old", "17 years old", "16 years old",
      "14-15 years old", "13 years old"  )))
```

#### Predictor Profile Graph (Standardized)

```{r}
#| label: profile-standardized
# Cluster sizes
cluster_sizes_z <- c("C1" = 35.6, "C2" = 33.7, "C3" = 15.8, "C4" = 14.9)

# Create the description for each cluster
cluster_descriptions <- c(
  "C1" = "High PR - Low NR",
  "C2" = "Low PR - Low NR",
  "C3" = "Low PR - High NR",
  "C4" = "High PR - High NR")

cluster_labels_z <- setNames(
  paste0(names(cluster_sizes_z), ": ", cluster_descriptions),
  names(cluster_sizes_z))

# Colors — pulls from cluster_colors defined in _setup.R
zscore_colors <- cluster_colors

# Create predictor profile plot
p_profile <- ggplot(profile_df, aes(x = display, y = value, fill = cluster)) +
  geom_vline(
    xintercept = seq(1.5, length(profile_df) - 0.5, by = 1),
    color      = "#DDDDDD",
    linewidth  = 0.4) +
  geom_col(position = position_dodge(width = dodge_width), width = bar_width) +
  geom_hline(yintercept = 0, color = ref_line_color,
             linewidth = ref_line_lwd, linetype = ref_line_type) +
 scale_fill_manual(values = cluster_colors, labels = cluster_labels_z) +
  scale_y_continuous(
    name   = "Mean (Standardized)",
    limits = c(-.5, 1.15),
    breaks = seq(-.5, 1, by = .5),
    labels = number_format(accuracy = 0.01)) +  
  labs(
    title   = "ASP and Covariate Levels Across Profiles",
    caption = str_wrap("Note. Age, Peer Substance Use, Pubertal Timing, and Autonomy Supportive Parenting were standardized (M = 0, SD = 1). Sex dummy-coded (0 = male, 1 = female); % Male shown as raw proportion.", width = 150)) +
  base_theme +
  theme(legend.position = "top",    
        plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16)) +
  guides(fill = guide_legend(nrow = 1))
```

```{r}
#| label: fig-profile-standardized
#| fig-width: 18
#| fig-height: 8
#| echo: false

print(p_profile)
```

```{r}
#| label: profile-standardized-save
#| include: false

ggsave(file.path(fig_predictor, "Predictor_Profile_Standardized.png"),
       plot = p_profile, width = 18, height = 8,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```

Older adolescents and adolescents with earlier pubertal timing were
overrepresented in the two alcohol-problem classes (C3, C4), consistent
with developmental increases in negative risk-taking across adolescence
([@duell2018], [@moffitt2017]) and with the early-timing hypothesis
linking earlier physical maturation to elevated risk propensity
([@natsuaki2009]). Male adolescents were more likely to fall in the
alcohol-problem classes, consistent with prior findings of higher
delinquency rates among males [@miller2010]. Peer substance use was the
single strongest and most consistent predictor, the only covariate with
all pairwise comparisons against both C1 and C3 significant, matching
developmental literature identifying peer behavior as a robust predictor
of adolescent delinquency ([@dishion2011; @dishion1999;
@henneberger2021]).

#### Age Distribution

Youngest (13) at bottom; oldest (18–19) at top.

```{r}
#| label: plot-age-distribution

# Create age distribution plot
p_age <- ggplot(age_df, aes(x = cluster, y = proportion, fill = age_group)) +
  geom_col(position = "stack", width = bar_width) +
  scale_fill_manual(
    values = age_colors, name = "Age Group",
    breaks = c("13 years old","14-15 years old","16 years old",
               "17 years old","18-19 years old") ) +
  scale_y_continuous(
    name   = "Proportion",
    limits = c(0, 1.01), expand = c(0, 0),
    breaks = seq(0, 1, by = 0.25),
    labels = number_format(accuracy = 0.01)) +
  labs(
    title   = "Age Distribution by Cluster",
    caption = str_wrap(
      "Note. Stacked bars show proportion within each age group per cluster (sum to 1.0). Ordered youngest (bottom) to oldest (top).",
      width = 150) ) +
    or_theme +
  theme(
    axis.title.x    = element_blank(),
    legend.position = "bottom" , 
    legend.text     = element_text(size = 14),
    legend.title    = element_text(size = 15, face = "bold"),
    legend.key.size = unit(1.1, "lines"),
     plot.caption = element_text(
                      hjust = 0,
                      lineheight = 1.2,
                      size   = 16)) +
  guides(fill = guide_legend(nrow =1))
```

```{r}
#| label: fig-age-distribution
#| fig-width: 18
#| fig-height: 8
#| echo: false
print(p_age)
```

```{r}
#| label: age-plot-save
#| include: false
ggsave(file.path(fig_predictor, "Age_Distribution.png"),
       plot = p_age, width = 18, height = 8,
       dpi = fig_dpi, units = "in",
       bg = "transparent")
```
